### 2014-10-30 2014/nature-review/Deep-Learning-Nature-Review.tex

\documentclass{article} % For LaTeX2e
\usepackage{times}
\usepackage{url}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{tabularx}
\usepackage{algorithm}
\usepackage{algorithmic}

\title{Deep Learning \\ NATURE Review Article}

\author{Target: 4500 words, 4 main figures and less than 100 citations (8 pages total)}

\begin{document}

%\maketitle

\begin{abstract}
Deep learning is currently the most exciting area of machine learning
because it has dramatically improved the state-of-the-art in speech
recognition, object recognition and object detection.  It is also being
remarkably successful at many other tasks such as language modeling
(predicting the next word in a sentence) and predicting the activity of
potential drug molecules.  In addition to its remarkable successes for
perceptual tasks, deep learning is currently on the verge of beating the
state-of-the-art in Machine Translation and is likely to become the method
of choice for this task over the next few years.
There is widespread interest in deep learning because it is so good at
extracting structure from very large datasets and it appears to be very
widely applicable. For example, it was named as one of the top ten new
technologies of 2013 by MIT Technology Review.
\end{abstract}

\begin{verbatim}
We plan to organize this review chronologically, but with forward looking
 comments. We will start with the excitement over the backpropagation
 learning algorithm for multilayer neural networks in the 1980's and
 explain why these networks did not live up to the very high expectations
 that people had for them.  Given our current knowledge, the real reasons
 for the disappointments of the 1990's are now clear: people radically
 underestimated the amount of data and the amount of computation that were
 required to demonstrate the power of this type of learning algorithm.
\end{verbatim}

\section{DISTRIBUTED REPRESENTATIONS AND DEEP LEARNING}

Traditional approaches to pattern recognition required engineers to design
a feature extractor that turns raw sensory data into representations that
are suitable for conventional machine learning algorithms. Deep learning
algorithms aim to replace these hand-crafted representations by learned
ones, taking advantage of the information implicitly available in
data. Deep learning research aims at discovering mechanisms for learning
multiple levels of representation of the data, with higher levels of
representation representing more abstract features that simplify further
processing, e.g., for tasks such as object recognition, speech recognition,
or translation from one language to another.

The kind of representations extracted by deep learning have two central
characteristics: they are distributed representations and they are deep
representations. A distributed representation is composed of features that
are not mutually exclusive, such that the number of configurations of
values of these features grows exponentially with the number of
features. Imagine for example that these features are binary, i.e., each
can take only one of two values. Then up to configurations of these bits
are possible, thus carving up to regions in the input space. Yet the number
of free parameters required to carve this very rich partition of the input
space grows only linearly with , since feature (e.g., a neuron) comes with
its set of parameters (e.g., synaptic weights into the neuron). This is in
contrast with the majority of other machine learning approaches, in which
the number of parameters (and thus of examples) necessary to distinguish
input regions grows linearly with , instead of logarithmically. This
includes most clustering techniques, nearest-neighbor techniques, most
kernel-based non-parametric approaches, and decision trees. These
approaches can be plagued by the curse of dimensionality [Bengio+Delalleau
  2006]: in rich domains such as images and text, the number of
configurations of interest of the observed variables is exponentially
large, making it very difficult to generalize. As argued in [Bengio 2009,
  Montufar xx, Pascanu xx], distributed representations potentially have a
very large statistical advantage, in the sense that fewer examples might be
needed to achieve the same ability to generalize to new cases, or that
markedly better generalization can be obtained for the same number of
training examples.

Similarly to a deep neural network where the output of one group of neurons
is the input of another group, etc, if the learned function is obtained as
a composition of many non-linearities, then very rich functions can be
represented which can carve out exponentially more regions in input space
than a shallow architecture [Pascanu xx]. In both cases (distributed
representations and depth of representation), the mathematical arguments
rely on the ability to compose many pieces in order to get access to a very
rich class of computations. More abstract features can be computed in terms
of a composition of lower-level ones more efficiently than if these
abstract features had to be spelled out more explicitly, by enumerating all
the cases or templates that correspond to that feature. Whereas previous
theoretical work on neural networks showed that two layers (i.e. a single
hidden layer) was enough to approximate any continuous function, recent
theoretical results show that a shallow architecture may need to be very
fat where a deep one can much more efficiently (with fewer computations)
represent the same function [cite].

However, one has to keep in mind that not every function can be efficiently
represented with a distributed representation or with a deep
architecture. In other words, using deep learning corresponds to relying on
a broad a priori belief about the underlying data in order to achieve good
generalization to new cases. The prior is basically that the data we
observe can be explained by the combination of a limited set of factors or
features, and that one can learn about each of these factors or features
without having to see all the configurations of the other
ones. Fortunately, this very broad assumption seems to hold well in AI
tasks involving images, sounds and language.

\section{THE BACK-PROPAGATION ALGORITHM AND ITS APPARENT LIMITATIONS}

\begin{verbatim}
We will explain the core issue of hand-engineering features versus learning
 features automatically from data and show how backpropagation learns
 features. We will describe the most successful application of the
 80's/90's which was reading hand-written digits. We will also explain
 "distributed representations[a]" and illustrate them by showing how
 backpropagation can be used to convert words to vectors of meaningful
 features which are now very widely used in natural language applications.
  We will show how backpropagation through time can be used to learn
 recurrent neural networks and we will explain the vanishing/exploding
 gradients problem that makes this type of learning difficult.
\end{verbatim}

\subsection{How backpropagation works}

Many machine learning algorithms are designed around the minimization of a
training criterion, such as the error made on a set of training
examples. Monitoring the average value of that training criterion over
training examples tells us if learning is progressing, and learning
algorithms involve a method to modify parameters of the model (such as the
synaptic efficacies of an artificial neural network) so as to gradually
improve this training criterion. The most basic approach relies on
computing or estimating the derivatives of the training error with respect
to parameters, which tells us, for each parameter, what would happen to the
error made by the neural network if that parameter was increased
slightly. We call all these derivatives the error gradient and it indicates
in which direction to move the parameters. If the gradient on a particular
synaptic weight is positive, it means that one should decrease the weight
in order to decrease the error made. Instead, if the gradient is negative,
it means that increasing the weight would decrease the error, and so that
is what we should do.

Backpropagation is an algorithm to compute these derivatives, when the
function computed by the neural network is the composition of many
elementary computations, as in the case of a deep multi-layer
architecture. It relies on the idea that if we know in what direction to
change the outputs of artificial neurons at one layer, we can figure out
how to change the synaptic weights into these neurons so as to reduce the
error. Importantly, these error gradients on the neurons outputs at one
layer also tell us how to obtain the error gradients at the previous
layer. This is obtained through application of the chain rule for
derivatives, which specifies how to compute derivatives through a
composition of functions.  word-embeddings.png

\subsection{Neural word representations}

A nice example of the use of backpropagation is how it can be used to learn
representations for symbols, such as words. [Hinton 86] introduced the idea
of learning a distributed representation for symbols and [Bengio et al
  2001, 2003] showed how to combine the idea of learning a distributed
representation for each word with the task of modeling the probability
distribution of sequences of words (sentences from a corpus of text). We
have basically the composition two components: (1) mapping each of the
previous words in the sentence to their representation, and (2) mapping the
concatenation of these representations to a prediction for the
representation of the next word, from which one can compute its
probability.  Like in other applications of back-propagation, there seems
to be a chicken-and-egg problem, in the sense that the mapping (2) relies
on the word representations in (1), while it is the objective of achieving
the task performed by (2) that pushes the word representations in (1) to be
meaningful. It works because both are learned together, by a large number
of small gradual changes. The result is that words that are semantically
similar in the sense of being exchangeable end up nearly with the same
representation: one can replace “funny” by “comical” in many sentences and
the sentence remains plausible, so these two words have almost the same
representation. Similarly, “Tuesday” and “Wednesday” are exchangeable in
most sentences, so they end up very close to each other, while “dog” and
“cat” are not always exchangeable but often are, so they are not too far
from each in that space. Furthermore, the way in which the representations
for related words differ ends up being also meaningful: the difference
between the representation for “king” and the representation for “queen” is
very close the difference between the representation for “man” and the
representation for “woman” [Mikolov et al]. The system has discovered
features, such as gender in the above example, and learned to map each word
to its proper features.

\begin{figure}[H]
\includegraphics{fig-hidden-recurrence-rnn.png}
\caption{(a better figure can be made)}
\end{figure}

\subsection{Training recurrent nets}

A particularly interesting and general form of neural network is the
recurrent neural network. It can process temporally structured data, by
scanning through it just like humans and animals do: at each time step the
neural network “sees” as input the next element of the sequence, and
changes its internal state , which is therefore a representation of all the
previously seen elements . We can think of the brain as an enormous
recurrent neural network: the activity of all the neurons at a given time
constitutes the state of the brain recurrent network. Your next visual
saccade may provide the next input to the brain’s recurrent network. A
recurrent network can also have outputs, corresponding to actions or
predictions, and the training objective for a recurrent networks is
obtained from these values. For example, a recurrent network for language
modeling has output units that predict the next word in the sequence. At
each time step it gets a new word as input and updates its state. Such a
network can also “imagine” a plausible sequence, which makes it a so-called
generative model.  At each time step, given the current state, the network
makes a probabilistic guess about the next word, say, given the previous
words. By randomly choosing the next word according to these probabilities,
one more element is added to the sequence, which becomes an input used to
update the state and predict the next word. Begin by thinking of the first
word to type, then type it. In general, given the words already typed,
think of the next word to type that would be compatible with those you
already wrote, etc. Maximizing the training criterion for such a recurrent
network also maximizes the probability that the recurrent network will
generate sequences such as those in the training corpus.

The challenge of training very deep nets and recurrent networks over long sequences

One can view a recurrent network as a very deep neural network if we
consider each time step as one layer (right hand side of Figure xx). Deep
networks and recurrent networks are both very powerful because they can
represent very complex and non-linear functions formed by a long chain of
compositions. However, for the same reason, they are intrinsically
difficult to train, from the point of view of numerical
optimization. Composing many non-linear functions yields an even more
non-linear function. In cases where most of the composed functions had
small derivatives (a small change in input yields a smaller change in
output), the error gradient vanishes because it is the product of the
derivatives computed for each layer or each time step. In other cases (but
more rarely), a very small change in the parameters will yield a large
change in the error, and we have exploding gradients. Both of these
conditions throw off gradient-based optimization, which relies on
back-propagation to compute gradients. Although there has been substantial
progress to train recurrent neural networks (see a review of these
techniques [Pascanu et al 2013]), it remains a challenge to deal with long
sequences where there are so-called long-term dependencies: a small change
in the input at some point in the sequence could have a strong impact on
the correct output much later in the sequence.

\section{DEEP UNSUPERVISED LEARNING}

\begin{verbatim}
We will explain the breakthrough that occurred in 2006 when an effective
 new way of learning multiple layers of features was discovered. The idea
 is to learn the layers one a time with the objective being to build a good
 model of the correlations between feature activities in the previous layer
 (or the input).  After this layer-by-layer unsupervised "pre-training" to
 discover features, backpropagation works very well for fine-tuning the
 features that have already been discovered.  Will we describe several
 variations of the pre-training methods.
\end{verbatim}

\begin{figure}[H]
\includegraphics{stacked-rbms.png}
\caption{stacking RBMs}
\end{figure}

\subsection{Unsupervised pre-training}

Prior to 2006, most attempts at training multi-layered neural networks with
more than a few layers had failed and did not have any impact. However, a
few researchers, including the authors of this paper, continued searching
for ways to train deep neural networks, and many of them joined forces in a
Canadian research program with international nodes, called NCAP (Neural
Computation and Adaptive Perception), funded by the Canadian Institute For
Advanced Research (CIFAR), and headed by Geoff Hinton (and since 2014, by
Yoshua Bengio and Yann LeCun). In 2006, these researchers put out several
papers [2006-DL-crop] outlining a general strategy for taking advantage of
unsupervised learning of individual layers of a deep neural network in
order to initialize (or pre-train) a deep supervised network. Unlike
supervised learning, unsupervised learning only needs to look at the input
of the neural network, not the targets that we want to associate with these
inputs, so it can take advantage of large quantities of unlabeled data (for
which no target is given, which have not been examined and tagged by
humans) and even data associated with other tasks (see ‘transfer learning’,
below).  These unsupervised learning algorithm produce a new representation
of the data by learning the synaptic weights of a new layer of neurons that
map the raw input (or the output of the previous layer) to a new set of
learned features. Whereas traditional machine learning is based on
hand-crafted features, deep learning aims at taking advantage of data
(potentially unlabeled) in order to discover good features. This raises a
very good and difficult question: what is a good representation? Different
deep learning algorithms answer this question in different ways, but the
general idea is that a good representation makes it easier to learn further
dependencies, either between the features themselves (e.g., they are
independent or just linearly related to each other), or between the
features and some target variables (if we want to predict some categories,
given these features). An important ingredient of these 2006 papers is
that, although the job of learning a good representation is difficult,
unsupervised learning allows to improve a representation by iteratively
transforming previous representations. In these experiments, a new
Restricted Boltzmann Machine or auto-encoder is trained for creating the
next layer, and the representation learned can be used as input to train
the next layer similarly. After adding several such layers, one can
consider the overall mapping from input to top-level representation as the
initialization for a deep supervised neural network, which is further
trained with respect to the supervised objective (e.g., minimizing
classification error). We call this stage supervised fine-tuning because
all the layers are adjusted with respect to a single global objective,
after having been initialized from layer-wise unsupervised pre-training.

\subsection{Semi-supervised learning and transfer learning}

Most data that would be available to a learner does not come with
human-verified labels providing the desired characterization of its
content. This is true for humans as well as for machines and has motivated
machine learning methods called semi-supervised learning, in which both
labeled data and unlabeled data are exploited. One way in which unlabeled
data can be exploited is by helping to learn good representations, that
capture the underlying factors of variation present in the data. An example
of this approach is the unsupervised pre-training scheme discussed above
(which does not require labels), followed by supervised fine-tuning (on the
labeled examples). An interesting side-effect of learning good
representations (whether in an unsupervised, semi-supervised, or supervised
way) is that they tend to be useful for learning new but related tasks. For
example, in 2011 we participated in two international transfer learning
competitions [cite] and won thanks to unsupervised learning of
representations. The representations were learned using a large dataset
corresponding to some categories, while they were to be used as input for
training a linear classifier on a very small number of examples of other
categories. Because the different categories shared underlying factors of
variation (for example, many different object categories share
similarly-shaped parts and sub-parts), the models that benefited from the
pre-trained representations performed considerably better than those which
did not, and yielded very good generalization with only a few labeled
examples of the new categories. We believe that such transfer works well
when the learned representations capture high-level abstractions, some of
which may be relevant for both the training tasks and the test tasks. The
same process proved successful in the context of domain adaptation [Glorot
  2011], where the task was the same (sentiment analysis) but the input
distribution was different (the training domain might include comments
about movies and books, while the test domain could include comments about
music or electronic products).

\section{DEEP SPEECH RECOGNITION}

{\bf Geoff}

\begin{verbatim}
We will explain the acoustic modeling problem in speech recognition and
 desribe how initial results in 2009 using pre-training on the TIMIT
 dataset led to huge improvement in acoustic modeling over the next few
 years. Acoustic models based on deep neural networks are now used in most
 smart phones.  We will address the issue of when pre-training is needed
 and when it is unneccessary.
\end{verbatim}

\section{DEEP CONVOLUTIONAL NETWORKS FOR OBJECT RECOGNITION AND DETECTION}

{\bf Yann}

\section{DEEP RECURRENT NEURAL NETWORKS FOR MACHINE TRANSLATION}

{\bf Yoshua}

Machine translation is an appealing task because it seems that in order to
do it well, one has to capture a lot of the subtleties of meaning. Unlike
object recognition and speech recognition, it goes beyond pure sensory
perception and into the world of symbols where many researchers believed
that neural networks would be unable to rival more conventional AI
approaches.

In 2014, deep learning techniques, in particular those based on recurrent
networks, made rapid progress [Devlin et al, 2014, Cho et al 2014,
  Sutskever et al 2014]. Systems based on deep learning have now reached
and in some cases passed the state-of-the-art, e.g., for English-French and
English-Arabic. Machine translation technology had matured into a fairly
complex pipeline over the previous decade [cite], involving many separately
trained pieces and much domain expertise. Only a few parameters were
trained to optimize a global objective, and the core elements of deep
learning, i.e., multiple layers of distributed representations, were
lacking. Instead, the core component of traditional machine translation, as
in earlier language modeling systems, relied on counting co-occurences of
short phrases with their presumed translated phrases. This makes it very
difficult for the learned system to generalize to semantically similar
phrases that have not been seen in the training set. Because the number of
ways of saying the same thing grows exponentially with the size of the
vocabulary and the length of the phrase, this approach suffers from the
curse of dimensionality: there are too many possible configurations of the
variables of interest (here, words) to be able to observe all of the
relevant configurations often enough. The initial goal of deep learning
researchers was to use distributed representations to replace or complement
some of the pieces of existing machine translation systems, but more recent
work aims to replace the entire system with a set of recurrent neural
networks that are trained end-to-end [cite]. We now give a simplified
sketch of how this can be done.

For each language there is a recurrent encoder network that reads the words
in the source phrase or sentence one at a time. Each word is converted to a
distributed representation and the hidden state of the encoder network is
then recomputed from this word representation and the previous hidden
state. The final hidden state of the encoder network then becomes the
initial hidden state of the decoder network for the target language and
this decoder network produces a probability distribution for the first word
in the translation. If we pick from this distribution and feed the chosen
word into the decoder network as input, the decoder network will produce a
probability distribution over possible second words conditioned on the
chosen first word. If we keep repeating this process, the decoder network
defines a probability distribution over complete translations and the
encoder and decoder network can both be trained by backpropagation through
time to maximize the probability assigned to the translation that is given
in the training set.

It turns out that word representations learned by recurrent neural networks
that are trained to translate are better at capturing the semantic
similarity judgements made by humans than other representations of words,
including those derived by learning a neural language model [Hill 2014].

\section{DEEP MINDS}

{\bf Geoff}

\section{FUTURE DIRECTIONS}

Widespread application of the existing DREDNET technology:  Predicting activities of molecules, detecting particles, controlling robot motion ...

Deep Boltzmann machines (just mention undirected models!).

Intelligent fixations for vision.

Convolutional capsules (just a couple of sentences).

PROVISIONAL LIST OF FIGURES, TABLES AND BOXES

One box explaining the backpropagation algorithm for computing error derivatives.

One box explaining stochastic gradient descent.

....

KEY REFERENCES (dotted lines separate refs for each heading above)
(some of the conference papers here are awaiting review as journal papers and those refs will be updated if accepted.)

Rumelhart, Hinton \& Williams backprop paper

Lecun et al Document analysis with CNN.

Robinson RNN for speech

Neuro probabilistic language models

----------------

Hinton et al  Deep belief net paper.

Hinton et. al. Speech recognition with DNNs, Signal Processing Magazine

----------------

al. et. Lecun  Overfeat paper

Krizhevsky, Sutskever and Hinton, Imagenet paper

al. et. Lecun face recognition paper?

----------------

Hochreiter ref?

Graves et. al. paper showing that LSTM's work for handwriting recognition.

al. et. Bengio machine translation paper

Sutskever Le and Vinyals machine translation paper

-----------------

Deep mind NIPS paper on videogame player.

------------------
MT papers:

Devlin et al, 2014, Cho et al 2014, Sutskever et al 2014

\bibliography{strings,strings-shorter,ml,aigaion}
\bibliographystyle{natbib}

\end{document}

### 2014-11-01 Deep-Learning-Nature-Review.tex

--- 2014-10-30 2014/nature-review/Deep-Learning-Nature-Review.tex
@@
-\documentclass{article} % For LaTeX2e
+\documentclass[11pt]{article} % For LaTeX2e
 \usepackage{times}
@@
 
-%\maketitle
+\maketitle
 
@@
 (predicting the next word in a sentence) and predicting the activity of
-potential drug molecules.  In addition to its remarkable successes for
+potential drug molecules.  In addition to its impressive successes for
 perceptual tasks, deep learning is currently on the verge of beating the
@@
 widely applicable. For example, it was named as one of the top ten new
-technologies of 2013 by MIT Technology Review.
+technologies of 2013 by MIT Technology Review. This review explains some
+of the main underlying concepts and current applications of deep learning,
+and closes with perspectives for future research in this area.
 \end{abstract}
@@
 kernel-based non-parametric approaches, and decision trees. These
-approaches can be plagued by the curse of dimensionality [Bengio+Delalleau
-  2006]: in rich domains such as images and text, the number of
-configurations of interest of the observed variables is exponentially
-large, making it very difficult to generalize. As argued in [Bengio 2009,
-  Montufar xx, Pascanu xx], distributed representations potentially have a
-very large statistical advantage, in the sense that fewer examples might be
-needed to achieve the same ability to generalize to new cases, or that
-markedly better generalization can be obtained for the same number of
-training examples.
+approaches can be plagued by the curse of
+dimensionality~\citep{Bengio-localfailure-NIPS-2006-small,Bengio-decision-trees10}:
+in rich domains such as images and text, the number of configurations of
+interest of the observed variables is exponentially large, making it very
+difficult to generalize. As argued in \citet{Bengio-book, Montufar+Morton-2014, Montufar-et-al-NIPS2014},
+distributed representations potentially have a very large
+statistical advantage, in the sense that fewer examples might be needed to
+achieve the same ability to generalize to new cases, or that markedly
+better generalization can be obtained for the same number of training
+examples.
 
@@
 represented which can carve out exponentially more regions in input space
-than a shallow architecture [Pascanu xx]. In both cases (distributed
+than a shallow architecture~\citep{Montufar-et-al-NIPS2014}. In both cases (distributed
 representations and depth of representation), the mathematical arguments
@@
 fat where a deep one can much more efficiently (with fewer computations)
-represent the same function [cite].
+represent the same function~\citep{Bengio+Delalleau-ALT-2011-small,Montufar-et-al-NIPS2014}.
 
@@
 derivatives, which specifies how to compute derivatives through a
-composition of functions.  word-embeddings.png
+composition of functions.  Many different types of fixed or learned
+modules can thus be composed and trained end-to-end~\citep{LeCun98} 
+to form complex systems that perform
+tasks such as speech recognition, document analysis, or machine
+translation.
+
+\begin{figure}[H]
+\centerline{\includegraphics[width=0.7\linewidth]{word-embeddings.png}}
+\caption{Illustration of word representations learned for modeling
+language, non-linearly projected to 2-D for visualization.}
+\end{figure}
 
@@
 A nice example of the use of backpropagation is how it can be used to learn
-representations for symbols, such as words. [Hinton 86] introduced the idea
-of learning a distributed representation for symbols and [Bengio et al
-  2001, 2003] showed how to combine the idea of learning a distributed
+representations for symbols, such as words. ~\citet{Hinton86b} introduced the idea
+of learning a distributed representation for symbols and \citet{BenDucVin01-short}
+showed how to combine the idea of learning a distributed
 representation for each word with the task of modeling the probability
@@
 similar in the sense of being exchangeable end up nearly with the same
-representation: one can replace “funny” by “comical” in many sentences and
+representation: one can replace ``funny'' by ``comical'' in many sentences and
 the sentence remains plausible, so these two words have almost the same
-representation. Similarly, “Tuesday” and “Wednesday” are exchangeable in
-most sentences, so they end up very close to each other, while “dog” and
-“cat” are not always exchangeable but often are, so they are not too far
+representation. Similarly, ``Tuesday'' and ``Wednesday'' are exchangeable in
+most sentences, so they end up very close to each other, while ``dog'' and
+``cat'' are not always exchangeable but often are, so they are not too far
 from each in that space. Furthermore, the way in which the representations
 for related words differ ends up being also meaningful: the difference
-between the representation for “king” and the representation for “queen” is
-very close the difference between the representation for “man” and the
-representation for “woman” [Mikolov et al]. The system has discovered
+between the representation for ``king'' and the representation for ``queen'' is
+very close the difference between the representation for ``man'' and the
+representation for ``woman'' \citep{Mikolov-et-al-ICLR2013}. The system has discovered
 features, such as gender in the above example, and learned to map each word
@@
 \begin{figure}[H]
-\includegraphics{fig-hidden-recurrence-rnn.png}
+\centerline{\includegraphics[width=0.7\linewidth]{fig-hidden-recurrence-rnn.png}}
 \caption{(a better figure can be made)}
@@
 progress to train recurrent neural networks (see a review of these
-techniques [Pascanu et al 2013]), it remains a challenge to deal with long
+techniques~\citep{Pascanu+al-ICML2013-small}, it remains a challenge to deal with long
 sequences where there are so-called long-term dependencies: a small change
@@
 \begin{figure}[H]
-\includegraphics{stacked-rbms.png}
+\centerline{\includegraphics[width=0.7\linewidth]{stacked-rbms.png}}
 \caption{stacking RBMs}
@@
 Yoshua Bengio and Yann LeCun). In 2006, these researchers put out several
-papers [2006-DL-crop] outlining a general strategy for taking advantage of
+papers~\citep{Hinton06, Bengio-nips-2006-small,ranzato-07-small} outlining a general strategy for taking advantage of
 unsupervised learning of individual layers of a deep neural network in
@@
 example, in 2011 we participated in two international transfer learning
-competitions [cite] and won thanks to unsupervised learning of
+competitions~\citep{UTLC+LISA-2011-small,Goodfellow+all-NIPS2011} and won thanks to unsupervised learning of
 representations. The representations were learned using a large dataset
@@
 which may be relevant for both the training tasks and the test tasks. The
-same process proved successful in the context of domain adaptation [Glorot
-  2011], where the task was the same (sentiment analysis) but the input
+same process proved successful in the context of domain adaptation~\citep{Glorot+al-ICML-2011-small},
+where the task was the same (sentiment analysis) but the input
 distribution was different (the training domain might include comments
@@
 In 2014, deep learning techniques, in particular those based on recurrent
-networks, made rapid progress [Devlin et al, 2014, Cho et al 2014,
-  Sutskever et al 2014]. Systems based on deep learning have now reached
-and in some cases passed the state-of-the-art, e.g., for English-French and
-English-Arabic. Machine translation technology had matured into a fairly
-complex pipeline over the previous decade [cite], involving many separately
-trained pieces and much domain expertise. Only a few parameters were
-trained to optimize a global objective, and the core elements of deep
-learning, i.e., multiple layers of distributed representations, were
-lacking. Instead, the core component of traditional machine translation, as
-in earlier language modeling systems, relied on counting co-occurences of
-short phrases with their presumed translated phrases. This makes it very
-difficult for the learned system to generalize to semantically similar
-phrases that have not been seen in the training set. Because the number of
-ways of saying the same thing grows exponentially with the size of the
-vocabulary and the length of the phrase, this approach suffers from the
-curse of dimensionality: there are too many possible configurations of the
-variables of interest (here, words) to be able to observe all of the
-relevant configurations often enough. The initial goal of deep learning
-researchers was to use distributed representations to replace or complement
-some of the pieces of existing machine translation systems, but more recent
-work aims to replace the entire system with a set of recurrent neural
-networks that are trained end-to-end [cite]. We now give a simplified
-sketch of how this can be done.
+networks, made rapid
+progress~\citep{Devlin-et-al-ACL2014,Bahdanau-et-al-arxiv2014,Sutskever-et-al-arxiv2014}. Systems
+based on deep learning have now reached and in some cases passed the
+state-of-the-art, e.g., for English-French and English-Arabic. Machine
+translation technology had matured into a fairly complex pipeline over the
+previous decade, involving many separately trained pieces and much
+domain expertise. Only a few parameters were trained to optimize a global
+objective, and the core elements of deep learning, i.e., multiple layers of
+distributed representations, were lacking. Instead, the core component of
+traditional machine translation, as in earlier language modeling systems,
+relied on counting co-occurences of short phrases with their presumed
+translated phrases. This makes it very difficult for the learned system to
+generalize to semantically similar phrases that have not been seen in the
+training set. Because the number of ways of saying the same thing grows
+exponentially with the size of the vocabulary and the length of the phrase,
+this approach suffers from the curse of dimensionality: there are too many
+possible configurations of the variables of interest (here, words) to be
+able to observe all of the relevant configurations often enough. The
+initial goal of deep learning researchers was to use distributed
+representations to replace or complement some of the pieces of existing
+machine translation systems, but more recent work aims to replace the
+entire system with a set of recurrent neural networks that are trained
+end-to-end~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-arxiv2014}.
+We now give a simplified sketch of how this can be done.
 
@@
 time to maximize the probability assigned to the translation that is given
-in the training set.
+in the training set. An attention mechanism can be defined to
+learn to make the connection between the input and output sequences
+and improve performance on long sentences~\citep{Bahdanau-et-al-arxiv2014}.
 
@@
 similarity judgements made by humans than other representations of words,
-including those derived by learning a neural language model [Hill 2014].
+including those derived by learning a neural language model~\citep{Hill-et-al-2014}.
 
@@
 Rumelhart, Hinton \& Williams backprop paper 
+\citep{Rumelhart86b}
 
 Lecun et al Document analysis with CNN.
+\citep{LeCun+98}
 
 Robinson RNN for speech
+\citep{Robinson+Fallside91}
 
 Neuro probabilistic language models
+\citep{BenDucVin01-short}
 
@@
 Hinton et al  Deep belief net paper.
+\citep{Hinton06}
 
 Hinton et. al. Speech recognition with DNNs, Signal Processing Magazine 
+\citep{Hinton-et-al-2012}
+
+Graves et al RNNs for speech
+\citep{Graves-et-al-ICASSP2013}
 
@@
 al. et. Lecun  Overfeat paper
+\citep{Sermanet14}
 
 Krizhevsky, Sutskever and Hinton, Imagenet paper
+\citep{Krizhevsky-2012-small}
 
@@
 Hochreiter ref?
+\citep{Ciresan-2010}
 
 Graves et. al. paper showing that LSTM's work for handwriting recognition.
+\citep{Graves-et-al-2009}
 
-al. et. Bengio machine translation paper
+------------------
+MT papers:
 
-Sutskever Le and Vinyals machine translation paper
+\citep{Devlin-et-al-ACL2014,Bahdanau-et-al-arxiv2014,Sutskever-et-al-arxiv2014}
 
@@
 Deep mind NIPS paper on videogame player.
-
-------------------
-MT papers:
-
-Devlin et al, 2014, Cho et al 2014, Sutskever et al 2014
+\citep{Deepmind-atari-arxiv2014}
 

### 2014-11-01 2014/nature-review/Deep-Learning-Nature-Review.tex

@@
 \usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{xspace}
-\usepackage{tabularx}
-\usepackage{algorithm}
-\usepackage{algorithmic}
 

### 2014-11-01 2014/nature-review/Deep-Learning-Nature-Review.tex

@@
 \usepackage{amssymb}
+\usepackage{graphicx}
+
+\addtolength{\textwidth}{2cm}
+%\addtolength{\textwidth}{1cm}
+\addtolength{\textheight}{2cm}
+%\addtolength{\textheight}{1cm}
+\addtolength{\topmargin}{-1cm}
+\addtolength{\oddsidemargin}{-1cm}
+\addtolength{\evensidemargin}{-1cm}
 
@@
 
-\begin{verbatim}
+{\em 
 We plan to organize this review chronologically, but with forward looking
@@
  required to demonstrate the power of this type of learning algorithm.
-\end{verbatim}
+}
 
@@
 interest of the observed variables is exponentially large, making it very
-difficult to generalize. As argued in \citet{Bengio-book, Montufar+Morton-2014, Montufar-et-al-NIPS2014},
+difficult to generalize. As argued in \citet{Bengio-book-2009,Montufar+Morton-2014,Montufar-et-al-NIPS2014},
 distributed representations potentially have a very large
@@
 
-\begin{verbatim}
+{\em 
 We will explain the core issue of hand-engineering features versus learning
@@
  gradients problem that makes this type of learning difficult.
-\end{verbatim}
+}
 
@@
 
+\iffalse
 \begin{figure}[H]
@@
 \end{figure}
+\fi
 
@@
 
-The challenge of training very deep nets and recurrent networks over long sequences
+\subsection{The challenge of training very deep nets and recurrent networks over long sequences}
 
@@
 
-\begin{verbatim}
+{\em
 We will explain the breakthrough that occurred in 2006 when an effective
@@
  variations of the pre-training methods.
-\end{verbatim}
+}
 
@@
 
-\begin{verbatim}
-We will explain the acoustic modeling problem in speech recognition and
+{\em We will explain the acoustic modeling problem in speech recognition and
  desribe how initial results in 2009 using pre-training on the TIMIT
@@
  and when it is unneccessary.
-\end{verbatim}
+}
 
@@
 
-Convolutional capsules (just a couple of sentences).
+Convolutional capsules (just a couple of sentences).\\
 
-PROVISIONAL LIST OF FIGURES, TABLES AND BOXES
+{\bf PROVISIONAL LIST OF FIGURES, TABLES AND BOXES}
 
-One box explaining the backpropagation algorithm for computing error derivatives.
+In addition to the above:\\
+
+One box explaining the backpropagation algorithm for computing error derivatives.\\
 
@@
 
-KEY REFERENCES (dotted lines separate refs for each heading above)
+{\bf KEY REFERENCES (dotted lines separate refs for each heading above)}\\
 (some of the conference papers here are awaiting review as journal papers and those refs will be updated if accepted.)
@@
 
-\bibliography{strings,strings-shorter,ml,aigaion}
+\bibliography{strings,ml,aigaion}
 \bibliographystyle{natbib}

### 2015-01-04 2014/nature-review/deepreview.tex

\documentclass[]{article}
\usepackage{epsfig}
\usepackage{graphicx}
\parskip 0.1in
\parindent 0.0in
\setcounter{secnumdepth}{1}
\title{Deep Learning}

\author{Yann LeCun, Yoshua Bengio, Geoffrey Hinton}

\begin{document}

\maketitle

\section{Preface}

Deep neural networks that learn many layers of feature detectors have dramatically improved the state-of-the-art for
speech recognition, object recognition, object detection, predicting the activity of drug molecules, and many other
tasks. Deep learning discovers intricate structure in large datasets by using the backpropagation algorithm to compute
error-derivatives for the weights that define each feature detector and then repeatedly changed these weights in
proportion to their derivatives averaged over small batches of training examples.  Deep recurrent neural networks, in which
each layer also receives input from its own state at the previous time-step, can now generate good captions for images
and are competitive for machine translation.

\section{Introduction}

Deep learning is currently making major advances in solving problems that have resisted the best attempts of the
Artificial Intelligence community for many years. Deep learning is very good at discovering intricate structure in
large, high-dimensional datasets and is therefore widely applicable in all of the sciences. For example, it beats all
previous machine learning techniques at predicting the activity of potential drug molecules~\cite{Dahl}, detecting new
fundamental particles~\cite{??}, classifying galaxies~\cite{??}, or predicting the effects of mutations in non-coding
DNA on gene expression and disease~\cite{Frey}.

So far, most applications of deep learning use feed-forward neural networks, but for tasks that involve sequential
inputs it is often better to use recurrent neural networks. These are given one term from an input sequence at each time
step and their recurrent internal ``hidden'' state depends on the whole history of the input sequence. Recurrent
networks are very good at predicting the next character in a string\cite{Sutskever} or the next word in a
sequence\cite{Mikolov} but they can also be used for more complex tasks.  After reading an English sentence one word at
a time, for example, an English``encoder'' network can be trained so that the final state vector of its hidden neurons
is a good representation of the thought expressed by the sentence.  This thought vector can then be used as the initial
hidden state of a trained French ``decoder'' network which outputs a probability distribution for the first word of the
French translation. If a particuar first word is chosen and provided as input to the decoder network it will then output
a probability distribution for the second word of the translation and so on until a full stop is chosen~\cite{Vinyals, Montreal}.

This rather naive way of performing machine translation has quickly become competitive with the state-of-the-art and
this raises serious doubts about whether understanding a sentence requires anything like internal symbolic expressions
that are manipulated by binding their symbols to the variables in judiciously chosen rules of inference.  The
encoder-decoder translation networks do not work by using internal rules of inference in which variables get bound to
symbols. They just use big state vectors, big weight matrices and scalar non-linearities to get the job done.  This
suggests that deriving conclusions using only the form of the premises is a powerful technique that people can
eventually master but it is not a good model of the way we normally reason or understand language. Recurrent neural
networks that represent thoughts as big vectors are more compatible with the view that everyday reasoning involves many
simultaneous analogies that each contribute plausibility to a conclusion\cite{metaphors}.

This review is organized chronlogically starting with the introduction of the backpropagation algortihm in the
1980's. After describing some of its early successes, we explain the actual and perceived limitations that prevented it
from living up to its early promise for so many years and how these limitatons were eventually overcome.

\section{Pattern recognition and the back-propagation algorithm}

Patterns can be classified by first converting each raw input vector into a vector of feature activations and then
collecting evidence for each possible class by computing a weighted sum of the features. If good features can be
designed by hand, it is relatively easy to learn class-specific weights that optimize
discrimination~\cite{perceptrons}. Widely-used example of such linear classification methods are logistic
 regression and linear support vector machines~\cite{logisticregression,svm}.

However, even with good hand-crafted features, many interesting tasks cannot be solved by linear classifiers. A common
method to get around this problem is to transform the feature vector into a higher dimensional vector through a
non-linear mapping, exploiting the fact that collections of vectors are more likely to be easily spearated when they
are spread around a high-dimensional space~\cite{cover}. One clever way to do so is to use a ``kernel function'' that
measures the similarity between two vectors.  Using the kernel function, any input vector, ${\bf x}_i$ in the training
set can be used to define a feature whose activation for a new input vector, ${\bf x}_{\rm new}$ is just the
similarity between ${\bf x}_i$ and ${\bf x}_{\rm new}$ . Given a kernel function, it is possible to jointly optimize
the choice of training inputs used to define features and the class-specific weights associated with the
features\cite{}. The training inputs used to define features are called ``support vectors'' and Support Vector
Machines\cite{Vapnik} have been very successful, especially for tasks with a relatively small number of training
examples in which features are individually informative. They do not, however, perform well for tasks such as object
recognition in which the input space is high-dimensional and the input-output mapping is so
complex that computing similarities with a simple kernel function is ineffective.

Problems such as image and speech recognition require the input-output function to be insensitive to irrelevant
variations of the input, such as variations in position, orientation or illumination of an object, or variation in
pitch or accent of speech, while being very sensitive to particular minute variations, such as the difference between
a white wolf and a samoyed dog. These tasks are better handled with multiple layers of progressively more
complicated feature detectors each of which is defined in terms of the simpler feature detectors in the previous
layers.

From the earliest days of pattern recognition\cite{selfridge, rosenblatt} researchers aimed to replace hand-engineered
features by multiple layers of learned features.  This was eventually achieved by using the backpropagation algorithm
which was discovered independently by several different groups in the 1970s and 1980s\cite{Werbos, Parker, Lecun,
  Rumelhart, schmidreference}.  Backpropagation is simply an application of chain rule, which yields an efficient way
of computing the gradient of a smooth objective function with respect to all of the weights in a layered network. The
objective function typically measures the discrepancy between the actual output of the network on a given training
case and the target output for that case.  The key insight is that the error derivatives with respect to the {\it
  outputs} of a layer can be computed by working backwards from the error derivatives with respect to the outputs of
the subsequent layer. Once this backwards recursion has been used to compute error derivatives for the outputs of all
layers, it is easy to compute error derivatives for the parameters (or weights) of each layer (see box 1).

%%%%%%% GEOFF %%%%%%%%%%
GEOFF:  Do box explaining backprop equatons
%%%%%%%%%%%%%%%%%%%%%%

\section{Stochastic gradient descent}

Backpropagation provides an efficient way of computing gradients for all of the weights in a layered network on a
single training case, but there are still many ways to make use of these gradients. The most obvious method is to
average the gradients over all of the cases in the training set and then to change each weight in proportion to its
negative gradient. When the training set contains millions of cases and is highly redundant, however, this is
needlessly slow. It is much faster to compute the average gradient on a small ``mini-batch'' of randomly chosen
training cases and to update each weight by this stochastic estimate of its full gradient times a learning rate which
decays as learning proceeds. Stochastic gradient descent can be improved by using the history of the individual weight
gradients to adapt the learning rate for each weight\cite{bottou-bousquet} or to provide additional increments that
depend on earlier gradients\cite{} but these techniques are still the subject of controversy and are beyond the scope
of this review.  However, it is now widely accepted that some form of stochastic gradient descent is the best bet for
large, redundant datasets.

\section{Distributed representations}

The hidden layers of a multi-layer neural network learn to represent the network's inputs in a way that
makes it easy to predict the target outputs. This is nicely demonstrated by training a multi-layer neural network to
predict the next word in a sequence from a local context of earlier words. Each context word is presented to the network
as a one-of-N vector in which one component has a value of $1$ and the rest are $0$. The output layer has one unit per
word and is forced to produce a probability distribution over words by using the ``softmax'' function:
\begin{equation}
p_j = \frac{exp(z_j)}{sum_i exp(z_i)}
\end{equation}
where $z_j$ is the total input received by output unit $j$. The hidden layers use either the logistic or tanh non-linearity:
\begin{equation}
y_j = \frac{1}{1+ exp(z_j)}, \ \ \ y_j = \frac{}{}
\end{equation}

The connectivity of the network is restricted so that each of the one-of-N input vectors must first be converted to a
``word-vector'' before being used by subsequent layers to give a high probability to the target next word.  The network
learns word-vectors that contain many active components each of which can be interpreted as a separate feature of the
word. This was first demonstrated\cite{RHW} using sequences of three words, such as (Colin, Mother, Victoria), derived
from two family trees. The learned word vector for Colin contained ``semantic'' features representing which family tree
Colin was in, what branch of that tree he was in, and what generation he was\cite{hinton86}. These semantic features
were not explicitly present in the input.  They were discovered by the learning procedure as a good way of factorizing
the structured relationships between the input and output symbols into multiple ``micro-rules''.  For example, given
Colin's generation and the learned feature of Mother that the output should be one generation higher than the input, we
can already restrict the answer to people in a specific generation. The ability of deep learning to discover many such
micro-rules allows it to generalize well to novel combinations of individually familiar context words.

The family tree demonstration used a small, clean dataset in which the regularities could be captured by micro-rules
that were never violated and the features were easy to interpret, but learned word-vectors also work very well when the word
sequences come from a large corpus of real text and the individual micro-rules are unreliable\cite{NPLM} . When trained
to predict the next word in a news story, for example, the learned word vectors for Monday and Tuesday are very similar,
as are the word vectors for Vietnam and Iraq.  Vector representations of words learned from text are now very widely
used in natural language applications\cite{Collobert, Mikolov, Socher?}.

The issue of representation lies at the heart of the debate between the logic-inspired and the neurally-inspired
paradigms for cognition. In the logic-inspired paradigm, an instance of a symbol is something whose only property is
that it is either identical or non-identical to other symbol instances. It has no internal structure that is
relevant to its use. By contrast, neural networks use large activity vectors instead of symbols and parallel matrix
operations followed by non-linearities instead of formal rules of inference. With sufficient data, computational power
and depth, the combination of backpropagation for computing gradients and stochastic gradient descent for updating the
weights on connections has proved to be very good at learning appropriate distributed representations. These
representations allow a lot of parallel computation at the feature level so that complex tasks can be solved in a very
small number of sequential steps.

\section{Convolutional neural networks}

%%%%%%%% YANN %%%%%%%

YANN: can you summarize convnets and their use for reading cheques in about 350 words?

%%%%%%%%%%%%%%%%%%%%

\section{Recurrent Neural Networks}

When backpropagation was first introduced, its most exciting use was for training recurrent neural networks. These
receive input vectors at each time-step and have one or more layers of hidden units that, in addition to having
connections to higher layers or ouput units, also have connections that allow the neurons within a layer to receive
weighted input from those same neurons at the previous time-step.  The gradient for a recurrent weight, $w_{ij}$ that
allows neuron $i$ at time $t$ to influence the activity of neuron $j$ at time $t+1$ can be computed by
backpropagation-through-time (BPTT) of the error derivative for $j$ at time $t+1$. The full gradient for $w_{ij}$ on a whole
input sequence is just the sum of its gradients at each time-step.

Recurrent neural networks are very powerful dynamical systems, but training them with BPTT proved to be problematic,
because the backpropagated gradients either grow or shrink at each time step, so over many time-steps they typically explode
or vanish\cite{Bengio, schmdhuber}.

\section{Overcoming the limitations of backpropagation in deep networks}

In the 1990s, the machine learning community decided that backpropagation in deep neural networks did not work well in
practice and it was largely abandoned.  At the time, people believed that learning many layers of non-linear feature
detectors with no prior knowledge was much too difficult for a simple gradient descent procedure and that the networks
were getting trapped in poor local minima. We now know that the main reasons for the limited success of backpropagation
in deep networks were quite different: The labeled datasets were far too small, the computers were far too slow, and the
weights were not initialised sensibly.

The obvious weakness of gradient descent is that it can get trapped in poor local minima, but in high-dimensional
weight-spaces this is much less common than getting stuck close to saddle points where the error surface curves upwards
in most directions but downwards in a few. Recent empirical\cite{Bengio} and theoretical results\cite{Yann} strongly
suggest that in deep networks almost all of these saddle points give very similar performance so it does not matter
which one we find.  This goes a long way towards explaining why stochastic gradient descent works so well for deep
neural networks.

Interest in deep feed-forward networks was revived in 2005\cite{IJCAI} by the introduction of an unsupervised learning
procedure that could create layers of feature detectors without requiring labeled data. The objective in learning each
layer of feature detectors was to be able to reconstruct or model the activities of feature detectors (or raw inputs) in the
layer below.  By ``pre-training''  several layers of progressively more complex feature detectors using this reconstruction
objective, the weights of a deep network could be initialized to sensible values.  A final layer of output units could
then be added to the top of the network and the whole deep system could be fine-tuned using standard
backpropagation\cite{Science}. This worked remarkably well for recognizing handwritten digits especially when the amount
of labeled data was very limited\cite{BengioNips06}.

The first major application of this pre-training approach was in speech recognition and it was made possible by the
advent of fast Graphics Processing Units (GPUs) that were convenient to program.  In 2009, the approach was used to map
short temporal windows of coefficients extracted from a soundwave to a set of probabilities for the various fragments of
speech that might be represented by the frame in the center of the window.  It achieved record-breaking results on a
standard speech recognition benchmark that used a small vocabulary\cite{TIMITpaper} and was quickly developed to give
record-breaking results on a large vocabulary task\cite{Dahl}.  By 2012, versions of the deep net from 2009 were being
developed by many of the major speech groups\cite{SPS} and were already being deployed in Android phones.  For smaller
datasets, unsupervised pre-training leads to significantly better generalization, but once deep
learning had been rehabilitated, it turned out that large labeled datasets did not need the pre-training stage, provided the
scales of the initial random weights were carefully chosen to avoid very large or very small initial gradients.

Interest in recurrent networks (which are deep in time) was revived by ``Long Short Term Memory'' networks (LSTMs) that
use special hidden units whose natural behaviour is to remember inputs for a long time\cite{Hochreiter}.  A
linear ``memory cell'' interacts with other units via gated connections whose weights are multiplied by the state of a
logistic gating unit that has an activation between $0$ and $1$. The input vector has little effect on the state of a
memory cell if the cell's input gate has a low activation so memory cells can learn to selectively ignore the input.  Similarly,
a memory cell has little effect on the output if its output gate has low activation. Each memory cell also has a
connection to itself at the next time-step that has a weight of 1 so it tends to copy its own real-valued state, but
this connection is gated by a ``keep gate'' so the stored value is quickly eroded unless the keep-gate has an activation
near $1$.  The gates themselves have their activities determined by weighted inputs that they receive from the input
vectors and all of the memory cells.

When LSTMs were first described in English\cite{HochSchmid} their complexity was off-putting and, despite their
impressive ability to learn very long-range temporal structure on toy problems, they were largely ignored until they
achieved excellent results in a competition to read cursive hand-writing\cite{Graves}. They have subsequently proved to
be a very effective for tasks that have sequential input.  They are now better than deep feed-forward networks for
speech recognition\cite{gravestimit} and unlike feed-forward nets they can implement an entire speech recognition system
that goes all the way from coefficents derived from the soundwave to the sequence of characters in the transcription~\cite{graves}.
They are also what is currently used for the encoder and decoder networks that perform so well at machine translation.

It is not yet clear which aspects of the complicated architecture of LSTMs are really necessary. Memory cells are a good
way of allowing the backpropagated error derivatives to neither explode nor vanish, but they are not the only way.  For
example, recurrent nets with rectified linear hidden units that use the non-linear function $y=max(0,x)$ are much
simpler than LSTMs and if the recurrent weight matrix is initialized to be close to the identity matrix they work about
as well as LSTMs both for toy problems involving very long-range temporal structure and for real tasks like predicting
the next word in text\cite{LeHinton}.

\section{Object recognition and detection with deep convolutional networks}

When deep convolutional networks were applied to a dataset of about a million images from the web that contained a
thousand different classes they achieved spectacular results, almost halving the error rates of the best computer vision
sytems~\cite{kriz}.  Some of the images and the network's predictions are shown in figure ??.  The major factors in this
success were the large labeled dataset~\cite{FeiFe}, the use of fast GPUs, and the convolutional early layers. were.
Rectified linear units made the training much faster and the fully connected final layers were prevented from
overfitting the peculiarities of the training set by randomly omitting hidden units during
training\cite{dropoutJMLR}. Deep convolutional nets have quickly replaced the carefully engineered pipelines full of
hand-designed features developed by the computer vision community and they are now deployed in many computer vision
applications, including image tagging, object detection, face recognition, facial expression recognition and scene parsing.

%%%%%%%% YANN %%%%%%%
YANN: Add about 300 words plus another part of the figure showing detected parts of people. 
%%%%%%%%%%%%%%%%%%%%

\section{Language processing with deep networks}

%%%%%%%% YOSHUA %%%%%%%
YOSHUA: Add about 300 words plus t-sne figure of words and/or phrases.

The intro already explained the simple version of translation using encoder and decoder nets and I  suggest we  save
your ``translating with attention'' for introducing attention in the 
the future section.

After your stuff on language, I would like to include some version of the paragraph below along with a figure.
%%%%%%%%%%%%%%%%%%%%%%%

When translating between languages, the hidden units of the decoder network that generates possible translations are
initalized with a ``thought vector'' which is the final hidden state of an encoder network.  But it is also possible to
initialize a decoder network using a linear transformation of the last hidden layer of a deep convolutonal network that
has already been trained to recognize objects in images. The decoder network can then be trained by BPTT on a dataset of
captioned images~\cite{NIC,StanfordNIC,more?} and it produces remarkably good captions (see figure ??).

\section{Deep reinforcement learning}

Learning which action will maximize the total future reward given a state of the world is difficult when the rewards are
delayed and the world has a very large number of possible states. One impressive recent development is a single system
that can learn to play many different video games~\cite{NatureAtari}. The system uses a deep convolutional network to
map from the pixels on the screen at several previous time steps to the expected future reward of each possible action
and its only inputs are the pixels and the current score in the game. The learning works by minimizing the difference
between the system's current estimate of future reward and a noisy but less biased estimate that is obtained by using
the same deep network to estimate the future reward of the next state, where the next state is chosen stochastically
based on the predicted future rewards of all the currently legal actions.

\section{Learning to process symbols}

Deep learning is now the method of choice for many perceptual tasks and is making major inroads in natural language
processing where a huge amount of messy commonsense knowledge is required. Rapid progress is also being made in applying deep

learning, in particular BPTT, to relatively pure symbol processing tasks.  An impressive recent example is the ``Neural
Turing Machine'' which is a recurrent neural network that can choose when and what to write to or read from an external
memory~\cite{NTM}. Among other things, this can learn to ouput a sorted list of symbols when its input consists of an
unsorted sequence in which each symbol is accompanied by a real value that indicates its priority in the list.  To
achieve this it must learn a sorting program purely from examples of the input and output sequences.  A different system
that also uses BPTT can learn to map the sequence of characters in a simple program written in Python to the seqeunce of
characters that the program would print~\cite{Woj}. Rather surprisingly, BPTT can learn about {\it for loops} and can
also learn to perform addition and multiplication of strings contining several decimal digits. Achieving this kind of
behaviour seemed ludicrously optimistic when BPTT was first proposed.

\section{The future of deep learning}

Human vision is an active process in which intelligent fixations are used to ensure that only a tiny fraction of the optic array is ever processed at high resolution. We expect much of the future progress in vision, especially for video understanding or robotics, to come from systemsthat are trained end-to-end and combine deep convolutional networks with recurrent nets that use reinforcement learning to decide where to look\cite{}.  Such systems already outperform passive vision systems\cite{ba\&mnih}.  Similarly, when understanding sentences or whole documents we expect future systems to learn strategies for selectively attending to one part at a time\cite{Yoshua translation}. We also expect that in the next few years deep recurrent neural networks will become far better at understanding the meaning of documents, including the way in which the thoughts expressed by the sentences are related to one another.

This review has focused on supervised learning because of its recent achievments and we have ignored much
interesting work on unsupervised learning procedures for deep neural networks~\cite{deep-boltz, wake-sleep, google-cat, denoising-auto, ranzato-pami, more recent}. In the long run, however, we believe unsupervised learning will be essential for allowing deep networks to generalize well by creating internal representations that untangle the many different factors that interact to produce images, speech  and documents~\cite{Yann, Yoshua: untangling papers}.

Finally, we believe that the recent successes of deep learning using stochastic gradient descent are only the beginning. 
End-to-end learning systems that have many parameters and few built-in assumptions can make use of more data and more compute power far more easily than systems that require hand-engineering of domain-specific knowledge.

\section{Key references}
{\it Quite a few of these are very recent and have not yet been
  published through a peer-review so we are giving their arxiv
  references}.

\subsection{Back-propagation and Stochastic Gradient Descent}
Rumelhart, D. E., Hinton, G. E., \& Williams, R. J. (1986)\\ Learning
representations by back-propagating errors.\\ {\it Nature}, {\bf 323}, 533-536.

LeCun, Y., Boser, B., Denker, J. S., Henderson, D., Howard, R.E., Hubbard, W., \& Jackel, L.D.\\ 
Backpropagation Applied to Handwritten Zip Code Recognition\\
Neural Computation, 1(4):541-551, Winter 1989

\subsection{Distributed representations}
Hinton, G. E., McClelland, J. L., \& Rumelhart, D. E. (1986)\\ Distributed
representations.\\ In Rumelhart, D. E. and McClelland, J. L., editors, {\it
  Parallel Distributed Processing: Explorations in the Microstructure of
  Cognition. Volume 1: Foundations}, MIT Press, Cambridge, MA. pp 77-109.

Bengio, Y., Ducharme, R., \& Vincent, P. (2001)\\ A neural probabilistic
language model.\\ In {\it NIPS 2000}, pages 932--938.

%% Collobert, R., Weston, J., Bottou, L., Karlen, M., Kavukcuoglu, K., Kuksa, P. (2011)\\
%% Natural language processing (almost) from scratch.\\
%% {\it The Journal of Machine Learning Research} 12, pages 2493-2537

Mikolov, T., Sutskever, I., Chen, K., Corrado, G.S., \& Dean J.\\
Distributed representations of words and phrases and their compositionality.\\
In {\it NIPS 2013}, pages 3111-3119.

Montufar, G.~F., Pascanu, R., Cho, K., and Bengio, Y. (2014).  On the number of
linear regions of deep neural networks.\\ In {\it NIPS'2014}.

\subsection{The apparent limitations of backpropagation in deep networks}

Bengio, Y., Simard, P., \& Frasconi, P. (1994)\\ Learning long-term dependencies
with gradient descent is difficult.\\ {\it IEEE Transactions on Neural
  Networks}, 5(2), 157-166.

%% Dauphin, Y., Pascanu, R., Gulcehre, C., Cho, K., Ganguli, S., and Bengio,
%% Y.\\ 
%% Identifying and attacking the saddle point problem in high-dimensional
%% non-convex optimization.\\ In {\it NIPS'2014}.

\subsection{Deep unsupervised learning}
Hinton, G., Osindero, S., \& Teh, Y. W. (2006)\\ A fast learning algorithm for
deep belief nets.\\ {\it Neural Computation}, 18(7), 1527-1554.

Erhan, D., Bengio, Y., Courville, A., Manzagol, P. A., Vincent, P., \& Bengio,
S. (2010)\\ Why does unsupervised pre-training help deep learning?\\ {\it The
  Journal of Machine Learning Research}, {\bf 11}, 625-660.

%% Mesnil, G., Dauphin, Y., Glorot, X., Rifai, S., Bengio, Y., Goodfellow, I.,
%% Lavoie, E., Muller, X., Desjardins, G., Warde-Farley, D., Vincent, P.,
%% Courville, A., and Bergstra, J. (2011).\\ Unsupervised and transfer learning
%% challenge: a deep learning approach.\\ In {\em JMLR W\&CP: Proc. Unsupervised
%%  and Transfer Learning\/}, volume~7.

\subsection{Deep speech recognition}

%% Mohamed, A. R., Dahl, G. E., \& Hinton, G. (2012)\\ Acoustic modeling using deep
%% belief networks.\\ {\it IEEE Transactions on Audio, Speech, and Language
%%  Processing}, 20(1), 14-22.

Hinton, G., Deng, L., Yu, D., Dahl, G. E., Mohamed, A. R., Jaitly, N. \&
Kingsbury, B. (2012)\\ Deep neural networks for acoustic modeling in speech
recognition: The shared views of four research groups.\\ {\it Signal Processing
  Magazine}, IEEE, 29(6), 82-97.

\subsection{Deep convolutional networks for image understanding}
LeCun, Y., Bottou, L., Bengio, Y., \& Haffner, P. (1998)\\ Gradient-based
learning applied to document recognition.\\ {\it Proceedings of the IEEE},
86(11), 2278-2324.

%% Jain, V., Seung, H.S., \& Turaga S.C. (2010)\\
%% Machines that learn to segment images: a crucial technology for connectomics\\
%% {\it Current opinion in neurobiology} 20 (5), 653-666.

Krizhevsky, A., Sutskever, I., \& Hinton, G. E. (2012)\\ Imagenet classification
with deep convolutional neural networks.\\ In Advances in Neural Information
Processing Systems (1097-1105).

%% Farabet, C., Couprie, C., Najman, L. \& LeCun, Y. (2013)\\
%% Learning Hierarchical Features for Scene Labeling,\\
%% IEEE Transactions on Pattern Analysis and Machine Intelligence, August 2013.

%% Sermanet, P., Eigen, D., Zhang, X., Mathieu, M., Fergus, R., \& LeCun,
%% Y. (2013)\\ Overfeat: Integrated recognition, localization and detection using
%% convolutional networks.\\ {\it Proceedings of ICLR 2014}, arXiv:1312.6229.

Tompson, J., Goroshin, R., Jain, A., LeCun, Y., \& Bregler, C. (2014)
Efficient Object Localization Using Convolutional Networks. arXiv:1411.4280.

\subsection{Deep recurrent neural networks for text understanding and machine translation}

Graves, A., Liwicki, M., Fernández, S., Bertolami, R., Bunke, H., \&
Schmidhuber, J. (2009)\\ A novel connectionist system for unconstrained
handwriting recognition.\\ {\it IEEE Transactions on Pattern Analysis and
  Machine Intelligence}, 31(5), 855-868.

%% Devlin, J., Zbib, R., Huang, Z., Lamar, T., Schwartz, R., \& Makhoul,
% J. (2014)\\ Fast and robust neural network joint models for statistical machine
%% translation.\\ In {\it 52nd Annual Meeting of the Association for Computational
%%   Linguistics}, Baltimore, MD, USA.

Sutskever, I., Vinyals, O., \& Le, Q. V. (2014)\\ Sequence to sequence learning
with neural networks.\\ {\it arXiv preprint}, arXiv:1409.3215.

Bahdanau, D., Cho, K., \& Bengio, Y. (2014)\\ Neural machine translation by
jointly learning to align and translate.  {\it arXiv preprint}, arXiv:1409.0473.

Vinyals, O., Toshev, A., Bengio, S., \& Erhan, D. (2014)\\ Show and Tell: A
Neural Image Caption Generator.\\ {\it arXiv preprint}, arXiv:1411.4555.

%% Weston, J., Chopra, S., Bordes, A. (2014)\\
%% Memory Networks. arXiv:1410.3916.

\subsection{Deep reinforcement learning}
Mnih, V., Kavukcuoglu, K., Silver, D., Graves, A., Antonoglou, I., Wierstra, D.,
\& Riedmiller, M. (2013)\\ Playing Atari with deep reinforcement
learning.\\ {\it arXiv preprint} arXiv:1312.5602.

\newpage

\section{Boxes}

A box for the forward propagation and backpropagation equations.

\section{Figures}

1. Diagram showing:\\ a) A feedforward neural net.\\ b) A recurrent neural net
expanded in time.\\ c) A convolutional neural net.

2. Maps showing:\\ a) that word vectors learned by backpropagation are very good
at capturing the meanings of the words.\\ b) that the thought vectors extracted
by a translation model are very good at capturing the thoughts expressed by
sentences.

3. A diagram showing how unsupervised learning can be used to pre-train a deep
feeedforward neural net.

4. Some images with:\\ a) The opinions of a deep convolutional net about the
classes of the prominent objects.\\ b) Several alternative captions generated by
using the output of the convolutional net to determine the initial hidden state
of a decoder recurrent net.

5. Some images showing the localizations of parts of a person found by a deep
convolutional neural net.\\

\end{document}

### 2015-01-04 2014/nature-review/deepreview.tex
fixed some typos, made some suggestions of changes, and added or set a few references

@@
 
-Deep neural networks that learn many layers of feature detectors have dramatically improved the state-of-the-art for
-speech recognition, object recognition, object detection, predicting the activity of drug molecules, and many other
-tasks. Deep learning discovers intricate structure in large datasets by using the backpropagation algorithm to compute
-error-derivatives for the weights that define each feature detector and then repeatedly changed these weights in
-proportion to their derivatives averaged over small batches of training examples.  Deep recurrent neural networks, in which
-each layer also receives input from its own state at the previous time-step, can now generate good captions for images
-and are competitive for machine translation.
+Deep neural networks that learn many layers of feature detectors have
+dramatically improved the state-of-the-art for speech recognition, object
+recognition, object detection, predicting the activity of drug molecules,
+and many other tasks. Deep learning discovers intricate structure in large
+datasets by using the backpropagation algorithm to compute
+error-derivatives for the weights that define each feature detector and
+then repeatedly change these weights in proportion to their derivatives
+averaged over small batches of training examples.  Deep recurrent neural
+networks, in which each layer also receives input from its own state at the
+previous time-step, can now generate good captions for images and are
+competitive for machine translation.
 
@@
 
-Deep learning is currently making major advances in solving problems that have resisted the best attempts of the
-Artificial Intelligence community for many years. Deep learning is very good at discovering intricate structure in
-large, high-dimensional datasets and is therefore widely applicable in all of the sciences. For example, it beats all
-previous machine learning techniques at predicting the activity of potential drug molecules~\cite{Dahl}, detecting new
-fundamental particles~\cite{??}, classifying galaxies~\cite{??}, or predicting the effects of mutations in non-coding
-DNA on gene expression and disease~\cite{Frey}.
+Deep learning is currently making major advances in solving problems that
+have resisted the best attempts of the Artificial Intelligence community
+for many years. Deep learning is very good at discovering intricate
+structure in large, high-dimensional datasets and is therefore widely
+applicable in all of the sciences. For example, it beats all previous
+machine learning techniques at predicting the activity of potential drug
+molecules~\cite{Dahl}, detecting new fundamental particles~\cite{??},
+classifying galaxies~\cite{??}, or predicting the effects of mutations in
+non-coding DNA on gene expression and disease~\cite{Frey}.
 
-So far, most applications of deep learning use feed-forward neural networks, but for tasks that involve sequential
-inputs it is often better to use recurrent neural networks. These are given one term from an input sequence at each time
-step and their recurrent internal ``hidden'' state depends on the whole history of the input sequence. Recurrent
-networks are very good at predicting the next character in a string\cite{Sutskever} or the next word in a
-sequence\cite{Mikolov} but they can also be used for more complex tasks.  After reading an English sentence one word at
-a time, for example, an English``encoder'' network can be trained so that the final state vector of its hidden neurons
-is a good representation of the thought expressed by the sentence.  This thought vector can then be used as the initial
-hidden state of a trained French ``decoder'' network which outputs a probability distribution for the first word of the
-French translation. If a particuar first word is chosen and provided as input to the decoder network it will then output
-a probability distribution for the second word of the translation and so on until a full stop is chosen~\cite{Vinyals, Montreal}.
+So far, most applications of deep learning use feed-forward neural
+networks, but for tasks that involve sequential inputs it is often better
+to use recurrent neural networks. These are given one element from an input % YB term-->element
+sequence at each time step and their recurrent internal ``hidden'' state
+depends on the whole history of the input sequence. Recurrent networks are
+very good at predicting the next character in a string\cite{Sutskever} or
+the next word in a sequence\cite{Mikolov} but they can also be used for
+more complex tasks.  After reading an English sentence one word at a time,
+for example, an English``encoder'' network can be trained so that the final
+state vector of its hidden neurons is a good representation of the thought
+expressed by the sentence.  This thought vector can then be used as the
+initial hidden state of a trained French ``decoder'' network which outputs
+a probability distribution for the first word of the French translation. If
+a particular first word is chosen (e.g., according to that probability)
+and provided as input to the decoder
+network it will then output a probability distribution for the second word
+of the translation and so on until a full stop is chosen~\cite{Vinyals,
+  Montreal}. 
+%YB added a clarification
+Overall, this process generates sequences of French words according to
+a probability distribution that depends on the English sentence.
 
-This rather naive way of performing machine translation has quickly become competitive with the state-of-the-art and
-this raises serious doubts about whether understanding a sentence requires anything like internal symbolic expressions
-that are manipulated by binding their symbols to the variables in judiciously chosen rules of inference.  The
-encoder-decoder translation networks do not work by using internal rules of inference in which variables get bound to
-symbols. They just use big state vectors, big weight matrices and scalar non-linearities to get the job done.  This
-suggests that deriving conclusions using only the form of the premises is a powerful technique that people can
-eventually master but it is not a good model of the way we normally reason or understand language. Recurrent neural
-networks that represent thoughts as big vectors are more compatible with the view that everyday reasoning involves many
-simultaneous analogies that each contribute plausibility to a conclusion\cite{metaphors}.
+This rather naive way of performing machine translation has quickly become
+competitive with the state-of-the-art and this raises serious doubts about
+whether understanding a sentence requires anything like internal symbolic
+expressions that are manipulated by binding their symbols to the variables
+in judiciously chosen rules of inference.  The encoder-decoder translation
+networks do not work by using internal rules of inference in which
+variables get bound to symbols. They just use big state vectors, big weight
+matrices and scalar non-linearities to get the job done.  This suggests
+that deriving conclusions using only the form of the premises is a powerful
+technique that people can eventually master but it is not a good model of
+the way we normally reason or understand language. Recurrent neural
+networks that represent thoughts as big vectors are more compatible with
+the view that everyday reasoning involves many simultaneous analogies that
+each contribute plausibility to a conclusion\cite{metaphors}.
 
-This review is organized chronlogically starting with the introduction of the backpropagation algortihm in the
-1980's. After describing some of its early successes, we explain the actual and perceived limitations that prevented it
-from living up to its early promise for so many years and how these limitatons were eventually overcome.
+This review is organized chronlogically starting with the introduction of
+the backpropagation algorithm in the 1980's. After describing some of its
+early successes, we explain the actual and perceived limitations that
+prevented it from living up to its early promise for so many years and how
+these limitations were eventually overcome.
 
@@
 
-Patterns can be classified by first converting each raw input vector into a vector of feature activations and then
-collecting evidence for each possible class by computing a weighted sum of the features. If good features can be
-designed by hand, it is relatively easy to learn class-specific weights that optimize
-discrimination~\cite{perceptrons}. Widely-used example of such linear classification methods are logistic
- regression and linear support vector machines~\cite{logisticregression,svm}.
+Patterns can be classified by first converting each raw input vector into a
+vector of feature activations and then collecting evidence for each
+possible class by computing a weighted sum of the features. If good
+features can be designed by hand, it is relatively easy to learn
+class-specific weights that optimize
+discrimination~\cite{perceptrons}. Widely-used examples of such linear
+classification methods are logistic regression and linear support vector
+machines~\cite{logisticregression,svm}.
 
-However, even with good hand-crafted features, many interesting tasks cannot be solved by linear classifiers. A common
-method to get around this problem is to transform the feature vector into a higher dimensional vector through a
-non-linear mapping, exploiting the fact that collections of vectors are more likely to be easily spearated when they
-are spread around a high-dimensional space~\cite{cover}. One clever way to do so is to use a ``kernel function'' that
-measures the similarity between two vectors.  Using the kernel function, any input vector, ${\bf x}_i$ in the training
-set can be used to define a feature whose activation for a new input vector, ${\bf x}_{\rm new}$ is just the
-similarity between ${\bf x}_i$ and ${\bf x}_{\rm new}$ . Given a kernel function, it is possible to jointly optimize
-the choice of training inputs used to define features and the class-specific weights associated with the
-features\cite{}. The training inputs used to define features are called ``support vectors'' and Support Vector
-Machines\cite{Vapnik} have been very successful, especially for tasks with a relatively small number of training
-examples in which features are individually informative. They do not, however, perform well for tasks such as object
-recognition in which the input space is high-dimensional and the input-output mapping is so
-complex that computing similarities with a simple kernel function is ineffective.
+However, even with good hand-crafted features, many interesting tasks
+cannot be solved by linear classifiers. A common method to get around this
+problem is to transform the feature vector into a higher dimensional vector
+through a non-linear mapping, exploiting the fact that collections of
+vectors are more likely to be easily spearated when they are spread around
+a high-dimensional space~\cite{cover}. One clever way to do so is to use a
+``kernel function'' that measures the similarity between two vectors.
+Using the kernel function, any input vector, ${\bf x}_i$ in the training
+set can be used to define a feature whose activation for a new input
+vector, ${\bf x}_{\rm new}$ is just the similarity between ${\bf x}_i$ and
+${\bf x}_{\rm new}$. Given a kernel function, it is possible to jointly
+optimize the choice of training inputs used to define features and the
+class-specific weights associated with the features~\cite{svm}. The training
+inputs used to define features are called ``support vectors'' and Support
+Vector Machines\cite{Vapnik} have been very successful, especially for
+tasks with a relatively small number of training examples in which features
+are individually informative. They do not, however, perform well for tasks
+such as object recognition in which the input space is high-dimensional and
+the input-output mapping is so complex that computing similarities with a
+simple kernel function is ineffective.
 
-Problems such as image and speech recognition require the input-output function to be insensitive to irrelevant
-variations of the input, such as variations in position, orientation or illumination of an object, or variation in
-pitch or accent of speech, while being very sensitive to particular minute variations, such as the difference between
-a white wolf and a samoyed dog. These tasks are better handled with multiple layers of progressively more
-complicated feature detectors each of which is defined in terms of the simpler feature detectors in the previous
-layers.
+Problems such as image and speech recognition require the input-output
+function to be insensitive to irrelevant variations of the input, such as
+variations in position, orientation or illumination of an object, or
+variation in pitch or accent of speech, while being very sensitive to
+particular minute variations, such as the difference between a white wolf
+and a samoyed dog. These tasks are better handled with multiple layers of
+progressively more complicated feature detectors each of which is defined
+in terms of the simpler feature detectors in the previous layers.
 
-From the earliest days of pattern recognition\cite{selfridge, rosenblatt} researchers aimed to replace hand-engineered
-features by multiple layers of learned features.  This was eventually achieved by using the backpropagation algorithm
-which was discovered independently by several different groups in the 1970s and 1980s\cite{Werbos, Parker, Lecun,
-  Rumelhart, schmidreference}.  Backpropagation is simply an application of chain rule, which yields an efficient way
-of computing the gradient of a smooth objective function with respect to all of the weights in a layered network. The
-objective function typically measures the discrepancy between the actual output of the network on a given training
-case and the target output for that case.  The key insight is that the error derivatives with respect to the {\it
-  outputs} of a layer can be computed by working backwards from the error derivatives with respect to the outputs of
-the subsequent layer. Once this backwards recursion has been used to compute error derivatives for the outputs of all
-layers, it is easy to compute error derivatives for the parameters (or weights) of each layer (see box 1).
+From the earliest days of pattern recognition~\cite{selfridge, rosenblatt}
+researchers aimed to replace hand-engineered features by multiple layers of
+learned features.  This was eventually achieved by using the
+backpropagation algorithm which was discovered independently by several
+different groups in the 1970s and 1980s\cite{Werbos, Parker, Lecun,
+  Rumelhart, schmidreference}.  Backpropagation is simply an application of
+the chain rule, which yields an efficient way of computing the gradient of a
+smooth objective function with respect to all of the weights in a layered
+network. The objective function typically measures the discrepancy between
+the actual output of the network on a given training case and the target
+output for that case.  The key insight is that the error derivatives with
+respect to the {\it outputs} of a layer can be computed by working
+backwards from the error derivatives with respect to the outputs of the
+subsequent layer. Once this backwards recursion has been used to compute
+error derivatives for the outputs of all layers, it is easy to compute
+error derivatives for the parameters (or weights) of each layer (see box
+1).
 
@@
 
-Backpropagation provides an efficient way of computing gradients for all of the weights in a layered network on a
-single training case, but there are still many ways to make use of these gradients. The most obvious method is to
-average the gradients over all of the cases in the training set and then to change each weight in proportion to its
-negative gradient. When the training set contains millions of cases and is highly redundant, however, this is
-needlessly slow. It is much faster to compute the average gradient on a small ``mini-batch'' of randomly chosen
-training cases and to update each weight by this stochastic estimate of its full gradient times a learning rate which
-decays as learning proceeds. Stochastic gradient descent can be improved by using the history of the individual weight
-gradients to adapt the learning rate for each weight\cite{bottou-bousquet} or to provide additional increments that
-depend on earlier gradients\cite{} but these techniques are still the subject of controversy and are beyond the scope
-of this review.  However, it is now widely accepted that some form of stochastic gradient descent is the best bet for
-large, redundant datasets.
+Backpropagation provides an efficient way of computing gradients for all of
+the weights in a layered network on a single training case, but there are
+still many ways to make use of these gradients. The most obvious method is
+to average the gradients over all of the cases in the training set and then
+to change each weight in proportion to its negative gradient, which is the
+direction where a small change will most improve the objective function. 
+When the training set contains millions of cases and is highly redundant, however,
+this is needlessly slow. It is much faster to compute the average gradient
+on a small ``mini-batch'' of randomly chosen training cases and to update
+each weight by this stochastic estimate of its full gradient times a
+learning rate which decays as learning proceeds. Stochastic gradient
+descent can be improved by using the history of the individual weight
+gradients to adapt the learning rate for each weight\cite{bottou-bousquet}
+or to provide additional increments that depend on earlier gradients\cite{momentum}
+but these techniques are still the subject of controversy and are beyond
+the scope of this review.  However, it is now widely accepted that some
+form of stochastic gradient descent is the best bet for large, redundant
+datasets.
 
@@
 
-The hidden layers of a multi-layer neural network learn to represent the network's inputs in a way that
-makes it easy to predict the target outputs. This is nicely demonstrated by training a multi-layer neural network to
-predict the next word in a sequence from a local context of earlier words. Each context word is presented to the network
-as a one-of-N vector in which one component has a value of $1$ and the rest are $0$. The output layer has one unit per
-word and is forced to produce a probability distribution over words by using the ``softmax'' function:
+The hidden layers of a multi-layer neural network learn to represent the
+network's inputs in a way that makes it easy to predict the target
+outputs. This is nicely demonstrated by training a multi-layer neural
+network to predict the next word in a sequence from a local context of
+earlier words~\citep{BenDucVin01-short}. 
+Each context word is presented to the network as a one-of-N
+vector in which one component has a value of $1$ and the rest are $0$. The
+output layer has one unit per word and is forced to produce a probability
+distribution over words by using the ``softmax'' function:
 \begin{equation}
@@
 \end{equation}
-where $z_j$ is the total input received by output unit $j$. The hidden layers use either the logistic or tanh non-linearity:
+where $z_j$ is the total input received by output unit $j$. The hidden
+layers use either the logistic or tanh non-linearity:
 \begin{equation}
-y_j = \frac{1}{1+ exp(z_j)}, \ \ \ y_j = \frac{}{}
+y_j = \frac{1}{1+ exp(z_j)}, \ \ \ y_j = \frac{}{} % YB: what was the intent here for the 2nd part?
 \end{equation}
 
-The connectivity of the network is restricted so that each of the one-of-N input vectors must first be converted to a
-``word-vector'' before being used by subsequent layers to give a high probability to the target next word.  The network
-learns word-vectors that contain many active components each of which can be interpreted as a separate feature of the
-word. This was first demonstrated\cite{RHW} using sequences of three words, such as (Colin, Mother, Victoria), derived
-from two family trees. The learned word vector for Colin contained ``semantic'' features representing which family tree
-Colin was in, what branch of that tree he was in, and what generation he was\cite{hinton86}. These semantic features
-were not explicitly present in the input.  They were discovered by the learning procedure as a good way of factorizing
-the structured relationships between the input and output symbols into multiple ``micro-rules''.  For example, given
-Colin's generation and the learned feature of Mother that the output should be one generation higher than the input, we
-can already restrict the answer to people in a specific generation. The ability of deep learning to discover many such
-micro-rules allows it to generalize well to novel combinations of individually familiar context words.
+The connectivity of the network is restricted so that each of the one-of-N
+input vectors must first be converted to a ``word-vector'' before being
+used by subsequent layers to give a high probability to the target next
+word.  The network learns word-vectors that contain many active components
+each of which can be interpreted as a separate feature of the word. This
+was first demonstrated\cite{RHW} using sequences of three words, such as
+(Colin, Mother, Victoria), derived from two family trees. The learned word
+vector for Colin contained ``semantic'' features representing which family
+tree Colin was in, what branch of that tree he was in, and from what generation
+he was~\cite{hinton86}. These semantic features were not explicitly present
+in the input.  They were discovered by the learning procedure as a good way
+of factorizing the structured relationships between the input and output
+symbols into multiple ``micro-rules''.  For example, given Colin's
+generation and the learned feature of Mother that the output should be one
+generation higher than the input, we can already restrict the answer to
+people in a specific generation. The ability of deep learning to discover
+many such micro-rules allows it to generalize well to novel combinations of
+individually familiar context words.
 
-The family tree demonstration used a small, clean dataset in which the regularities could be captured by micro-rules
-that were never violated and the features were easy to interpret, but learned word-vectors also work very well when the word
-sequences come from a large corpus of real text and the individual micro-rules are unreliable\cite{NPLM} . When trained
-to predict the next word in a news story, for example, the learned word vectors for Monday and Tuesday are very similar,
-as are the word vectors for Vietnam and Iraq.  Vector representations of words learned from text are now very widely
-used in natural language applications\cite{Collobert, Mikolov, Socher?}.
+The family tree demonstration used a small, clean dataset in which the
+regularities could be captured by micro-rules that were never violated and
+the features were easy to interpret, but learned word-vectors also work
+very well when the word sequences come from a large corpus of real text and
+the individual micro-rules are unreliable\cite{BenDucVin01-short}. When trained to
+predict the next word in a news story, for example, the learned word
+vectors for Monday and Tuesday are very similar, as are the word vectors
+for Vietnam and Iraq.  Vector representations of words learned from text
+are now very widely used in natural language applications\cite{Collobert,
+  Mikolov, Socher?}.
 
-The issue of representation lies at the heart of the debate between the logic-inspired and the neurally-inspired
-paradigms for cognition. In the logic-inspired paradigm, an instance of a symbol is something whose only property is
-that it is either identical or non-identical to other symbol instances. It has no internal structure that is
-relevant to its use. By contrast, neural networks use large activity vectors instead of symbols and parallel matrix
-operations followed by non-linearities instead of formal rules of inference. With sufficient data, computational power
-and depth, the combination of backpropagation for computing gradients and stochastic gradient descent for updating the
-weights on connections has proved to be very good at learning appropriate distributed representations. These
-representations allow a lot of parallel computation at the feature level so that complex tasks can be solved in a very
-small number of sequential steps.
+The issue of representation lies at the heart of the debate between the
+logic-inspired and the neurally-inspired paradigms for cognition. In the
+logic-inspired paradigm, an instance of a symbol is something whose only
+property is that it is either identical or non-identical to other symbol
+instances. It has no internal structure that is relevant to its use. By
+contrast, neural networks use large activity vectors instead of symbols and
+parallel matrix operations followed by non-linearities instead of formal
+rules of inference. With sufficient data, computational power and depth,
+the combination of backpropagation for computing gradients and stochastic
+gradient descent for updating the weights on connections has proved to be
+very good at learning appropriate distributed representations. These
+representations allow a lot of parallel computation at the feature level so
+that complex tasks can be solved in a very small number of sequential
+steps. In fact, it can be shown that a measure of the richness of the functions
+that can be represented by a deep neural network grows exponentially
+with depth~\citep{Montufar-et-al-NIPS2014}.
 
@@
 
-YANN: can you summarize convnets and their use for reading cheques in about 350 words?
+YANN: can you summarize convnets and their use for reading cheques in about
+350 words?
 
@@
 
-When backpropagation was first introduced, its most exciting use was for training recurrent neural networks. These
-receive input vectors at each time-step and have one or more layers of hidden units that, in addition to having
-connections to higher layers or ouput units, also have connections that allow the neurons within a layer to receive
-weighted input from those same neurons at the previous time-step.  The gradient for a recurrent weight, $w_{ij}$ that
-allows neuron $i$ at time $t$ to influence the activity of neuron $j$ at time $t+1$ can be computed by
-backpropagation-through-time (BPTT) of the error derivative for $j$ at time $t+1$. The full gradient for $w_{ij}$ on a whole
-input sequence is just the sum of its gradients at each time-step.
+When backpropagation was first introduced, its most exciting use was for
+training recurrent neural networks. These receive input vectors at each
+time-step and have one or more layers of hidden units that, in addition to
+having connections to higher layers or ouput units, also have connections
+that allow the neurons within a layer to receive weighted input from those
+same neurons at the previous time-step.  The gradient for a recurrent
+weight, $w_{ij}$ that allows neuron $i$ at time $t$ to influence the
+activity of neuron $j$ at time $t+1$ can be computed by
+backpropagation-through-time (BPTT) of the error derivative for $j$ at time
+$t+1$. The full gradient for $w_{ij}$ on a whole input sequence is just the
+sum of its gradients at each time-step.
 
-Recurrent neural networks are very powerful dynamical systems, but training them with BPTT proved to be problematic,
-because the backpropagated gradients either grow or shrink at each time step, so over many time-steps they typically explode
-or vanish\cite{Bengio, schmdhuber}.
+Recurrent neural networks are very powerful dynamical systems, but training
+them with BPTT proved to be problematic, because the backpropagated
+gradients either grow or shrink at each time step, so over many time-steps
+they typically explode or vanish~\cite{Bengio-trnn93-small,Hochreiter91-small}.
 
@@
 
-In the 1990s, the machine learning community decided that backpropagation in deep neural networks did not work well in
-practice and it was largely abandoned.  At the time, people believed that learning many layers of non-linear feature
-detectors with no prior knowledge was much too difficult for a simple gradient descent procedure and that the networks
-were getting trapped in poor local minima. We now know that the main reasons for the limited success of backpropagation
-in deep networks were quite different: The labeled datasets were far too small, the computers were far too slow, and the
-weights were not initialised sensibly.
+In the 1990s, the machine learning community decided that backpropagation
+in deep neural networks did not work well in practice and it was largely
+abandoned.  At the time, people believed that learning many layers of
+non-linear feature detectors with no prior knowledge was much too difficult
+for a simple gradient descent procedure and that the networks were getting
+trapped in poor local minima. We now know that the main reasons for the
+limited success of backpropagation in deep networks were quite different:
+The labeled datasets were far too small, the computers were far too slow,
+and the weights were not initialised sensibly.
 
-The obvious weakness of gradient descent is that it can get trapped in poor local minima, but in high-dimensional
-weight-spaces this is much less common than getting stuck close to saddle points where the error surface curves upwards
-in most directions but downwards in a few. Recent empirical\cite{Bengio} and theoretical results\cite{Yann} strongly
-suggest that in deep networks almost all of these saddle points give very similar performance so it does not matter
-which one we find.  This goes a long way towards explaining why stochastic gradient descent works so well for deep
+The obvious weakness of gradient descent is that it can get trapped in poor
+local minima, but in high-dimensional weight-spaces this is much less
+common than getting stuck close to saddle points where the error surface
+curves upwards in most directions but downwards in a few. Recent
+results~\cite{Dauphin-et-al-NIPS2014-small,Yann} strongly suggest
+that in deep networks almost all of these local minima give very similar
+performance so it does not matter which one we find.  This goes a long way
+towards explaining why stochastic gradient descent works so well for deep
 neural networks.
 
-Interest in deep feed-forward networks was revived in 2005\cite{IJCAI} by the introduction of an unsupervised learning
-procedure that could create layers of feature detectors without requiring labeled data. The objective in learning each
-layer of feature detectors was to be able to reconstruct or model the activities of feature detectors (or raw inputs) in the
-layer below.  By ``pre-training''  several layers of progressively more complex feature detectors using this reconstruction
-objective, the weights of a deep network could be initialized to sensible values.  A final layer of output units could
-then be added to the top of the network and the whole deep system could be fine-tuned using standard
-backpropagation\cite{Science}. This worked remarkably well for recognizing handwritten digits especially when the amount
-of labeled data was very limited\cite{BengioNips06}.
+Interest in deep feed-forward networks was revived in 2005\cite{IJCAI} by
+the introduction of an unsupervised learning procedure that could create
+layers of feature detectors without requiring labeled data. The objective
+in learning each layer of feature detectors was to be able to reconstruct
+or model the activities of feature detectors (or raw inputs) in the layer
+below.  By ``pre-training'' several layers of progressively more complex
+feature detectors using this reconstruction objective, the weights of a
+deep network could be initialized to sensible values.  A final layer of
+output units could then be added to the top of the network and the whole
+deep system could be fine-tuned using standard
+backpropagation\cite{Science}. This worked remarkably well for recognizing
+handwritten digits especially when the amount of labeled data was very
+limited\cite{Bengio-nips-2006-small}.
 
-The first major application of this pre-training approach was in speech recognition and it was made possible by the
-advent of fast Graphics Processing Units (GPUs) that were convenient to program.  In 2009, the approach was used to map
-short temporal windows of coefficients extracted from a soundwave to a set of probabilities for the various fragments of
-speech that might be represented by the frame in the center of the window.  It achieved record-breaking results on a
-standard speech recognition benchmark that used a small vocabulary\cite{TIMITpaper} and was quickly developed to give
-record-breaking results on a large vocabulary task\cite{Dahl}.  By 2012, versions of the deep net from 2009 were being
-developed by many of the major speech groups\cite{SPS} and were already being deployed in Android phones.  For smaller
-datasets, unsupervised pre-training leads to significantly better generalization, but once deep
-learning had been rehabilitated, it turned out that large labeled datasets did not need the pre-training stage, provided the
-scales of the initial random weights were carefully chosen to avoid very large or very small initial gradients.
+The first major application of this pre-training approach was in speech
+recognition and it was made possible by the advent of fast Graphics
+Processing Units (GPUs) that were convenient to program~\citep{RainaICML09-small}.  
+In 2009, the approach was used to map short temporal windows of coefficients extracted
+from a soundwave to a set of probabilities for the various fragments of
+speech that might be represented by the frame in the center of the window.
+It achieved record-breaking results on a standard speech recognition
+benchmark that used a small vocabulary\cite{TIMITpaper} and was quickly
+developed to give record-breaking results on a large vocabulary
+task~\cite{Dahl}.  By 2012, versions of the deep net from 2009 were being
+developed by many of the major speech groups\cite{SPS} and were already
+being deployed in Android phones.  For smaller datasets, unsupervised
+pre-training leads to significantly better generalization~\citep{Erhan+al-2010-small}, 
+but once deep
+learning had been rehabilitated, it turned out that large labeled datasets
+did not need the pre-training stage, provided the scales of the initial
+random weights were carefully chosen to avoid very large or very small
+initial gradients.
 
-Interest in recurrent networks (which are deep in time) was revived by ``Long Short Term Memory'' networks (LSTMs) that
-use special hidden units whose natural behaviour is to remember inputs for a long time\cite{Hochreiter}.  A
-linear ``memory cell'' interacts with other units via gated connections whose weights are multiplied by the state of a
-logistic gating unit that has an activation between $0$ and $1$. The input vector has little effect on the state of a
-memory cell if the cell's input gate has a low activation so memory cells can learn to selectively ignore the input.  Similarly,
-a memory cell has little effect on the output if its output gate has low activation. Each memory cell also has a
-connection to itself at the next time-step that has a weight of 1 so it tends to copy its own real-valued state, but
-this connection is gated by a ``keep gate'' so the stored value is quickly eroded unless the keep-gate has an activation
-near $1$.  The gates themselves have their activities determined by weighted inputs that they receive from the input
-vectors and all of the memory cells.
+Interest in recurrent networks (which are deep in time) was revived by
+``Long Short Term Memory'' networks (LSTMs) that use special hidden units
+whose natural behaviour is to remember inputs for a long
+time~\cite{Hochreiter+Schmidhuber-1997}.  
+A linear ``memory cell'' interacts with other units
+via gated connections whose weights are multiplied by the state of a
+logistic gating unit that has an activation between $0$ and $1$. The input
+vector has little effect on the state of a memory cell if the cell's input
+gate has a low activation so memory cells can learn to selectively ignore
+the input.  Similarly, a memory cell has little effect on the output if its
+output gate has low activation. Each memory cell also has a connection to
+itself at the next time-step that has a weight of 1 so it tends to copy its
+own real-valued state, but this connection is gated by a ``keep gate'' so
+the stored value is quickly eroded unless the keep-gate has an activation
+near $1$.  The gates themselves have their activities determined by
+weighted inputs that they receive from the input vectors and all of the
+memory cells.
 
-When LSTMs were first described in English\cite{HochSchmid} their complexity was off-putting and, despite their
-impressive ability to learn very long-range temporal structure on toy problems, they were largely ignored until they
-achieved excellent results in a competition to read cursive hand-writing\cite{Graves}. They have subsequently proved to
-be a very effective for tasks that have sequential input.  They are now better than deep feed-forward networks for
-speech recognition\cite{gravestimit} and unlike feed-forward nets they can implement an entire speech recognition system
-that goes all the way from coefficents derived from the soundwave to the sequence of characters in the transcription~\cite{graves}.
-They are also what is currently used for the encoder and decoder networks that perform so well at machine translation.
+When LSTMs were first described in
+English~\cite{Hochreiter+Schmidhuber-1997} their complexity was off-putting
+and, despite their impressive ability to learn very long-range temporal
+structure on toy problems, they were largely ignored until they achieved
+excellent results in a competition to read cursive
+hand-writing~\cite{Graves}. They have subsequently proved to be very
+effective for tasks that have sequential input.  They are now better than
+deep feed-forward networks for speech recognition~\cite{gravestimit} and
+unlike feed-forward nets they can implement an entire speech recognition
+system that goes all the way from coefficents derived from the soundwave to
+the sequence of characters in the transcription~\cite{gravestimit}.  They
+are also what is currently used for the encoder and decoder networks that
+perform so well at machine
+translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
 
-It is not yet clear which aspects of the complicated architecture of LSTMs are really necessary. Memory cells are a good
-way of allowing the backpropagated error derivatives to neither explode nor vanish, but they are not the only way.  For
-example, recurrent nets with rectified linear hidden units that use the non-linear function $y=max(0,x)$ are much
-simpler than LSTMs and if the recurrent weight matrix is initialized to be close to the identity matrix they work about
-as well as LSTMs both for toy problems involving very long-range temporal structure and for real tasks like predicting
-the next word in text\cite{LeHinton}.
+% YB: I would skip the next paragraph; we've tried this and failed
+% so maybe it is better to wait and see before we put this in a review
+It is not yet clear which aspects of the complicated architecture of LSTMs
+are really necessary. Memory cells are a good way of allowing the
+backpropagated error derivatives to neither explode nor vanish, but they
+are not the only way.  For example, recurrent nets with rectified linear
+hidden units that use the non-linear function $y=max(0,x)$ are much simpler
+than LSTMs and if the recurrent weight matrix is initialized to be close to
+the identity matrix they work about as well as LSTMs both for toy problems
+involving very long-range temporal structure and for real tasks like
+predicting the next word in text\cite{LeHinton}.
 
@@
 
-When deep convolutional networks were applied to a dataset of about a million images from the web that contained a
-thousand different classes they achieved spectacular results, almost halving the error rates of the best computer vision
-sytems~\cite{kriz}.  Some of the images and the network's predictions are shown in figure ??.  The major factors in this
-success were the large labeled dataset~\cite{FeiFe}, the use of fast GPUs, and the convolutional early layers. were.
-Rectified linear units made the training much faster and the fully connected final layers were prevented from
-overfitting the peculiarities of the training set by randomly omitting hidden units during
-training\cite{dropoutJMLR}. Deep convolutional nets have quickly replaced the carefully engineered pipelines full of
-hand-designed features developed by the computer vision community and they are now deployed in many computer vision
-applications, including image tagging, object detection, face recognition, facial expression recognition and scene parsing.
+When deep convolutional networks were applied to a dataset of about a
+million images from the web that contained a thousand different classes
+they achieved spectacular results, almost halving the error rates of the
+best computer vision sytems~\cite{kriz}.  Some of the images and the
+network's predictions are shown in figure ??.  The major factors in this
+success were the large labeled dataset~\cite{FeiFe}, the use of fast GPUs,
+and the convolutional early layers.  It had already been shown that deep
+supervised nets with rectified linear units could be trained without
+unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
+These rectified linear units made the
+training much faster and the fully connected final layers were prevented
+from overfitting the peculiarities of the training set by randomly omitting
+hidden units during training~\cite{dropoutJMLR}. Deep convolutional nets
+have quickly replaced the carefully engineered pipelines full of
+hand-designed features developed by the computer vision community and they
+are now deployed in many computer vision applications, including image
+tagging, object detection, face recognition, facial expression recognition
+and scene parsing.
 
 %%%%%%%% YANN %%%%%%%
-YANN: Add about 300 words plus another part of the figure showing detected parts of people. 
+YANN: Add about 300 words plus another part of the figure showing detected
+parts of people.
 %%%%%%%%%%%%%%%%%%%%
@@
 
-The intro already explained the simple version of translation using encoder and decoder nets and I  suggest we  save
-your ``translating with attention'' for introducing attention in the 
-the future section.
+The intro already explained the simple version of translation using encoder
+and decoder nets and I suggest we save your ``translating with attention''
+for introducing attention in the the future section.
 
-After your stuff on language, I would like to include some version of the paragraph below along with a figure.
+After your stuff on language, I would like to include some version of the
+paragraph below along with a figure.
 %%%%%%%%%%%%%%%%%%%%%%%
 
-When translating between languages, the hidden units of the decoder network that generates possible translations are
-initalized with a ``thought vector'' which is the final hidden state of an encoder network.  But it is also possible to
-initialize a decoder network using a linear transformation of the last hidden layer of a deep convolutonal network that
-has already been trained to recognize objects in images. The decoder network can then be trained by BPTT on a dataset of
-captioned images~\cite{NIC,StanfordNIC,more?} and it produces remarkably good captions (see figure ??).
+When translating between languages, the hidden units of the decoder network
+that generates possible translations are initalized with a ``thought
+vector'' which is the final hidden state of an encoder network.  But it is
+also possible to initialize a decoder network using a linear transformation
+of the last hidden layer of a deep convolutonal network that has already
+been trained to recognize objects in images. The decoder network can then
+be trained by BPTT on a dataset of captioned
+images~\cite{NIC,StanfordNIC,more?} and it produces remarkably good
+captions (see figure ??).
 
@@
 
-Learning which action will maximize the total future reward given a state of the world is difficult when the rewards are
-delayed and the world has a very large number of possible states. One impressive recent development is a single system
-that can learn to play many different video games~\cite{NatureAtari}. The system uses a deep convolutional network to
-map from the pixels on the screen at several previous time steps to the expected future reward of each possible action
-and its only inputs are the pixels and the current score in the game. The learning works by minimizing the difference
-between the system's current estimate of future reward and a noisy but less biased estimate that is obtained by using
-the same deep network to estimate the future reward of the next state, where the next state is chosen stochastically
-based on the predicted future rewards of all the currently legal actions.
+Learning which action will maximize the total future reward given a state
+of the world is difficult when the rewards are delayed and the world has a
+very large number of possible states. One impressive recent development is
+a single system that can learn to play many different video
+games~\cite{NatureAtari}. The system uses a deep convolutional network to
+map from the pixels on the screen at several previous time steps to the
+expected future reward of each possible action and its only inputs are the
+pixels and the current score in the game. The learning works by minimizing
+the difference between the system's current estimate of future reward and a
+noisy but less biased estimate that is obtained by using the same deep
+network to estimate the future reward of the next state, where the next
+state is chosen stochastically based on the predicted future rewards of all
+the currently legal actions.
 
@@
 
-Deep learning is now the method of choice for many perceptual tasks and is making major inroads in natural language
-processing where a huge amount of messy commonsense knowledge is required. Rapid progress is also being made in applying deep
-
-learning, in particular BPTT, to relatively pure symbol processing tasks.  An impressive recent example is the ``Neural
-Turing Machine'' which is a recurrent neural network that can choose when and what to write to or read from an external
-memory~\cite{NTM}. Among other things, this can learn to ouput a sorted list of symbols when its input consists of an
-unsorted sequence in which each symbol is accompanied by a real value that indicates its priority in the list.  To
-achieve this it must learn a sorting program purely from examples of the input and output sequences.  A different system
-that also uses BPTT can learn to map the sequence of characters in a simple program written in Python to the seqeunce of
-characters that the program would print~\cite{Woj}. Rather surprisingly, BPTT can learn about {\it for loops} and can
-also learn to perform addition and multiplication of strings contining several decimal digits. Achieving this kind of
-behaviour seemed ludicrously optimistic when BPTT was first proposed.
+Deep learning is now the method of choice for many perceptual tasks and is
+making major inroads in natural language processing where a huge amount of
+messy commonsense knowledge is required. Rapid progress is also being made
+in applying deep learning, in particular BPTT, to relatively pure symbol processing tasks.
+An impressive recent example is the ``Neural Turing Machine'' which is a
+recurrent neural network that can choose when and what to write to or read
+from an external memory~\cite{NTM}. Among other things, this can learn to
+ouput a sorted list of symbols when its input consists of an unsorted
+sequence in which each symbol is accompanied by a real value that indicates
+its priority in the list.  To achieve this it must learn a sorting program
+purely from examples of the input and output sequences.  A different system
+that also uses BPTT can learn to map the sequence of characters in a simple
+program written in Python to the sequence of characters that the program
+would print~\cite{Woj}. Rather surprisingly, BPTT can learn about {\it for
+  loops} and can also learn to perform addition and multiplication of
+strings containing several decimal digits. Achieving this kind of behaviour
+seemed ludicrously optimistic when BPTT was first proposed.
 
@@
 
-Human vision is an active process in which intelligent fixations are used to ensure that only a tiny fraction of the optic array is ever processed at high resolution.
-We expect much of the future progress in vision, especially for video understanding or robotics, to come from systemsthat are trained end-to-end and combine deep convolutional networks with recurrent nets that use reinforcement learning to decide where to look\cite{}.
-Such systems already outperform passive vision systems\cite{ba\&mnih}.
-Similarly, when understanding sentences or whole documents we expect future systems to learn strategies for selectively attending to one part at a time\cite{Yoshua translation}.
-We also expect that in the next few years deep recurrent neural networks will become far better at understanding the meaning of documents, including the way in which the thoughts expressed by the sentences are related to one another.
+Human vision is an active process in which intelligent fixations are used
+to ensure that only a tiny fraction of the optic array is ever processed at
+high resolution. We expect much of the future progress in vision,
+especially for video understanding or robotics, to come from systems that
+are trained end-to-end and combine deep convolutional networks with
+recurrent nets that use reinforcement learning to decide where to
+look\cite{}.  Such systems already outperform passive vision
+systems\cite{ba\&mnih}.  Similarly, when understanding sentences or whole
+documents we expect future systems to learn strategies for selectively
+attending to one part at a time~\cite{Bahdanau-et-al-arxiv2014}. We also expect
+that in the next few years deep recurrent neural networks will become far
+better at understanding the meaning of documents, including the way in
+which the thoughts expressed by the sentences are related to one another.
 
-This review has focused on supervised learning because of its recent achievments and we have ignored much
-interesting work on unsupervised learning procedures for deep neural networks~\cite{deep-boltz, wake-sleep, google-cat, denoising-auto, ranzato-pami, more recent}.
-In the long run, however, we believe unsupervised learning will be essential for allowing deep networks to generalize well by creating internal representations that untangle the many different factors that interact to produce images, speech  and documents~\cite{Yann, Yoshua: untangling papers}.
+This review has focused on supervised learning because of its recent
+achievements and we have ignored much interesting work on unsupervised
+learning procedures for deep neural networks~\cite{deep-boltz, wake-sleep,
+  google-cat, denoising-auto, ranzato-pami, more recent}. In the long run,
+however, we believe unsupervised learning will be essential for allowing
+deep networks to generalize well by creating internal representations that
+untangle the many different factors that interact to produce images, speech
+and documents~\cite{Yann, Yoshua: untangling papers,Bengio-Courville-Vincent-TPAMI2013}.
 
-Finally, we believe that the recent successes of deep learning using stochastic gradient descent are only the beginning. 
-End-to-end learning systems that have many parameters and few built-in assumptions can make use of more data and more compute power far more easily than systems that require hand-engineering of domain-specific knowledge.
+Finally, we believe that the recent successes of deep learning using
+stochastic gradient descent are only the beginning.  As pointed out
+in~\citet{Bengio+Lecun-chapter2007-small}, end-to-end learning
+systems that have many parameters and few built-in assumptions can make use
+of more data and more compute power far more easily than systems that
+require hand-engineering of domain-specific knowledge.
 
 \section{Key references}
-{\it Quite a few of these are very recent and have not yet been
-  published through a peer-review so we are giving their arxiv
-  references}.
+{\it Quite a few of these are very recent and have not yet been published
+  through a peer-review so we are giving their arxiv references}.
 
@@
 Rumelhart, D. E., Hinton, G. E., \& Williams, R. J. (1986)\\ Learning
-representations by back-propagating errors.\\ {\it Nature}, {\bf 323}, 533-536.
+representations by back-propagating errors.\\ {\it Nature}, {\bf 323},
+533-536.
 
-LeCun, Y., Boser, B., Denker, J. S., Henderson, D., Howard, R.E., Hubbard, W., \& Jackel, L.D.\\ 
-Backpropagation Applied to Handwritten Zip Code Recognition\\
-Neural Computation, 1(4):541-551, Winter 1989
+LeCun, Y., Boser, B., Denker, J. S., Henderson, D., Howard, R.E., Hubbard,
+W., \& Jackel, L.D.\\ Backpropagation Applied to Handwritten Zip Code
+Recognition\\ Neural Computation, 1(4):541-551, Winter 1989
 
@@
 
-Mikolov, T., Sutskever, I., Chen, K., Corrado, G.S., \& Dean J.\\
-Distributed representations of words and phrases and their compositionality.\\
-In {\it NIPS 2013}, pages 3111-3119.
+Mikolov, T., Sutskever, I., Chen, K., Corrado, G.S., \& Dean
+J.\\ Distributed representations of words and phrases and their
+compositionality.\\ In {\it NIPS 2013}, pages 3111-3119.
 
-Montufar, G.~F., Pascanu, R., Cho, K., and Bengio, Y. (2014).  On the number of
-linear regions of deep neural networks.\\ In {\it NIPS'2014}.
+Montufar, G.~F., Pascanu, R., Cho, K., and Bengio, Y. (2014).  On the
+number of linear regions of deep neural networks.\\ In {\it NIPS'2014}.
 
@@
 
-Bengio, Y., Simard, P., \& Frasconi, P. (1994)\\ Learning long-term dependencies
-with gradient descent is difficult.\\ {\it IEEE Transactions on Neural
-  Networks}, 5(2), 157-166.
+Bengio, Y., Simard, P., \& Frasconi, P. (1994)\\ Learning long-term
+dependencies with gradient descent is difficult.\\ {\it IEEE Transactions
+  on Neural Networks}, 5(2), 157-166.
 
@@
 \subsection{Deep unsupervised learning}
-Hinton, G., Osindero, S., \& Teh, Y. W. (2006)\\ A fast learning algorithm for
-deep belief nets.\\ {\it Neural Computation}, 18(7), 1527-1554.
+Hinton, G., Osindero, S., \& Teh, Y. W. (2006)\\ A fast learning algorithm
+for deep belief nets.\\ {\it Neural Computation}, 18(7), 1527-1554.
 
-Erhan, D., Bengio, Y., Courville, A., Manzagol, P. A., Vincent, P., \& Bengio,
-S. (2010)\\ Why does unsupervised pre-training help deep learning?\\ {\it The
-  Journal of Machine Learning Research}, {\bf 11}, 625-660.
+Erhan, D., Bengio, Y., Courville, A., Manzagol, P. A., Vincent, P., \&
+Bengio, S. (2010)\\ Why does unsupervised pre-training help deep
+learning?\\ {\it The Journal of Machine Learning Research}, {\bf 11},
+625-660.
 
@@
 Kingsbury, B. (2012)\\ Deep neural networks for acoustic modeling in speech
-recognition: The shared views of four research groups.\\ {\it Signal Processing
-  Magazine}, IEEE, 29(6), 82-97.
+recognition: The shared views of four research groups.\\ {\it Signal
+  Processing Magazine}, IEEE, 29(6), 82-97.
 
@@
 
-Krizhevsky, A., Sutskever, I., \& Hinton, G. E. (2012)\\ Imagenet classification
-with deep convolutional neural networks.\\ In Advances in Neural Information
-Processing Systems (1097-1105).
+Krizhevsky, A., Sutskever, I., \& Hinton, G. E. (2012)\\ Imagenet
+classification with deep convolutional neural networks.\\ In Advances in
+Neural Information Processing Systems (1097-1105).
 
@@
 Tompson, J., Goroshin, R., Jain, A., LeCun, Y., \& Bregler, C. (2014)
-Efficient Object Localization Using Convolutional Networks. arXiv:1411.4280.
+Efficient Object Localization Using Convolutional
+Networks. arXiv:1411.4280.
 
@@
 
-Sutskever, I., Vinyals, O., \& Le, Q. V. (2014)\\ Sequence to sequence learning
-with neural networks.\\ {\it arXiv preprint}, arXiv:1409.3215.
+Sutskever, I., Vinyals, O., \& Le, Q. V. (2014)\\ Sequence to sequence
+learning with neural networks.\\ {\it arXiv preprint}, arXiv:1409.3215.
 
 Bahdanau, D., Cho, K., \& Bengio, Y. (2014)\\ Neural machine translation by
-jointly learning to align and translate.  {\it arXiv preprint}, arXiv:1409.0473.
+jointly learning to align and translate.  {\it arXiv preprint},
+arXiv:1409.0473.
 
@@
 \subsection{Deep reinforcement learning}
-Mnih, V., Kavukcuoglu, K., Silver, D., Graves, A., Antonoglou, I., Wierstra, D.,
-\& Riedmiller, M. (2013)\\ Playing Atari with deep reinforcement
-learning.\\ {\it arXiv preprint} arXiv:1312.5602.
+Mnih, V., Kavukcuoglu, K., Silver, D., Graves, A., Antonoglou, I.,
+Wierstra, D., \& Riedmiller, M. (2013)\\ Playing Atari with deep
+reinforcement learning.\\ {\it arXiv preprint} arXiv:1312.5602.
 
@@
 
-1. Diagram showing:\\ a) A feedforward neural net.\\ b) A recurrent neural net
-expanded in time.\\ c) A convolutional neural net.
+1. Diagram showing:\\ a) A feedforward neural net.\\ b) A recurrent neural
+net expanded in time.\\ c) A convolutional neural net.
 
-2. Maps showing:\\ a) that word vectors learned by backpropagation are very good
-at capturing the meanings of the words.\\ b) that the thought vectors extracted
-by a translation model are very good at capturing the thoughts expressed by
-sentences.
+2. Maps showing:\\ a) that word vectors learned by backpropagation are very
+good at capturing the meanings of the words.\\ b) that the thought vectors
+extracted by a translation model are very good at capturing the thoughts
+expressed by sentences.
 
-3. A diagram showing how unsupervised learning can be used to pre-train a deep
-feeedforward neural net.
+3. A diagram showing how unsupervised learning can be used to pre-train a
+deep feeedforward neural net.
 
-4. Some images with:\\ a) The opinions of a deep convolutional net about the
-classes of the prominent objects.\\ b) Several alternative captions generated by
-using the output of the convolutional net to determine the initial hidden state
-of a decoder recurrent net.
+4. Some images with:\\ a) The opinions of a deep convolutional net about
+the classes of the prominent objects.\\ b) Several alternative captions
+generated by using the output of the convolutional net to determine the
+initial hidden state of a decoder recurrent net.
 
-5. Some images showing the localizations of parts of a person found by a deep
-convolutional neural net.\\
+5. Some images showing the localizations of parts of a person found by a
+deep convolutional neural net.\\
 

### 2015-01-05 deepreview.tex
reformatted deepreview.tex into 80-line paragraphs to make it easier to see changes with diff
small changes as discussed, LSTM and beating previous methods, updated bib file

--- 2015-01-04 2014/nature-review/deepreview.tex
@@
 
-Deep learning is currently making major advances in solving problems that have resisted the best attempts of the
-Artificial Intelligence community for many years. Deep learning is very good at discovering intricate structure in
-large, high-dimensional datasets and is therefore widely applicable in all of the sciences. For example, it beats all
-previous machine learning techniques at predicting the activity of potential drug molecules~\cite{Dahl}, detecting new
-fundamental particles~\cite{??}, classifying galaxies~\cite{??}, or predicting the effects of mutations in non-coding
-DNA on gene expression and disease~\cite{Frey}.
+Deep learning is currently making major advances in solving problems that
+have resisted the best attempts of the Artificial Intelligence community
+for many years. Deep learning is very good at discovering intricate
+structure in large, high-dimensional datasets and is therefore widely
+applicable in all of the sciences. For example, it has beaten previous
+machine learning techniques at predicting the activity of potential drug
+molecules~\cite{Dahl}, detecting new fundamental particles~\cite{??},
+classifying galaxies~\cite{??}, or predicting the effects of mutations in
+non-coding DNA on gene expression and disease~\cite{Frey}.
 
-So far, most applications of deep learning use feed-forward neural networks, but for tasks that involve sequential
-inputs it is often better to use recurrent neural networks. These are given one term from an input sequence at each time
-step and their recurrent internal ``hidden'' state depends on the whole history of the input sequence. Recurrent
-networks are very good at predicting the next character in a string\cite{Sutskever} or the next word in a
-sequence\cite{Mikolov} but they can also be used for more complex tasks.  After reading an English sentence one word at
-a time, for example, an English``encoder'' network can be trained so that the final state vector of its hidden neurons
-is a good representation of the thought expressed by the sentence.  This thought vector can then be used as the initial
-hidden state of a trained French ``decoder'' network which outputs a probability distribution for the first word of the
-French translation. If a particuar first word is chosen and provided as input to the decoder network it will then output
-a probability distribution for the second word of the translation and so on until a full stop is chosen~\cite{Vinyals, Montreal}.
+So far, most applications of deep learning use feed-forward neural
+networks, but for tasks that involve sequential inputs it is often better
+to use recurrent neural networks. These are given one element from an input % YB term-->element
+sequence at each time step and their recurrent internal ``hidden'' state
+depends on the whole history of the input sequence. Recurrent networks are
+very good at predicting the next character in a string\cite{Sutskever} or
+the next word in a sequence\cite{Mikolov} but they can also be used for
+more complex tasks.  After reading an English sentence one word at a time,
+for example, an English``encoder'' network can be trained so that the final
+state vector of its hidden neurons is a good representation of the thought
+expressed by the sentence.  This thought vector can then be used as the
+initial hidden state of a trained French ``decoder'' network which outputs
+a probability distribution for the first word of the French translation. If
+a particular first word is chosen (e.g., according to that probability)
+and provided as input to the decoder
+network it will then output a probability distribution for the second word
+of the translation and so on until a full stop is chosen~\cite{Vinyals,
+  Montreal}. 
+%YB added a clarification
+Overall, this process generates sequences of French words according to
+a probability distribution that depends on the English sentence.
 
-This rather naive way of performing machine translation has quickly become competitive with the state-of-the-art and
-this raises serious doubts about whether understanding a sentence requires anything like internal symbolic expressions
-that are manipulated by binding their symbols to the variables in judiciously chosen rules of inference.  The
-encoder-decoder translation networks do not work by using internal rules of inference in which variables get bound to
-symbols. They just use big state vectors, big weight matrices and scalar non-linearities to get the job done.  This
-suggests that deriving conclusions using only the form of the premises is a powerful technique that people can
-eventually master but it is not a good model of the way we normally reason or understand language. Recurrent neural
-networks that represent thoughts as big vectors are more compatible with the view that everyday reasoning involves many
-simultaneous analogies that each contribute plausibility to a conclusion\cite{metaphors}.
+This rather naive way of performing machine translation has quickly become
+competitive with the state-of-the-art and this raises serious doubts about
+whether understanding a sentence requires anything like internal symbolic
+expressions that are manipulated by binding their symbols to the variables
+in judiciously chosen rules of inference.  The encoder-decoder translation
+networks do not work by using internal rules of inference in which
+variables get bound to symbols. They just use big state vectors, big weight
+matrices and scalar non-linearities to get the job done.  This suggests
+that deriving conclusions using only the form of the premises is a powerful
+technique that people can eventually master but it is not a good model of
+the way we normally reason or understand language. Recurrent neural
+networks that represent thoughts as big vectors are more compatible with
+the view that everyday reasoning involves many simultaneous analogies that
+each contribute plausibility to a conclusion\cite{metaphors}.
 
-This review is organized chronlogically starting with the introduction of the backpropagation algortihm in the
-1980's. After describing some of its early successes, we explain the actual and perceived limitations that prevented it
-from living up to its early promise for so many years and how these limitatons were eventually overcome.
+This review is organized chronlogically starting with the introduction of
+the backpropagation algorithm in the 1980's. After describing some of its
+early successes, we explain the actual and perceived limitations that
+prevented it from living up to its early promise for so many years and how
+these limitations were eventually overcome.
 
@@
 
-In the 1990s, the machine learning community decided that backpropagation in deep neural networks did not work well in
-practice and it was largely abandoned.  At the time, people believed that learning many layers of non-linear feature
-detectors with no prior knowledge was much too difficult for a simple gradient descent procedure and that the networks
-were getting trapped in poor local minima. We now know that the main reasons for the limited success of backpropagation
-in deep networks were quite different: The labeled datasets were far too small, the computers were far too slow, and the
-weights were not initialised sensibly.
+In the 1990s, the machine learning community decided that backpropagation
+in deep neural networks did not work well in practice and it was largely
+abandoned.  At the time, people believed that learning many layers of
+non-linear feature detectors with no prior knowledge was much too difficult
+for a simple gradient descent procedure and that the networks were getting
+trapped in poor local minima. We now know that the main reasons for the
+limited success of backpropagation in deep networks were quite different:
+The labeled datasets were far too small, the computers were far too slow,
+and the weights were not initialised sensibly.
 
-The obvious weakness of gradient descent is that it can get trapped in poor local minima, but in high-dimensional
-weight-spaces this is much less common than getting stuck close to saddle points where the error surface curves upwards
-in most directions but downwards in a few. Recent empirical\cite{Bengio} and theoretical results\cite{Yann} strongly
-suggest that in deep networks almost all of these saddle points give very similar performance so it does not matter
-which one we find.  This goes a long way towards explaining why stochastic gradient descent works so well for deep
+The obvious weakness of gradient descent is that it can get trapped in poor
+local minima, but in high-dimensional weight-spaces this is much less
+common than getting stuck close to saddle points where the error surface
+curves upwards in most directions but downwards in a few. Recent
+results~\cite{Dauphin-et-al-NIPS2014-small,Yann} strongly suggest
+that in deep networks almost all of these local minima give very similar
+performance so it does not matter which one we find.  This goes a long way
+towards explaining why stochastic gradient descent works so well for deep
 neural networks.
 
-Interest in deep feed-forward networks was revived in 2005\cite{IJCAI} by the introduction of an unsupervised learning
-procedure that could create layers of feature detectors without requiring labeled data. The objective in learning each
-layer of feature detectors was to be able to reconstruct or model the activities of feature detectors (or raw inputs) in the
-layer below.  By ``pre-training''  several layers of progressively more complex feature detectors using this reconstruction
-objective, the weights of a deep network could be initialized to sensible values.  A final layer of output units could
-then be added to the top of the network and the whole deep system could be fine-tuned using standard
-backpropagation\cite{Science}. This worked remarkably well for recognizing handwritten digits especially when the amount
-of labeled data was very limited\cite{BengioNips06}.
+Interest in deep feed-forward networks was revived in 2005\cite{IJCAI} by
+the introduction of an unsupervised learning procedure that could create
+layers of feature detectors without requiring labeled data. The objective
+in learning each layer of feature detectors was to be able to reconstruct
+or model the activities of feature detectors (or raw inputs) in the layer
+below.  By ``pre-training'' several layers of progressively more complex
+feature detectors using this reconstruction objective, the weights of a
+deep network could be initialized to sensible values.  A final layer of
+output units could then be added to the top of the network and the whole
+deep system could be fine-tuned using standard
+backpropagation\cite{Science}. This worked remarkably well for recognizing
+handwritten digits especially when the amount of labeled data was very
+limited\cite{Bengio-nips-2006-small}.
 
-The first major application of this pre-training approach was in speech recognition and it was made possible by the
-advent of fast Graphics Processing Units (GPUs) that were convenient to program.  In 2009, the approach was used to map
-short temporal windows of coefficients extracted from a soundwave to a set of probabilities for the various fragments of
-speech that might be represented by the frame in the center of the window.  It achieved record-breaking results on a
-standard speech recognition benchmark that used a small vocabulary\cite{TIMITpaper} and was quickly developed to give
-record-breaking results on a large vocabulary task\cite{Dahl}.  By 2012, versions of the deep net from 2009 were being
-developed by many of the major speech groups\cite{SPS} and were already being deployed in Android phones.  For smaller
-datasets, unsupervised pre-training leads to significantly better generalization, but once deep
-learning had been rehabilitated, it turned out that large labeled datasets did not need the pre-training stage, provided the
-scales of the initial random weights were carefully chosen to avoid very large or very small initial gradients.
+The first major application of this pre-training approach was in speech
+recognition and it was made possible by the advent of fast Graphics
+Processing Units (GPUs) that were convenient to program~\citep{RainaICML09-small}.  
+In 2009, the approach was used to map short temporal windows of coefficients extracted
+from a soundwave to a set of probabilities for the various fragments of
+speech that might be represented by the frame in the center of the window.
+It achieved record-breaking results on a standard speech recognition
+benchmark that used a small vocabulary\cite{TIMITpaper} and was quickly
+developed to give record-breaking results on a large vocabulary
+task~\cite{Dahl}.  By 2012, versions of the deep net from 2009 were being
+developed by many of the major speech groups\cite{SPS} and were already
+being deployed in Android phones.  For smaller datasets, unsupervised
+pre-training leads to significantly better generalization~\citep{Erhan+al-2010-small}, 
+but once deep
+learning had been rehabilitated, it turned out that large labeled datasets
+did not need the pre-training stage, provided the scales of the initial
+random weights were carefully chosen to avoid very large or very small
+initial gradients.
 
-Interest in recurrent networks (which are deep in time) was revived by ``Long Short Term Memory'' networks (LSTMs) that
-use special hidden units whose natural behaviour is to remember inputs for a long time\cite{Hochreiter}.  A
-linear ``memory cell'' interacts with other units via gated connections whose weights are multiplied by the state of a
-logistic gating unit that has an activation between $0$ and $1$. The input vector has little effect on the state of a
-memory cell if the cell's input gate has a low activation so memory cells can learn to selectively ignore the input.  Similarly,
-a memory cell has little effect on the output if its output gate has low activation. Each memory cell also has a
-connection to itself at the next time-step that has a weight of 1 so it tends to copy its own real-valued state, but
-this connection is gated by a ``keep gate'' so the stored value is quickly eroded unless the keep-gate has an activation
-near $1$.  The gates themselves have their activities determined by weighted inputs that they receive from the input
-vectors and all of the memory cells.
+Interest in recurrent networks (which are deep in time) was revived by
+``Long Short Term Memory'' networks (LSTMs) that use special hidden units
+whose natural behaviour is to remember inputs for a long
+time~\cite{Hochreiter+Schmidhuber-1997}.  
+A linear ``memory cell'' interacts with other units
+via gated connections whose weights are multiplied by the state of a
+logistic gating unit that has an activation between $0$ and $1$. The input
+vector has little effect on the state of a memory cell if the cell's input
+gate has a low activation so memory cells can learn to selectively ignore
+the input.  Similarly, a memory cell has little effect on the output if its
+output gate has low activation. Each memory cell also has a connection to
+itself (it is a leaky neuron) at the next time-step that has a weight of 1 so it tends to copy its
+own real-valued state, but this self-connection is gated by a ``keep gate'' so
+the stored value leaks quickly unless the keep-gate has an activation
+near $1$.  The gates themselves have their activities determined by
+weighted inputs that they receive from the input vectors and all of the
+memory cells.
 
-When LSTMs were first described in English\cite{HochSchmid} their complexity was off-putting and, despite their
-impressive ability to learn very long-range temporal structure on toy problems, they were largely ignored until they
-achieved excellent results in a competition to read cursive hand-writing\cite{Graves}. They have subsequently proved to
-be a very effective for tasks that have sequential input.  They are now better than deep feed-forward networks for
-speech recognition\cite{gravestimit} and unlike feed-forward nets they can implement an entire speech recognition system
-that goes all the way from coefficents derived from the soundwave to the sequence of characters in the transcription~\cite{graves}.
-They are also what is currently used for the encoder and decoder networks that perform so well at machine translation.
+When LSTMs were first described in
+English~\cite{Hochreiter+Schmidhuber-1997} their complexity was off-putting
+and, despite their impressive ability to learn very long-range temporal
+structure on toy problems, they were largely ignored until they achieved
+excellent results in a competition to read cursive
+hand-writing~\cite{Graves}. They have subsequently proved to be very
+effective for tasks that have sequential input.  They are now better than
+deep feed-forward networks for speech recognition~\cite{gravestimit} and
+unlike feed-forward nets they can implement an entire speech recognition
+system that goes all the way from coefficents derived from the soundwave to
+the sequence of characters in the transcription~\cite{gravestimit}.  They
+are also what is currently used for the encoder and decoder networks that
+perform so well at machine
+translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
 
-It is not yet clear which aspects of the complicated architecture of LSTMs are really necessary. Memory cells are a good
-way of allowing the backpropagated error derivatives to neither explode nor vanish, but they are not the only way.  For
-example, recurrent nets with rectified linear hidden units that use the non-linear function $y=max(0,x)$ are much
-simpler than LSTMs and if the recurrent weight matrix is initialized to be close to the identity matrix they work about
-as well as LSTMs both for toy problems involving very long-range temporal structure and for real tasks like predicting
-the next word in text\cite{LeHinton}.
+Recent work~\citep{Chung-et-al-NIPSDL2014} suggests that not all the elements of the LSTM
+architecture are required, the crucial element being the presence of
+of the gate controlling the leaking rate (or time constant) of the cell.
+%It is not yet clear which aspects of the complicated architecture of LSTMs
+%are really necessary. Memory cells are a good way of allowing the
+%backpropagated error derivatives to neither explode nor vanish, but they
+%are not the only way.  For example, recurrent nets with rectified linear
+%hidden units that use the non-linear function $y=max(0,x)$ are much simpler
+%than LSTMs and if the recurrent weight matrix is initialized to be close to
+%the identity matrix they work about as well as LSTMs both for toy problems
+%involving very long-range temporal structure and for real tasks like
+%predicting the next word in text\cite{LeHinton}.
 

### 2015-01-06 2014/nature-review/deepreview.tex
word embedding figure + start working on refs

@@
 \usepackage{graphicx}
+\usepackage{natbib}
+
 \parskip 0.1in
@@
 structure in large, high-dimensional datasets and is therefore widely
-applicable in all of the sciences. For example, it beats all previous
+applicable in all of the sciences. For example, it has beaten previous
 machine learning techniques at predicting the activity of potential drug
@@
 output gate has low activation. Each memory cell also has a connection to
-itself at the next time-step that has a weight of 1 so it tends to copy its
-own real-valued state, but this connection is gated by a ``keep gate'' so
-the stored value is quickly eroded unless the keep-gate has an activation
+itself (it is a leaky neuron) at the next time-step that has a weight of 1 so it tends to copy its
+own real-valued state, but this self-connection is gated by a ``keep gate'' so
+the stored value leaks quickly unless the keep-gate has an activation
 near $1$.  The gates themselves have their activities determined by
@@
 
-% YB: I would skip the next paragraph; we've tried this and failed
-% so maybe it is better to wait and see before we put this in a review
-It is not yet clear which aspects of the complicated architecture of LSTMs
-are really necessary. Memory cells are a good way of allowing the
-backpropagated error derivatives to neither explode nor vanish, but they
-are not the only way.  For example, recurrent nets with rectified linear
-hidden units that use the non-linear function $y=max(0,x)$ are much simpler
-than LSTMs and if the recurrent weight matrix is initialized to be close to
-the identity matrix they work about as well as LSTMs both for toy problems
-involving very long-range temporal structure and for real tasks like
-predicting the next word in text\cite{LeHinton}.
+Recent work~\citep{Chung-et-al-NIPSDL2014} suggests that not all the elements of the LSTM
+architecture are required, the crucial element being the presence of
+of the gate controlling the leaking rate (or time constant) of the cell.
+%It is not yet clear which aspects of the complicated architecture of LSTMs
+%are really necessary. Memory cells are a good way of allowing the
+%backpropagated error derivatives to neither explode nor vanish, but they
+%are not the only way.  For example, recurrent nets with rectified linear
+%hidden units that use the non-linear function $y=max(0,x)$ are much simpler
+%than LSTMs and if the recurrent weight matrix is initialized to be close to
+%the identity matrix they work about as well as LSTMs both for toy problems
+%involving very long-range temporal structure and for real tasks like
+%predicting the next word in text\cite{LeHinton}.
 
@@
 
+\iffalse
 %%%%%%%% YOSHUA %%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%
+\fi
+
+\begin{figure}[H]
+\centerline{\includegraphics[width=0.7\linewidth]{word-embeddings.png}}
+\caption{Illustration of word representations learned for modeling
+language, non-linearly projected to 2-D for visualization.}
+\end{figure}
 
@@
 look\cite{}.  Such systems already outperform passive vision
-systems\cite{ba\&mnih}.  Similarly, when understanding sentences or whole
+systems\cite{ba+mnih}.  Similarly, when understanding sentences or whole
 documents we expect future systems to learn strategies for selectively
@@
 untangle the many different factors that interact to produce images, speech
-and documents~\cite{Yann, Yoshua: untangling papers,Bengio-Courville-Vincent-TPAMI2013}.
+and documents~\cite{Yann,Bengio-Courville-Vincent-TPAMI2013}.
 
@@
 
+\bibliography{strings,ml,aigaion}
+\bibliographystyle{natbib}
+
 \end{document}

### 2015-01-07 deepreview.tex
added language paragraph and working on cleaning references

@@
 
-From the earliest days of pattern recognition~\cite{selfridge, rosenblatt}
+From the earliest days of pattern recognition~\cite{selfridge,Rosenblatt57}
 researchers aimed to replace hand-engineered features by multiple layers of
@@
 backpropagation algorithm which was discovered independently by several
-different groups in the 1970s and 1980s\cite{Werbos, Parker, Lecun,
-  Rumelhart, schmidreference}.  Backpropagation is simply an application of
+different groups in the 1970s and 1980s\cite{Werbos74,Parker85,LeCun85,Rumelhart86b,schmidreference}.  Backpropagation is simply an application of
 the chain rule, which yields an efficient way of computing the gradient of a
@@
 \begin{equation}
-p_j = \frac{exp(z_j)}{sum_i exp(z_i)}
+p_j = \frac{exp(z_j)}{\sum_i exp(z_i)}
 \end{equation}
@@
 \begin{equation}
-y_j = \frac{1}{1+ exp(z_j)}, \ \ \ y_j = \frac{}{} % YB: what was the intent here for the 2nd part?
+y_j = \frac{1}{1+ \exp(z_j)}, \ \ \ y_j = \frac{}{} % YB: what was the intent here for the 2nd part?
 \end{equation}
@@
 very well when the word sequences come from a large corpus of real text and
-the individual micro-rules are unreliable\cite{BenDucVin01-short}. When trained to
+the individual micro-rules are unreliable~\citep{BenDucVin01-short}. When trained to
 predict the next word in a news story, for example, the learned word
@@
 for Vietnam and Iraq.  Vector representations of words learned from text
-are now very widely used in natural language applications\cite{Collobert,
+are now very widely used in natural language applications~\cite{Collobert,
   Mikolov, Socher?}.
@@
 parallel matrix operations followed by non-linearities instead of formal
-rules of inference. With sufficient data, computational power and depth,
+rules of inference. The elements of such vectors can be understood as
+the values of learned attributes (or features) associated with some object. 
+With sufficient data, computational power and depth,
 the combination of backpropagation for computing gradients and stochastic
@@
 %%%%%%%%%%%%%%%%%%%%%%%
+\fi
+
+Prior to the introduction of neural language
+models~\citet{BenDucVin01-short} the standard approach to statistical
+modeling of language was based on counting frequencies of occurences of
+short symbol sequences (called n-grams). One disadvantage is that it
+taking into account context beyond a handful of words would require
+training corpora of exponentially growing size (with respect to the
+desired context size), since the number of configurations of $n$
+words from a vocabulary of size $V$ grows as $V^n$. Related to this
+problem and the very notion of symbol introduced above, n-grams
+do not generalize across the many semantically related sequences 
+of words (for example the many ways of expressing roughly the same thing).
+Instead, neural language models learn to associate each word to a
+distributed representation (a vector) such that semantically related
+words end up close to each other in that space, as illustrated in
+Figure~\ref{fig:word-embeddings}. To obtain that effect, one
+{\em jointly learns} a representation for each word as well as
+a function that computes a quantity of interest that is relevant
+to a language related task, such as predicting the next word
+in a sequence (neural language models) or predicting the sequence
+of translated words given the sequence of source words, i.e., neural
+machine translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
+
+\begin{figure}[H]
+\centerline{\includegraphics[width=0.7\linewidth]{word-embeddings.png}}
+\caption{Illustration of word representations learned for modeling
+language, non-linearly projected to 2-D for visualization using the
+t-SNE algorithm~\citep{VanDerMaaten08}.}
+\label{fig:word-embeddings}
+\end{figure}
 
@@
 images~\cite{NIC,StanfordNIC,more?} and it produces remarkably good
-captions (see figure ??).
+captions (see Figure~\ref{fig:caption-generation}).
+
+\begin{figure}[ht]
+%\centerline{\includegraphics[width=0.7\linewidth]{}}
+\caption{Example of captions generated by a recurrent network taking
+as extra input the representation extracted by a deep convolutional
+net from an image, with the whole system trained to ``translate''
+images into captions. Reproduced from~\citet{?}.
+}
+\label{fig:caption-generation}
+\end{figure}
 
@@
 \subsection{Back-propagation and Stochastic Gradient Descent}
-Rumelhart, D. E., Hinton, G. E., \& Williams, R. J. (1986)\\ Learning
-representations by back-propagating errors.\\ {\it Nature}, {\bf 323},
-533-536.
+
+%Rumelhart, D. E., Hinton, G. E., \& Williams, R. J. (1986)\\ Learning
+%representations by back-propagating errors.\\ {\it Nature}, {\bf 323},
+%533-536.
+\citep{Rumelhart86b}
 

### 2015-01-07 deepreview.tex
updating bib entries and references cleanup

@@
 descent can be improved by using the history of the individual weight
-gradients to adapt the learning rate for each weight\cite{bottou-bousquet}
+gradients to adapt the learning rate for each weight~\citep{bottou-bousquet-2008-small}
 or to provide additional increments that depend on earlier gradients\cite{momentum}
@@
 tree Colin was in, what branch of that tree he was in, and from what generation
-he was~\cite{hinton86}. These semantic features were not explicitly present
+he was~\cite{Hinton-et-al-PDP1986}. These semantic features were not explicitly present
 in the input.  They were discovered by the learning procedure as a good way
@@
 for Vietnam and Iraq.  Vector representations of words learned from text
-are now very widely used in natural language applications~\cite{Collobert,
-  Mikolov, Socher?}.
+are now very widely used in natural language applications~\cite{Schwenk-2007,CollobertR2008-short,
+Mikolov-Interspeech-2011-small,Socher-2011}.
 
@@
 350 words?
+
+\citep{LeCun89-small}
 
@@
 network's predictions are shown in figure ??.  The major factors in this
-success were the large labeled dataset~\cite{FeiFe}, the use of fast GPUs,
+success were the large labeled dataset~\cite{imagenet_cvpr09}, the use of fast GPUs,
 and the convolutional early layers.  It had already been shown that deep
@@
 
-\section{Key references}
-{\it Quite a few of these are very recent and have not yet been published
-  through a peer-review so we are giving their arxiv references}.
+%\section{Key references}
+%{\it Quite a few of these are very recent and have not yet been published
+%  through a peer-review so we are giving their arxiv references}.
 
-\subsection{Back-propagation and Stochastic Gradient Descent}
+%\subsection{Back-propagation and Stochastic Gradient Descent}
 
@@
 %533-536.
-\citep{Rumelhart86b}
+% --> \citep{Rumelhart86b}
 
-LeCun, Y., Boser, B., Denker, J. S., Henderson, D., Howard, R.E., Hubbard,
-W., \& Jackel, L.D.\\ Backpropagation Applied to Handwritten Zip Code
-Recognition\\ Neural Computation, 1(4):541-551, Winter 1989
+%LeCun, Y., Boser, B., Denker, J. S., Henderson, D., Howard, R.E., Hubbard,
+%W., \& Jackel, L.D.\\ Backpropagation Applied to Handwritten Zip Code
+%Recognition\\ Neural Computation, 1(4):541-551, Winter 1989
+%--> \citep{LeCun89-small}
 
-\subsection{Distributed representations}
-Hinton, G. E., McClelland, J. L., \& Rumelhart, D. E. (1986)\\ Distributed
-representations.\\ In Rumelhart, D. E. and McClelland, J. L., editors, {\it
-  Parallel Distributed Processing: Explorations in the Microstructure of
-  Cognition. Volume 1: Foundations}, MIT Press, Cambridge, MA. pp 77-109.
+%\subsection{Distributed representations}
+
+%Hinton, G. E., McClelland, J. L., \& Rumelhart, D. E. (1986)\\ Distributed
+%representations.\\ In Rumelhart, D. E. and McClelland, J. L., editors, {\it
+%  Parallel Distributed Processing: Explorations in the Microstructure of
+%  Cognition. Volume 1: Foundations}, MIT Press, Cambridge, MA. pp 77-109.
+% --> \citep{Hinton-et-al-PDP1986}
 

### 2015-01-07 2014/nature-review/deepreview.tex
updating bib entries and references cleanup

@@
 very good at predicting the next character in a string\cite{Sutskever} or
-the next word in a sequence\cite{Mikolov} but they can also be used for
+the next word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
 more complex tasks.  After reading an English sentence one word at a time,
@@
 for Vietnam and Iraq.  Vector representations of words learned from text
-are now very widely used in natural language applications\cite{Collobert,
-  Mikolov, Socher?}.
+are now very widely used in natural language 
+applications~\citep{Schwenk-2007,collobert:2011b,Socher-2011,Mikolov-et-al-NIPS2013}.
 
@@
 gradients either grow or shrink at each time step, so over many time-steps
-they typically explode or vanish~\cite{Bengio-trnn93-small,Hochreiter91-small}.
+they typically explode or vanish~\cite{Bengio-et-al-TNN1994,Hochreiter91-small}.
 
@@
 from overfitting the peculiarities of the training set by randomly omitting
-hidden units during training~\cite{dropoutJMLR}. Deep convolutional nets
+hidden units during training~\cite{Srivastava14}. Deep convolutional nets
 have quickly replaced the carefully engineered pipelines full of
@@
 
+Prior to the introduction of neural language
+models~\citet{BenDucVin01-short} the standard approach to statistical
+modeling of language was based on counting frequencies of occurences of
+short symbol sequences (called n-grams). One disadvantage is that it
+taking into account context beyond a handful of words would require
+training corpora of exponentially growing size (with respect to the
+desired context size), since the number of configurations of $n$
+words from a vocabulary of size $V$ grows as $V^n$. Related to this
+problem and the very notion of symbol introduced above, n-grams
+do not generalize across the many semantically related sequences 
+of words (for example the many ways of expressing roughly the same thing).
+Instead, neural language models learn to associate each word to a
+distributed representation (a vector) such that semantically related
+words end up close to each other in that space, as illustrated in
+Figure~\ref{fig:word-embeddings}. To obtain that effect, one
+{\em jointly learns} a representation for each word as well as
+a function that computes a quantity of interest that is relevant
+to a language related task, such as predicting the next word
+in a sequence (neural language models) or predicting the sequence
+of translated words given the sequence of source words, i.e., neural
+machine translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
+
 \begin{figure}[H]
@@
 \caption{Illustration of word representations learned for modeling
-language, non-linearly projected to 2-D for visualization.}
+language, non-linearly projected to 2-D for visualization using the
+t-SNE algorithm~\citep{VanDerMaaten08}.}
+\label{fig:word-embeddings}
 \end{figure}
@@
 
-\section{Key references}
-{\it Quite a few of these are very recent and have not yet been published
-  through a peer-review so we are giving their arxiv references}.
+%\section{Key references}
+%{\it Quite a few of these are very recent and have not yet been published
+%  through a peer-review so we are giving their arxiv references}.
 
-\subsection{Back-propagation and Stochastic Gradient Descent}
-Rumelhart, D. E., Hinton, G. E., \& Williams, R. J. (1986)\\ Learning
-representations by back-propagating errors.\\ {\it Nature}, {\bf 323},
-533-536.
+%\subsection{Back-propagation and Stochastic Gradient Descent}
 
-LeCun, Y., Boser, B., Denker, J. S., Henderson, D., Howard, R.E., Hubbard,
-W., \& Jackel, L.D.\\ Backpropagation Applied to Handwritten Zip Code
-Recognition\\ Neural Computation, 1(4):541-551, Winter 1989
+%Rumelhart, D. E., Hinton, G. E., \& Williams, R. J. (1986)\\ Learning
+%representations by back-propagating errors.\\ {\it Nature}, {\bf 323},
+%533-536.
+% --> \citep{Rumelhart86b}
 
-\subsection{Distributed representations}
-Hinton, G. E., McClelland, J. L., \& Rumelhart, D. E. (1986)\\ Distributed
-representations.\\ In Rumelhart, D. E. and McClelland, J. L., editors, {\it
-  Parallel Distributed Processing: Explorations in the Microstructure of
-  Cognition. Volume 1: Foundations}, MIT Press, Cambridge, MA. pp 77-109.
+%LeCun, Y., Boser, B., Denker, J. S., Henderson, D., Howard, R.E., Hubbard,
+%W., \& Jackel, L.D.\\ Backpropagation Applied to Handwritten Zip Code
+%Recognition\\ Neural Computation, 1(4):541-551, Winter 1989
+%--> \citep{LeCun89-small}
 
-Bengio, Y., Ducharme, R., \& Vincent, P. (2001)\\ A neural probabilistic
-language model.\\ In {\it NIPS 2000}, pages 932--938.
+%\subsection{Distributed representations}
+
+%Hinton, G. E., McClelland, J. L., \& Rumelhart, D. E. (1986)\\ Distributed
+%representations.\\ In Rumelhart, D. E. and McClelland, J. L., editors, {\it
+%  Parallel Distributed Processing: Explorations in the Microstructure of
+%  Cognition. Volume 1: Foundations}, MIT Press, Cambridge, MA. pp 77-109.
+% --> \citep{Hinton-et-al-PDP1986}
+
+%Bengio, Y., Ducharme, R., \& Vincent, P. (2001)\\ A neural probabilistic
+%language model.\\ In {\it NIPS 2000}, pages 932--938.
+% --> \citep{BenDucVin01-short}
 
@@
 %% {\it The Journal of Machine Learning Research} 12, pages 2493-2537
+% --> \citep{collobert:2011b}
 
-Mikolov, T., Sutskever, I., Chen, K., Corrado, G.S., \& Dean
-J.\\ Distributed representations of words and phrases and their
-compositionality.\\ In {\it NIPS 2013}, pages 3111-3119.
+%Mikolov, T., Sutskever, I., Chen, K., Corrado, G.S., \& Dean
+%J.\\ Distributed representations of words and phrases and their
+%compositionality.\\ In {\it NIPS 2013}, pages 3111-3119.
+% --> \citep{Mikolov-et-al-NIPS2013}
 
-Montufar, G.~F., Pascanu, R., Cho, K., and Bengio, Y. (2014).  On the
-number of linear regions of deep neural networks.\\ In {\it NIPS'2014}.
+%Montufar, G.~F., Pascanu, R., Cho, K., and Bengio, Y. (2014).  On the
+%number of linear regions of deep neural networks.\\ In {\it NIPS'2014}.
+% --> \citep{Montufar-et-al-NIPS2014}.
 
-\subsection{The apparent limitations of backpropagation in deep networks}
+%\subsection{The apparent limitations of backpropagation in deep networks}
 
-Bengio, Y., Simard, P., \& Frasconi, P. (1994)\\ Learning long-term
-dependencies with gradient descent is difficult.\\ {\it IEEE Transactions
-  on Neural Networks}, 5(2), 157-166.
+%Bengio, Y., Simard, P., \& Frasconi, P. (1994)\\ Learning long-term
+%dependencies with gradient descent is difficult.\\ {\it IEEE Transactions
+%  on Neural Networks}, 5(2), 157-166.
+% --> \citep{Bengio-et-al-TRNN1994}
 

### 2015-01-07 2014/nature-review/deepreview.tex
working on references

@@
 depends on the whole history of the input sequence. Recurrent networks are
-very good at predicting the next character in a string\cite{Sutskever} or
+very good at predicting the next character in a string~\citep{Sutskever-et-al-ICML2011} or
 the next word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
@@
 network it will then output a probability distribution for the second word
-of the translation and so on until a full stop is chosen~\cite{Vinyals,
-  Montreal}. 
+of the translation and so on until a full stop is 
+chosen~\cite{Vinyals-et-al-arxiv2014,Bahdanau-et-al-arxiv2014}. 
 %YB added a clarification
@@
 class-specific weights that optimize
-discrimination~\cite{perceptrons}. Widely-used examples of such linear
+discrimination~\citep{Rosenblatt57}. Widely-used examples of such linear
 classification methods are logistic regression and linear support vector
-machines~\cite{logisticregression,svm}.
+machines~\citep{logisticregression,Boser92}.
 
@@
 optimize the choice of training inputs used to define features and the
-class-specific weights associated with the features~\cite{svm}. The training
+class-specific weights associated with the features~\cite{Boser92}. The training
 inputs used to define features are called ``support vectors'' and Support
-Vector Machines\cite{Vapnik} have been very successful, especially for
+Vector Machines~\citep{Boser92} have been very successful, especially for
 tasks with a relatively small number of training examples in which features
@@
 backpropagation algorithm which was discovered independently by several
-different groups in the 1970s and 1980s\cite{Werbos74,Parker85,LeCun85,Rumelhart86b,schmidreference}.  Backpropagation is simply an application of
+different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,Rumelhart86b,schmidreference}.  
+Backpropagation is simply an application of
 the chain rule, which yields an efficient way of computing the gradient of a
@@
 gradients to adapt the learning rate for each weight~\citep{bottou-bousquet-2008-small}
-or to provide additional increments that depend on earlier gradients\cite{momentum}
+or to provide additional increments that depend on earlier gradients~\cite{momentum}
 but these techniques are still the subject of controversy and are beyond
@@
 tree Colin was in, what branch of that tree he was in, and from what generation
-he was~\cite{Hinton-et-al-PDP1986}. These semantic features were not explicitly present
+he was~\citep{Hinton-et-al-PDP1986}. These semantic features were not explicitly present
 in the input.  They were discovered by the learning procedure as a good way
@@
 gradients either grow or shrink at each time step, so over many time-steps
-they typically explode or vanish~\cite{Bengio-et-al-TNN1994,Hochreiter91-small}.
+they typically explode or vanish~\citep{Bengio-et-al-TNN1994,Hochreiter91-small}.
 
@@
 curves upwards in most directions but downwards in a few. Recent
-results~\cite{Dauphin-et-al-NIPS2014-small,Yann} strongly suggest
+results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-arxiv2014} strongly suggest
 that in deep networks almost all of these local minima give very similar
@@
 
-Interest in deep feed-forward networks was revived in 2005\cite{IJCAI} by
-the introduction of an unsupervised learning procedure that could create
-layers of feature detectors without requiring labeled data. The objective
-in learning each layer of feature detectors was to be able to reconstruct
-or model the activities of feature detectors (or raw inputs) in the layer
-below.  By ``pre-training'' several layers of progressively more complex
-feature detectors using this reconstruction objective, the weights of a
-deep network could be initialized to sensible values.  A final layer of
-output units could then be added to the top of the network and the whole
-deep system could be fine-tuned using standard
-backpropagation\cite{Science}. This worked remarkably well for recognizing
+Interest in deep feed-forward networks was revived in
+2005 and 2006~\citep{IJCAI,Hinton06-small,Bengio-nips-2006-small,ranzato-07-small} 
+by the introduction
+of unsupervised learning procedures that could create layers of feature
+detectors without requiring labeled data. The objective in learning each
+layer of feature detectors was to be able to reconstruct or model the
+activities of feature detectors (or raw inputs) in the layer below.  By
+``pre-training'' several layers of progressively more complex feature
+detectors using this reconstruction objective, the weights of a deep
+network could be initialized to sensible values.  A final layer of output
+units could then be added to the top of the network and the whole deep
+system could be fine-tuned using standard
+backpropagation\cite{Hinton-Science2006}. This worked remarkably well for recognizing
 handwritten digits especially when the amount of labeled data was very
-limited\cite{Bengio-nips-2006-small}.
+limited~\citep{Bengio-nips-2006-small,Erhan+al-2010-small}.
 
@@
 being deployed in Android phones.  For smaller datasets, unsupervised
-pre-training leads to significantly better generalization~\citep{Erhan+al-2010-small}, 
-but once deep
+pre-training helps to prevent overfitting~\citep{Erhan+al-2010-small}, 
+leading to significantly better generalization when the number of
+labeled examples is small. Once deep
 learning had been rehabilitated, it turned out that large labeled datasets
@@
 whose natural behaviour is to remember inputs for a long
-time~\cite{Hochreiter+Schmidhuber-1997}.  
+time~\citep{Hochreiter+Schmidhuber-1997}.  
 A linear ``memory cell'' interacts with other units
@@
 When LSTMs were first described in
-English~\cite{Hochreiter+Schmidhuber-1997} their complexity was off-putting
+English~\citep{Hochreiter+Schmidhuber-1997} their complexity was off-putting
 and, despite their impressive ability to learn very long-range temporal
@@
 they achieved spectacular results, almost halving the error rates of the
-best computer vision sytems~\cite{kriz}.  Some of the images and the
-network's predictions are shown in figure ??.  The major factors in this
-success were the large labeled dataset~\cite{imagenet_cvpr09}, the use of fast GPUs,
+best computer vision sytems~\citep{Krizhevsky-2012-small}.  Some of the images and the
+network's predictions are shown in figure~\ref{fig:imagenet-outputs}.  
+The major factors in this
+success were the large labeled dataset~\citep{imagenet_cvpr09}, the use of fast GPUs,
 and the convolutional early layers.  It had already been shown that deep
@@
 from overfitting the peculiarities of the training set by randomly omitting
-hidden units during training~\cite{Srivastava14}. Deep convolutional nets
+hidden units during training~\citep{Srivastava14}. Deep convolutional nets
 have quickly replaced the carefully engineered pipelines full of
@@
 net from an image, with the whole system trained to ``translate''
-images into captions. Reproduced from~\citet{?}.
+images into captions. Reproduced from~\cite{?}.
 }
@@
 recurrent neural network that can choose when and what to write to or read
-from an external memory~\cite{NTM}. Among other things, this can learn to
+from an external memory~\citep{Graves-et-al-arxiv2014}. 
+Among other things, this can learn to
 ouput a sorted list of symbols when its input consists of an unsorted
@@
 program written in Python to the sequence of characters that the program
-would print~\cite{Woj}. Rather surprisingly, BPTT can learn about {\it for
+would print~\citep{Zaremba+Sutskever-arxiv2014}. 
+Rather surprisingly, BPTT can learn about {\it for
   loops} and can also learn to perform addition and multiplication of
@@
 achievements and we have ignored much interesting work on unsupervised
-learning procedures for deep neural networks~\cite{deep-boltz, wake-sleep,
+learning procedures for deep neural 
+networks~\citep{Salakhutdinov2009-small,wake-sleep,
   google-cat, denoising-auto, ranzato-pami, more recent}. In the long run,
@@
 
-\subsection{Deep unsupervised learning}
-Hinton, G., Osindero, S., \& Teh, Y. W. (2006)\\ A fast learning algorithm
-for deep belief nets.\\ {\it Neural Computation}, 18(7), 1527-1554.
+%\subsection{Deep unsupervised learning}
+%Hinton, G., Osindero, S., \& Teh, Y. W. (2006)\\ A fast learning algorithm
+%for deep belief nets.\\ {\it Neural Computation}, 18(7), 1527-1554.
+% --> \citep{Hinton06-small}
 
-Erhan, D., Bengio, Y., Courville, A., Manzagol, P. A., Vincent, P., \&
-Bengio, S. (2010)\\ Why does unsupervised pre-training help deep
-learning?\\ {\it The Journal of Machine Learning Research}, {\bf 11},
-625-660.
+%Erhan, D., Bengio, Y., Courville, A., Manzagol, P. A., Vincent, P., \&
+%Bengio, S. (2010)\\ Why does unsupervised pre-training help deep
+%learning?\\ {\it The Journal of Machine Learning Research}, {\bf 11},
+%625-660.
+% --> \citep{Erhan+al-2010-small}
 
@@
 
-\subsection{Deep speech recognition}
+%\subsection{Deep speech recognition}
 
@@
 
-Krizhevsky, A., Sutskever, I., \& Hinton, G. E. (2012)\\ Imagenet
-classification with deep convolutional neural networks.\\ In Advances in
-Neural Information Processing Systems (1097-1105).
+%Krizhevsky, A., Sutskever, I., \& Hinton, G. E. (2012)\\ Imagenet
+%classification with deep convolutional neural networks.\\ In Advances in
+%Neural Information Processing Systems (1097-1105).
+% --> \cite{Krizhevsky-2012-small}
 
@@
 
-Sutskever, I., Vinyals, O., \& Le, Q. V. (2014)\\ Sequence to sequence
-learning with neural networks.\\ {\it arXiv preprint}, arXiv:1409.3215.
+%Sutskever, I., Vinyals, O., \& Le, Q. V. (2014)\\ Sequence to sequence
+%learning with neural networks.\\ {\it arXiv preprint}, arXiv:1409.3215.
+% --> \citep{Sutskever-et-al-NIPS2014}
 
-Bahdanau, D., Cho, K., \& Bengio, Y. (2014)\\ Neural machine translation by
-jointly learning to align and translate.  {\it arXiv preprint},
-arXiv:1409.0473.
+%Bahdanau, D., Cho, K., \& Bengio, Y. (2014)\\ Neural machine translation by
+%jointly learning to align and translate.  {\it arXiv preprint},
+%arXiv:1409.0473.
+% --> \citep{Bahdanau-et-al-arxiv2014}
 
-Vinyals, O., Toshev, A., Bengio, S., \& Erhan, D. (2014)\\ Show and Tell: A
-Neural Image Caption Generator.\\ {\it arXiv preprint}, arXiv:1411.4555.
+%Vinyals, O., Toshev, A., Bengio, S., \& Erhan, D. (2014)\\ Show and Tell: A
+%Neural Image Caption Generator.\\ {\it arXiv preprint}, arXiv:1411.4555.
+% --> \citep{Vinyals-et-al-arxiv2014}
 

### 2015-01-08 deepreview.tex
test pushing to new repo

@@
 machine learning techniques at predicting the activity of potential drug
-molecules~\cite{Dahl}, detecting new fundamental particles~\cite{??},
+molecules~\cite{Dahl}, detecting new fundamental particles~\cite{particles},
 classifying galaxies~\cite{??}, or predicting the effects of mutations in
@@
 depends on the whole history of the input sequence. Recurrent networks are
-very good at predicting the next character in a string\cite{Sutskever} or
-the next word in a sequence\cite{Mikolov} but they can also be used for
+very good at predicting the next character in a string~\citep{Sutskever-et-al-ICML2011} or
+the next word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
 more complex tasks.  After reading an English sentence one word at a time,
@@
 for Vietnam and Iraq.  Vector representations of words learned from text
-are now very widely used in natural language applications~\cite{Schwenk-2007,CollobertR2008-short,
-Mikolov-Interspeech-2011-small,Socher-2011}.
+are now very widely used in natural language 
+applications~\citep{Schwenk-2007,collobert:2011b,Socher-2011,Mikolov-et-al-NIPS2013}.
 
@@
 gradients either grow or shrink at each time step, so over many time-steps
-they typically explode or vanish~\cite{Bengio-trnn93-small,Hochreiter91-small}.
+they typically explode or vanish~\citep{Bengio-et-al-TNN1994,Hochreiter91-small}.
 
@@
 from overfitting the peculiarities of the training set by randomly omitting
-hidden units during training~\cite{dropoutJMLR}. Deep convolutional nets
+hidden units during training~\citep{Srivastava14}. Deep convolutional nets
 have quickly replaced the carefully engineered pipelines full of
@@
 
-Bengio, Y., Ducharme, R., \& Vincent, P. (2001)\\ A neural probabilistic
-language model.\\ In {\it NIPS 2000}, pages 932--938.
+%Bengio, Y., Ducharme, R., \& Vincent, P. (2001)\\ A neural probabilistic
+%language model.\\ In {\it NIPS 2000}, pages 932--938.
+% --> \citep{BenDucVin01-short}
 

### 2015-01-11 deepreview.tex
Update deepreview.tex
added a sentence and 2 references on the advantage of deep representations over classical non-parametric methods

@@
 
+%% this comment is geoff practising using github %%%
 \section{Introduction}
@@
 that can be represented by a deep neural network grows exponentially
-with depth~\citep{Montufar-et-al-NIPS2014}.
+with depth~\citep{Montufar-et-al-NIPS2014}, whereas according to the
+same measure, classical (shallow) non-parametric statistical methods such as
+Gaussian Support Vector Machines~\citep{Bengio-localfailure-NIPS-2006-small} 
+or decision trees~\citep{Bengio-decision-trees10}
+would require an exponentially large number
+of training examples to represent the same kinds of functions.
 

### 2015-01-11 deepreview.tex
added figure with phrase embeddings

@@
 \begin{figure}[H]
-\centerline{\includegraphics[width=0.7\linewidth]{word-embeddings.png}}
-\caption{Illustration of word representations learned for modeling
+\centerline{
+\includegraphics[width=0.49\linewidth]{word-embeddings.png}
+\includegraphics[width=0.49\linewidth]{phrase_zoom2.png}
+}
+\caption{Left: Illustration of word representations learned for modeling
 language, non-linearly projected to 2-D for visualization using the
-t-SNE algorithm~\citep{VanDerMaaten08}.}
+t-SNE algorithm~\citep{VanDerMaaten08}. Right: 2-D representation
+of phrases learned by an English-to-French encoder-decoder recurrent
+neural network~\citep{Cho-et-al-EMNLP2014}. One can observe that
+semantically similar words or sequences of words are  mapped to
+nearby representations.
+}
 \label{fig:word-embeddings}

### 2015-01-11 deepreview.tex
added caption generation figure, caption, and text

@@
 \begin{figure}[ht]
-%\centerline{\includegraphics[width=0.7\linewidth]{}}
+\centerline{\includegraphics[width=\linewidth]{caption-generation-Vinyals-et-al.png}}
 \caption{Example of captions generated by a recurrent network taking
@@
 net from an image, with the whole system trained to ``translate''
-images into captions. Reproduced from~\cite{?}.
+images into captions. Reproduced from~\citet{Vinyals-et-al-arxiv2014}.
 }
@@
 \end{figure}
+
+Similarly, instead of translating the meaning of a French sentence to an English sentence,
+one can learn to ``translate'' the meaning of an image into an English sentence,
+as illustrated in Figure~\ref{fig:caption-generation}. The ``encoder'' here is
+a convolutional neural network (see below), a deep architecture that is particularly
+well suited to extract a meaningful representation from an image, while the ``decoder''
+is a recurrent neural network similar to the ones used for machine translation
+and neural language modeling. There has been a surge of interest in such
+systems recently~\citep{Vinyals-et-al-arxiv2014,OTHER-CAPTION-GENERATION-PAPERS}.
 
@@
 
+\iffalse
+
 Hinton, G., Deng, L., Yu, D., Dahl, G. E., Mohamed, A. R., Jaitly, N. \&
@@
 
+\fi
+
 \newpage
 
-\section{Boxes}
+\section{Boxes (TODO)}
 

### 2015-01-11 deepreview.tex
tentatively fixed the SPS citation

@@
 task~\cite{Dahl}.  By 2012, versions of the deep net from 2009 were being
-developed by many of the major speech groups\cite{SPS} and were already
+developed by many of the major speech groups\cite{Hinton-et-al-2012} and were already
 being deployed in Android phones.  For smaller datasets, unsupervised
@@
 
-Hinton, G., Deng, L., Yu, D., Dahl, G. E., Mohamed, A. R., Jaitly, N. \&
-Kingsbury, B. (2012)\\ Deep neural networks for acoustic modeling in speech
-recognition: The shared views of four research groups.\\ {\it Signal
-  Processing Magazine}, IEEE, 29(6), 82-97.
+%Hinton, G., Deng, L., Yu, D., Dahl, G. E., Mohamed, A. R., Jaitly, N. \&
+%Kingsbury, B. (2012)\\ Deep neural networks for acoustic modeling in speech
+%recognition: The shared views of four research groups.\\ {\it Signal
+%  Processing Magazine}, IEEE, 29(6), 82-97.
+% --> \citep{Hinton-et-al-2012}
 

### 2015-01-11 deepreview.tex
fixed a few references from the key references list

@@
 excellent results in a competition to read cursive
-hand-writing~\cite{Graves}. They have subsequently proved to be very
+hand-writing~\cite{Graves-et-al-2009}. They have subsequently proved to be very
 effective for tasks that have sequential input.  They are now better than
@@
 %%%%%%%%%%%%%%%%%%%%
+
+Cite at least \citep{LeCun+98} and \citep{Tompson-et-al-arxiv2014}
 
@@
 a single system that can learn to play many different video
-games~\cite{NatureAtari}. The system uses a deep convolutional network to
+games~\cite{Deepmind-atari-arxiv2013}. The system uses a deep convolutional network to
 map from the pixels on the screen at several previous time steps to the
@@
 
-\subsection{Deep convolutional networks for image understanding}
-LeCun, Y., Bottou, L., Bengio, Y., \& Haffner, P. (1998)\\ Gradient-based
-learning applied to document recognition.\\ {\it Proceedings of the IEEE},
-86(11), 2278-2324.
+%\subsection{Deep convolutional networks for image understanding}
+%LeCun, Y., Bottou, L., Bengio, Y., \& Haffner, P. (1998)\\ Gradient-based
+%learning applied to document recognition.\\ {\it Proceedings of the IEEE},
+%86(11), 2278-2324.
+% --> \citep{LeCun+98}
 
@@
 
-Tompson, J., Goroshin, R., Jain, A., LeCun, Y., \& Bregler, C. (2014)
-Efficient Object Localization Using Convolutional
-Networks. arXiv:1411.4280.
+%Tompson, J., Goroshin, R., Jain, A., LeCun, Y., \& Bregler, C. (2014)
+%Efficient Object Localization Using Convolutional
+%Networks. arXiv:1411.4280.
+% --> \citep{Tompson-et-al-arxiv2014}
 
-\subsection{Deep recurrent neural networks for text understanding and machine translation}
+%\subsection{Deep recurrent neural networks for text understanding and machine translation}
 
-Graves, A., Liwicki, M., Fernández, S., Bertolami, R., Bunke, H., \&
-Schmidhuber, J. (2009)\\ A novel connectionist system for unconstrained
-handwriting recognition.\\ {\it IEEE Transactions on Pattern Analysis and
-  Machine Intelligence}, 31(5), 855-868.
+%Graves, A., Liwicki, M., Fernández, S., Bertolami, R., Bunke, H., \&
+%Schmidhuber, J. (2009)\\ A novel connectionist system for unconstrained
+%handwriting recognition.\\ {\it IEEE Transactions on Pattern Analysis and
+%  Machine Intelligence}, 31(5), 855-868.
+% --> \citep{Graves-et-al-2009}
 
@@
 %learning with neural networks.\\ {\it arXiv preprint}, arXiv:1409.3215.
+
 % --> \citep{Sutskever-et-al-NIPS2014}
@@
 
-\subsection{Deep reinforcement learning}
-Mnih, V., Kavukcuoglu, K., Silver, D., Graves, A., Antonoglou, I.,
-Wierstra, D., \& Riedmiller, M. (2013)\\ Playing Atari with deep
-reinforcement learning.\\ {\it arXiv preprint} arXiv:1312.5602.
+%\subsection{Deep reinforcement learning}
+%Mnih, V., Kavukcuoglu, K., Silver, D., Graves, A., Antonoglou, I.,
+%Wierstra, D., \& Riedmiller, M. (2013)\\ Playing Atari with deep
+%reinforcement learning.\\ {\it arXiv preprint} arXiv:1312.5602.
+% --> \citep{Deepmind-atari-arxiv2014}
 
-\fi
+%\fi
 

### 2015-01-11 deepreview.tex
cover reference and a little explanation

@@
 
-This review is organized chronlogically starting with the introduction of
+This review is organized chronologically starting with the introduction of
 the backpropagation algorithm in the 1980's. After describing some of its
@@
 through a non-linear mapping, exploiting the fact that collections of
-vectors are more likely to be easily spearated when they are spread around
-a high-dimensional space~\cite{cover}. One clever way to do so is to use a
+vectors are more likely to be easily separated when they are spread around
+a high-dimensional space~\cite{Cover65}. In the limit when there are at least
+as many dimensions as examples, a linear classifier can always separate the
+training examples, and one clever way to obtain a very high-dimensional space is to use a
 ``kernel function'' that measures the similarity between two vectors.
@@
 models~\citet{BenDucVin01-short} the standard approach to statistical
-modeling of language was based on counting frequencies of occurences of
+modeling of language was based on counting frequencies of occurrences of
 short symbol sequences (called n-grams). One disadvantage is that it
@@
 When translating between languages, the hidden units of the decoder network
-that generates possible translations are initalized with a ``thought
+that generates possible translations are initialized with a ``thought
 vector'' which is the final hidden state of an encoder network.  But it is

### 2015-01-11 deepreview.tex
fixed eqn 2, assuming the missing defn was of tanh

@@
 machine learning techniques at predicting the activity of potential drug
-molecules~\cite{Dahl}, detecting new fundamental particles~\cite{particles},
-classifying galaxies~\cite{??}, or predicting the effects of mutations in
-non-coding DNA on gene expression and disease~\cite{Frey}.
+molecules~\citep{Dahl}, detecting new fundamental particles~\citep{particles},
+classifying galaxies~\citep{??}, or predicting the effects of mutations in
+non-coding DNA on gene expression and disease~\citep{Frey}.
 
@@
 of the translation and so on until a full stop is 
-chosen~\cite{Vinyals-et-al-arxiv2014,Bahdanau-et-al-arxiv2014}. 
+chosen~\citep{Vinyals-et-al-arxiv2014,Bahdanau-et-al-arxiv2014}. 
 %YB added a clarification
@@
 the view that everyday reasoning involves many simultaneous analogies that
-each contribute plausibility to a conclusion\cite{metaphors}.
+each contribute plausibility to a conclusion~\citep{metaphors}.
 
@@
 vectors are more likely to be easily separated when they are spread around
-a high-dimensional space~\cite{Cover65}. In the limit when there are at least
+a high-dimensional space~\citep{Cover65}. In the limit when there are at least
 as many dimensions as examples, a linear classifier can always separate the
@@
 optimize the choice of training inputs used to define features and the
-class-specific weights associated with the features~\cite{Boser92}. The training
+class-specific weights associated with the features~\citep{Boser92}. The training
 inputs used to define features are called ``support vectors'' and Support
@@
 
-From the earliest days of pattern recognition~\cite{selfridge,Rosenblatt57}
+From the earliest days of pattern recognition~\citep{selfridge,Rosenblatt57}
 researchers aimed to replace hand-engineered features by multiple layers of
@@
 gradients to adapt the learning rate for each weight~\citep{bottou-bousquet-2008-small}
-or to provide additional increments that depend on earlier gradients~\cite{momentum}
+or to provide additional increments that depend on earlier gradients~\citep{momentum}
 but these techniques are still the subject of controversy and are beyond
@@
 \begin{equation}
-p_j = \frac{exp(z_j)}{\sum_i exp(z_i)}
+p_j = \frac{e^{z_j}}{\sum_i e^{z_i}}
 \end{equation}
 where $z_j$ is the total input received by output unit $j$. The hidden
-layers use either the logistic or tanh non-linearity:
+layers use either the logistic or tanh non-linearity, respectively:
 \begin{equation}
-y_j = \frac{1}{1+ \exp(z_j)}, \ \ \ y_j = \frac{}{} % YB: what was the intent here for the 2nd part?
+y_j = \frac{1}{1+ e^{z_j}}, \ \ \ y_j = \frac{e^{z_j}-e^{-z_j}}{e^{z_j}+e^{-z_j}} % YB: what was the intent here for the 2nd part?
 \end{equation}
@@
 each of which can be interpreted as a separate feature of the word. This
-was first demonstrated\cite{RHW} using sequences of three words, such as
+was first demonstrated~\citep{RHW} using sequences of three words, such as
 (Colin, Mother, Victoria), derived from two family trees. The learned word
@@
 system could be fine-tuned using standard
-backpropagation\cite{Hinton-Science2006}. This worked remarkably well for recognizing
+backpropagation\citep{Hinton-Science2006}. This worked remarkably well for recognizing
 handwritten digits especially when the amount of labeled data was very
@@
 It achieved record-breaking results on a standard speech recognition
-benchmark that used a small vocabulary\cite{TIMITpaper} and was quickly
+benchmark that used a small vocabulary~\citep{TIMITpaper} and was quickly
 developed to give record-breaking results on a large vocabulary
-task~\cite{Dahl}.  By 2012, versions of the deep net from 2009 were being
-developed by many of the major speech groups\cite{Hinton-et-al-2012} and were already
+task~\citep{Dahl}.  By 2012, versions of the deep net from 2009 were being
+developed by many of the major speech groups\citep{Hinton-et-al-2012} and were already
 being deployed in Android phones.  For smaller datasets, unsupervised
@@
 excellent results in a competition to read cursive
-hand-writing~\cite{Graves-et-al-2009}. They have subsequently proved to be very
+hand-writing~\citep{Graves-et-al-2009}. They have subsequently proved to be very
 effective for tasks that have sequential input.  They are now better than
-deep feed-forward networks for speech recognition~\cite{gravestimit} and
+deep feed-forward networks for speech recognition~\citep{gravestimit} and
 unlike feed-forward nets they can implement an entire speech recognition
 system that goes all the way from coefficents derived from the soundwave to
-the sequence of characters in the transcription~\cite{gravestimit}.  They
+the sequence of characters in the transcription~\citep{gravestimit}.  They
 are also what is currently used for the encoder and decoder networks that
@@
 %involving very long-range temporal structure and for real tasks like
-%predicting the next word in text\cite{LeHinton}.
+%predicting the next word in text\citep{LeHinton}.
 
@@
 be trained by BPTT on a dataset of captioned
-images~\cite{NIC,StanfordNIC,more?} and it produces remarkably good
+images~\citep{NIC,StanfordNIC,more?} and it produces remarkably good
 captions (see Figure~\ref{fig:caption-generation}).
@@
 a single system that can learn to play many different video
-games~\cite{Deepmind-atari-arxiv2013}. The system uses a deep convolutional network to
+games~\citep{Deepmind-atari-arxiv2013}. The system uses a deep convolutional network to
 map from the pixels on the screen at several previous time steps to the
@@
 recurrent nets that use reinforcement learning to decide where to
-look\cite{}.  Such systems already outperform passive vision
-systems\cite{ba+mnih}.  Similarly, when understanding sentences or whole
+look~\citep{rl+attention}.  Such systems already outperform passive vision
+systems~\citep{ba+mnih}.  Similarly, when understanding sentences or whole
 documents we expect future systems to learn strategies for selectively
-attending to one part at a time~\cite{Bahdanau-et-al-arxiv2014}. We also expect
+attending to one part at a time~\citep{Bahdanau-et-al-arxiv2014}. We also expect
 that in the next few years deep recurrent neural networks will become far
@@
 untangle the many different factors that interact to produce images, speech
-and documents~\cite{Yann,Bengio-Courville-Vincent-TPAMI2013}.
+and documents~\citep{Yann,Bengio-Courville-Vincent-TPAMI2013}.
 
@@
 
-\iffalse
-
 %Hinton, G., Deng, L., Yu, D., Dahl, G. E., Mohamed, A. R., Jaitly, N. \&
@@
 %Neural Information Processing Systems (1097-1105).
-% --> \cite{Krizhevsky-2012-small}
+% --> \citep{Krizhevsky-2012-small}
 
@@
 
-%\fi
-
 \newpage
@@
 
-2. Maps showing:\\ a) that word vectors learned by backpropagation are very
+DONE 2. Maps showing:\\ a) that word vectors learned by backpropagation are very
 good at capturing the meanings of the words.\\ b) that the thought vectors
@@
 4. Some images with:\\ a) The opinions of a deep convolutional net about
-the classes of the prominent objects.\\ b) Several alternative captions
+the classes of the prominent objects.\\ 
+DONE b) Several alternative captions
 generated by using the output of the convolutional net to determine the

### 2015-01-11 deepreview.tex
removed the 'in English' about LSTMs

@@
 
-Backpropagation provides an efficient way of computing gradients for all of
+Backpropagation provides an efficient way of computing gradients with respect to all of
 the weights in a layered network on a single training case, but there are
@@
 that allow the neurons within a layer to receive weighted input from those
-same neurons at the previous time-step.  The gradient for a recurrent
+same neurons at the previous time-step.  The gradient with respect to a recurrent
 weight, $w_{ij}$ that allows neuron $i$ at time $t$ to influence the
@@
 backpropagation-through-time (BPTT) of the error derivative for $j$ at time
-$t+1$. The full gradient for $w_{ij}$ on a whole input sequence is just the
+$t+1$. The full gradient with respect to $w_{ij}$ on a whole input sequence is just the
 sum of its gradients at each time-step.
@@
 
-When LSTMs were first described in
-English~\citep{Hochreiter+Schmidhuber-1997} their complexity was off-putting
+When LSTMs were first described % in English
+\citep{Hochreiter+Schmidhuber-1997} their complexity was off-putting
 and, despite their impressive ability to learn very long-range temporal

### 2015-01-11 deepreview.tex
added a reference on the importance of the gating of the self-loop gate for LSTMs

@@
 
-Recent work~\citep{Chung-et-al-NIPSDL2014} suggests that not all the elements of the LSTM
+Recent work~\citep{Chung-et-al-NIPSDL2014,Yao-et-al-SLU-workshop2014} 
+suggests that not all the elements of the LSTM
 architecture are required, the crucial element being the presence of

### 2015-01-11 deepreview.tex
added a reference on the importance of the gating of the self-loop gate for LSTMs

--- 2015-01-11 deepreview.tex
@@
 
-Recent work~\citep{Chung-et-al-NIPSDL2014} suggests that not all the elements of the LSTM
-architecture are required, the crucial element being the presence of
-of the gate controlling the leaking rate (or time constant) of the cell.
+Recent work~\citep{Chung-et-al-NIPSDL2014,Yao-et-al-SLU-workshop2014} 
+suggests that not all the elements of the LSTM
+architecture are required, the crucial element being the presence and learnability 
+of the so-called forget gate, controlling the leaking rate (or time constant) of the cell.
 %It is not yet clear which aspects of the complicated architecture of LSTMs
@@
 short symbol sequences (called n-grams). One disadvantage is that it
-taking into account context beyond a handful of words would require
+takes into account context beyond a handful of words would require
 training corpora of exponentially growing size (with respect to the

### 2015-01-11 deepreview.tex
clarifying a few points

@@
 error-derivatives for the weights that define each feature detector and
-then repeatedly change these weights in proportion to their derivatives
+then repeatedly changing these weights in proportion to their derivatives
 averaged over small batches of training examples.  Deep recurrent neural
@@
 molecules~\citep{Dahl}, detecting new fundamental particles~\citep{particles},
-classifying galaxies~\citep{??}, or predicting the effects of mutations in
+classifying galaxies~\citep{??}, and predicting the effects of mutations in
 non-coding DNA on gene expression and disease~\citep{Frey}.
@@
 matrices and scalar non-linearities to get the job done.  This suggests
-that deriving conclusions using only the form of the premises is a powerful
+that deriving conclusions using only the form of the premises (which is
+what symbolic systems do) is a powerful
 technique that people can eventually master but it is not a good model of
@@
 Backpropagation is simply an application of
-the chain rule, which yields an efficient way of computing the gradient of a
+the chain rule of derivatives,  % clarifying which chain rule, not the probabilities one
+which yields an efficient way of computing the gradient of a
 smooth objective function with respect to all of the weights in a layered
@@
 direction where a small change will most improve the objective function. 
-When the training set contains millions of cases and is highly redundant, however,
+When the training set contains millions of cases and is highly redundant
+% explaining what ``redundant'' means, here
+(with many examples illustrating variations of the the same underlying 
+aspect or concept), however,
 this is needlessly slow. It is much faster to compute the average gradient
@@
 logistic gating unit that has an activation between $0$ and $1$. The input
-vector has little effect on the state of a memory cell if the cell's input
+vector has little effect on the state of that cell if the cell's input
 gate has a low activation so memory cells can learn to selectively ignore
@@
 modeling of language was based on counting frequencies of occurrences of
-short symbol sequences (called n-grams). One disadvantage is that it
-takes into account context beyond a handful of words would require
+short symbol sequences (called n-grams). One disadvantage of the n-gram
+approach is that taking into account context more than a handful of words would require
 training corpora of exponentially growing size (with respect to the
@@
 problem and the very notion of symbol introduced above, n-grams
-do not generalize across the many semantically related sequences 
-of words (for example the many ways of expressing roughly the same thing).
-Instead, neural language models learn to associate each word to a
+treat each word as an atomic unit, so cannot generalize across
+semantically related sequences of words,
+while we know that there are many ways of expressing roughly the same thing.
+Instead, neural language models can generalize across semantically related words
+and semantically related phrases by learning to associate each word to a
 distributed representation (a vector) such that semantically related
@@
 When translating between languages, the hidden units of the decoder network
-that generates possible translations are initialized with a ``thought
+that generates possible translations are conditioned on (i.e., receive as extra
+input or as initialization) a ``thought
 vector'' which is the final hidden state of an encoder network.  But it is

### 2015-01-11 deepreview.tex
added box on back-prop

@@
 %%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[H]
+\centerline{
+\includegraphics[width=0.33\linewidth]{chain-rule.pdf}
+\includegraphics[width=0.33\linewidth]{forward-graph.pdf}
+\includegraphics[width=0.33\linewidth]{backward-graph.pdf}
+}
+\caption{
+Left: the chain rule of derivatives tells us how
+two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
+are composed. A small change $\Delta x$ in $x$
+gets transformed first into a small change $\Delta y$ in $y$
+by getting multiplied by $\frac{\partial y}{\partial x}$
+(that is the definition of partial derivative). Similarly
+that creates a change $\Delta z$ in $z$. Inserting one equation
+into the other gives the chain rule of derivatives, i.e., how
+$\Delta x$ gets turned into $\Delta z$ through multiplication
+by the product of $\frac{\partial y}{\partial x}$ and
+$\frac{\partial z}{\partial x}$. It also works when $x$, $y$
+and $z$ are vectors (and the derivatives are Jacobian matrices). \newline
+Middle: many computations, such as the succession of linear and non-linear 
+transformations of multi-layer neural networks, can be described as
+a graph where each node $i$ computes a quantity $u_i$ (e.g. the
+output of an artificial neuron or the value of a parameter such as
+a synaptic weight), based on the set of values $a_i$, composed of elements $u_j$ of
+predecessors in the graph, i.e., $u_i = f_i(a_i)$. \newline
+Right: the chain rule can be generalized to such computational
+flow graphs by recursively computing the derivative of the final
+node $u_N$ (e.g., the error to be minimized or the output of the neural network)
+with respect to every internal node. The basic recursive equation
+is
+$\frac{\partial u_N}{\partial u_j} = \sum_{i: j \in a_i} \frac{\partial u_N}{\partial u_i}
+\frac{\partial f(a_i)}{\partial u_j}.$
+}
+\label{fig:backprop-box}
+\end{figure}
 

### 2015-01-11 deepreview.tex
added figures and caption for fprop and bprop in mlp

@@
 they typically explode or vanish~\citep{Bengio-et-al-TNN1994,Hochreiter91-small}.
+
+\begin{figure}[H]
+\centerline{
+\includegraphics[width=0.49\linewidth]{bp-mlp1.pdf}
+\includegraphics[width=0.49\linewidth]{bp-mlp5.pdf}
+}
+\caption{Top left: graph of feedforward computations in a multi-layer
+neural network with one input layer ($x$), two hidden layers ($h_1$ and $h_2$),
+one output layer (out), a target $y$, and an error that compares the output
+and the target. The synaptic weight matrices $W_1$, $W_2$ and $W_3$ are
+the parameters of the neural network.\newline
+Top right: back-propagation through the same multi-layer neural network
+(downward arrows compute the partial derivative of the error with respect
+to every node, i.e., the units value and the parameters.\newline
+Bottom:
+}
+\label{fig:mlp}
+\end{figure}
 

### 2015-01-12 deepreview.tex
added RNN figure and caption

@@
 }
+\centerline{
+\includegraphics[width=0.89\linewidth]{fig-hidden-recurrence-rnn.pdf}
+}
 \caption{Top left: graph of feedforward computations in a multi-layer
@@
 (downward arrows compute the partial derivative of the error with respect
-to every node, i.e., the units value and the parameters.\newline
-Bottom:
+to every node, i.e., the units value and the parameters).\newline
+Bottom: a recurrent neural network (left) and the unfolding in time of the
+computation involved in its forward computation. The artificial neurons
+(e.g. hidden units grouped under node $s$ with values $s_t$ at time $t$) 
+get inputs from other neurons at previous time steps
+(this is represented with the black square, for delay of
+one time step, on the left). In this way, a recurrent net can map an input sequence
+with elements $x_t$ into an output sequence with elements $o_t$, with
+each $o_t$ depending on all the previous $x_{t'}$ (for $t'\leq t$). The
+same parameters (matrices $U$, $V$, $W$) are used at each time step.
+Many other architectures are possible, including a variant in which
+the outputs (e.g. words) can be generated according to a distribution 
+(the probability of the next word) and used as inputs for the next time step.
+The back-propagation algorithm (Figure~\ref{fig:backprop-box})
+can be directly applied to the computational graph of the unfolded network
+on the right hand side, to compute the derivative of a total error
+(at all time steps) with respect to all the states $s_t$ and all the parameters.
 }

### 2015-01-12 deepreview.tex
added boxes around the figures
fixed the {Frey} reference

@@
 \usepackage{natbib}
-
+\usepackage{float}
 \parskip 0.1in
@@
 classifying galaxies~\citep{??}, and predicting the effects of mutations in
-non-coding DNA on gene expression and disease~\citep{Frey}.
+non-coding DNA on gene expression and disease~\citep{Keung-et-al-2014}. % Frey
 
@@
 %%%%%%% GEOFF %%%%%%%%%%
-GEOFF: Do box explaining backprop equatons
+%GEOFF: Do box explaining backprop equatons
 %%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[H]
+\fbox{
+\begin{minipage}{\textwidth}
 \centerline{
@@
 \frac{\partial f(a_i)}{\partial u_j}.$
+}
+\end{minipage}
 }
@@
 \begin{figure}[H]
+\fbox{
+\begin{minipage}{\textwidth}
 \centerline{
@@
 (at all time steps) with respect to all the states $s_t$ and all the parameters.
+}
+\end{minipage}
 }
@@
 nearby representations.
+}
+\end{minipage}
 }
@@
 \begin{figure}[ht]
+\fbox{
+\begin{minipage}{\textwidth}
 \centerline{\includegraphics[width=\linewidth]{caption-generation-Vinyals-et-al.png}}
@@
 images into captions. Reproduced from~\citet{Vinyals-et-al-arxiv2014}.
+}
+\end{minipage}
 }
@@
 
+\iffalse
 \section{Boxes (TODO)}
@@
 
+\fi
+
 \bibliography{strings,ml,aigaion}

### 2015-01-12 deepreview.tex
fixing the wake-sleep, google-cat, and 'more recent' unsupervised citations

@@
 
-Recent work~\citep{Chung-et-al-NIPSDL2014,Yao-et-al-SLU-workshop2014} 
+Recent work~\citep{Chung-et-al-NIPSDL2014-small,Yao-et-al-SLU-workshop2014} 
 suggests that not all the elements of the LSTM
@@
 learning procedures for deep neural 
-networks~\citep{Salakhutdinov2009-small,wake-sleep,
-  google-cat, denoising-auto, ranzato-pami, more recent}. In the long run,
+networks~\citep{Salakhutdinov2009-small,Hinton95,QuocLe-ICML2012,VincentPLarochelleH2008-small,ranzato-pami,Kingma-et-al-NIPS2014}. 
+In the long run,
 however, we believe unsupervised learning will be essential for allowing

### 2015-01-12 deepreview.tex
added the caption generation citations

@@
 effective for tasks that have sequential input.  They are now better than
-deep feed-forward networks for speech recognition~\citep{gravestimit} and
+deep feed-forward networks for speech recognition~\citep{Graves-et-al-ICASSP2013} and
 unlike feed-forward nets they can implement an entire speech recognition
@@
 be trained by BPTT on a dataset of captioned
-images~\citep{NIC,StanfordNIC,more?} and it produces remarkably good
+images~\citep{Kiros-et-al-ICML2014,Mao+al-arxiv2014,Donahue-et-al-arxiv2014,
+              Vinyals-et-al-arxiv2014,
+             Chen+Zitnick-arxiv2014,Karpathy+Li-arxiv2014,Venugopalan-et-al-arxiv2014} 
+and such models produce remarkably good
 captions (see Figure~\ref{fig:caption-generation}).
@@
 and neural language modeling. There has been a surge of interest in such
-systems recently~\citep{Vinyals-et-al-arxiv2014,OTHER-CAPTION-GENERATION-PAPERS}.
+systems recently~\citep{Kiros-et-al-ICML2014,Mao+al-arxiv2014,Donahue-et-al-arxiv2014,
+              Vinyals-et-al-arxiv2014,Fang-et-al-arxiv2014,
+             Chen+Zitnick-arxiv2014,Karpathy+Li-arxiv2014,Venugopalan-et-al-arxiv2014}.
 

### 2015-01-12 deepreview.tex
added more image-to-caption examples

@@
 \centerline{\includegraphics[width=\linewidth]{caption-generation-Vinyals-et-al.png}}
-\caption{Example of captions generated by a recurrent network taking
+\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}
+\includegraphics[width=0.49\linewidth]{Examples_im2txt_002.png}}
+\caption{Examples of captions generated by a recurrent network taking
 as extra input the representation extracted by a deep convolutional
 net from an image, with the whole system trained to ``translate''
-images into captions. Reproduced from~\citet{Vinyals-et-al-arxiv2014}.
+images into captions. Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
 }

### 2015-01-12 deepreview.tex
Update deepreview.tex

@@
 each of which can be interpreted as a separate feature of the word. This
-was first demonstrated~\citep{RHW} using sequences of three words, such as
-(Colin, Mother, Victoria), derived from two family trees. The learned word
+was first demonstrated using sequences of three words, such as
+(Colin, Mother, Victoria), derived from two family trees~\citep{RHW}. The learned word
 vector for Colin contained ``semantic'' features representing which family
 tree Colin was in, what branch of that tree he was in, and from what generation
-he was~\citep{Hinton-et-al-PDP1986}. These semantic features were not explicitly present
+he was. These semantic features were not explicitly present
 in the input.  They were discovered by the learning procedure as a good way

### 2015-01-12 deepreview.tex
fixed incorrect citation to deep MT papers

--- 2015-01-12 deepreview.tex
@@
 expressed by the sentence.  This thought vector can then be used as the
-initial hidden state of a trained French ``decoder'' network which outputs
+initial hidden state (or as extra input) of a trained French ``decoder'' 
+network which outputs
 a probability distribution for the first word of the French translation. If
@@
 of the translation and so on until a full stop is 
-chosen~\citep{Vinyals-et-al-arxiv2014,Bahdanau-et-al-arxiv2014}. 
-%YB added a clarification
+chosen~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}. 
 Overall, this process generates sequences of French words according to

### 2015-01-12 deepreview.tex
added an explanation and citation about the trouble with kernel machines

@@
 classification methods are logistic regression and linear support vector
-machines~\citep{logisticregression,Boser92}.
+machines~\citep{bishop-book95,Boser92}.
 
@@
 vectors are more likely to be easily separated when they are spread around
-a high-dimensional space~\citep{Cover65}. In the limit when there are at least
+a high-dimensional space~\citep{Cover65}. When there are at least
 as many dimensions as examples, a linear classifier can always separate the
@@
 ${\bf x}_{\rm new}$. Given a kernel function, it is possible to jointly
-optimize the choice of training inputs used to define features and the
+optimize the choice of training examplars used to define features and the
 class-specific weights associated with the features~\citep{Boser92}. The training
-inputs used to define features are called ``support vectors'' and Support
+examplars used to define features are called ``support vectors'' and Support
 Vector Machines~\citep{Boser92} have been very successful, especially for
@@
 the input-output mapping is so complex that computing similarities with a
-simple kernel function is ineffective.
+simple kernel function is ineffective. With a general-purpose feature space
+and kernel (such as the Gaussian kernel), generalization far from the
+examples is poor~\citep{Bengio-localfailure-NIPS-2006-small}, which is problematic
+in high-dimensional spaces. However, kernel machines would perform well if a
+very good set of features were provided: learning very good
+features is what deep learning can do.
 

### 2015-01-12 deepreview.tex
avoiding strong claims, pointing to figures

@@
 particular minute variations, such as the difference between a white wolf
-and a samoyed dog. These tasks are better handled with multiple layers of
+and a samoyed dog. Experiments and theory of deep learning both
+suggest that these tasks are better handled with multiple layers of
 progressively more complicated feature detectors each of which is defined
@@
 Backpropagation is simply an application of
-the chain rule of derivatives,  % clarifying which chain rule, not the probabilities one
+the chain rule of derivatives (left of Figure~\ref{fig:backprop-box}),
 which yields an efficient way of computing the gradient of a
 smooth objective function with respect to all of the weights in a layered
-network. The objective function typically measures the discrepancy between
+network (Figures~\ref{fig:backprop-box} and~\ref{fig:mlp}). 
+The objective function typically measures an error to be minimized: 
+the discrepancy between
 the actual output of the network on a given training case and the target
@@
 error derivatives for the outputs of all layers, it is easy to compute
-error derivatives for the parameters (or weights) of each layer (see box
-1).
+error derivatives for the parameters (or weights) of each layer 
+(Figure~\ref{fig:mlp}).
 

### 2015-01-12 deepreview.tex
added the tricks of the trade book citation

@@
 by getting multiplied by $\frac{\partial y}{\partial x}$
-(that is the definition of partial derivative). Similarly
-that creates a change $\Delta z$ in $z$. Inserting one equation
+(that is the definition of partial derivative). Similarly,
+the change $\Delta y$ creates a change $\Delta z$ in $z$. Substituting one equation
 into the other gives the chain rule of derivatives, i.e., how
@@
 a synaptic weight), based on the set of values $a_i$, composed of elements $u_j$ of
-predecessors in the graph, i.e., $u_i = f_i(a_i)$. \newline
-Right: the chain rule can be generalized to such computational
+predecessors in the graph, i.e., $u_i = f_i(a_i)$. This can be readily
+generalized to nodes whose output is a vector or a matrix.\newline
+Right: the chain rule can be applied efficiently to computational
 flow graphs by recursively computing the derivative of the final
@@
 % explaining what ``redundant'' means, here
-(with many examples illustrating variations of the the same underlying 
+(with many examples illustrating variations of the same underlying 
 aspect or concept), however,
@@
 on a small ``mini-batch'' of randomly chosen training cases and to update
-each weight by this stochastic estimate of its full gradient times a
+each weight by this stochastic estimate of its full gradient, times a
 learning rate which decays as learning proceeds. Stochastic gradient
@@
 but these techniques are still the subject of controversy and are beyond
-the scope of this review.  However, it is now widely accepted that some
+the scope of this review (but see~\citet{Montavon2012} for tricks of the trade). 
+However, it is now widely accepted that some
 form of stochastic gradient descent is the best bet for large, redundant

### 2015-01-12 deepreview.tex
added the rectifier nonlinearity

@@
 where $z_j$ is the total input received by output unit $j$. The hidden
-layers use either the logistic or tanh non-linearity, respectively:
+layers use a non-linearity such as the logistic, tanh, or rectifier, respectively:
 \begin{equation}
-y_j = \frac{1}{1+ e^{z_j}}, \ \ \ y_j = \frac{e^{z_j}-e^{-z_j}}{e^{z_j}+e^{-z_j}} % YB: what was the intent here for the 2nd part?
+y_j = \frac{1}{1+ e^{z_j}}, \ \ \ y_j = \frac{e^{z_j}-e^{-z_j}}{e^{z_j}+e^{-z_j}}, \ \ \
+ y_j = \max(0, z_j)
 \end{equation}

### 2015-01-12 deepreview.tex
Update deepreview.tex
Geoff removed two \cite{schmidreference}  and \cite{momentum}

@@
 backpropagation algorithm which was discovered independently by several
-different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,Rumelhart86b,schmidreference}.  
+different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}.  
 Backpropagation is simply an application of
@@
 gradients to adapt the learning rate for each weight~\citep{bottou-bousquet-2008-small}
-or to provide additional increments that depend on earlier gradients~\citep{momentum}
+or to provide additional increments that depend on earlier gradients 
 but these techniques are still the subject of controversy and are beyond

### 2015-01-12 deepreview.tex
Update deepreview.tex
got rid of several cites:  
gravestimit (already there with different name)
rl+attention
Yann

Am adding ba+mnih to bib file

The Dahl is for the Kaggel competition. Lets just forget to put that in the refs for now.

@@
 machine learning techniques at predicting the activity of potential drug
-molecules~\citep{Dahl}, detecting new fundamental particles~\citep{particles},
-classifying galaxies~\citep{??}, and predicting the effects of mutations in
+molecules~\citep{DahlKaggle}, detecting new fundamental particles~\citep{YANNparticles},
+classifying galaxies~\citep{YANNgalaxies}, and predicting the effects of mutations in
 non-coding DNA on gene expression and disease~\citep{Keung-et-al-2014}. % Frey
@@
 system that goes all the way from coefficents derived from the soundwave to
-the sequence of characters in the transcription~\citep{gravestimit}.  They
+the sequence of characters in the transcription.  They
 are also what is currently used for the encoder and decoder networks that
@@
 recurrent nets that use reinforcement learning to decide where to
-look~\citep{rl+attention}.  Such systems already outperform passive vision
+look.  Such systems already outperform passive vision
 systems~\citep{ba+mnih}.  Similarly, when understanding sentences or whole
@@
 untangle the many different factors that interact to produce images, speech
-and documents~\citep{Yann,Bengio-Courville-Vincent-TPAMI2013}.
+and documents~\citep{Bengio-Courville-Vincent-TPAMI2013}.
 

### 2015-01-12 deepreview.tex
moved Geoff's refs from aigaion.bib to ml.bib and added a GSN citation

@@
 learning procedures for deep neural 
-networks~\citep{Salakhutdinov2009-small,Hinton95,QuocLe-ICML2012,VincentPLarochelleH2008-small,ranzato-pami,Kingma-et-al-NIPS2014}. 
+networks~\citep{Salakhutdinov2009-small,Hinton95,QuocLe-ICML2012,VincentPLarochelleH2008-small,ranzato-pami,Bengio-et-al-ICML-2014,Kingma-et-al-NIPS2014}. 
 In the long run,

### 2015-01-13 deepreview.tex
found the Dahl Kaggle paper and put it in ml.bib

@@
 machine learning techniques at predicting the activity of potential drug
-molecules~\citep{DahlKaggle}, detecting new fundamental particles~\citep{YANNparticles},
+molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental particles~\citep{YANNparticles},
 classifying galaxies~\citep{YANNgalaxies}, and predicting the effects of mutations in
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[H]
+\begin{figure}[ht]
 \fbox{
@@
 gradients either grow or shrink at each time step, so over many time-steps
-they typically explode or vanish~\citep{Bengio-et-al-TNN1994,Hochreiter91-small}.
+they typically explode or vanish~\citep{Hochreiter91-small,Bengio-et-al-TNN1994},
+and the conditions which allow stable dynamics also correspond to
+vanishing gradients~\citep{Bengio-et-al-TNN1994},
 
-\begin{figure}[H]
+\begin{figure}[ht]
 \fbox{
@@
 get inputs from other neurons at previous time steps
-(this is represented with the black square, for delay of
+(this is represented with the black square, standing for a delay of
 one time step, on the left). In this way, a recurrent net can map an input sequence
@@
 Many other architectures are possible, including a variant in which
-the outputs (e.g. words) can be generated according to a distribution 
-(the probability of the next word) and used as inputs for the next time step.
+the sampled outputs (e.g. words) can be generated according to an output 
+distribution (the probability of the next word, for every possible next word) 
+and used as inputs for the next time step.
 The back-propagation algorithm (Figure~\ref{fig:backprop-box})
@@
 on the right hand side, to compute the derivative of a total error
-(at all time steps) with respect to all the states $s_t$ and all the parameters.
+(e.g., the log-probability of generating the right sequence of
+words) with respect to all the states $s_t$ and all the parameters.
 }
@@
 The labeled datasets were far too small, the computers were far too slow,
-and the weights were not initialised sensibly.
+and the weights were not initialised sensibly~\citep{GlorotAISTATS2010-small}.
 
@@
 results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-arxiv2014} strongly suggest
-that in deep networks almost all of these local minima give very similar
-performance so it does not matter which one we find.  This goes a long way
+the vast majority of critical points (where derivatives of the total
+objective function with respect to parameters vanish) are saddle points, not
+local minima, and that in large deep networks almost all of these local minima 
+give very similar performance so it does not matter which one we find.  This goes a long way
 towards explaining why stochastic gradient descent works so well for deep
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
 \fbox{
@@
 \includegraphics[width=0.49\linewidth]{Examples_im2txt_002.png}}
-\caption{Examples of captions generated by a recurrent network taking
+\caption{Three examples of captions generated by a recurrent network taking
 as extra input the representation extracted by a deep convolutional
-net from an image, with the whole system trained to ``translate''
-images into captions. Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
+net from a test image, with the whole system trained to ``translate''
+images into captions. GT stands for ``Ground Truth'' and was labeled by humans
+whereas the other text is machine-generated.
+Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
 }

### 2015-01-13 deepreview.tex
found and plugged the other missing Dahl reference for large-vocabulary ASR

@@
 developed to give record-breaking results on a large vocabulary
-task~\citep{Dahl}.  By 2012, versions of the deep net from 2009 were being
+task~\citep{Dahl2012}.  By 2012, versions of the deep net from 2009 were being
 developed by many of the major speech groups\citep{Hinton-et-al-2012} and were already

### 2015-01-13 deepreview.tex
stub for missing fig:imagenet-output

@@
 recognition and it was made possible by the advent of fast Graphics
-Processing Units (GPUs) that were convenient to program~\citep{RainaICML09-small}.  
-In 2009, the approach was used to map short temporal windows of coefficients extracted
-from a soundwave to a set of probabilities for the various fragments of
-speech that might be represented by the frame in the center of the window.
-It achieved record-breaking results on a standard speech recognition
-benchmark that used a small vocabulary~\citep{TIMITpaper} and was quickly
-developed to give record-breaking results on a large vocabulary
-task~\citep{Dahl2012}.  By 2012, versions of the deep net from 2009 were being
-developed by many of the major speech groups\citep{Hinton-et-al-2012} and were already
-being deployed in Android phones.  For smaller datasets, unsupervised
+Processing Units (GPUs) that were convenient to
+program~\citep{RainaICML09-small}.  In 2009, the approach was used to map
+short temporal windows of coefficients extracted from a soundwave to a set
+of probabilities for the various fragments of speech that might be
+represented by the frame in the center of the window.  It achieved
+record-breaking results on a standard speech recognition benchmark that
+used a small vocabulary~\citep{TIMITpaper} and was quickly developed to
+give record-breaking results on a large vocabulary task~\citep{Dahl2012}.
+By 2012, versions of the deep net from 2009 were being developed by many of
+the major speech groups~\citep{Hinton-et-al-2012} and were already being
+deployed in Android phones.  For smaller datasets, unsupervised
 pre-training helps to prevent overfitting~\citep{Erhan+al-2010-small},
-leading to significantly better generalization when the number of
-labeled examples is small. Once deep
+leading to significantly better generalization when the number of labeled
+examples is small, or in a transfer setting, i.e., when we have lots of
+examples for some ``source'' tasks but very few for some ``target''
+task. In this context, unsupervised representation learning helped to win
+several transfer learning
+competitions~\citep{UTLC+LISA-2011-small,Goodfellow-icml2012}.  Once deep
 learning had been rehabilitated, it turned out that large labeled datasets
@@
 random weights were carefully chosen to avoid very large or very small
-initial gradients.
+initial gradients~\citep{GlorotAISTATS2010-small} and that even better
+results (especially for very deep networks)
+could be obtained by using rectifying non-linearities rather than
+sigmoid or tanh
+non-linearities~\citep{Glorot+al-AI-2011-small,Krizhevsky-2012-small}.
 
@@
 the input.  Similarly, a memory cell has little effect on the output if its
-output gate has low activation. Each memory cell also has a connection to
-itself (it is a leaky neuron) at the next time-step that has a weight of 1 so it tends to copy its
-own real-valued state, but this self-connection is gated by a ``keep gate'' so
-the stored value leaks quickly unless the keep-gate has an activation
+output gate has low activation. Each memory cell is a gated leaky neuron, i.e.,
+it also has a connection to
+itself at the next time-step that has a weight of 1 (so it tends to copy its
+own real-valued state and accumulate the external signal), 
+but this self-connection is {\em gated} by a ``forget gate'' so
+the stored value leaks quickly unless the forget gate has an activation
 near $1$.  The gates themselves have their activities determined by
@@
 When LSTMs were first described % in English
-\citep{Hochreiter+Schmidhuber-1997} their complexity was off-putting
+\citep{Hochreiter+Schmidhuber-1997},
+their complexity might have been off-putting
 and, despite their impressive ability to learn very long-range temporal
@@
 hand-writing~\citep{Graves-et-al-2009}. They have subsequently proved to be very
-effective for tasks that have sequential input.  They are now better than
+effective for tasks that have sequential input.  Combined with depth of
+representation, they are now better than
 deep feed-forward networks for speech recognition~\citep{Graves-et-al-ICASSP2013} and
@@
 system that goes all the way from coefficents derived from the soundwave to
-the sequence of characters in the transcription.  They
-are also what is currently used for the encoder and decoder networks that
+the sequence of characters in the transcription.  LSTMs or related forms of gated
+units are also what is currently used for the encoder and decoder networks that
 perform so well at machine
@@
 architecture are required, the crucial element being the presence and learnability 
-of the so-called forget gate, controlling the leaking rate (or time constant) of the cell.
+of the forget gate, controlling the leaking rate (or time constant) of the cell.
 %It is not yet clear which aspects of the complicated architecture of LSTMs
@@
 Cite at least \citep{LeCun+98} and \citep{Tompson-et-al-arxiv2014}
+
+\begin{figure}[ht]
+\fbox{
+\begin{minipage}{\textwidth}
+\centerline{
+\includegraphics[width=0.89\linewidth]{}
+}
+\caption{
+}
+\end{minipage}
+}
+\label{fig:imagenet-outputs}
+\end{figure}
 

### 2015-01-13 deepreview.tex
formatting changes

@@
-\documentclass[]{article}
+\documentclass[10pts]{article}
+\usepackage{times}
 \usepackage{epsfig}
@@
 \usepackage{float}
+
+\addtolength{\textwidth}{2cm}
+%\addtolength{\textwidth}{1cm}
+\addtolength{\textheight}{2cm}
+%\addtolength{\textheight}{1cm}
+\addtolength{\topmargin}{-1cm}
+\addtolength{\oddsidemargin}{-1cm}
+\addtolength{\evensidemargin}{-1cm}
+
 \parskip 0.1in
@@
 they achieved spectacular results, almost halving the error rates of the
-best computer vision sytems~\citep{Krizhevsky-2012-small}.  Some of the images and the
-network's predictions are shown in figure~\ref{fig:imagenet-outputs}.  
+best computer vision sytems~\citep{Krizhevsky-2012-small}.  
+%Some of the images and the
+%network's predictions are shown in figure~\ref{fig:imagenet-outputs}.  
+The same kind of network is used to generate the captions shown in Figure~\ref{fig:caption-generation}.
 The major factors in this
@@
 
+\iffalse
 \begin{figure}[ht]
@@
 \centerline{
-\includegraphics[width=0.89\linewidth]{}
+%\includegraphics[width=0.89\linewidth]{}
 }
@@
 Prior to the introduction of neural language
-models~\citet{BenDucVin01-short} the standard approach to statistical
+models~\citep{BenDucVin01-short}, the standard approach to statistical
 modeling of language was based on counting frequencies of occurrences of
 short symbol sequences (called n-grams). One disadvantage of the n-gram
-approach is that taking into account context more than a handful of words would require
+approach is that taking into account context of more than a handful of words would require
 training corpora of exponentially growing size (with respect to the
 desired context size), since the number of configurations of $n$
-words from a vocabulary of size $V$ grows as $V^n$. Related to this
+words from a vocabulary of size $V$ grows as $V^n$. Although this can
+be mitigated by only considering the more frequent of the longer sequences of 
+words, it means that less frequent sequences of words are modeled in a 
+context-blind way. Related to this
 problem and the very notion of symbol introduced above, n-grams
@@
 Instead, neural language models can generalize across semantically related words
-and semantically related phrases by learning to associate each word to a
+and semantically related phrases by learning to associate each word with a
 distributed representation (a vector) such that semantically related
@@
 of translated words given the sequence of source words, i.e., neural
-machine translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
+machine translation~\citep{Devlin-et-al-ACL2014,Bahdanau-et-al-arxiv2014,
+Sutskever-et-al-NIPS2014}.
 
@@
 also possible to initialize a decoder network using a linear transformation
-of the last hidden layer of a deep convolutonal network that has already
+of the last hidden layer of a deep convolutional network that has already
 been trained to recognize objects in images. The decoder network can then

### 2015-01-13 deepreview-v2.tex
made v2

--- 2015-01-13 deepreview.tex
@@
 
-Deep neural networks that learn many layers of feature detectors have
-dramatically improved the state-of-the-art for speech recognition, object
-recognition, object detection, predicting the activity of drug molecules,
-and many other tasks. Deep learning discovers intricate structure in large
-datasets by using the backpropagation algorithm to compute
-error-derivatives for the weights that define each feature detector and
-then repeatedly changing these weights in proportion to their derivatives
-averaged over small batches of training examples.  Deep recurrent neural
-networks, in which each layer also receives input from its own state at the
-previous time-step, can now generate good captions for images and are
-competitive for machine translation.
+Deep neural networks that learn to represent data in multiple layers
+of increasing abstration have dramatically improved the
+state-of-the-art for speech recognition, object recognition, object
+detection, predicting the activity of drug molecules, and many other
+tasks. Deep learning discovers intricate structure in large datasets
+by using the backpropagation algorithm to compute error-derivatives
+for the weights that define each layer of feature detector and then
+repeatedly changing these weights in proportion to their derivatives
+averaged over small batches of training examples.  Deep recurrent
+neural networks, in which each layer also receives input from its own
+state at the previous time-step, can now generate good captions for
+images and are competitive for language translation.
 
@@
 
-Deep learning is currently making major advances in solving problems that
-have resisted the best attempts of the Artificial Intelligence community
-for many years. Deep learning is very good at discovering intricate
-structure in large, high-dimensional datasets and is therefore widely
-applicable in all of the sciences. For example, it has beaten previous
-machine learning techniques at predicting the activity of potential drug
-molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental particles~\citep{YANNparticles},
-classifying galaxies~\citep{YANNgalaxies}, and predicting the effects of mutations in
-non-coding DNA on gene expression and disease~\citep{Keung-et-al-2014}. % Frey
+Deep learning is currently making major advances in solving problems
+that have resisted the best attempts of the Artificial Intelligence
+community for many years. Deep learning is very good at discovering
+intricate structure in large, high-dimensional datasets and is
+therefore applicable to many domains of science and business. For
+example, it has beaten previous machine learning techniques at
+predicting the activity of potential drug
+molecules~\citep{DahlKaggle}, detecting new fundamental
+particles~\citep{YANNparticles}, classifying
+galaxies~\citep{YANNgalaxies}, and predicting the effects of mutations
+in non-coding DNA on gene expression and
+disease~\citep{Keung-et-al-2014}. % Frey
 
 So far, most applications of deep learning use feed-forward neural
-networks, but for tasks that involve sequential inputs it is often better
-to use recurrent neural networks. These are given one element from an input % YB term-->element
-sequence at each time step and their recurrent internal ``hidden'' state
-depends on the whole history of the input sequence. Recurrent networks are
-very good at predicting the next character in a string~\citep{Sutskever-et-al-ICML2011} or
-the next word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
-more complex tasks.  After reading an English sentence one word at a time,
-for example, an English``encoder'' network can be trained so that the final
-state vector of its hidden neurons is a good representation of the thought
-expressed by the sentence.  This thought vector can then be used as the
-initial hidden state (or as extra input) of a trained French ``decoder'' 
-network which outputs
-a probability distribution for the first word of the French translation. If
-a particular first word is chosen (e.g., according to that probability)
-and provided as input to the decoder
-network it will then output a probability distribution for the second word
-of the translation and so on until a full stop is 
+networks, but for tasks that involve sequential inputs, such as speech
+and language, it is often better to use recurrent neural
+networks. These are given one element from an input 
+% YB term-->element 
+sequence at each time step and their recurrent internal
+``hidden'' state depends on the whole history of the input
+sequence. Recurrent networks are very good at predicting the next
+character in text~\citep{Sutskever-et-al-ICML2011} or the next
+word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be
+used for more complex tasks.  After reading an English sentence one
+word at a time, for example, an English``encoder'' network can be
+trained so that the final state vector of its hidden units is a good
+representation of the thought expressed by the sentence.  This thought
+vector can then be used as the initial hidden state (or as extra
+input) of a trained French ``decoder'' network which outputs a
+probability distribution for the first word of the French
+translation. If a particular first word is chosen (e.g., according to
+that probability) and provided as input to the decoder network it will
+then output a probability distribution for the second word of the
+translation and so on until a full stop is
 chosen~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
@@
 which yields an efficient way of computing the gradient of a
-smooth objective function with respect to all of the weights in a layered
-network (Figures~\ref{fig:backprop-box} and~\ref{fig:mlp}). 
+smooth objective function with respect to all of the weights and parameters in 
+a layered network (Figures~\ref{fig:backprop-box} and~\ref{fig:mlp}). 
 The objective function typically measures an error to be minimized: 
@@
 output for that case.  The key insight is that the error derivatives with
-respect to the {\it outputs} of a layer can be computed by working
-backwards from the error derivatives with respect to the outputs of the
-subsequent layer. Once this backwards recursion has been used to compute
+respect to the {\it inputs} of a layer can be computed by working
+backwards from the error derivatives with respect to its {\em outputs}. 
+Once this backwards recursion has been used to compute
 error derivatives for the outputs of all layers, it is easy to compute
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
+\begin{figure}[H]
 \fbox{
@@
 to change each weight in proportion to its negative gradient, which is the
-direction where a small change will most improve the objective function. 
+direction where a small change will most decrease the objective function. 
 When the training set contains millions of cases and is highly redundant
@@
 
-\section{Convolutional neural networks}
+\section{Convolutional networks}
 
 %%%%%%%% YANN %%%%%%% 
+% YANN: can you summarize convnets and their use for reading cheques in about
+% 350 words?
+% I couldn't do it in 350 words.
 
-YANN: can you summarize convnets and their use for reading cheques in about
-350 words?
+\vspace{-0.12cm}
+\subsection*{Convolutional Networks}
+\vspace{-0.12cm}
 
-\citep{LeCun89-small}
+\begin{figure}
+\begin{center}
+  \vspace{0.2cm}
+  % \includegraphics[width=0.99\linewidth]{convnet-diagram}
+  nice convnet figure goes here
+\end{center}
+\vspace{-0.6cm}
+\caption{A typical ConvNet architecture}
+\vspace{-0.8cm}
+\label{fig:convnet}
+\end{figure}
+
+One of the most widely used deep learning architectures is the
+convolutional networt, or ConvNet~\cite{lecun-90c,lecun-98}. ConvNets
+are designed to process data that come in the form of arrays, for
+example a color image composed of three 2D arrays containing pixel
+intensities in the three color channels. Many data modalities are in
+the form of multiple arrays: 1D for signals and language, 2D for
+images or audio spectrograms, 3D for video or volumetric images.
+There are four key ideas behind ConvNets that take advantage of the
+properties of many such natural signals: local connections, shared
+weights, pooling, and the use of many layers.
+
+The architecture of a typical ConvNet is shown in
+figure~\ref{fig-convnet}. The first few stages are composed of two
+types of layers: convolutional layers and pooling layers. Units in a
+convolutional layer are organized in {\em faeture maps}, within which
+each unit is connected to local patches in the feature maps of the
+previous layer through a set of weights called a {\em filter
+  bank}. The result of this local weight sum is then passed through a
+non-linearity.  All units in a feature map share the same filter
+bank. Multiple feature maps in a layer use different filter banks. The
+reason for this architecture is twofold. First, In array data such as
+images, local groups of values are often highly correlated forming
+local motifs that are easily detected. Second, the local statistics of
+images and other signals are invariant to location. In other word, if
+a motif can appear in one part of the image, it could appear anywhere,
+hence the idea of units at different locations sharing the same
+weights.  Mathematically, the filtering operation performed by a
+feature map is a {\em discrete convolution}, hence the name
+ConvNet. While the role of the convolutional layer is to detect local
+conjunctions of features in the previous layer, the role of the
+pooling layer is to merge semantically similar features into
+one. Since the relative positions of the features forming a motif can
+vary somewhat, reliably detecting the motif can be done by
+coarse-graining the position of each feature. A typical pooling unit
+computes the max of a local patch of units in one feature map (or in a
+few). Neighboring pooling units take input from patches that are
+shifted by more than one row or column, thereby reducing the dimension
+of the representation and creating an invariance to small shifts.  Two
+or three stages of convolutionas, non-linearity, and pooling are
+stacked, followed by more convolutional and fully-connected layers.
+Back-propagating gradients through a ConvNet is as simple
+as through a regular deep network, allowing all the weights in all the
+filter banks to be trained.
+
+This architecture exploits the property that many natural signals are
+{\em compositional hierarchies}. In images, local combinations of
+edges form motifs, motifs assemble into parts, and parts form objects.
+Similar hierarchies exists in speech and text from sounds to phones,
+phonemes, syllables, words, and sentences. The elements that form
+elements in the next level can vary in relative position and
+appearance. The convolutional and pooling layers in ConvNets are
+directly inspired by the classic notions of simple cells and complex
+cells in visual neuroscience~\cite{Hubel1962}, and the overall
+architecture is reminiscent of visual cortex ventral pathway hierarchy
+LGN-V1-V2-V4-IT~\cite{VanEssen}. ConvNets have their roots in
+Fukushima's Neocognitron~\cite{fukushima-82} whose architecture was
+somewhat similar but did not have an end-to-end supervised learning
+algorithm such as backpropagation+SGD.
+
+There have been numerous applications of convolutional networks going
+back to the early 1990s. Perhaps the earliest success was a bank check
+reading system which was commercially deployed by AT\&T and NCR in
+1994~\cite{lecun-98}. The system used a ConvNets trained jointly with
+a graphical model that implemented language constraints. By the late
+1990s this system was reading over 10\% of all the checks in the
+US. Microsoft later deploy ConvNets in a number of OCR and handwriting
+recognition
+systems~\cite{simard-03,chellapilla-ist-06,chellapilla-iwfhr-06b}
+including for Arabic and
+Chinese~\cite{abdulkader-iwfhr-06,chellapilla-iwfhr-06a}. ConvNets
+have also been used since the early 1990s for object detection in
+natural images, including faces and
+pedestrians~\cite{vaillant-monrocq-lecun-94,garcia-delakis-04,osadchy-07,nasse-09,sermanet-cvpr-13}.
 
@@
 gradients either grow or shrink at each time step, so over many time-steps
-they typically explode or vanish~\citep{Hochreiter91-small,Bengio-et-al-TNN1994},
-and the conditions which allow stable dynamics also correspond to
-vanishing gradients~\citep{Bengio-et-al-TNN1994},
+they typically explode or vanish~\citep{Bengio-et-al-TNN1994,Hochreiter91-small}.
 
-\begin{figure}[ht]
+\begin{figure}[H]
 \fbox{
@@
 get inputs from other neurons at previous time steps
-(this is represented with the black square, standing for a delay of
+(this is represented with the black square, for delay of
 one time step, on the left). In this way, a recurrent net can map an input sequence
@@
 Many other architectures are possible, including a variant in which
-the sampled outputs (e.g. words) can be generated according to an output 
-distribution (the probability of the next word, for every possible next word) 
-and used as inputs for the next time step.
+the outputs (e.g. words) can be generated according to a distribution 
+(the probability of the next word) and used as inputs for the next time step.
 The back-propagation algorithm (Figure~\ref{fig:backprop-box})
@@
 on the right hand side, to compute the derivative of a total error
-(e.g., the log-probability of generating the right sequence of
-words) with respect to all the states $s_t$ and all the parameters.
+(at all time steps) with respect to all the states $s_t$ and all the parameters.
 }
@@
 The labeled datasets were far too small, the computers were far too slow,
-and the weights were not initialised sensibly~\citep{GlorotAISTATS2010-small}.
+and the weights were not initialised sensibly.
 
@@
 results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-arxiv2014} strongly suggest
-the vast majority of critical points (where derivatives of the total
-objective function with respect to parameters vanish) are saddle points, not
-local minima, and that in large deep networks almost all of these local minima 
-give very similar performance so it does not matter which one we find.  This goes a long way
+that in deep networks almost all of these local minima give very similar
+performance so it does not matter which one we find.  This goes a long way
 towards explaining why stochastic gradient descent works so well for deep
@@
 developed to give record-breaking results on a large vocabulary
-task~\citep{Dahl2012}.  By 2012, versions of the deep net from 2009 were being
+task~\citep{Dahl}.  By 2012, versions of the deep net from 2009 were being
 developed by many of the major speech groups\citep{Hinton-et-al-2012} and were already
@@
 
-\section{Object recognition and detection with deep convolutional networks}
+\section{Image understanding with deep convolutional networks}
+
+Since the early 2000s, there have been many tasks in which ConvNets
+have been applied with success to the detection, segmentation, and
+recognition of objects and regions in images. These were invariably
+tasks in which labeled data was relatively abundant, such as traffic
+sign recognition~\cite{sermanet-ijcnn-11,schmidhuber-traffic}, the
+segmentation of biological images~\cite{ning-05} particularly for
+connectomics~\cite{seung-xx}, the detection of faces, text, and human bodies in natural
+images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,nasse-09,taylor-xx,sermanet-cvpr-13,thompson-14}.
+
+An important application is the labeling of images at the pixel level,
+whose applications include driving mobile robots and self-driving
+cars~\cite{hadsell-jfr-09,farabet-icml-12},
+
+Other applications of ConvNets that are gaining importance concern
+natural language understanding~\cite{collobert-weston-xx}, and speech
+recognition~\cite{sainath-xx}.
+
+But works were largely ignored by the computer vision and machine
+learning community until the ImageNet competition in 2012.
 
@@
 %%%%%%%% YANN %%%%%%%
-YANN: Add about 300 words plus another part of the figure showing detected
-parts of people.
-%%%%%%%%%%%%%%%%%%%%
-
-Cite at least \citep{LeCun+98} and \citep{Tompson-et-al-arxiv2014}
+MORE TEXT on other applications of convnets goes here, including Thompson and such.\citep{Tompson-et-al-arxiv2014}
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
 \fbox{
@@
 \includegraphics[width=0.49\linewidth]{Examples_im2txt_002.png}}
-\caption{Three examples of captions generated by a recurrent network taking
+\caption{Examples of captions generated by a recurrent network taking
 as extra input the representation extracted by a deep convolutional
-net from a test image, with the whole system trained to ``translate''
-images into captions. GT stands for ``Ground Truth'' and was labeled by humans
-whereas the other text is machine-generated.
-Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
+net from an image, with the whole system trained to ``translate''
+images into captions. Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
 }

### 2015-01-13 deepreview.tex
fixed figure numbers that were wrong
merged Yann's version

@@
 \usepackage{natbib}
-\usepackage{float}
+%\usepackage{float}
 
@@
 
-Deep learning is currently making major advances in solving problems that
-have resisted the best attempts of the Artificial Intelligence community
-for many years. Deep learning is very good at discovering intricate
-structure in large, high-dimensional datasets and is therefore widely
-applicable in all of the sciences. For example, it has beaten previous
-machine learning techniques at predicting the activity of potential drug
-molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental particles~\citep{YANNparticles},
-classifying galaxies~\citep{YANNgalaxies}, and predicting the effects of mutations in
-non-coding DNA on gene expression and disease~\citep{Keung-et-al-2014}. % Frey
+Deep learning is currently making major advances in solving problems
+that have resisted the best attempts of the Artificial Intelligence
+community for many years. Deep learning is very good at discovering
+intricate structure in large, high-dimensional datasets and is
+therefore applicable to many domains of science and business. For
+example, it has beaten previous machine learning techniques at
+predicting the activity of potential drug
+molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental
+particles~\citep{YANNparticles}, classifying
+galaxies~\citep{YANNgalaxies}, and predicting the effects of mutations
+in non-coding DNA on gene expression and
+disease~\citep{Keung-et-al-2014}. % Frey
 
 So far, most applications of deep learning use feed-forward neural
-networks, but for tasks that involve sequential inputs it is often better
-to use recurrent neural networks. These are given one element from an input % YB term-->element
-sequence at each time step and their recurrent internal ``hidden'' state
-depends on the whole history of the input sequence. Recurrent networks are
-very good at predicting the next character in a string~\citep{Sutskever-et-al-ICML2011} or
-the next word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
-more complex tasks.  After reading an English sentence one word at a time,
-for example, an English``encoder'' network can be trained so that the final
-state vector of its hidden neurons is a good representation of the thought
-expressed by the sentence.  This thought vector can then be used as the
-initial hidden state (or as extra input) of a trained French ``decoder'' 
-network which outputs
-a probability distribution for the first word of the French translation. If
-a particular first word is chosen (e.g., according to that probability)
-and provided as input to the decoder
-network it will then output a probability distribution for the second word
-of the translation and so on until a full stop is 
+networks, but for tasks that involve sequential inputs, such as speech
+and language, it is often better to use recurrent neural
+networks. These are given one element from an input 
+% YB term-->element 
+sequence at each time step and their recurrent internal
+``hidden'' state depends on the whole history of the input
+sequence. Recurrent networks are very good at predicting the next
+character in text~\citep{Sutskever-et-al-ICML2011} or the next
+word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be
+used for more complex tasks.  After reading an English sentence one
+word at a time, for example, an English``encoder'' network can be
+trained so that the final state vector of its hidden units is a good
+representation of the thought expressed by the sentence.  This thought
+vector can then be used as the initial hidden state (or as extra
+input) of a trained French ``decoder'' network which outputs a
+probability distribution for the first word of the French
+translation. If a particular first word is chosen (e.g., according to
+that probability) and provided as input to the decoder network it will
+then output a probability distribution for the second word of the
+translation and so on until a full stop is
 chosen~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
+\begin{figure}[htp]
 \fbox{
@@
 }
+\label{fig:backprop-box}
 \end{minipage}
 }
-\label{fig:backprop-box}
 \end{figure}
@@
 
-\section{Convolutional neural networks}
+\section{Convolutional networks}
 
 %%%%%%%% YANN %%%%%%% 
+% YANN: can you summarize convnets and their use for reading cheques in about
+% 350 words?
+% I couldn't do it in 350 words.
 
-YANN: can you summarize convnets and their use for reading cheques in about
-350 words?
+\vspace{-0.12cm}
+\subsection*{Convolutional Networks}
+\vspace{-0.12cm}
 
-\citep{LeCun89-small}
+\iffalse
+\begin{figure}
+\begin{center}
+  \vspace{0.2cm}
+  % \includegraphics[width=0.99\linewidth]{convnet-diagram}
+  nice convnet figure goes here
+\end{center}
+\vspace{-0.6cm}
+\caption{A typical ConvNet architecture}
+\vspace{-0.8cm}
+\label{fig:convnet}
+\end{figure}
+\fi
+
+One of the most widely used deep learning architectures is the
+convolutional networt, or ConvNet~\cite{lecun-90c,lecun-98}. ConvNets
+are designed to process data that come in the form of arrays, for
+example a color image composed of three 2D arrays containing pixel
+intensities in the three color channels. Many data modalities are in
+the form of multiple arrays: 1D for signals and language, 2D for
+images or audio spectrograms, 3D for video or volumetric images.
+There are four key ideas behind ConvNets that take advantage of the
+properties of many such natural signals: local connections, shared
+weights, pooling, and the use of many layers.
+
+The architecture of a typical ConvNet is shown in
+figure~\ref{fig-convnet}. The first few stages are composed of two
+types of layers: convolutional layers and pooling layers. Units in a
+convolutional layer are organized in {\em faeture maps}, within which
+each unit is connected to local patches in the feature maps of the
+previous layer through a set of weights called a {\em filter
+  bank}. The result of this local weight sum is then passed through a
+non-linearity.  All units in a feature map share the same filter
+bank. Multiple feature maps in a layer use different filter banks. The
+reason for this architecture is twofold. First, In array data such as
+images, local groups of values are often highly correlated forming
+local motifs that are easily detected. Second, the local statistics of
+images and other signals are invariant to location. In other word, if
+a motif can appear in one part of the image, it could appear anywhere,
+hence the idea of units at different locations sharing the same
+weights.  Mathematically, the filtering operation performed by a
+feature map is a {\em discrete convolution}, hence the name
+ConvNet. While the role of the convolutional layer is to detect local
+conjunctions of features in the previous layer, the role of the
+pooling layer is to merge semantically similar features into
+one. Since the relative positions of the features forming a motif can
+vary somewhat, reliably detecting the motif can be done by
+coarse-graining the position of each feature. A typical pooling unit
+computes the max of a local patch of units in one feature map (or in a
+few). Neighboring pooling units take input from patches that are
+shifted by more than one row or column, thereby reducing the dimension
+of the representation and creating an invariance to small shifts.  Two
+or three stages of convolutionas, non-linearity, and pooling are
+stacked, followed by more convolutional and fully-connected layers.
+Back-propagating gradients through a ConvNet is as simple
+as through a regular deep network, allowing all the weights in all the
+filter banks to be trained.
+
+This architecture exploits the property that many natural signals are
+{\em compositional hierarchies}. In images, local combinations of
+edges form motifs, motifs assemble into parts, and parts form objects.
+Similar hierarchies exists in speech and text from sounds to phones,
+phonemes, syllables, words, and sentences. The elements that form
+elements in the next level can vary in relative position and
+appearance. The convolutional and pooling layers in ConvNets are
+directly inspired by the classic notions of simple cells and complex
+cells in visual neuroscience~\cite{Hubel1962}, and the overall
+architecture is reminiscent of visual cortex ventral pathway hierarchy
+LGN-V1-V2-V4-IT~\cite{VanEssen}. ConvNets have their roots in
+Fukushima's Neocognitron~\cite{fukushima-82} whose architecture was
+somewhat similar but did not have an end-to-end supervised learning
+algorithm such as backpropagation+SGD.
+
+There have been numerous applications of convolutional networks going
+back to the early 1990s. Perhaps the earliest success was a bank check
+reading system which was commercially deployed by AT\&T and NCR in
+1994~\cite{lecun-98}. The system used a ConvNets trained jointly with
+a graphical model that implemented language constraints. By the late
+1990s this system was reading over 10\% of all the checks in the
+US. Microsoft later deploy ConvNets in a number of OCR and handwriting
+recognition
+systems~\cite{simard-03,chellapilla-ist-06,chellapilla-iwfhr-06b}
+including for Arabic and
+Chinese~\cite{abdulkader-iwfhr-06,chellapilla-iwfhr-06a}. ConvNets
+have also been used since the early 1990s for object detection in
+natural images, including faces and
+pedestrians~\cite{vaillant-monrocq-lecun-94,garcia-delakis-04,osadchy-07,nasse-09,sermanet-cvpr-13}.
 
@@
 }
+\label{fig:mlp}
 \end{minipage}
 }
-\label{fig:mlp}
 \end{figure}
@@
 best computer vision sytems~\citep{Krizhevsky-2012-small}.  
-%Some of the images and the
-%network's predictions are shown in figure~\ref{fig:imagenet-outputs}.  
 The same kind of network is used to generate the captions shown in Figure~\ref{fig:caption-generation}.
+
 The major factors in this
@@
 %%%%%%%% YANN %%%%%%%
-YANN: Add about 300 words plus another part of the figure showing detected
-parts of people.
-%%%%%%%%%%%%%%%%%%%%
-
-Cite at least \citep{LeCun+98} and \citep{Tompson-et-al-arxiv2014}
-
-\iffalse
-\begin{figure}[ht]
-\fbox{
-\begin{minipage}{\textwidth}
-\centerline{
-%\includegraphics[width=0.89\linewidth]{}
-}
-\caption{
-}
-\end{minipage}
-}
-\label{fig:imagenet-outputs}
-\end{figure}
-\fi
+% MORE TEXT on other applications of convnets goes here, including Thompson and such.\citep{Tompson-et-al-arxiv2014}
 
@@
 }
+\label{fig:word-embeddings}
 \end{minipage}
 }
-\label{fig:word-embeddings}
 \end{figure}
@@
 }
+\label{fig:caption-generation}
 \end{minipage}
 }
-\label{fig:caption-generation}
 \end{figure}
@@
 
-\bibliography{strings,ml,aigaion}
+\bibliography{strings,ml,aigaion,convnetyann,lecun}
 \bibliographystyle{natbib}

### 2015-01-13 deepreview.tex
figured a few of the missing references

@@
 directly inspired by the classic notions of simple cells and complex
-cells in visual neuroscience~\cite{Hubel1962}, and the overall
+cells in visual neuroscience~\cite{Hubel62}, and the overall
 architecture is reminiscent of visual cortex ventral pathway hierarchy
@@
 tasks in which labeled data was relatively abundant, such as traffic
-sign recognition~\cite{sermanet-ijcnn-11,schmidhuber-traffic}, the
+sign recognition~\cite{sermanet-ijcnn-11,Ciresan-et-al-2012}, the
 segmentation of biological images~\cite{ning-05} particularly for
-connectomics~\cite{seung-xx}, the detection of faces, text, and human bodies in natural
-images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,nasse-09,taylor-xx,sermanet-cvpr-13,thompson-14}.
+connectomics~\cite{Turaga2010}, the detection of faces, text, and human bodies in natural
+images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,nasse-09,taylor-xx,sermanet-cvpr-13,Tompson-et-al-arxiv2014}.
 
@@
 Other applications of ConvNets that are gaining importance concern
-natural language understanding~\cite{collobert-weston-xx}, and speech
+natural language understanding~\citep{collobert:2011b}, and speech
 recognition~\cite{sainath-xx}.

### 2015-01-13 deepreview.tex
updated the 'loss surface' paper to AISTATS

@@
 curves upwards in most directions but downwards in a few. Recent
-results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-arxiv2014} strongly suggest
+results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015} strongly suggest
 the vast majority of critical points (where derivatives of the total

### 2015-01-13 deepreview.tex
fixed the missing references

@@
 molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental
-particles~\citep{YANNparticles}, classifying
-galaxies~\citep{YANNgalaxies}, and predicting the effects of mutations
+particles~\citep{Melis-Higgs-boson-competition-2014}, % TODO ~\citep{YANNparticles}, 
+classifying galaxies, % %TODO ~\citep{YANNgalaxies}, 
+and predicting the effects of mutations
 in non-coding DNA on gene expression and
@@
 
-\iffalse
-\begin{figure}
-\begin{center}
-  \vspace{0.2cm}
-  % \includegraphics[width=0.99\linewidth]{convnet-diagram}
-  nice convnet figure goes here
-\end{center}
-\vspace{-0.6cm}
-\caption{A typical ConvNet architecture}
-\vspace{-0.8cm}
-\label{fig:convnet}
-\end{figure}
-\fi
+%\begin{figure}
+%\begin{center}
+%  \vspace{0.2cm}
+%  % \includegraphics[width=0.99\linewidth]{convnet-diagram}
+%  nice convnet figure goes here
+%\end{center}
+%\vspace{-0.6cm}
+%\caption{A typical ConvNet architecture}
+%\vspace{-0.8cm}
+%\label{fig:convnet}
+%\end{figure}
 
@@
 
-The architecture of a typical ConvNet is shown in
-figure~\ref{fig-convnet}. The first few stages are composed of two
+The architecture of a typical ConvNet is 
+%shown in
+%figure~\ref{fig-convnet}. 
+structured as a series of stages. The first few stages are composed of two
 types of layers: convolutional layers and pooling layers. Units in a
@@
 architecture is reminiscent of visual cortex ventral pathway hierarchy
-LGN-V1-V2-V4-IT~\cite{VanEssen}. ConvNets have their roots in
+LGN-V1-V2-V4-IT~\cite{Felleman+VanEssen-1991}. ConvNets have their roots in
 Fukushima's Neocognitron~\cite{fukushima-82} whose architecture was
@@
 connectomics~\cite{Turaga2010}, the detection of faces, text, and human bodies in natural
-images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,nasse-09,taylor-xx,sermanet-cvpr-13,Tompson-et-al-arxiv2014}.
+images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,
+nasse-09,Jain-et-al-arxiv2013,sermanet-cvpr-13,Tompson-et-al-arxiv2014}.
 
@@
 natural language understanding~\citep{collobert:2011b}, and speech
-recognition~\cite{sainath-xx}.
+recognition~\cite{Sainath-et-al-ICASSP2013}.
 

### 2015-01-22 deepreview.tex
fixed more bib problems
typo

@@
 Deep neural networks that learn to represent data in multiple layers
-of increasing abstration have dramatically improved the
+of increasing abstraction have dramatically improved the
 state-of-the-art for speech recognition, object recognition, object
@@
 % I couldn't do it in 350 words.
-
-\vspace{-0.12cm}
-\subsection*{Convolutional Networks}
-\vspace{-0.12cm}
 

### 2015-01-27 deepreview.tex
GEOFF's EDITIONS SENT VIA deepreview_jan26.tex BY EMAIL

@@
 
-Deep neural networks that learn to represent data in multiple layers
-of increasing abstraction have dramatically improved the
+Deep neural networks that learn to represent data using layers
+of increasing abstract features have dramatically improved the
 state-of-the-art for speech recognition, object recognition, object
@@
 by using the backpropagation algorithm to compute error-derivatives
-for the weights that define each layer of feature detector and then
+for the weights that define each layer of feature detectors and then
 repeatedly changing these weights in proportion to their derivatives
@@
 
-Deep learning is currently making major advances in solving problems
+Deep learning is making major advances in solving problems
 that have resisted the best attempts of the Artificial Intelligence
@@
 used for more complex tasks.  After reading an English sentence one
-word at a time, for example, an English``encoder'' network can be
+word at a time, for example, an English ``encoder'' network can be
 trained so that the final state vector of its hidden units is a good
 representation of the thought expressed by the sentence.  This thought
-vector can then be used as the initial hidden state (or as extra
-input) of a trained French ``decoder'' network which outputs a
+vector can then be used as the initial hidden state of (or as extra
+input to)  a jointly trained French ``decoder'' network which outputs a
 probability distribution for the first word of the French
-translation. If a particular first word is chosen (e.g., according to
-that probability) and provided as input to the decoder network it will
+translation. If a particular first word is chosen from this distribution and provided as input to the decoder network it will
 then output a probability distribution for the second word of the
@@
 variables get bound to symbols. They just use big state vectors, big weight
-matrices and scalar non-linearities to get the job done.  This suggests
-that deriving conclusions using only the form of the premises (which is
-what symbolic systems do) is a powerful
-technique that people can eventually master but it is not a good model of
+matrices and scalar non-linearities to get the job done.  Deriving conclusions using only the form of the premises (which is
+what formal logic does) is a powerful
+technique that people can eventually master, but the translation
+results with recurrent nets suggest that it may not be a good model of
 the way we normally reason or understand language. Recurrent neural
@@
 the view that everyday reasoning involves many simultaneous analogies that
-each contribute plausibility to a conclusion~\citep{metaphors}.
-
-This review is organized chronologically starting with the introduction of
-the backpropagation algorithm in the 1980's. After describing some of its
-early successes, we explain the actual and perceived limitations that
+each contribute plausibility to a conclusion~\citep{metaphors,rogers}.
+
+This review is organized chronologically. We start with the introduction of
+the backpropagation algorithm in the 1980's and early demonstrations
+that it could learn good features for language and vision tasks.  We
+then explain the actual and perceived limitations that
 prevented it from living up to its early promise for so many years and how
@@
 vector of feature activations and then collecting evidence for each
-possible class by computing a weighted sum of the features. If good
+different class by computing a differently weighted sum of the feature
+activations. If good
 features can be designed by hand, it is relatively easy to learn
@@
 such as object recognition in which the input space is high-dimensional and
-the input-output mapping is so complex that computing similarities with a
-simple kernel function is ineffective. With a general-purpose feature space
-and kernel (such as the Gaussian kernel), generalization far from the
-examples is poor~\citep{Bengio-localfailure-NIPS-2006-small}, which is problematic
-in high-dimensional spaces. However, kernel machines would perform well if a
-very good set of features were provided: learning very good
-features is what deep learning can do.
+the input-output mapping is so complex that computing similarities using a
+simple kernel function is ineffective. 
+%% THE TEXT BELOW IS NICE BUT TOO MUCH DETAIL ABOUT  SVMS FOR THIS
+%% REVIEW.  
+%% With a general-purpose feature space
+%% and kernel (such as the Gaussian kernel), generalization far from the
+%% examples is poor~\citep{Bengio-localfailure-NIPS-2006-small}, which is problematic
+%% in high-dimensional spaces. However, kernel machines would perform well if a
+%% very good set of features were provided: learning very good
+%% features is what deep learning can do.
 
@@
 particular minute variations, such as the difference between a white wolf
-and a samoyed dog. Experiments and theory of deep learning both
+and a samoyed dog. Experimental and theoretical deep learning results 
 suggest that these tasks are better handled with multiple layers of
@@
 which yields an efficient way of computing the gradient of a
-smooth objective function with respect to all of the weights and parameters in 
-a layered network (Figures~\ref{fig:backprop-box} and~\ref{fig:mlp}). 
+smooth objective function with respect to all of the weights 
+%% and parameters 
+%% FOR THE NOVICE READER INCLUDING BOTH WEIGHTS AND "PARAMTERS" IS CONFUSING.
+%% ITS BETTER TO SAY SOMETHING CLEARER EVEN IF IT IS LESS GENERAL OR CORRECT.
+in  a layered network (Figures~\ref{fig:backprop-box} and~\ref{fig:mlp}). 
 The objective function typically measures an error to be minimized: 
@@
 output for that case.  The key insight is that the error derivatives with
-respect to the {\it inputs} of a layer can be computed by working
-backwards from the error derivatives with respect to its {\em outputs}. 
+respect to the {\it ouputs} of a layer can be computed by working
+backwards from the error derivatives with respect to its {\em inputs}
+to the subsequent layer.
+%% THE TRICKY STEP IS COMPUTING dE/dOUT FROM dE/dIN IN THE SUBSEQUENT
+%% LAYER.  I THINK WE SHOULD FOCUS ON GETTING OVER THAT IDEA FIRST.
+%%  I PLAN TO REDO THE FIGURE TO MAKE IT SIMPLER.  ITS MUCH TOO
+%%  CONFUSING FOR A NOVICE AT PRESENT.  WE SHOULD HAVE A SEPARATE
+%%  DISCUSSION OF THAT.
 Once this backwards recursion has been used to compute
 error derivatives for the outputs of all layers, it is easy to compute
-error derivatives for the parameters (or weights) of each layer 
+error derivatives for the incoming weights of the neurons in each layer 
 (Figure~\ref{fig:mlp}).
 
-%%%%%%% GEOFF %%%%%%%%%%
-%GEOFF: Do box explaining backprop equatons
-%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[htp]
@@
 to average the gradients over all of the cases in the training set and then
-to change each weight in proportion to its negative gradient, which is the
-direction where a small change will most decrease the objective function. 
-When the training set contains millions of cases and is highly redundant
-% explaining what ``redundant'' means, here
-(with many examples illustrating variations of the same underlying 
-aspect or concept), however,
-this is needlessly slow. It is much faster to compute the average gradient
+to change each weight in proportion to its negative gradient. This
+moves the weight vector in the direction of steepest descent in the
+objective function. However, this is needlessly slow when the 
+training set contains millions of cases and is highly redundant
+with most examples conveying no new information about the mapping from
+inputs to outputs.  It is much faster to compute the average gradient
 on a small ``mini-batch'' of randomly chosen training cases and to update
@@
 but these techniques are still the subject of controversy and are beyond
-the scope of this review (but see~\citet{Montavon2012} for tricks of the trade). 
+the scope of this review (see~\citet{Montavon2012} for tricks of the trade). 
 However, it is now widely accepted that some
@@
 vector for Colin contained ``semantic'' features representing which family
-tree Colin was in, what branch of that tree he was in, and from what generation
-he was. These semantic features were not explicitly present
+tree Colin was in, what branch of that tree he was in, and what generation
+he was from. 
+%% TWO NATIVE ENGLISH SPEAKERS AGREE THAT THE "FROM" SHOULD STAY AT THE
+%% END IN ORDER NOT TO SOUND STILTED.
+These semantic features were not explicitly present
 in the input.  They were discovered by the learning procedure as a good way
@@
 or decision trees~\citep{Bengio-decision-trees10}
-would require an exponentially large number
-of training examples to represent the same kinds of functions.
+would require exponential storage 
+%% an exponentially large number of training examples 
+%% THIS WOULD BE VERY CONFUSING FOR MOST READERS WHO WILL NOT KNOW
+%% ABOUT NON-PARAMETERIC METHODS.
+to represent the same kinds of functions.
 
 \section{Convolutional networks}
-
-%%%%%%%% YANN %%%%%%% 
-% YANN: can you summarize convnets and their use for reading cheques in about
-% 350 words?
-% I couldn't do it in 350 words.
 
@@
 One of the most widely used deep learning architectures is the
-convolutional networt, or ConvNet~\cite{lecun-90c,lecun-98}. ConvNets
+convolutional neural network, or ConvNet~\cite{lecun-90c,lecun-98}. ConvNets
 are designed to process data that come in the form of arrays, for
@@
 types of layers: convolutional layers and pooling layers. Units in a
-convolutional layer are organized in {\em faeture maps}, within which
+convolutional layer are organized in {\em feature maps}, within which
 each unit is connected to local patches in the feature maps of the
@@
 non-linearity.  All units in a feature map share the same filter
-bank. Multiple feature maps in a layer use different filter banks. The
-reason for this architecture is twofold. First, In array data such as
+bank. Different feature maps in a layer use different filter banks. The
+reason for this architecture is twofold. First, in array data such as
 images, local groups of values are often highly correlated forming
@@
 feature map is a {\em discrete convolution}, hence the name
-ConvNet. While the role of the convolutional layer is to detect local
+ConvNet.
+
+While the role of the convolutional layer is to detect local
 conjunctions of features in the previous layer, the role of the
@@
 of the representation and creating an invariance to small shifts.  Two
-or three stages of convolutionas, non-linearity, and pooling are
+or three stages of convolution, non-linearity, and pooling are
 stacked, followed by more convolutional and fully-connected layers.
@@
 as through a regular deep network, allowing all the weights in all the
-filter banks to be trained.
+filter banks to be trained. The error derivative for a shared weight
+is just the sum the derivatives for all the instances of that weight.
 
@@
 Similar hierarchies exists in speech and text from sounds to phones,
-phonemes, syllables, words, and sentences. The elements that form
-elements in the next level can vary in relative position and
+phonemes, syllables, words, and sentences. The features that form
+elements at the next level can vary in relative position and
 appearance. The convolutional and pooling layers in ConvNets are
@@
 architecture is reminiscent of visual cortex ventral pathway hierarchy
-LGN-V1-V2-V4-IT~\cite{Felleman+VanEssen-1991}. ConvNets have their roots in
-Fukushima's Neocognitron~\cite{fukushima-82} whose architecture was
+LGN-V1-V2-V4-IT~\citep{Felleman+VanEssen-1991}. ConvNets have their roots in
+Fukushima's Neocognitron~\citep{fukushima-82} whose architecture was
 somewhat similar but did not have an end-to-end supervised learning
@@
 There have been numerous applications of convolutional networks going
-back to the early 1990s. Perhaps the earliest success was a bank check
+back to the early 1990s. A major early success
+%% THIS REWORDING AVOIDS COMPLAINTS FROM HECHT NIELSEN ETC.
+was a bank check
 reading system which was commercially deployed by AT\&T and NCR in
@@
 gradients either grow or shrink at each time step, so over many time-steps
-they typically explode or vanish~\citep{Hochreiter91-small,Bengio-et-al-TNN1994},
-and the conditions which allow stable dynamics also correspond to
-vanishing gradients~\citep{Bengio-et-al-TNN1994},
+they typically explode or
+vanish~\citep{Hochreiter91-small,Bengio-et-al-TNN1994}.
+%% THE TEXT BELOW IS TOO COMPLICATED AND ALSO TOO SIMPLE: THERE ARE
+%% STABLE CONDITIONS IN WHICH THE GRADIENTS DO NOT VANISH (ORTHONORMAL
+%% RECURRENT MATRICES). I THINK ITS BETTER NOT TO GET INTO THIS.
+%% and the conditions which allow stable dynamics also correspond to
+%% vanishing gradients~\citep{Bengio-et-al-TNN1994},
 
@@
 results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015} strongly suggest
-the vast majority of critical points (where derivatives of the total
-objective function with respect to parameters vanish) are saddle points, not
-local minima, and that in large deep networks almost all of these local minima 
+the vast majority of the critical points (where derivatives of the total
+objective function with respect to the parameters vanish) are saddle points, not
+local minima, and that in large, deep networks almost all of the 
+saddle points with predominantly upwards curvature  
 give very similar performance so it does not matter which one we find.  This goes a long way
@@
 Interest in deep feed-forward networks was revived in
-2005 and 2006~\citep{IJCAI,Hinton06-small,Bengio-nips-2006-small,ranzato-07-small} 
+%% Even though I had a conference  paper in 2005, the revival was in 2006!
+2006~\citep{IJCAI,Hinton06-small,Bengio-nips-2006-small,ranzato-07-small} 
 by the introduction
@@
 system could be fine-tuned using standard
-backpropagation\citep{Hinton-Science2006}. This worked remarkably well for recognizing
+backpropagation \citep{Hinton-Science2006}. This worked remarkably well for recognizing
 handwritten digits especially when the amount of labeled data was very
@@
 leading to significantly better generalization when the number of labeled
-examples is small, or in a transfer setting, i.e., when we have lots of
+examples is small, or in a transfer setting where we have lots of
 examples for some ``source'' tasks but very few for some ``target''
@@
 sigmoid or tanh
-non-linearities~\citep{Glorot+al-AI-2011-small,Krizhevsky-2012-small}.
+non-linearities~\citep{NairRelu,Glorot+al-AI-2011-small,Krizhevsky-2012-small,DahlRelu}.
 
@@
 the input.  Similarly, a memory cell has little effect on the output if its
-output gate has low activation. Each memory cell is a gated leaky neuron, i.e.,
-it also has a connection to
-itself at the next time-step that has a weight of 1 (so it tends to copy its
-own real-valued state and accumulate the external signal), 
+output gate has low activation. Each memory cell is a gated leaky neuron: 
+it has a connection to
+itself at the next time-step that has a weight of 1 so it copies its
+own real-valued state and accumulates the external signal, 
 but this self-connection is {\em gated} by a ``forget gate'' so
-the stored value leaks quickly unless the forget gate has an activation
+the stored value quickly leaks away unless the forget gate has an activation
 near $1$.  The gates themselves have their activities determined by
@@
 \citep{Hochreiter+Schmidhuber-1997},
-their complexity might have been off-putting
+their complexity may have been off-putting
 and, despite their impressive ability to learn very long-range temporal
@@
 hand-writing~\citep{Graves-et-al-2009}. They have subsequently proved to be very
-effective for tasks that have sequential input.  Combined with depth of
-representation, they are now better than
+effective for tasks that have sequential input.  When they use several
+hidden layers, they are better than
 deep feed-forward networks for speech recognition~\citep{Graves-et-al-ICASSP2013} and
@@
 architecture are required, the crucial element being the presence and learnability 
-of the forget gate, controlling the leaking rate (or time constant) of the cell.
+of the forget gates which dynamically control the leaking rates  of
+the memory cells.
 %It is not yet clear which aspects of the complicated architecture of LSTMs
@@
 whose applications include driving mobile robots and self-driving
-cars~\cite{hadsell-jfr-09,farabet-icml-12},
-
-Other applications of ConvNets that are gaining importance concern
+cars~\cite{hadsell-jfr-09,farabet-icml-12}. Other applications of ConvNets that are gaining importance concern
 natural language understanding~\citep{collobert:2011b}, and speech
@@
 
-But works were largely ignored by the computer vision and machine
+ConvNets were largely ignored by the mainline computer vision and machine
 learning community until the ImageNet competition in 2012.
-
 When deep convolutional networks were applied to a dataset of about a
@@
 success were the large labeled dataset~\citep{imagenet_cvpr09}, the use of fast GPUs,
-and the convolutional early layers.  It had already been shown that deep
-supervised nets with rectified linear units could be trained without
-unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
-These rectified linear units made the
+and the convolutional early layers.  
+%% REPETITION
+%% It had already been shown that deep
+%% supervised nets with rectified linear units could be trained without
+%% unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
+Rectified linear units made the
 training much faster and the fully connected final layers were prevented
@@
 
-%%%%%%%% YANN %%%%%%%
-% MORE TEXT on other applications of convnets goes here, including Thompson and such.\citep{Tompson-et-al-arxiv2014}
+\iffalse
+YANN:
+CAN WE MOVE SOME OF THE LATER REFS IN THE EARLIER CONVNET SECTION TO
+HERE?
+\fi
 
@@
 
-\iffalse
-%%%%%%%% YOSHUA %%%%%%%
-YOSHUA: Add about 300 words plus t-sne figure of words and/or phrases.
-
-The intro already explained the simple version of translation using encoder
-and decoder nets and I suggest we save your ``translating with attention''
-for introducing attention in the the future section.
-
-After your stuff on language, I would like to include some version of the
-paragraph below along with a figure.
-%%%%%%%%%%%%%%%%%%%%%%%
-\fi
-
-Prior to the introduction of neural language
+Before the introduction of neural language
 models~\citep{BenDucVin01-short}, the standard approach to statistical
 modeling of language was based on counting frequencies of occurrences of
-short symbol sequences (called n-grams). One disadvantage of the n-gram
-approach is that taking into account context of more than a handful of words would require
-training corpora of exponentially growing size (with respect to the
-desired context size), since the number of configurations of $n$
-words from a vocabulary of size $V$ grows as $V^n$. Although this can
+short symbol sequences (called N-grams). One disadvantage of the n-gram
+approach is that the number of possible N-grams is $V^N$ where $V$ is
+the vocabulary size, so taking
+into account a context of more than a handful of words would require
+very large training corpora. Although this can
 be mitigated by only considering the more frequent of the longer sequences of 
-words, it means that less frequent sequences of words are modeled in a 
-context-blind way. Related to this
-problem and the very notion of symbol introduced above, n-grams
-treat each word as an atomic unit, so cannot generalize across
-semantically related sequences of words,
-while we know that there are many ways of expressing roughly the same thing.
-Instead, neural language models can generalize across semantically related words
-and semantically related phrases by learning to associate each word with a
-distributed representation (a vector) such that semantically related
-words end up close to each other in that space, as illustrated in
-Figure~\ref{fig:word-embeddings}. To obtain that effect, one
-{\em jointly learns} a representation for each word as well as
-a function that computes a quantity of interest that is relevant
-to a language related task, such as predicting the next word
-in a sequence (neural language models) or predicting the sequence
-of translated words given the sequence of source words, i.e., neural
-machine translation~\citep{Devlin-et-al-ACL2014,Bahdanau-et-al-arxiv2014,
+words, it means that less frequent sequences are modeled in a 
+context-blind way. N-grams
+treat each word as an atomic unit, so they cannot generalize across
+semantically related sequences of words. By contrast,
+neural language models can generalize across semantically related
+contexts because they associate each word with a vector of real
+valued features
+and semantically related
+words end up close to each other in that vector space, as illustrated in
+Figure~\ref{fig:word-embeddings}. The distributed representations of
+words are obtained by using backpropagation to jointly learn 
+a representation for each word and
+a function that predicts a target quantity such as the next word
+in a sequence (for language modeling) or a whole sequence
+of translated words (for machine translation)~\citep{Devlin-et-al-ACL2014,Bahdanau-et-al-arxiv2014,
 Sutskever-et-al-NIPS2014}.
@@
 
-When translating between languages, the hidden units of the decoder network
-that generates possible translations are conditioned on (i.e., receive as extra
-input or as initialization) a ``thought
-vector'' which is the final hidden state of an encoder network.  But it is
-also possible to initialize a decoder network using a linear transformation
-of the last hidden layer of a deep convolutional network that has already
-been trained to recognize objects in images. The decoder network can then
-be trained by BPTT on a dataset of captioned
-images~\citep{Kiros-et-al-ICML2014,Mao+al-arxiv2014,Donahue-et-al-arxiv2014,
-              Vinyals-et-al-arxiv2014,
-             Chen+Zitnick-arxiv2014,Karpathy+Li-arxiv2014,Venugopalan-et-al-arxiv2014} 
-and such models produce remarkably good
-captions (see Figure~\ref{fig:caption-generation}).
+Instead of translating the meaning of a French sentence to an English sentence,
+one can learn to ``translate'' the meaning of an image into an English sentence,
+as illustrated in Figure~\ref{fig:caption-generation}. The ``encoder'' here is
+a deep convolutional neural network which converts the pixels into an
+activiy vector in its last hidden layer. The ``decoder''
+is a recurrent neural network similar to the ones used for machine translation
+and neural language modeling. There has been a surge of interest in such
+systems recently~\citep{Kiros-et-al-ICML2014,Mao+al-arxiv2014,Donahue-et-al-arxiv2014,
+              Vinyals-et-al-arxiv2014,Fang-et-al-arxiv2014,
+             Chen+Zitnick-arxiv2014,Karpathy+Li-arxiv2014,Venugopalan-et-al-arxiv2014}.
 
@@
 as extra input the representation extracted by a deep convolutional
-net from a test image, with the whole system trained to ``translate''
-images into captions. GT stands for ``Ground Truth'' and was labeled by humans
+net from a test image, with the recurrent network trained to ``translate''
+high-level reprsentations of images into captions. GT stands for ``Ground Truth'' and was labeled by humans
 whereas the other text is machine-generated.
@@
 \end{figure}
-
-Similarly, instead of translating the meaning of a French sentence to an English sentence,
-one can learn to ``translate'' the meaning of an image into an English sentence,
-as illustrated in Figure~\ref{fig:caption-generation}. The ``encoder'' here is
-a convolutional neural network (see below), a deep architecture that is particularly
-well suited to extract a meaningful representation from an image, while the ``decoder''
-is a recurrent neural network similar to the ones used for machine translation
-and neural language modeling. There has been a surge of interest in such
-systems recently~\citep{Kiros-et-al-ICML2014,Mao+al-arxiv2014,Donahue-et-al-arxiv2014,
-              Vinyals-et-al-arxiv2014,Fang-et-al-arxiv2014,
-             Chen+Zitnick-arxiv2014,Karpathy+Li-arxiv2014,Venugopalan-et-al-arxiv2014}.
 
@@
 games~\citep{Deepmind-atari-arxiv2013}. The system uses a deep convolutional network to
-map from the pixels on the screen at several previous time steps to the
+map from the pixels on the screen at several time steps to the
 expected future reward of each possible action and its only inputs are the

### 2015-01-27 deepreview.tex
small changes on Geoff's big commit

@@
 and the convolutional early layers.  
-%% REPETITION
-%% It had already been shown that deep
-%% supervised nets with rectified linear units could be trained without
-%% unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
+%% REPETITION : YB, actually it was not said earlier, only that ReLU was better than sigmoid or tanh
+It had already been shown that deep
+supervised nets with rectified linear units could be trained without
+unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
 Rectified linear units made the
@@
 a deep convolutional neural network which converts the pixels into an
-activiy vector in its last hidden layer. The ``decoder''
+activity vector in its last hidden layer. The ``decoder''
 is a recurrent neural network similar to the ones used for machine translation
@@
 net from a test image, with the recurrent network trained to ``translate''
-high-level reprsentations of images into captions. GT stands for ``Ground Truth'' and was labeled by humans
+high-level representations of images into captions. GT stands for ``Ground Truth'' and was labeled by humans
 whereas the other text is machine-generated.

### 2015-01-31 deepreview.tex
changed the redundancy explanation and fixed the Nair reference

@@
 moves the weight vector in the direction of steepest descent in the
-objective function. However, this is needlessly slow when the 
-training set contains millions of cases and is highly redundant
-with most examples conveying no new information about the mapping from
-inputs to outputs.  It is much faster to compute the average gradient
+objective function. 
+%However, this is needlessly slow when the 
+%training set contains millions of cases and is highly redundant
+%with most examples conveying no new information about the mapping from
+%inputs to outputs.  
+However, this is needlessly slow when the 
+training set contains millions of cases because each example or small batch of
+examples can be seen as a noisy (but correct in average) version of the true total gradient: a very large
+number of these noisy steps can converge much faster than very few but expensive clean steps because even
+the clean gradient is not the optimal direction in which to go.
+%
+It is much faster to compute the average gradient
 on a small ``mini-batch'' of randomly chosen training cases and to update
@@
 could be obtained by using rectifying non-linearities rather than
-sigmoid or tanh
-non-linearities~\citep{NairRelu,Glorot+al-AI-2011-small,Krizhevsky-2012-small,DahlRelu}.
+sigmoid or tanh non-linearities
+~\citep{Nair-2010-small,Glorot+al-AI-2011-small,Krizhevsky-2012-small,Dahl-et-al-ICASSP2013}.
 

### 2015-01-31 deepreview.tex
edited and inserted Geoff's backprop figures and captions

@@
 the view that everyday reasoning involves many simultaneous analogies that
-each contribute plausibility to a conclusion~\citep{metaphors,rogers}.
+each contribute plausibility to a conclusion~\citep{metaphors,Rogers+McClelland-book2004}.
 
@@
 \centerline{
-\includegraphics[width=0.33\linewidth]{chain-rule.pdf}
-\includegraphics[width=0.33\linewidth]{forward-graph.pdf}
-\includegraphics[width=0.33\linewidth]{backward-graph.pdf}
+\includegraphics[width=0.3\linewidth]{chain-rule.pdf}}
+\centerline{
+\hspace*{.5mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_forward.pdf}}
+\hspace*{0mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_backward.pdf}}
 }
 \caption{
-Left: the chain rule of derivatives tells us how
+{\bf Top.} The chain rule of derivatives tells us how
 two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
@@
 and $z$ are vectors (and the derivatives are Jacobian matrices). \newline
-Middle: many computations, such as the succession of linear and non-linear 
-transformations of multi-layer neural networks, can be described as
-a graph where each node $i$ computes a quantity $u_i$ (e.g. the
-output of an artificial neuron or the value of a parameter such as
-a synaptic weight), based on the set of values $a_i$, composed of elements $u_j$ of
-predecessors in the graph, i.e., $u_i = f_i(a_i)$. This can be readily
-generalized to nodes whose output is a vector or a matrix.\newline
-Right: the chain rule can be applied efficiently to computational
-flow graphs by recursively computing the derivative of the final
-node $u_N$ (e.g., the error to be minimized or the output of the neural network)
-with respect to every internal node. The basic recursive equation
-is
-$\frac{\partial u_N}{\partial u_j} = \sum_{i: j \in a_i} \frac{\partial u_N}{\partial u_i}
-\frac{\partial f(a_i)}{\partial u_j}.$
+{\bf Bottom left.} The equations used for computing the forward pass in a neural net
+with two hidden layers. At each layer, we first compute the total input,
+$z$, to each unit which is a weighted sum of the outputs of the units in
+the layer below. Then a non-linear function $f(.)$ is applied to $z$ to get
+the output of the unit.  For simplicity, we have omitted bias terms. \newline
+{\bf Bottom right.} The equations used for computing the backward pass. At each hidden
+layer we compute the error-derivative w.r.t. the output of each unit which
+is a weighted sum of the error-derivatives w.r.t. the total inputs to the
+units in the layer above. We then convert the error-derivative w.r.t. the
+output into the error-derivative w.r.t. the input by multipling it by the
+derivative of $f(z)$.  At the output layer, the error-derivative w.r.t. the
+output of a unit is computed by differentiating the cost function. This
+gives $y_l-t_l$ If the cost function for unit $l$ is $0.5(y_l-t_l)^2$ where
+$t_l$ is the target value.
 }

### 2015-01-31 deepreview.tex
higher-resolution figures, removed old bp figure, rewrote preface

@@
 
-Deep neural networks that learn to represent data using layers
-of increasing abstract features have dramatically improved the
+Deep neural networks that learn to represent data using layers of
+increasingly abstract features have dramatically improved the
 state-of-the-art for speech recognition, object recognition, object
 detection, predicting the activity of drug molecules, and many other
-tasks. Deep learning discovers intricate structure in large datasets
-by using the backpropagation algorithm to compute error-derivatives
-for the weights that define each layer of feature detectors and then
-repeatedly changing these weights in proportion to their derivatives
-averaged over small batches of training examples.  Deep recurrent
-neural networks, in which each layer also receives input from its own
-state at the previous time-step, can now generate good captions for
-images and are competitive for language translation.
+tasks. Deep learning discovers intricate structure in large datasets by
+using the {\em backpropagation algorithm} to compute how the neural network
+should change in order to make less errors.  Deep {\em recurrent neural
+  networks} can process a sequence of inputs and produce a sequence of
+outputs and can now generate good captions for images and are competitive
+for language translation.
 

### 2015-01-31 deepreview.tex
higher-resolution figures, removed old bp figure, rewrote preface

@@
 %% ITS BETTER TO SAY SOMETHING CLEARER EVEN IF IT IS LESS GENERAL OR CORRECT.
-in  a layered network (Figures~\ref{fig:backprop-box} and~\ref{fig:mlp}). 
+in  a layered network (Figure~\ref{fig:backprop-box}). 
 The objective function typically measures an error to be minimized: 
@@
 error derivatives for the incoming weights of the neurons in each layer 
-(Figure~\ref{fig:mlp}).
+(Figure~\ref{fig:backprop-box}).
 
@@
 that allow the neurons within a layer to receive weighted input from those
-same neurons at the previous time-step.  The gradient with respect to a recurrent
+same neurons at the previous time-step.  As illustrated in
+Figure~\ref{fig:rnn}, the gradient with respect to a recurrent
 weight, $w_{ij}$ that allows neuron $i$ at time $t$ to influence the
@@
 \begin{minipage}{\textwidth}
-\centerline{
-\includegraphics[width=0.49\linewidth]{bp-mlp1.pdf}
-\includegraphics[width=0.49\linewidth]{bp-mlp5.pdf}
-}
+%\centerline{
+%\includegraphics[width=0.49\linewidth]{bp-mlp1.pdf}
+%\includegraphics[width=0.49\linewidth]{bp-mlp5.pdf}
+%}
 \centerline{
@@
 }
-\caption{Top left: graph of feedforward computations in a multi-layer
-neural network with one input layer ($x$), two hidden layers ($h_1$ and $h_2$),
-one output layer (out), a target $y$, and an error that compares the output
-and the target. The synaptic weight matrices $W_1$, $W_2$ and $W_3$ are
-the parameters of the neural network.\newline
-Top right: back-propagation through the same multi-layer neural network
-(downward arrows compute the partial derivative of the error with respect
-to every node, i.e., the units value and the parameters).\newline
-Bottom: a recurrent neural network (left) and the unfolding in time of the
+\caption{
+%Top left: graph of feedforward computations in a multi-layer
+%neural network with one input layer ($x$), two hidden layers ($h_1$ and $h_2$),
+%one output layer (out), a target $y$, and an error that compares the output
+%and the target. The synaptic weight matrices $W_1$, $W_2$ and $W_3$ are
+%the parameters of the neural network.\newline
+%Top right: back-propagation through the same multi-layer neural network
+%(downward arrows compute the partial derivative of the error with respect
+%to every node, i.e., the units value and the parameters).\newline
+A recurrent neural network (left) and the unfolding in time of the
 computation involved in its forward computation. The artificial neurons
@@
 }
-\label{fig:mlp}
+\label{fig:rnn}
 \end{minipage}

### 2015-01-31 deepreview.tex
revised introduction to make it more accessible

@@
 So far, most applications of deep learning use feed-forward neural
-networks, but for tasks that involve sequential inputs, such as speech
+networks, illustrated in Figure~\ref{fig:backprop-box}, which learn
+to map a fixed-size input (e.g., an image) to a fixed-size output
+(e.g. a probability for each one of several candidate object categories). 
+They do so by successive transformations corresponding to different
+layers of artificial neurons (called {\em hidden units}),
+which can be interpreted as learned derived features of the raw input. However, 
+for tasks that involve sequential inputs, such as speech
 and language, it is often better to use recurrent neural
-networks. These are given one element from an input 
-% YB term-->element 
-sequence at each time step and their recurrent internal
-``hidden'' state depends on the whole history of the input
-sequence. Recurrent networks are very good at predicting the next
+networks, illustrated in Figure~\ref{fig:rnn} and discussed in Section~\ref{sec:rnn}. 
+Recurrent neural networks process an input sequence one element at a time,
+maintaining in their hidden units a ``state vector'' that implicitly contains information
+about the history all the past elements of the sequence.
+Recurrent networks are very good at predicting the next
 character in text~\citep{Sutskever-et-al-ICML2011} or the next
@@
 probability distribution for the first word of the French
-translation. If a particular first word is chosen from this distribution and provided as input to the decoder network it will
+translation. If a particular first word is chosen 
+from this distribution and provided as input to the decoder network it will
 then output a probability distribution for the second word of the
@@
 whether understanding a sentence requires anything like internal symbolic
-expressions that are manipulated by binding their symbols to the variables
-in judiciously chosen rules of inference.  The encoder-decoder translation
+expressions that are manipulated by binding their symbols to variables
+in judiciously chosen rules of inference, as was previously thought by
+many early AI researchers.  The encoder-decoder translation
 networks do not work by using internal rules of inference in which
 variables get bound to symbols. They just use big state vectors, big weight
-matrices and scalar non-linearities to get the job done.  Deriving conclusions using only the form of the premises (which is
+matrices and scalar non-linearities to get the job done.  
+Deriving conclusions using only the form of the premises (which is
 what formal logic does) is a powerful
 technique that people can eventually master, but the translation
-results with recurrent nets suggest that it may not be a good model of
+results with recurrent neural networks suggest that it may not be a good model of
 the way we normally reason or understand language. Recurrent neural
@@
 \section{Recurrent Neural Networks}
+\label{sec:rnn}
 

### 2015-01-31 deepreview.tex
got rid of SVMs, added MLP space distortion figure and caption

@@
 \usepackage{natbib}
-%\usepackage{float}
+\usepackage{hyperref}
 
@@
 
+A neural network has input neurons that can be driven by raw input
+signals such as the pixels of an image, forming an {\em input vector},
+similarly to how sensory neurons are driven by outside signals like photons or sounds.
 Patterns can be classified by first converting each raw input vector into a
-vector of feature activations and then collecting evidence for each
+vector of feature activations (by analogy to the response of brain neurons
+to these raw signals) and then collecting evidence for each
 different class by computing a differently weighted sum of the feature
@@
 class-specific weights that optimize
-discrimination~\citep{Rosenblatt57}. Widely-used examples of such linear
-classification methods are logistic regression and linear support vector
-machines~\citep{bishop-book95,Boser92}.
+discrimination~\citep{Rosenblatt57}. Widely-used examples of such {\em linear
+classifiers} are {\em logistic regression} and {\em linear support vector
+machines}~\citep{bishop-book95,Boser92}.
 
 However, even with good hand-crafted features, many interesting tasks
-cannot be solved by linear classifiers. A common method to get around this
+cannot be solved by linear classifiers. By distorting the space
+as in Figure~\ref{fig:space-distortion}, a linear discrimination
+becomes possible.
+\iffalse
+A common method to get around this
 problem is to transform the feature vector into a higher dimensional vector
+(i.e., corresponding to a much larger set of features)
 through a non-linear mapping, exploiting the fact that collections of
@@
 %% features is what deep learning can do.
+\fi
 
@@
 Backpropagation is simply an application of
-the chain rule of derivatives (left of Figure~\ref{fig:backprop-box}),
+the chain rule of derivatives (Figure~\ref{fig:backprop-box}),
 which yields an efficient way of computing the gradient of a
@@
 \centerline{
-\includegraphics[width=0.3\linewidth]{chain-rule.pdf}}
+\fbox{\includegraphics[width=0.69\linewidth]{netvis-simple-2S.png}}
+\fbox{\includegraphics[width=0.28\linewidth]{chain-rule.pdf}}}
 \centerline{
@@
 \caption{
-{\bf Top.} The chain rule of derivatives tells us how
+{\bf Top left.} A multi-layer neural network can distort the
+input space to make the classes (whose examples are supposed to
+be on the red and blue lines, here) linearly separable. Note how
+a regular grid in input space is also transformed. This is an
+illustrative example with only two input units, two hidden units
+and one output unit, but the networks used for object recognition
+or natural language processing contain tens or hundreds of thousand
+of units. Reproduced with permission by Chris Olah from~\url{http://colah.github.io/}.
+\newline
+{\bf Top right.} The chain rule of derivatives tells us how
 two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)

### 2015-01-31 deepreview.tex
more audience-adaptation, sec. 3

@@
 different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}.  
+The gradient of an objective function with respect to the network
+weights tells us how a change in each of the weights would influence
+the final error (top left of Figure~\ref{fig:backprop-box}).
+Gradient descent changes the weights accordingly, in the direction of the
+gradient, towards reducing the error.
 Backpropagation is simply an application of
-the chain rule of derivatives (Figure~\ref{fig:backprop-box}),
-which yields an efficient way of computing the gradient of a
-smooth objective function with respect to all of the weights 
-%% and parameters 
+the chain rule of derivatives 
+which yields an efficient way of computing this gradient in
+a deep neural network (or any other composition of differentiable operations).
 %% FOR THE NOVICE READER INCLUDING BOTH WEIGHTS AND "PARAMTERS" IS CONFUSING.
 %% ITS BETTER TO SAY SOMETHING CLEARER EVEN IF IT IS LESS GENERAL OR CORRECT.
-in  a layered network (Figure~\ref{fig:backprop-box}). 
-The objective function typically measures an error to be minimized: 
+The objective function being minimized typically measures an error to be minimized: 
 the discrepancy between

### 2015-01-31 deepreview.tex
fixed the 'metaphors' citation (now Lakoff+Johnson-2008)

@@
 the view that everyday reasoning involves many simultaneous analogies that
-each contribute plausibility to a conclusion~\citep{metaphors,Rogers+McClelland-book2004}.
+each contribute plausibility to a 
+conclusion~\citep{Lakoff+Johnson-2008,Rogers+McClelland-book2004}.
 
@@
 cannot be solved by linear classifiers. By distorting the space
-as in Figure~\ref{fig:space-distortion}, a linear discrimination
+as in Figure~\ref{fig:backprop-box} (top left), a linear discrimination
 becomes possible.

### 2015-01-31 deepreview.tex
many small changes and a big reorganization of sections

@@
 
-So far, most applications of deep learning use feed-forward neural
+So far, most applications of deep learning use feedforward neural
 networks, illustrated in Figure~\ref{fig:backprop-box}, which learn
@@
 {\bf Bottom left.} The equations used for computing the forward pass in a neural net
-with two hidden layers. At each layer, we first compute the total input,
+with two hidden layers (each composed of a group of hidden units that are
+similarly connected). At each layer, we first compute the total input,
 $z$, to each unit which is a weighted sum of the outputs of the units in
@@
 examples can be seen as a noisy (but correct in average) version of the true total gradient: a very large
-number of these noisy steps can converge much faster than very few but expensive clean steps because even
+number of these noisy steps can converge to a good solution
+much faster than very few but expensive clean steps, because even
 the clean gradient is not the optimal direction in which to go.
 %
-It is much faster to compute the average gradient
-on a small ``mini-batch'' of randomly chosen training cases and to update
+It is much faster to proceed one small ``mini-batch'' examples at a time,
+i.e., compute the average gradient
+on a small set of randomly chosen training cases (say 128) and to update
 each weight by this stochastic estimate of its full gradient, times a
@@
 However, it is now widely accepted that some
-form of stochastic gradient descent is the best bet for large, redundant
+form of stochastic gradient descent is the best bet for large
 datasets.
@@
 Each context word is presented to the network as a one-of-N
-vector in which one component has a value of $1$ and the rest are $0$. The
+vector, i.e., in which one component has a value of $1$ and the rest are $0$. The
 output layer has one unit per word and is forced to produce a probability
@@
 \end{equation}
-where $z_j$ is the total input received by output unit $j$. The hidden
+where $z_j$ is the total input received by output unit $j$. 
+\iffalse
+The hidden
 layers use a non-linearity such as the logistic, tanh, or rectifier, respectively:
@@
 y_j = \frac{1}{1+ e^{z_j}}, \ \ \ y_j = \frac{e^{z_j}-e^{-z_j}}{e^{z_j}+e^{-z_j}}, \ \ \
- y_j = \max(0, z_j)
+ y_j = \max(0, z_j).
 \end{equation}
-
-The connectivity of the network is restricted so that each of the one-of-N
-input vectors must first be converted to a ``word-vector'' before being
-used by subsequent layers to give a high probability to the target next
-word.  The network learns word-vectors that contain many active components
+\fi
+
+The presence of a particular word at the $i$-th position in the input of the network 
+sets the value of specialized hidden units to values that correspond to the
+representation of that word. Different words create a different pattern of
+activations, or word-vector, illustrated in Figure~\ref{fig:word-embeddings}.
+In a language model, the other layers of the network learn to convert the input word-vectors
+into an output word-vector for the predicted next-word, which can be used
+to predict the probability for any word in the vocabulary to appear as the next word.
+The network learns word-vectors that contain many active components
 each of which can be interpreted as a separate feature of the word. This
@@
 predict the next word in a news story, for example, the learned word
-vectors for Monday and Tuesday are very similar, as are the word vectors
-for Vietnam and Iraq.  Vector representations of words learned from text
+vectors for Tuesday and Wednesday are very similar, as are the word vectors
+for Vietnam and Iraq. This is not to say that Tuesday and Wednesday are
+put in the same category, but rather that they share many semantic and
+grammatical attributes, so that what the neural network learns from one day of
+the week can often be generalized to the other, while individual days remain
+different (for example one imagines that the neural network also
+learns to some features that distinguises week-end days and weekdays, while
+Saturday and Sunday might have special roles in some religions, etc.).
+Such representations are called {\em distributed representations} because
+its elements (the features) are not mutually exclusive and there exists
+a rich and vast set of possible configurations of values of these features,
+corresponding to the variations seen in the observed data.
+These word-vectors are thus {\em learned features} that were not determined
+ahead of time by humans, but automatically discovered by the neural network.
+Vector representations of words learned from text
 are now very widely used in natural language 
-applications~\citep{Schwenk-2007,collobert:2011b,Socher-2011,Mikolov-et-al-NIPS2013}.
+applications~\citep{Schwenk-2007,collobert:2011b,Socher-2011,
+Mikolov-et-al-NIPS2013,Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
 
@@
 
+\section{Language processing with deep networks}
+
+Before the introduction of neural language
+models~\citep{BenDucVin01-short}, the standard approach to statistical
+modeling of language was based on counting frequencies of occurrences of
+short symbol sequences (called N-grams). One disadvantage of the n-gram
+approach is that the number of possible N-grams is $V^N$ where $V$ is
+the vocabulary size, so taking
+into account a context of more than a handful of words would require
+very large training corpora. Although this can
+be mitigated by only considering the more frequent of the longer sequences of 
+words, it means that less frequent sequences are modeled in a 
+context-blind way. N-grams
+treat each word as an atomic unit, so they cannot generalize across
+semantically related sequences of words. By contrast,
+neural language models can generalize across semantically related
+contexts because they associate each word with a vector of real
+valued features
+and semantically related
+words end up close to each other in that vector space, as illustrated in
+Figure~\ref{fig:word-embeddings}. The distributed representations of
+words are obtained by using backpropagation to jointly learn 
+a representation for each word and
+a function that predicts a target quantity such as the next word
+in a sequence (for language modeling) or a whole sequence
+of translated words (for machine translation)~\citep{Devlin-et-al-ACL2014,Bahdanau-et-al-arxiv2014,
+Sutskever-et-al-NIPS2014}.
+
+\begin{figure}[ht]
+\fbox{
+\begin{minipage}{\textwidth}
+\centerline{
+\includegraphics[width=0.49\linewidth]{word-embeddings.png}
+\includegraphics[width=0.49\linewidth]{phrase_zoom2.png}
+}
+\caption{Left: Illustration of word representations learned for modeling
+language, non-linearly projected to 2-D for visualization using the
+t-SNE algorithm~\citep{VanDerMaaten08}. Right: 2-D representation
+of phrases learned by an English-to-French encoder-decoder recurrent
+neural network~\citep{Cho-et-al-EMNLP2014}. One can observe that
+semantically similar words or sequences of words are  mapped to
+nearby representations.
+}
+\label{fig:word-embeddings}
+\end{minipage}
+}
+\end{figure}
+
+Instead of translating the meaning of a French sentence to an English sentence,
+one can learn to ``translate'' the meaning of an image into an English sentence,
+as illustrated in Figure~\ref{fig:caption-generation}. The ``encoder'' here is
+a deep convolutional neural network which converts the pixels into an
+activity vector in its last hidden layer. The ``decoder''
+is a recurrent neural network similar to the ones used for machine translation
+and neural language modeling. There has been a surge of interest in such
+systems recently~\citep{Kiros-et-al-ICML2014,Mao+al-arxiv2014,Donahue-et-al-arxiv2014,
+              Vinyals-et-al-arxiv2014,Fang-et-al-arxiv2014,
+             Chen+Zitnick-arxiv2014,Karpathy+Li-arxiv2014,Venugopalan-et-al-arxiv2014}.
+
+\begin{figure}[ht]
+\fbox{
+\begin{minipage}{\textwidth}
+\centerline{\includegraphics[width=\linewidth]{caption-generation-Vinyals-et-al.png}}
+\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}
+\includegraphics[width=0.49\linewidth]{Examples_im2txt_002.png}}
+\caption{Three examples of captions generated by a recurrent network taking
+as extra input the representation extracted by a deep convolutional
+net from a test image, with the recurrent network trained to ``translate''
+high-level representations of images into captions. GT stands for ``Ground Truth'' and was labeled by humans
+whereas the other text is machine-generated.
+Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
+}
+\label{fig:caption-generation}
+\end{minipage}
+}
+\end{figure}
+
 \section{Convolutional networks}
@@
 images, local groups of values are often highly correlated forming
-local motifs that are easily detected. Second, the local statistics of
+local motifs that are easily detected. Second, the spatially local statistics of
 images and other signals are invariant to location. In other word, if
@@
 directly inspired by the classic notions of simple cells and complex
-cells in visual neuroscience~\cite{Hubel62}, and the overall
-architecture is reminiscent of visual cortex ventral pathway hierarchy
-LGN-V1-V2-V4-IT~\citep{Felleman+VanEssen-1991}. ConvNets have their roots in
+cells in visual neuroscience~\citep{Hubel62}, and the overall
+architecture is reminiscent of the LGN-V1-V2-V4-IT visual cortex ventral pathway hierarchy
+~\citep{Felleman+VanEssen-1991}. ConvNets have their roots in
 Fukushima's Neocognitron~\citep{fukushima-82} whose architecture was
 somewhat similar but did not have an end-to-end supervised learning
-algorithm such as backpropagation+SGD.
+algorithm such as backpropagation plus stochastic gradient descent.
 
@@
 reading system which was commercially deployed by AT\&T and NCR in
-1994~\cite{lecun-98}. The system used a ConvNets trained jointly with
-a graphical model that implemented language constraints. By the late
+1994~\citep{lecun-98}. The system used a ConvNet trained jointly with
+a probabilistic model that implemented language constraints. By the late
 1990s this system was reading over 10\% of all the checks in the
-US. Microsoft later deploy ConvNets in a number of OCR and handwriting
+US. Microsoft later deployed ConvNets in a number of OCR and handwriting
 recognition
-systems~\cite{simard-03,chellapilla-ist-06,chellapilla-iwfhr-06b}
+systems~\citep{simard-03,chellapilla-ist-06,chellapilla-iwfhr-06b}
 including for Arabic and
-Chinese~\cite{abdulkader-iwfhr-06,chellapilla-iwfhr-06a}. ConvNets
+Chinese~\citep{abdulkader-iwfhr-06,chellapilla-iwfhr-06a}. ConvNets
 have also been used since the early 1990s for object detection in
@@
 pedestrians~\cite{vaillant-monrocq-lecun-94,garcia-delakis-04,osadchy-07,nasse-09,sermanet-cvpr-13}.
+
+\section{Image understanding with deep convolutional networks}
+
+Since the early 2000s, there have been many tasks in which ConvNets
+have been applied with success to the detection, segmentation, and
+recognition of objects and regions in images. These were invariably
+tasks in which labeled data was relatively abundant, such as traffic
+sign recognition~\cite{sermanet-ijcnn-11,Ciresan-et-al-2012}, the
+segmentation of biological images~\cite{ning-05} particularly for
+connectomics~\cite{Turaga2010}, the detection of faces, text, and human bodies in natural
+images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,
+nasse-09,Jain-et-al-arxiv2013,sermanet-cvpr-13,Tompson-et-al-arxiv2014}.
+
+An important application is the labeling of images at the pixel level,
+whose applications include driving mobile robots and self-driving
+cars~\cite{hadsell-jfr-09,farabet-icml-12}. Other applications of ConvNets that are gaining importance concern
+natural language understanding~\citep{collobert:2011b}, and speech
+recognition~\cite{Sainath-et-al-ICASSP2013}.
+
+ConvNets were largely ignored by the mainline computer vision and machine
+learning community until the ImageNet competition in 2012.
+When deep convolutional networks were applied to a dataset of about a
+million images from the web that contained a thousand different classes
+they achieved spectacular results, almost halving the error rates of the
+best computer vision sytems~\citep{Krizhevsky-2012-small}.  
+The same kind of network is used to generate the captions shown in Figure~\ref{fig:caption-generation}.
+
+The major factors in this
+success were the large labeled dataset~\citep{imagenet_cvpr09}, the use of fast GPUs,
+and the convolutional early layers.  
+%% REPETITION : YB, actually it was not said earlier, only that ReLU was better than sigmoid or tanh
+It had already been shown that deep
+supervised nets with rectified linear units could be trained without
+unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
+Rectified linear units made the
+training much faster and the fully connected final layers were prevented
+from overfitting the peculiarities of the training set by randomly omitting
+hidden units during training~\citep{Srivastava14}. Deep convolutional nets
+have quickly replaced the carefully engineered pipelines full of
+hand-designed features developed by the computer vision community and they
+are now deployed in many computer vision applications, including image
+tagging, object detection, face recognition, facial expression recognition
+and scene parsing.
+
+\iffalse
+YANN:
+CAN WE MOVE SOME OF THE LATER REFS IN THE EARLIER CONVNET SECTION TO
+HERE?
+\fi
 
@@
 same neurons at the previous time-step.  As illustrated in
-Figure~\ref{fig:rnn}, the gradient with respect to a recurrent
+Figure~\ref{fig:rnn}, when we consider the outputs of the hidden units at
+different discrete time steps as if they were the outputs of different neurons
+in a deep multi-layer network, it becomes clear how we can apply backpropagation
+to train recurrent networks. The
+the gradient with respect to a recurrent
 weight, $w_{ij}$ that allows neuron $i$ at time $t$ to influence the
@@
 for a simple gradient descent procedure and that the networks were getting
-trapped in poor local minima. We now know that the main reasons for the
+trapped in poor local minima, weight configurations for which no small
+change would reduce the average error. We now know that the main reasons for the
 limited success of backpropagation in deep networks were quite different:
@@
 
-The obvious weakness of gradient descent is that it can get trapped in poor
+The obvious perceived weakness of gradient descent is that it can get trapped in poor
 local minima, but in high-dimensional weight-spaces this is much less
-common than getting stuck close to saddle points where the error surface
+common than getting stuck close to {\em saddle points}, where the error surface
 curves upwards in most directions but downwards in a few. Recent
 results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015} strongly suggest
-the vast majority of the critical points (where derivatives of the total
-objective function with respect to the parameters vanish) are saddle points, not
-local minima, and that in large, deep networks almost all of the 
+that the vast majority of the critical points\footnote{where derivatives of the total
+objective function with respect to the parameters vanish, and include minima,
+maxima and saddle points} are saddle points, not
+local minima, and that in large and deep networks almost all of the 
 saddle points with predominantly upwards curvature  
-give very similar performance so it does not matter which one we find.  This goes a long way
+give very similar performance, so it does not matter which one we find.  This goes a long way
 towards explaining why stochastic gradient descent works so well for deep
@@
 
-Interest in deep feed-forward networks was revived in
+Interest in deep feedforward networks was revived in
 %% Even though I had a conference  paper in 2005, the revival was in 2006!
@@
 system could be fine-tuned using standard
-backpropagation \citep{Hinton-Science2006}. This worked remarkably well for recognizing
+backpropagation \citep{Hinton-Science2006,Bengio-nips-2006-small,ranzato-07-small}. 
+This worked remarkably well for recognizing
 handwritten digits especially when the amount of labeled data was very
-limited~\citep{Bengio-nips-2006-small,Erhan+al-2010-small}.
+limited~\citep{Erhan+al-2010-small}.
 
@@
 Processing Units (GPUs) that were convenient to
-program~\citep{RainaICML09-small}.  In 2009, the approach was used to map
+program~\citep{RainaICML09-small} and allowed researchers to train
+networks 10 or 20 times faster.  In 2009, the approach was used to map
 short temporal windows of coefficients extracted from a soundwave to a set
@@
 random weights were carefully chosen to avoid very large or very small
-initial gradients~\citep{GlorotAISTATS2010-small} and that even better
-results (especially for very deep networks)
+initial gradients~\citep{GlorotAISTATS2010-small}. Even better
+results, especially for very deep networks,
 could be obtained by using rectifying non-linearities rather than
@@
 
-Interest in recurrent networks (which are deep in time) was revived by
+Interest in recurrent networks (which correspond to a very deep network
+when unfolded in time as in Figure~\ref{fig:rnn}) was revived by
 ``Long Short Term Memory'' networks (LSTMs) that use special hidden units
@@
 via gated connections whose weights are multiplied by the state of a
-logistic gating unit that has an activation between $0$ and $1$. The input
-vector has little effect on the state of that cell if the cell's input
-gate has a low activation so memory cells can learn to selectively ignore
-the input.  Similarly, a memory cell has little effect on the output if its
-output gate has low activation. Each memory cell is a gated leaky neuron: 
-it has a connection to
+logistic gating unit that takes a value between $0$ and $1$. An input gating
+unit can learn when to isolate the cell from the effect of the outside world,
+and an output gating unit prevents the content of the memory cell to be
+broadcast to the rest of the neural network. More importantly, the
+memory cell itself acts like an accumulator or a gated leaky neuron:  it has a connection to
 itself at the next time-step that has a weight of 1 so it copies its
@@
 hidden layers, they are better than
-deep feed-forward networks for speech recognition~\citep{Graves-et-al-ICASSP2013} and
-unlike feed-forward nets they can implement an entire speech recognition
+deep feedforward networks for speech recognition~\citep{Graves-et-al-ICASSP2013} and
+unlike feedforward nets they can implement an entire speech recognition
 system that goes all the way from coefficents derived from the soundwave to
@@
 %predicting the next word in text\citep{LeHinton}.
-
-\section{Image understanding with deep convolutional networks}
-
-Since the early 2000s, there have been many tasks in which ConvNets
-have been applied with success to the detection, segmentation, and
-recognition of objects and regions in images. These were invariably
-tasks in which labeled data was relatively abundant, such as traffic
-sign recognition~\cite{sermanet-ijcnn-11,Ciresan-et-al-2012}, the
-segmentation of biological images~\cite{ning-05} particularly for
-connectomics~\cite{Turaga2010}, the detection of faces, text, and human bodies in natural
-images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,
-nasse-09,Jain-et-al-arxiv2013,sermanet-cvpr-13,Tompson-et-al-arxiv2014}.
-
-An important application is the labeling of images at the pixel level,
-whose applications include driving mobile robots and self-driving
-cars~\cite{hadsell-jfr-09,farabet-icml-12}. Other applications of ConvNets that are gaining importance concern
-natural language understanding~\citep{collobert:2011b}, and speech
-recognition~\cite{Sainath-et-al-ICASSP2013}.
-
-ConvNets were largely ignored by the mainline computer vision and machine
-learning community until the ImageNet competition in 2012.
-When deep convolutional networks were applied to a dataset of about a
-million images from the web that contained a thousand different classes
-they achieved spectacular results, almost halving the error rates of the
-best computer vision sytems~\citep{Krizhevsky-2012-small}.  
-The same kind of network is used to generate the captions shown in Figure~\ref{fig:caption-generation}.
-
-The major factors in this
-success were the large labeled dataset~\citep{imagenet_cvpr09}, the use of fast GPUs,
-and the convolutional early layers.  
-%% REPETITION : YB, actually it was not said earlier, only that ReLU was better than sigmoid or tanh
-It had already been shown that deep
-supervised nets with rectified linear units could be trained without
-unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
-Rectified linear units made the
-training much faster and the fully connected final layers were prevented
-from overfitting the peculiarities of the training set by randomly omitting
-hidden units during training~\citep{Srivastava14}. Deep convolutional nets
-have quickly replaced the carefully engineered pipelines full of
-hand-designed features developed by the computer vision community and they
-are now deployed in many computer vision applications, including image
-tagging, object detection, face recognition, facial expression recognition
-and scene parsing.
-
-\iffalse
-YANN:
-CAN WE MOVE SOME OF THE LATER REFS IN THE EARLIER CONVNET SECTION TO
-HERE?
-\fi
-
-\section{Language processing with deep networks}
-
-Before the introduction of neural language
-models~\citep{BenDucVin01-short}, the standard approach to statistical
-modeling of language was based on counting frequencies of occurrences of
-short symbol sequences (called N-grams). One disadvantage of the n-gram
-approach is that the number of possible N-grams is $V^N$ where $V$ is
-the vocabulary size, so taking
-into account a context of more than a handful of words would require
-very large training corpora. Although this can
-be mitigated by only considering the more frequent of the longer sequences of 
-words, it means that less frequent sequences are modeled in a 
-context-blind way. N-grams
-treat each word as an atomic unit, so they cannot generalize across
-semantically related sequences of words. By contrast,
-neural language models can generalize across semantically related
-contexts because they associate each word with a vector of real
-valued features
-and semantically related
-words end up close to each other in that vector space, as illustrated in
-Figure~\ref{fig:word-embeddings}. The distributed representations of
-words are obtained by using backpropagation to jointly learn 
-a representation for each word and
-a function that predicts a target quantity such as the next word
-in a sequence (for language modeling) or a whole sequence
-of translated words (for machine translation)~\citep{Devlin-et-al-ACL2014,Bahdanau-et-al-arxiv2014,
-Sutskever-et-al-NIPS2014}.
-
-\begin{figure}[ht]
-\fbox{
-\begin{minipage}{\textwidth}
-\centerline{
-\includegraphics[width=0.49\linewidth]{word-embeddings.png}
-\includegraphics[width=0.49\linewidth]{phrase_zoom2.png}
-}
-\caption{Left: Illustration of word representations learned for modeling
-language, non-linearly projected to 2-D for visualization using the
-t-SNE algorithm~\citep{VanDerMaaten08}. Right: 2-D representation
-of phrases learned by an English-to-French encoder-decoder recurrent
-neural network~\citep{Cho-et-al-EMNLP2014}. One can observe that
-semantically similar words or sequences of words are  mapped to
-nearby representations.
-}
-\label{fig:word-embeddings}
-\end{minipage}
-}
-\end{figure}
-
-Instead of translating the meaning of a French sentence to an English sentence,
-one can learn to ``translate'' the meaning of an image into an English sentence,
-as illustrated in Figure~\ref{fig:caption-generation}. The ``encoder'' here is
-a deep convolutional neural network which converts the pixels into an
-activity vector in its last hidden layer. The ``decoder''
-is a recurrent neural network similar to the ones used for machine translation
-and neural language modeling. There has been a surge of interest in such
-systems recently~\citep{Kiros-et-al-ICML2014,Mao+al-arxiv2014,Donahue-et-al-arxiv2014,
-              Vinyals-et-al-arxiv2014,Fang-et-al-arxiv2014,
-             Chen+Zitnick-arxiv2014,Karpathy+Li-arxiv2014,Venugopalan-et-al-arxiv2014}.
-
-\begin{figure}[ht]
-\fbox{
-\begin{minipage}{\textwidth}
-\centerline{\includegraphics[width=\linewidth]{caption-generation-Vinyals-et-al.png}}
-\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}
-\includegraphics[width=0.49\linewidth]{Examples_im2txt_002.png}}
-\caption{Three examples of captions generated by a recurrent network taking
-as extra input the representation extracted by a deep convolutional
-net from a test image, with the recurrent network trained to ``translate''
-high-level representations of images into captions. GT stands for ``Ground Truth'' and was labeled by humans
-whereas the other text is machine-generated.
-Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
-}
-\label{fig:caption-generation}
-\end{minipage}
-}
-\end{figure}
 

### 2015-01-31 deepreview.tex
many small changes and a big reorganization of sections

@@
 networks~\citep{Salakhutdinov2009-small,Hinton95,QuocLe-ICML2012,VincentPLarochelleH2008-small,ranzato-pami,Bengio-et-al-ICML-2014,Kingma-et-al-NIPS2014}. 
+However, most data have not been hand-labeled by humans, and just like
+children are not told what they should have done at each instant, we need
+learning machines that can learn from mostly unlabeled examples.
 In the long run,
-however, we believe unsupervised learning will be essential for allowing
+we believe unsupervised learning will be essential for allowing
 deep networks to generalize well by creating internal representations that
@@
 
-\newpage
+%\newpage
 

### 2015-02-01 deepreview.tex
modified the fprop/bprop box captions as suggested by Geoff to include the 3 non-linearities

@@
 $\frac{\partial z}{\partial x}$. It also works when $x$, $y$
-and $z$ are vectors (and the derivatives are Jacobian matrices). \newline
-{\bf Bottom left.} The equations used for computing the forward pass in a neural net
-with two hidden layers (each composed of a group of hidden units that are
-similarly connected). At each layer, we first compute the total input,
-$z$, to each unit which is a weighted sum of the outputs of the units in
-the layer below. Then a non-linear function $f(.)$ is applied to $z$ to get
-the output of the unit.  For simplicity, we have omitted bias terms. \newline
-{\bf Bottom right.} The equations used for computing the backward pass. At each hidden
-layer we compute the error-derivative w.r.t. the output of each unit which
-is a weighted sum of the error-derivatives w.r.t. the total inputs to the
-units in the layer above. We then convert the error-derivative w.r.t. the
-output into the error-derivative w.r.t. the input by multipling it by the
-derivative of $f(z)$.  At the output layer, the error-derivative w.r.t. the
-output of a unit is computed by differentiating the cost function. This
-gives $y_l-t_l$ If the cost function for unit $l$ is $0.5(y_l-t_l)^2$ where
-$t_l$ is the target value.
+and $z$ are vectors (and the derivatives are Jacobian matrices). 
+\newline
+{\bf Bottom left.} 
+The equations used for computing the forward pass in a neural
+net with two hidden layers. At each layer, we first compute the total
+input $z$ to each unit, which is a weighted sum of the outputs of the
+units in the layer below. Then a non-linear function $f(.)$ is applied
+to $z$ to get the output of the unit.  For simplicity, we have omitted
+bias terms. The functions commonly used in neural networks include the
+logistic sigmoid, $f(z) = 1/(1+\exp(-z))$, the hyberbolic tangent,
+$f(z) = (\exp(z)-\exp(-z))/(\exp(z)+\exp(-z))$ and, more recently, the
+rectified linear unit, $f(z) = max(0, z)$.
+{\bf Bottom right.} 
+The equations used for computing the backward pass. At each
+hidden layer we compute the error-derivative w.r.t. the output of each
+unit which is a weighted sum of the error-derivatives w.r.t. the total
+inputs to the units in the layer above. We then convert the
+error-derivative w.r.t. the output into the error-derivative w.r.t.
+the input by multipling it by the gradient of $f(z)$.  At the output
+layer, the error-derivative w.r.t. the output of a unit is computed by
+differentiating the cost function. This gives $y_l-t_l$ if the cost
+function for unit $l$ is $0.5(y_l-t_l)^2$ where $t_l$ is the target
+value. Once the $\partial E/\partial z_k$ is known, the
+error-derivative for the weight $w_{jk}$ on the connection from unit
+$j$ in the layer below is just $y_j \partial E/\partial z_k$.
 }

### 2015-02-17 deepreview.tex
Yann's rewrite

@@
 
-Deep neural networks that learn to represent data using layers of
-increasingly abstract features have dramatically improved the
-state-of-the-art for speech recognition, object recognition, object
-detection, predicting the activity of drug molecules, and many other
-tasks. Deep learning discovers intricate structure in large datasets by
-using the {\em backpropagation algorithm} to compute how the neural network
-should change in order to make less errors.  Deep {\em recurrent neural
-  networks} can process a sequence of inputs and produce a sequence of
-outputs and can now generate good captions for images and are competitive
-for language translation.
-
-%% this comment is geoff practising using github %%%
+Deep Learning techniques allow a machine composed of multiple layers
+of processing to learn to represent data, so that each layer transform
+the representation of the previous layer in a more abstract
+fashion. These methods have dramatically improved the state-of-the-art
+for speech recognition, object recognition, object detection,
+predicting the activity of drug molecules, and many other tasks. Deep
+learning discovers intricate structure in large datasets by using the
+{\em backpropagation algorithm} to compute how the machine should
+change in order to make less errors.  Deep {\em convolutional nets}
+have brought about dramatic improvements in the processing of array
+data, such as images, video, speech and audio, while {\em recurrent
+  net} have shown excellent performance for analyzing sequential
+inputs and producing sequential outputs such as text and
+speech. Convolutional net has quickly become the leading for image
+recognition and speech recognition, while recurrent nets are
+competitive in several natural language processing tasks, including
+language translation. Combinations of recurrent and convolutional nets
+can generate good textual captions of images.
+
 \section{Introduction}
 
-Deep learning is making major advances in solving problems
-that have resisted the best attempts of the Artificial Intelligence
-community for many years. Deep learning is very good at discovering
-intricate structure in large, high-dimensional datasets and is
-therefore applicable to many domains of science and business. For
-example, it has beaten previous machine learning techniques at
-predicting the activity of potential drug
-molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental
-particles~\citep{Melis-Higgs-boson-competition-2014}, % TODO ~\citep{YANNparticles}, 
-classifying galaxies, % %TODO ~\citep{YANNgalaxies}, 
-and predicting the effects of mutations
-in non-coding DNA on gene expression and
-disease~\citep{Keung-et-al-2014}. % Frey
-
-So far, most applications of deep learning use feedforward neural
-networks, illustrated in Figure~\ref{fig:backprop-box}, which learn
-to map a fixed-size input (e.g., an image) to a fixed-size output
-(e.g. a probability for each one of several candidate object categories). 
-They do so by successive transformations corresponding to different
-layers of artificial neurons (called {\em hidden units}),
-which can be interpreted as learned derived features of the raw input. However, 
-for tasks that involve sequential inputs, such as speech
-and language, it is often better to use recurrent neural
-networks, illustrated in Figure~\ref{fig:rnn} and discussed in Section~\ref{sec:rnn}. 
-Recurrent neural networks process an input sequence one element at a time,
-maintaining in their hidden units a ``state vector'' that implicitly contains information
-about the history all the past elements of the sequence.
-Recurrent networks are very good at predicting the next
-character in text~\citep{Sutskever-et-al-ICML2011} or the next
-word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be
-used for more complex tasks.  After reading an English sentence one
-word at a time, for example, an English ``encoder'' network can be
-trained so that the final state vector of its hidden units is a good
-representation of the thought expressed by the sentence.  This thought
-vector can then be used as the initial hidden state of (or as extra
-input to)  a jointly trained French ``decoder'' network which outputs a
-probability distribution for the first word of the French
-translation. If a particular first word is chosen 
-from this distribution and provided as input to the decoder network it will
-then output a probability distribution for the second word of the
-translation and so on until a full stop is
-chosen~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
-Overall, this process generates sequences of French words according to
-a probability distribution that depends on the English sentence.
-
-This rather naive way of performing machine translation has quickly become
-competitive with the state-of-the-art and this raises serious doubts about
-whether understanding a sentence requires anything like internal symbolic
-expressions that are manipulated by binding their symbols to variables
-in judiciously chosen rules of inference, as was previously thought by
-many early AI researchers.  The encoder-decoder translation
-networks do not work by using internal rules of inference in which
-variables get bound to symbols. They just use big state vectors, big weight
-matrices and scalar non-linearities to get the job done.  
-Deriving conclusions using only the form of the premises (which is
-what formal logic does) is a powerful
-technique that people can eventually master, but the translation
-results with recurrent neural networks suggest that it may not be a good model of
-the way we normally reason or understand language. Recurrent neural
-networks that represent thoughts as big vectors are more compatible with
-the view that everyday reasoning involves many simultaneous analogies that
-each contribute plausibility to a 
-conclusion~\citep{Lakoff+Johnson-2008,Rogers+McClelland-book2004}.
-
-This review is organized chronologically. We start with the introduction of
-the backpropagation algorithm in the 1980's and early demonstrations
-that it could learn good features for language and vision tasks.  We
-then explain the actual and perceived limitations that
-prevented it from living up to its early promise for so many years and how
-these limitations were eventually overcome.
-
-\section{Pattern recognition and the back-propagation algorithm}
-
-A neural network has input neurons that can be driven by raw input
-signals such as the pixels of an image, forming an {\em input vector},
-similarly to how sensory neurons are driven by outside signals like photons or sounds.
-Patterns can be classified by first converting each raw input vector into a
-vector of feature activations (by analogy to the response of brain neurons
-to these raw signals) and then collecting evidence for each
-different class by computing a differently weighted sum of the feature
-activations. If good
-features can be designed by hand, it is relatively easy to learn
-class-specific weights that optimize
-discrimination~\citep{Rosenblatt57}. Widely-used examples of such {\em linear
-classifiers} are {\em logistic regression} and {\em linear support vector
-machines}~\citep{bishop-book95,Boser92}.
-
-However, even with good hand-crafted features, many interesting tasks
-cannot be solved by linear classifiers. By distorting the space
-as in Figure~\ref{fig:backprop-box} (top left), a linear discrimination
-becomes possible.
-\iffalse
-A common method to get around this
-problem is to transform the feature vector into a higher dimensional vector
-(i.e., corresponding to a much larger set of features)
-through a non-linear mapping, exploiting the fact that collections of
-vectors are more likely to be easily separated when they are spread around
-a high-dimensional space~\citep{Cover65}. When there are at least
-as many dimensions as examples, a linear classifier can always separate the
-training examples, and one clever way to obtain a very high-dimensional space is to use a
-``kernel function'' that measures the similarity between two vectors.
-Using the kernel function, any input vector, ${\bf x}_i$ in the training
-set can be used to define a feature whose activation for a new input
-vector, ${\bf x}_{\rm new}$ is just the similarity between ${\bf x}_i$ and
-${\bf x}_{\rm new}$. Given a kernel function, it is possible to jointly
-optimize the choice of training examplars used to define features and the
-class-specific weights associated with the features~\citep{Boser92}. The training
-examplars used to define features are called ``support vectors'' and Support
-Vector Machines~\citep{Boser92} have been very successful, especially for
-tasks with a relatively small number of training examples in which features
-are individually informative. They do not, however, perform well for tasks
-such as object recognition in which the input space is high-dimensional and
-the input-output mapping is so complex that computing similarities using a
-simple kernel function is ineffective. 
-%% THE TEXT BELOW IS NICE BUT TOO MUCH DETAIL ABOUT  SVMS FOR THIS
-%% REVIEW.  
-%% With a general-purpose feature space
-%% and kernel (such as the Gaussian kernel), generalization far from the
-%% examples is poor~\citep{Bengio-localfailure-NIPS-2006-small}, which is problematic
-%% in high-dimensional spaces. However, kernel machines would perform well if a
-%% very good set of features were provided: learning very good
-%% features is what deep learning can do.
-\fi
-
-Problems such as image and speech recognition require the input-output
-function to be insensitive to irrelevant variations of the input, such as
+% what is machine learning
+Machine learning technology (ML) powers many aspects of modern
+society: from web search to content filtering on social networks, and
+to recommendations on e-commerce websites. It is also increasingly
+present in consumer products such as cameras and smartphones. ML
+systems are used to identify objects in images, transcribe speech into
+text, match news items, posts, or products with users' interests, and
+select relevant results of search. Increasingly, these applications
+make use a new class of techniques: {\em deep learning}.
+
+% handcrafted features
+Traditional ML techniques are limited in their ability to process
+natural data in raw form. For decades, constructing a pattern
+recogntion or ML system has required engineering efforts and domain
+expertise to design a {\em feature extractor} that transforms the raw
+data, such as the pixel values of an image, into a suitable {\em
+  internal representation}, or feature vector, from which the learning
+subsystem, often a classifier, can detect, recognize, or classify
+patterns in the input.
+
+% deep learning: learning representations
+Deep learning is a set of methods that allows a machine to be fed with
+raw data and to automatically discover suitable representations of the
+data. The name comes from the fact that deep learning machines are
+multilayer stacks of trainable modules, each of which turns the
+representation of the data in its input into a more abstract and
+higher-level representation on its output in which relevant
+characteristics and amplified and irrelevant variabilities are
+attenuated. For example, an image comes in the form of an array of
+pixel values, features in the first layer would represent the presence
+or absence of edges of a particular orientation at all locations in
+the image, the second layer would detect motifs by spotting particular
+arrangements of edges, regardless of small variations in the positions
+of the edges.  The third layer would assemble motifs into parts of
+object, and subsequent layers would detect objects as combinations of
+parts. The key aspect of deep learning is that {\em these features and
+  representations are not designed by human engineers, but are
+  automatically learned from data.}
+
+% preview of deep learning results
+Deep learning is making major advances in solving problems that have
+resisted the best attempts of the Artificial Intelligence community
+for many years. It has turned out to be very good at discovering
+intricate structure in large, high-dimensional data and is therefore
+applicable to many domains of science, business, and government. In
+addition to beating records in image
+recognition~\citep{Krizhevsky-2012-small,farabet-pami-13,tompson-nips-14,szegedy-2014}
+and speech
+recognition~\citep{Hinton-et-al-2012,Sainath-et-al-ICASSP2013}, it has
+beaten other ML techniques at predicting the activity of potential
+drug molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental
+particles~\citep{Melis-Higgs-boson-competition-2014},
+~\citep{YANNparticles}, classifying galaxies, ~\citep{YANNgalaxies},
+reconstructing brain circuits~\citep{helmstaedter-nature-2013}, and
+predicting the effects of mutations in non-coding DNA on gene
+expression and disease~\citep{Keung-et-al-2014}.
+
+Perhaps more interestingly, deep learning has produced extremely
+promising results for a variety of tasks in natural language
+understanding~\citep{collobert:2011b}, particularly topic
+classification~\citep{topic-classification}, sentiment
+analysis~\citep{Glorot+al-ICML-2011-small},
+question-answering~\cite{bordes-14}, and language
+translation~\cite{Sutskever-et-al-NIPS2014}.
+
+% supervised learning, objective function, gradient.
+The most common form of machine learning, deep or not, is {\em
+  supervised learning}. Imagine that we want to build a system that
+can classify images as containing, say, a house, a car, a person, or a
+pet. We first collect a large dataset of images of houses, cars,
+persons and pets, each labeled with its category. During training, the
+machine is shown an image and produces an output label in the form of
+a vector of scores, one for each category. We want the desired
+category to have the highest score of all categories, but this is
+unlikely to happen before training.  We compute an {\em objective
+  function}, that measures the error (or distance) between the output
+scores and the desired pattern of scores. The machine then modifies
+its internal adjustable parameters so as to reduce this error. These
+adjustable parameters, often called weights, are real numbers which
+can be seen as ``knobs'' that define the input-output function of the
+machine. In a typical deep learning system, there may be hundreds of
+millions of these adjustable weights, and tens or hundreds of millions
+of examples with which to train the machine.
+
+To properly adjust the weight vector, the learning algorithm computes
+a {\em gradient vector} which, for each weight, indicates by what
+amount the error would increase or decrease if the weight were
+increased by a tiny amount. The weight vector is then adjusted in the
+opposite direction of the gradient vector.
+
+% GD
+The objective function, averaged over all the training examples, can
+be seen as a kind hilly landscape in the high-dimensional space of
+weight values. The negative gradient vector indicates the direction of
+steepest descent in this landscape, taking is closer to a minimum
+where the output error is low on average.
+
+% SGD
+In practice, most practitioners use a procedure called {\em Stochastic
+  Gradient Descent} or SGD. It consists in showing an input, computing
+the output and the error, computing the gradient, and adjusting the
+weight vector right away. The process is repeated with each example in
+the training set, again and again, until the average of the objective
+function stops decreasing. It is called stochastic because each sample
+gives a noisy estimate of the average gradient. The procedure gets
+close to a good minimum very quickly in
+practice~\cite{bottou-bousquet-2008-small} Following this, the
+performance of the system is measured on a different set of examples,
+not seen during training, called a {\em test set}. This serves to test
+the {\em generalization ability} of the machine, i.e. its ability to
+produce sensible answers on new inputs that is has never seen during
+training.
+
+% shallow linear classifiers
+Many practical applications of ML use linear classifiers on top of
+hand-engineered features. A two-class linear classifier computes a
+weighted sum of the feature vector components. If the weighted sum is
+above a threshold, the input is classified as belonging to a
+particular category. Multiple classes can be handled by using multiple
+two-class classifiers, each classifying one class against all
+others. The learning procedure adjusts the weights of the linear
+combination, for which the gradient is very easy to compute. Linear
+classifiers have been widely used for many decades, starting with the
+Perceptron model~\citep{Rosenblatt57}, the still-popular logistic
+regression~\cite{Hastie2001}, and the linear support vector
+machine~\cite{Boser92}.
+
+% features: solving the selectivity-invariance dilemma
+It has been known since the 1960s that linear classifiers can only
+carve out their input space into very simple regions, namely
+half-spaces separated by a hyperplane~\citep{Duda-Hart}.  But problems
+such as image and speech recognition require the input-output function
+to be insensitive to irrelevant variations of the input, such as
 variations in position, orientation or illumination of an object, or
 variation in pitch or accent of speech, while being very sensitive to
-particular minute variations, such as the difference between a white wolf
-and a samoyed dog. Experimental and theoretical deep learning results 
-suggest that these tasks are better handled with multiple layers of
-progressively more complicated feature detectors each of which is defined
-in terms of the simpler feature detectors in the previous layers.
-
-From the earliest days of pattern recognition~\citep{selfridge,Rosenblatt57}
-researchers aimed to replace hand-engineered features by multiple layers of
-learned features.  This was eventually achieved by using the
-backpropagation algorithm which was discovered independently by several
-different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}.  
-The gradient of an objective function with respect to the network
-weights tells us how a change in each of the weights would influence
-the final error (top left of Figure~\ref{fig:backprop-box}).
-Gradient descent changes the weights accordingly, in the direction of the
-gradient, towards reducing the error.
-Backpropagation is simply an application of
-the chain rule of derivatives 
-which yields an efficient way of computing this gradient in
-a deep neural network (or any other composition of differentiable operations).
-%% FOR THE NOVICE READER INCLUDING BOTH WEIGHTS AND "PARAMTERS" IS CONFUSING.
-%% ITS BETTER TO SAY SOMETHING CLEARER EVEN IF IT IS LESS GENERAL OR CORRECT.
-The objective function being minimized typically measures an error to be minimized: 
-the discrepancy between
-the actual output of the network on a given training case and the target
-output for that case.  The key insight is that the error derivatives with
-respect to the {\it ouputs} of a layer can be computed by working
-backwards from the error derivatives with respect to its {\em inputs}
-to the subsequent layer.
-%% THE TRICKY STEP IS COMPUTING dE/dOUT FROM dE/dIN IN THE SUBSEQUENT
-%% LAYER.  I THINK WE SHOULD FOCUS ON GETTING OVER THAT IDEA FIRST.
-%%  I PLAN TO REDO THE FIGURE TO MAKE IT SIMPLER.  ITS MUCH TOO
-%%  CONFUSING FOR A NOVICE AT PRESENT.  WE SHOULD HAVE A SEPARATE
-%%  DISCUSSION OF THAT.
-Once this backwards recursion has been used to compute
-error derivatives for the outputs of all layers, it is easy to compute
-error derivatives for the incoming weights of the neurons in each layer 
-(Figure~\ref{fig:backprop-box}).
+particular minute variations, such as the difference between a white
+wolf and a samoyed (a breed of wolf-like white dog).  At the pixel
+level, images of two samoyeds in different poses and in different
+environments may be very different from each other, while two images
+of a samoyed and a wolf in the same position and on similar
+backgrounds may be quite similar to each other. A linear classifier,
+or any other ``shallow'' classifiers operating on raw pixels could not
+possibly distinguish the latter two, while putting the former two in
+the same category. This is why shallow classifiers require a good
+feature extractor that solve the {\em selectivity-invariance dilemma},
+i.e. that produces representations that are {\em selective} to the
+important aspects of the image, while being {\em invariant} to its
+irrelevant aspects, like the pose of the animal.
+
+% classical extensions of simple classifiers
+There have been numerous attempts, going back to the early 1960s, to
+increase the flexibility of classifiers so as to allow them to
+separate categories from one another by more complex surfaces than
+flat hyperplanes. A popular example is the class of kernel methods,
+particularly the Support Vector Machine~\citep{Boser92} which
+represents inputs by a vector whose components are similarities to
+each of the training samples computed through a so-called {\em kernel
+  function}. While these methods are good general-purpose classifiers,
+they do not perform well for tasks such as image or speech recognition
+in which the input space is very high-dimensional and the input-output
+mapping is so complex that computing similarities using a simple
+kernel function is ineffective. Designing feature extractors that are
+sensitive to the relevant information and invariant to the irrelevant
+one is what generations of computer vision and pattern recognition
+researchers have spent their careers on.
+
+% deep architectures
+But what if features could be learned? That is the key idea of deep
+learning.
+
+A deep learning architecture is a multilayer stack of modules, all (or
+most) of which are subject to learning, and many of which compute
+non-linear input-output mappings.  Each module in the stack transforms
+its input so as to increase both the selectivity and the invariance a
+little bit.  With multiple non-linear layers, say 5 to 20, a system
+can implement extremely intricate functions of its inputs that are
+simultaneously sensitive to minute details-- distinguishing samoyeds
+from wolves-- and insensitive to irrelevant variations such as the
+background, the pose, the lighting, or the surrounding objects.
+
+% backprop
+How do we train multilayer architectures?
+
+From the earliest days of pattern
+recognition~\citep{selfridge,Rosenblatt57} researchers have aimed to
+replace hand-engineered features by trainable multilayer networks.
+But despite its extreme simplicity, the solution did not appear until
+the mid 1980s and was largely forsaken by the ML community until
+recently. As it turns out, training multilayer architectures can be
+done by simple stochastic gradient descent. As long as the modules are
+relatively smooth functions of their inputs and of their internal
+weights, one can compute gradients through the {\em backpropagation
+  procedure}.
+
+The idea that this could be done, and that it could work at all, took
+a surprisingly long time to emerge. It was discovered independently by
+everal different groups in the 1970s and
+1980s~\citep{Werbos74,Parker85,LeCun85,RHW}, and became quite popular
+in the late 1980s, spurring the second wave of interest for neural
+networks. But for reasons that historians of science will have to
+explain, the method was the subject of scorn by the ML community
+between the late 1990s and the late 2000s.
+
+The backpropagtion procedure to compute gradient of an objective
+function with respect to the weights of a multilayer stack of modules
+is nothing more than a practical application of the chain rule for
+derivatives. The key insight is that the derivative (or gradient) of
+the objective with respect to the {\it input} of a module can be
+computed by working backwards from the gradient with respect to the
+{\em output} of that module (or the input of the subsequent
+module). Basically, take a module $M$ inside the network, whose input
+is $x$, weight vector is $w$, and output is $y = M(x,w)$. If one knows
+the gradient $\deri{E}{y}$ of the objective function $E$ with respect
+to the module's output $y$, one can blindly apply chain rule to
+compute the gradient with respect to the input:
+\[
+  \deri{E}{x} = \deri{y}{x}\deri{E}{y}
+\]
+This equation works even when $x$ and $y$ are vectors, in which case
+$\deri{E}{x}$ and $\deri{E}{y}$ are gradient vectors, and
+$\deri{y}{x}$ is a {\em Jacobian matrix} of $M$ where element $(i,j)$
+indicates by how much output $j$ would increase proportionally if
+input $i$ were increased by a tiny amount $\Delta y_j =
+$\left(\deri{y}{x}\right)_{ij} \Delta x_i$. The above equation
+constitutes a recurrence formula that can be used to propagate
+gradients from the top layer all the way to the bottom. Once all these
+gradients are computed, one can similarly compute the gradients with
+respect to the weights of the module. Figure~\ref{fig:backprop-box})
+shows how this works out when the network is composed of layers of
+weight matrices interspersed with scalar non-linearities.
+%% the notation in the figure should be updated.
+
+% deep neural nets
+Many applications of deep learning use feedforward neural network
+architectures, illustrated in Figure~\ref{fig:backprop-box}, which
+learn to map a fixed-size input (e.g., an image) to a fixed-size
+output (e.g. a probability for each one of several categories). Each
+pair of layer can be seen as a set of units that compute a weighted
+sum of their inputs from the previous layer and pass the result
+through a non-linear function. The most common non-linear function is
+the so-called {\em ReLU} (Rectified Linear Unit) which is simply the
+half-wave rectifier $f(z) = \max(z,0)$. In past decades, neural nets
+used smoother non-linearities, such as $\tanh(z)$, but the ReLU seems
+to work better in networks with many layers. Units that are not in the
+output layer are traditionally called {\em hidden units}. The first
+layers can be seen as distorting the input in a non-linear way so that
+categories become linearly separable by the last layer, as indicated
+in figure~\ref{fig:backprop-box}.
 
@@
 to $z$ to get the output of the unit.  For simplicity, we have omitted
-bias terms. The functions commonly used in neural networks include the
-logistic sigmoid, $f(z) = 1/(1+\exp(-z))$, the hyberbolic tangent,
-$f(z) = (\exp(z)-\exp(-z))/(\exp(z)+\exp(-z))$ and, more recently, the
-rectified linear unit, $f(z) = max(0, z)$.
+bias terms. The non-linear function used in neural networks include 
+the rectified linear unit (ReLU) $f(z) = max(0, z)$, commonly-used in 
+recent years, as well as the more traditional sigmoids, such as the 
+hyberbolic tangent, $f(z) = (\exp(z)-\exp(-z))/(\exp(z)+\exp(-z))$ 
+and logistic function logistic, $f(z) = 1/(1+\exp(-z))$.
 {\bf Bottom right.} 
@@
 
-\section{Stochastic gradient descent}
-
-Backpropagation provides an efficient way of computing gradients with respect to all of
-the weights in a layered network on a single training case, but there are
-still many ways to make use of these gradients. The most obvious method is
-to average the gradients over all of the cases in the training set and then
-to change each weight in proportion to its negative gradient. This
-moves the weight vector in the direction of steepest descent in the
-objective function. 
-%However, this is needlessly slow when the 
-%training set contains millions of cases and is highly redundant
-%with most examples conveying no new information about the mapping from
-%inputs to outputs.  
-However, this is needlessly slow when the 
-training set contains millions of cases because each example or small batch of
-examples can be seen as a noisy (but correct in average) version of the true total gradient: a very large
-number of these noisy steps can converge to a good solution
-much faster than very few but expensive clean steps, because even
-the clean gradient is not the optimal direction in which to go.
-%
-It is much faster to proceed one small ``mini-batch'' examples at a time,
-i.e., compute the average gradient
-on a small set of randomly chosen training cases (say 128) and to update
-each weight by this stochastic estimate of its full gradient, times a
-learning rate which decays as learning proceeds. Stochastic gradient
-descent can be improved by using the history of the individual weight
-gradients to adapt the learning rate for each weight~\citep{bottou-bousquet-2008-small}
-or to provide additional increments that depend on earlier gradients 
-but these techniques are still the subject of controversy and are beyond
-the scope of this review (see~\citet{Montavon2012} for tricks of the trade). 
-However, it is now widely accepted that some
-form of stochastic gradient descent is the best bet for large
-datasets.
+\section{Convolutional networks}
+
+\begin{figure}
+\begin{center}
+  \vspace{0.2cm}
+  % \includegraphics[width=0.99\linewidth]{convnet-diagram}
+  nice convnet figure goes here
+\end{center}
+\vspace{-0.6cm}
+\caption{A typical convolutional network architecture}
+\vspace{-0.8cm}
+\label{fig:convnet}
+\end{figure}
+
+One of the most widely used deep learning architectures is the
+convolutional neural network, or
+ConvNet~\cite{lecun-90c,lecun-98}. ConvNets are designed to process
+data that come in the form of multiple arrays, for example a color
+image composed of three 2D arrays containing pixel intensities in the
+three color channels. Many data modalities are in the form of multiple
+arrays: 1D for signals and sequences including language, 2D for images
+or audio spectrograms, 3D for video or volumetric images.  There are
+four key ideas behind ConvNets that take advantage of the properties
+of many such natural signals: local connections, shared weights,
+pooling, and the use of many layers.
+
+The architecture of a typical ConvNet is shown in
+figure~\ref{fig-convnet}.  It is structured as a series of stages. The
+first few stages are composed of two types of layers: convolutional
+layers and pooling layers. Units in a convolutional layer are
+organized in {\em feature maps}, within which each unit is connected
+to local patches in the feature maps of the previous layer through a
+set of weights called a {\em filter bank}. The result of this local
+weight sum is then passed through a non-linearity such as a ReLU.  All
+units in a feature map share the same filter bank. Different feature
+maps in a layer use different filter banks. The reason for this
+architecture is twofold. First, in array data such as images, local
+groups of values are often highly correlated forming distinctive local
+motifs that are easily detected. Second, the local statistics of
+images and other signals are invariant to location. In other word, if
+a motif can appear in one part of the image, it could appear anywhere,
+hence the idea of units at different locations sharing the same
+weights, and detecting the same pattern in different parts of the
+array.  Mathematically, the filtering operation performed by a feature
+map is a {\em discrete convolution}, hence the name ConvNet.
+
+While the role of the convolutional layer is to detect local
+conjunctions of features from the previous layer, the role of the
+pooling layer is to merge semantically similar features into
+one. Since the relative positions of the features forming a motif can
+vary somewhat, reliably detecting the motif can be done by
+coarse-graining the position of each feature. A typical pooling unit
+computes the max of a local patch of units in one feature map (or in a
+few feature maps). Neighboring pooling units take input from patches
+that are shifted by more than one row or column, thereby reducing the
+dimension of the representation and creating an {\em invariance} to
+small shifts and distortions.  Two or three stages of convolution,
+non-linearity, and pooling are stacked, followed by more convolutional
+and fully-connected layers.  Back-propagating gradients through a
+ConvNet is as simple as through a regular deep network, allowing all
+the weights in all the filter banks to be trained. The error
+derivative for a shared weight is just the sum the derivatives for all
+the instances of that weight.
+
+This ConvNet architecture exploits the property that many natural
+signals are {\em compositional hierarchies}. In images, local
+combinations of edges form motifs, motifs assemble into parts, and
+parts form objects in various combinations.  Similar hierarchies
+exists in speech and text from sounds to phones, phonemes, syllables,
+words, and sentences. The pooling allow features that form elements at
+the next level to vary in relative position and appearance.
+
+The convolutional and pooling layers in ConvNets are directly inspired
+by the classic notions of simple cells and complex cells in visual
+neuroscience~\citep{Hubel62}, and the overall architecture is
+reminiscent of the LGN-V1-V2-V4-IT visual cortex ventral pathway
+hierarchy ~\citep{Felleman+VanEssen-1991}. When ConvNet models and
+monkeys are shown the same picture, the activations of high-level
+units in the ConvNet explains half the variance of random sets of 160
+neurans in the monkey's
+inferro-temporal~\citep{cadieu-plos-2014}. ConvNets have their roots
+in Fukushima's Neocognitron~\citep{fukushima-82} whose architecture
+was somewhat similar but did not have an end-to-end supervised
+learning algorithm such as backpropagation plus stochastic gradient
+descent.
+
+There have been numerous applications of convolutional networks going
+back to the early 1990s. A major early success was a bank check
+reading system which was commercially deployed by AT\&T and NCR in
+1994~\citep{lecun-98}. The system used a ConvNet trained jointly with
+a probabilistic model that implemented language constraints. By the
+late 1990s this system was reading over 10\% of all the checks in the
+US. A number of ConvNet-based OCR and handwriting recognition systems
+were later deployed by
+Microsoft~\citep{simard-03,chellapilla-ist-06,chellapilla-iwfhr-06b}
+including for Arabic and
+Chinese~\citep{abdulkader-iwfhr-06,chellapilla-iwfhr-06a}. ConvNets
+have also been used since the early 1990s for object detection in
+natural images, including faces and
+hands~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95} and for face
+recognition~\cite{lawrence-tnn-1997}.
+
+\section{Image understanding with deep convolutional networks}
+
+Since the early 2000s, there have been many tasks in which ConvNets
+have been applied with success to the detection, segmentation, and
+recognition of objects and regions in images. These were invariably
+tasks in which labeled data was relatively abundant, such as traffic
+sign recognition~\cite{sermanet-ijcnn-11,Ciresan-et-al-2012}, the
+segmentation of biological images~\cite{ning-05} particularly for
+connectomics~\cite{Turaga2010}, and the detection of faces, text,
+pedestrians, and human bodies in natural
+images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,
+  nasse-09,fan-pami-2010,Jain-et-al-arxiv2013,sermanet-cvpr-13,Tompson-et-al-arxiv2014}.
+A major practical success of ConvNets in recent times is face
+detection~\citep{taigman-cvpr-2014}.
+
+An important application is the labeling of images at the pixel level,
+whose applications include driving mobile robots and self-driving
+cars~\cite{hadsell-jfr-09,farabet-icml-12}. Other applications of
+ConvNets that are gaining importance concern natural language
+understanding~\citep{collobert:2011b}, and speech
+recognition~\cite{Sainath-et-al-ICASSP2013}.
+
+Depsite these successes, ConvNets were largely forsaken by the
+mainstream computer vision and machine learning communities until the
+ImageNet competition in 2012.  When deep convolutional networks were
+applied to a dataset of about a million images from the web that
+contained a thousand different classes they achieved spectacular
+results, almost halving the error rates of the best competing
+approaches~\citep{Krizhevsky-2012-small}.  Recent ConvNet
+architectures have 10 to 20 layers of ReLUs, hundreds of millions of
+weights, and billions of connections between units. The same kind of
+network is used to generate the captions shown in
+Figure~\ref{fig:caption-generation}. This success has brought about a
+revolution in computer vision, and ConvNets are now the dominant
+approach for almost all recognition and detection
+tasks~\citep{sermanet-iclr-14,girshick-cvpr-2014,taigman-cvpr-2014,simonyan-arxiv-2014,szegedy-2014,Tompson-et-al-arxiv2014}.
+The performance of ConvNet-based vision systems has caused most major
+technology companies to start research and development projects and to
+deploy ConvNet-based image understanding products and services
+including Google, Facebook, Microsoft, Yahoo!, Twitter, Adobe, and a
+quickly-growing number of startups.
+
+If ConvNets existed in the 1990s, why did they become so popular and
+dominant only now? Certainly, the prejudice of the ML and vision
+communities against neural nets played a role. But a number of recent
+ideas and technologies enabled the success. First, ConvNets strive
+when training data is abundant, and large labeled image datasets such
+as ImageNet~\citep{imagenet_cvpr09} only appeared recently. Second,
+the appearance of GPU cards capable of over 1Tflops, together with
+GPU-optimized software for ConvNet made it possible to train these
+large networks in a few weeks (naw days) of computer time. Third, the
+use of rectifying non-linearities instead of sigmoids allowed to train
+much deeper networks than previously possible without unsupervised
+pre-training~\cite{jarrett-iccv-09,Glorot+al-AI-2011-small}. Lastly,
+the ``drop out'' regularization method~\citep{Srivastava14} was shown
+to mitigate the effects of over-fitting in very large networks.
+
+Deep convolutional nets have quickly replaced the carefully engineered
+pipelines full of hand-designed features developed by the computer
+vision community and they are now deployed in many computer vision
+applications, including image tagging, object detection, face
+recognition, facial expression recognition and scene parsing.
+
+ConvNets have been implemented in efficient hardware, using
+Field-Programmable Gate Arrays~\citep{farabet-suml-11} and are
+currently the subject of chip development in several companies that
+could enable real-time vision applications in embedded devices,
+robots, cars, and mobile devices.
+
+%% YLC: I find that this section would be cleared if it followed
+%% the section on NLP applications.
 
@@
 
-The hidden layers of a multi-layer neural network learn to represent the
-network's inputs in a way that makes it easy to predict the target
+The hidden layers of a multi-layer neural network learn to represent
+the network's inputs in a way that makes it easy to predict the target
 outputs. This is nicely demonstrated by training a multi-layer neural
 network to predict the next word in a sequence from a local context of
-earlier words~\citep{BenDucVin01-short}. 
-Each context word is presented to the network as a one-of-N
-vector, i.e., in which one component has a value of $1$ and the rest are $0$. The
-output layer has one unit per word and is forced to produce a probability
+earlier words~\citep{BenDucVin01-short}.  Each context word is
+presented to the network as a one-of-N vector, i.e., in which one
+component has a value of $1$ and the rest are $0$. The output layer
+has one unit per word and is forced to produce a probability
 distribution over words by using the ``softmax'' function:
@@
 
-The presence of a particular word at the $i$-th position in the input of the network 
-sets the value of specialized hidden units to values that correspond to the
-representation of that word. Different words create a different pattern of
-activations, or word-vector, illustrated in Figure~\ref{fig:word-embeddings}.
-In a language model, the other layers of the network learn to convert the input word-vectors
-into an output word-vector for the predicted next-word, which can be used
-to predict the probability for any word in the vocabulary to appear as the next word.
-The network learns word-vectors that contain many active components
-each of which can be interpreted as a separate feature of the word. This
-was first demonstrated using sequences of three words, such as
-(Colin, Mother, Victoria), derived from two family trees~\citep{RHW}. The learned word
-vector for Colin contained ``semantic'' features representing which family
-tree Colin was in, what branch of that tree he was in, and what generation
-he was from. 
+The presence of a particular word at the $i$-th position in the input
+of the network sets the value of specialized hidden units to values
+that correspond to the representation of that word. Different words
+create a different pattern of activations, or word-vector, illustrated
+in Figure~\ref{fig:word-embeddings}.  In a language model, the other
+layers of the network learn to convert the input word-vectors into an
+output word-vector for the predicted next-word, which can be used to
+predict the probability for any word in the vocabulary to appear as
+the next word.  The network learns word-vectors that contain many
+active components each of which can be interpreted as a separate
+feature of the word. This was first demonstrated using sequences of
+three words, such as (Colin, Mother, Victoria), derived from two
+family trees~\citep{RHW}. The learned word vector for Colin contained
+``semantic'' features representing which family tree Colin was in,
+what branch of that tree he was in, and what generation he was from.
 %% TWO NATIVE ENGLISH SPEAKERS AGREE THAT THE "FROM" SHOULD STAY AT THE
@@
 The family tree demonstration used a small, clean dataset in which the
-regularities could be captured by micro-rules that were never violated and
-the features were easy to interpret, but learned word-vectors also work
-very well when the word sequences come from a large corpus of real text and
-the individual micro-rules are unreliable~\citep{BenDucVin01-short}. When trained to
-predict the next word in a news story, for example, the learned word
-vectors for Tuesday and Wednesday are very similar, as are the word vectors
-for Vietnam and Iraq. This is not to say that Tuesday and Wednesday are
+regularities could be captured by micro-rules that were never violated
+and the features were easy to interpret, but learned word-vectors also
+work very well when the word sequences come from a large corpus of
+real text and the individual micro-rules are
+unreliable~\citep{BenDucVin01-short}. When trained to predict the next
+word in a news story, for example, the learned word vectors for
+Tuesday and Wednesday are very similar, as are the word vectors for
+Vietnam and Iraq. This is not to say that Tuesday and Wednesday are
 put in the same category, but rather that they share many semantic and
-grammatical attributes, so that what the neural network learns from one day of
-the week can often be generalized to the other, while individual days remain
-different (for example one imagines that the neural network also
-learns to some features that distinguises week-end days and weekdays, while
-Saturday and Sunday might have special roles in some religions, etc.).
-Such representations are called {\em distributed representations} because
-its elements (the features) are not mutually exclusive and there exists
-a rich and vast set of possible configurations of values of these features,
-corresponding to the variations seen in the observed data.
-These word-vectors are thus {\em learned features} that were not determined
-ahead of time by humans, but automatically discovered by the neural network.
-Vector representations of words learned from text
-are now very widely used in natural language 
+grammatical attributes, so that what the neural network learns from
+one day of the week can often be generalized to the other, while
+individual days remain different (for example one imagines that the
+neural network also learns to some features that distinguises week-end
+days and weekdays, while Saturday and Sunday might have special roles
+in some religions, etc.).  Such representations are called {\em
+  distributed representations} because its elements (the features) are
+not mutually exclusive and there exists a rich and vast set of
+possible configurations of values of these features, corresponding to
+the variations seen in the observed data.  These word-vectors are thus
+{\em learned features} that were not determined ahead of time by
+humans, but automatically discovered by the neural network.  Vector
+representations of words learned from text are now very widely used in
+natural language
 applications~\citep{Schwenk-2007,collobert:2011b,Socher-2011,
@@
 
-The issue of representation lies at the heart of the debate between the
-logic-inspired and the neurally-inspired paradigms for cognition. In the
-logic-inspired paradigm, an instance of a symbol is something whose only
-property is that it is either identical or non-identical to other symbol
-instances. It has no internal structure that is relevant to its use. By
-contrast, neural networks use large activity vectors instead of symbols and
-parallel matrix operations followed by non-linearities instead of formal
-rules of inference. The elements of such vectors can be understood as
-the values of learned attributes (or features) associated with some object. 
-With sufficient data, computational power and depth,
-the combination of backpropagation for computing gradients and stochastic
-gradient descent for updating the weights on connections has proved to be
-very good at learning appropriate distributed representations. These
-representations allow a lot of parallel computation at the feature level so
-that complex tasks can be solved in a very small number of sequential
-steps. In fact, it can be shown that a measure of the richness of the functions
-that can be represented by a deep neural network grows exponentially
-with depth~\citep{Montufar-et-al-NIPS2014}, whereas according to the
-same measure, classical (shallow) non-parametric statistical methods such as
-Gaussian Support Vector Machines~\citep{Bengio-localfailure-NIPS-2006-small} 
-or decision trees~\citep{Bengio-decision-trees10}
-would require exponential storage 
+The issue of representation lies at the heart of the debate between
+the logic-inspired and the neurally-inspired paradigms for
+cognition. In the logic-inspired paradigm, an instance of a symbol is
+something whose only property is that it is either identical or
+non-identical to other symbol instances. It has no internal structure
+that is relevant to its use. By contrast, neural networks use large
+activity vectors instead of symbols and parallel matrix operations
+followed by non-linearities instead of formal rules of inference. The
+elements of such vectors can be understood as the values of learned
+attributes (or features) associated with some object.  With sufficient
+data, computational power and depth, the combination of
+backpropagation for computing gradients and stochastic gradient
+descent for updating the weights on connections has proved to be very
+good at learning appropriate distributed representations. These
+representations allow a lot of parallel computation at the feature
+level so that complex tasks can be solved in a very small number of
+sequential steps. In fact, it can be shown that a measure of the
+richness of the functions that can be represented by a deep neural
+network grows exponentially with
+depth~\citep{Montufar-et-al-NIPS2014}, whereas according to the same
+measure, classical (shallow) non-parametric statistical methods such
+as Gaussian Support Vector
+Machines~\citep{Bengio-localfailure-NIPS-2006-small} or decision
+trees~\citep{Bengio-decision-trees10} would require exponential
+storage
 %% an exponentially large number of training examples 
@@
 \section{Language processing with deep networks}
+
+%% YLC: we need to talk about word2vec somewhere around here.
+%% it's so popular that we can't just not mention it.
 
@@
 
-\section{Convolutional networks}
-
-%\begin{figure}
-%\begin{center}
-%  \vspace{0.2cm}
-%  % \includegraphics[width=0.99\linewidth]{convnet-diagram}
-%  nice convnet figure goes here
-%\end{center}
-%\vspace{-0.6cm}
-%\caption{A typical ConvNet architecture}
-%\vspace{-0.8cm}
-%\label{fig:convnet}
-%\end{figure}
-
-One of the most widely used deep learning architectures is the
-convolutional neural network, or ConvNet~\cite{lecun-90c,lecun-98}. ConvNets
-are designed to process data that come in the form of arrays, for
-example a color image composed of three 2D arrays containing pixel
-intensities in the three color channels. Many data modalities are in
-the form of multiple arrays: 1D for signals and language, 2D for
-images or audio spectrograms, 3D for video or volumetric images.
-There are four key ideas behind ConvNets that take advantage of the
-properties of many such natural signals: local connections, shared
-weights, pooling, and the use of many layers.
-
-The architecture of a typical ConvNet is 
-%shown in
-%figure~\ref{fig-convnet}. 
-structured as a series of stages. The first few stages are composed of two
-types of layers: convolutional layers and pooling layers. Units in a
-convolutional layer are organized in {\em feature maps}, within which
-each unit is connected to local patches in the feature maps of the
-previous layer through a set of weights called a {\em filter
-  bank}. The result of this local weight sum is then passed through a
-non-linearity.  All units in a feature map share the same filter
-bank. Different feature maps in a layer use different filter banks. The
-reason for this architecture is twofold. First, in array data such as
-images, local groups of values are often highly correlated forming
-local motifs that are easily detected. Second, the spatially local statistics of
-images and other signals are invariant to location. In other word, if
-a motif can appear in one part of the image, it could appear anywhere,
-hence the idea of units at different locations sharing the same
-weights.  Mathematically, the filtering operation performed by a
-feature map is a {\em discrete convolution}, hence the name
-ConvNet.
-
-While the role of the convolutional layer is to detect local
-conjunctions of features in the previous layer, the role of the
-pooling layer is to merge semantically similar features into
-one. Since the relative positions of the features forming a motif can
-vary somewhat, reliably detecting the motif can be done by
-coarse-graining the position of each feature. A typical pooling unit
-computes the max of a local patch of units in one feature map (or in a
-few). Neighboring pooling units take input from patches that are
-shifted by more than one row or column, thereby reducing the dimension
-of the representation and creating an invariance to small shifts.  Two
-or three stages of convolution, non-linearity, and pooling are
-stacked, followed by more convolutional and fully-connected layers.
-Back-propagating gradients through a ConvNet is as simple
-as through a regular deep network, allowing all the weights in all the
-filter banks to be trained. The error derivative for a shared weight
-is just the sum the derivatives for all the instances of that weight.
-
-This architecture exploits the property that many natural signals are
-{\em compositional hierarchies}. In images, local combinations of
-edges form motifs, motifs assemble into parts, and parts form objects.
-Similar hierarchies exists in speech and text from sounds to phones,
-phonemes, syllables, words, and sentences. The features that form
-elements at the next level can vary in relative position and
-appearance. The convolutional and pooling layers in ConvNets are
-directly inspired by the classic notions of simple cells and complex
-cells in visual neuroscience~\citep{Hubel62}, and the overall
-architecture is reminiscent of the LGN-V1-V2-V4-IT visual cortex ventral pathway hierarchy
-~\citep{Felleman+VanEssen-1991}. ConvNets have their roots in
-Fukushima's Neocognitron~\citep{fukushima-82} whose architecture was
-somewhat similar but did not have an end-to-end supervised learning
-algorithm such as backpropagation plus stochastic gradient descent.
-
-There have been numerous applications of convolutional networks going
-back to the early 1990s. A major early success
-%% THIS REWORDING AVOIDS COMPLAINTS FROM HECHT NIELSEN ETC.
-was a bank check
-reading system which was commercially deployed by AT\&T and NCR in
-1994~\citep{lecun-98}. The system used a ConvNet trained jointly with
-a probabilistic model that implemented language constraints. By the late
-1990s this system was reading over 10\% of all the checks in the
-US. Microsoft later deployed ConvNets in a number of OCR and handwriting
-recognition
-systems~\citep{simard-03,chellapilla-ist-06,chellapilla-iwfhr-06b}
-including for Arabic and
-Chinese~\citep{abdulkader-iwfhr-06,chellapilla-iwfhr-06a}. ConvNets
-have also been used since the early 1990s for object detection in
-natural images, including faces and
-pedestrians~\cite{vaillant-monrocq-lecun-94,garcia-delakis-04,osadchy-07,nasse-09,sermanet-cvpr-13}.
-
-\section{Image understanding with deep convolutional networks}
-
-Since the early 2000s, there have been many tasks in which ConvNets
-have been applied with success to the detection, segmentation, and
-recognition of objects and regions in images. These were invariably
-tasks in which labeled data was relatively abundant, such as traffic
-sign recognition~\cite{sermanet-ijcnn-11,Ciresan-et-al-2012}, the
-segmentation of biological images~\cite{ning-05} particularly for
-connectomics~\cite{Turaga2010}, the detection of faces, text, and human bodies in natural
-images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,
-nasse-09,Jain-et-al-arxiv2013,sermanet-cvpr-13,Tompson-et-al-arxiv2014}.
-
-An important application is the labeling of images at the pixel level,
-whose applications include driving mobile robots and self-driving
-cars~\cite{hadsell-jfr-09,farabet-icml-12}. Other applications of ConvNets that are gaining importance concern
-natural language understanding~\citep{collobert:2011b}, and speech
-recognition~\cite{Sainath-et-al-ICASSP2013}.
-
-ConvNets were largely ignored by the mainline computer vision and machine
-learning community until the ImageNet competition in 2012.
-When deep convolutional networks were applied to a dataset of about a
-million images from the web that contained a thousand different classes
-they achieved spectacular results, almost halving the error rates of the
-best computer vision sytems~\citep{Krizhevsky-2012-small}.  
-The same kind of network is used to generate the captions shown in Figure~\ref{fig:caption-generation}.
-
-The major factors in this
-success were the large labeled dataset~\citep{imagenet_cvpr09}, the use of fast GPUs,
-and the convolutional early layers.  
-%% REPETITION : YB, actually it was not said earlier, only that ReLU was better than sigmoid or tanh
-It had already been shown that deep
-supervised nets with rectified linear units could be trained without
-unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
-Rectified linear units made the
-training much faster and the fully connected final layers were prevented
-from overfitting the peculiarities of the training set by randomly omitting
-hidden units during training~\citep{Srivastava14}. Deep convolutional nets
-have quickly replaced the carefully engineered pipelines full of
-hand-designed features developed by the computer vision community and they
-are now deployed in many computer vision applications, including image
-tagging, object detection, face recognition, facial expression recognition
-and scene parsing.
-
-\iffalse
-YANN:
-CAN WE MOVE SOME OF THE LATER REFS IN THE EARLIER CONVNET SECTION TO
-HERE?
-\fi
-
 %%%%%%%%%%%%%%%%%%%%
@@
 
-When backpropagation was first introduced, its most exciting use was for
-training recurrent neural networks. These receive input vectors at each
-time-step and have one or more layers of hidden units that, in addition to
-having connections to higher layers or ouput units, also have connections
-that allow the neurons within a layer to receive weighted input from those
-same neurons at the previous time-step.  As illustrated in
-Figure~\ref{fig:rnn}, when we consider the outputs of the hidden units at
-different discrete time steps as if they were the outputs of different neurons
-in a deep multi-layer network, it becomes clear how we can apply backpropagation
-to train recurrent networks. The
-the gradient with respect to a recurrent
-weight, $w_{ij}$ that allows neuron $i$ at time $t$ to influence the
-activity of neuron $j$ at time $t+1$ can be computed by
-backpropagation-through-time (BPTT) of the error derivative for $j$ at time
-$t+1$. The full gradient with respect to $w_{ij}$ on a whole input sequence is just the
-sum of its gradients at each time-step.
-
-Recurrent neural networks are very powerful dynamical systems, but training
-them with BPTT proved to be problematic, because the backpropagated
-gradients either grow or shrink at each time step, so over many time-steps
-they typically explode or
+%% YLC: I moved this section down from the intro. It was redundant there.
+%% Some of it is probably redundant with much of what follows
+For tasks that involve sequential inputs, such as speech and language,
+it is often better to use recurrent neural networks, illustrated in
+Figure~\ref{fig:rnn}.  Recurrent neural networks process an input
+sequence one element at a time, maintaining in their hidden units a
+``state vector'' that implicitly contains information about the
+history all the past elements of the sequence.  Recurrent networks are
+very good at predicting the next character in
+text~\citep{Sutskever-et-al-ICML2011} or the next word in a
+sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
+more complex tasks.  After reading an English sentence one word at a
+time, for example, an English ``encoder'' network can be trained so
+that the final state vector of its hidden units is a good
+representation of the thought expressed by the sentence.  This thought
+vector can then be used as the initial hidden state of (or as extra
+input to) a jointly trained French ``decoder'' network which outputs a
+probability distribution for the first word of the French
+translation. If a particular first word is chosen from this
+distribution and provided as input to the decoder network it will then
+output a probability distribution for the second word of the
+translation and so on until a full stop is
+chosen~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
+Overall, this process generates sequences of French words according to
+a probability distribution that depends on the English sentence.
+
+This rather naive way of performing machine translation has quickly become
+competitive with the state-of-the-art and this raises serious doubts about
+whether understanding a sentence requires anything like internal symbolic
+expressions that are manipulated by binding their symbols to variables
+in judiciously chosen rules of inference, as was previously thought by
+many early AI researchers.  The encoder-decoder translation
+networks do not work by using internal rules of inference in which
+variables get bound to symbols. They just use big state vectors, big weight
+matrices and scalar non-linearities to get the job done.  
+Deriving conclusions using only the form of the premises (which is
+what formal logic does) is a powerful
+technique that people can eventually master, but the translation
+results with recurrent neural networks suggest that it may not be a good model of
+the way we normally reason or understand language. Recurrent neural
+networks that represent thoughts as big vectors are more compatible with
+the view that everyday reasoning involves many simultaneous analogies that
+each contribute plausibility to a 
+conclusion~\citep{Lakoff+Johnson-2008,Rogers+McClelland-book2004}.
+
+%% YLC: This is where the original recurrent net section begins.
+%% There should be some merging between the stuff above and the stuff below.
+
+When backpropagation was first introduced, its most exciting use was
+for training recurrent neural networks. These receive input vectors at
+each time-step and have one or more layers of hidden units that, in
+addition to having connections to higher layers or ouput units, also
+have connections that allow the neurons within a layer to receive
+weighted input from those same neurons at the previous time-step.  As
+illustrated in Figure~\ref{fig:rnn}, when we consider the outputs of
+the hidden units at different discrete time steps as if they were the
+outputs of different neurons in a deep multi-layer network, it becomes
+clear how we can apply backpropagation to train recurrent
+networks. The the gradient with respect to a recurrent weight,
+$w_{ij}$ that allows neuron $i$ at time $t$ to influence the activity
+of neuron $j$ at time $t+1$ can be computed by
+backpropagation-through-time (BPTT) of the error derivative for $j$ at
+time $t+1$. The full gradient with respect to $w_{ij}$ on a whole
+input sequence is just the sum of its gradients at each time-step.
+
+Recurrent neural networks are very powerful dynamical systems, but
+training them with BPTT proved to be problematic, because the
+backpropagated gradients either grow or shrink at each time step, so
+over many time-steps they typically explode or
 vanish~\citep{Hochreiter91-small,Bengio-et-al-TNN1994}.
@@
 
-\section{Overcoming the limitations of backpropagation in deep networks}
-
-In the 1990s, the machine learning community decided that backpropagation
-in deep neural networks did not work well in practice and it was largely
-abandoned.  At the time, people believed that learning many layers of
-non-linear feature detectors with no prior knowledge was much too difficult
-for a simple gradient descent procedure and that the networks were getting
-trapped in poor local minima, weight configurations for which no small
-change would reduce the average error. We now know that the main reasons for the
-limited success of backpropagation in deep networks were quite different:
-The labeled datasets were far too small, the computers were far too slow,
-and the weights were not initialised sensibly~\citep{GlorotAISTATS2010-small}.
-
-The obvious perceived weakness of gradient descent is that it can get trapped in poor
-local minima, but in high-dimensional weight-spaces this is much less
-common than getting stuck close to {\em saddle points}, where the error surface
-curves upwards in most directions but downwards in a few. Recent
-results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015} strongly suggest
-that the vast majority of the critical points\footnote{where derivatives of the total
-objective function with respect to the parameters vanish, and include minima,
-maxima and saddle points} are saddle points, not
-local minima, and that in large and deep networks almost all of the 
-saddle points with predominantly upwards curvature  
-give very similar performance, so it does not matter which one we find.  This goes a long way
-towards explaining why stochastic gradient descent works so well for deep
-neural networks.
-
-Interest in deep feedforward networks was revived in
-%% Even though I had a conference  paper in 2005, the revival was in 2006!
-2006~\citep{IJCAI,Hinton06-small,Bengio-nips-2006-small,ranzato-07-small} 
-by the introduction
-of unsupervised learning procedures that could create layers of feature
-detectors without requiring labeled data. The objective in learning each
-layer of feature detectors was to be able to reconstruct or model the
-activities of feature detectors (or raw inputs) in the layer below.  By
-``pre-training'' several layers of progressively more complex feature
-detectors using this reconstruction objective, the weights of a deep
-network could be initialized to sensible values.  A final layer of output
-units could then be added to the top of the network and the whole deep
-system could be fine-tuned using standard
-backpropagation \citep{Hinton-Science2006,Bengio-nips-2006-small,ranzato-07-small}. 
-This worked remarkably well for recognizing
-handwritten digits especially when the amount of labeled data was very
-limited~\citep{Erhan+al-2010-small}.
-
-The first major application of this pre-training approach was in speech
-recognition and it was made possible by the advent of fast Graphics
-Processing Units (GPUs) that were convenient to
-program~\citep{RainaICML09-small} and allowed researchers to train
-networks 10 or 20 times faster.  In 2009, the approach was used to map
-short temporal windows of coefficients extracted from a soundwave to a set
-of probabilities for the various fragments of speech that might be
-represented by the frame in the center of the window.  It achieved
-record-breaking results on a standard speech recognition benchmark that
-used a small vocabulary~\citep{TIMITpaper} and was quickly developed to
-give record-breaking results on a large vocabulary task~\citep{Dahl2012}.
-By 2012, versions of the deep net from 2009 were being developed by many of
-the major speech groups~\citep{Hinton-et-al-2012} and were already being
-deployed in Android phones.  For smaller datasets, unsupervised
-pre-training helps to prevent overfitting~\citep{Erhan+al-2010-small},
-leading to significantly better generalization when the number of labeled
-examples is small, or in a transfer setting where we have lots of
-examples for some ``source'' tasks but very few for some ``target''
-task. In this context, unsupervised representation learning helped to win
-several transfer learning
-competitions~\citep{UTLC+LISA-2011-small,Goodfellow-icml2012}.  Once deep
-learning had been rehabilitated, it turned out that large labeled datasets
-did not need the pre-training stage, provided the scales of the initial
-random weights were carefully chosen to avoid very large or very small
-initial gradients~\citep{GlorotAISTATS2010-small}. Even better
-results, especially for very deep networks,
-could be obtained by using rectifying non-linearities rather than
-sigmoid or tanh non-linearities
-~\citep{Nair-2010-small,Glorot+al-AI-2011-small,Krizhevsky-2012-small,Dahl-et-al-ICASSP2013}.
-
-Interest in recurrent networks (which correspond to a very deep network
-when unfolded in time as in Figure~\ref{fig:rnn}) was revived by
-``Long Short Term Memory'' networks (LSTMs) that use special hidden units
-whose natural behaviour is to remember inputs for a long
-time~\citep{Hochreiter+Schmidhuber-1997}.  
-A linear ``memory cell'' interacts with other units
-via gated connections whose weights are multiplied by the state of a
-logistic gating unit that takes a value between $0$ and $1$. An input gating
-unit can learn when to isolate the cell from the effect of the outside world,
-and an output gating unit prevents the content of the memory cell to be
-broadcast to the rest of the neural network. More importantly, the
-memory cell itself acts like an accumulator or a gated leaky neuron:  it has a connection to
-itself at the next time-step that has a weight of 1 so it copies its
-own real-valued state and accumulates the external signal, 
-but this self-connection is {\em gated} by a ``forget gate'' so
-the stored value quickly leaks away unless the forget gate has an activation
-near $1$.  The gates themselves have their activities determined by
-weighted inputs that they receive from the input vectors and all of the
-memory cells.
-
-When LSTMs were first described % in English
-\citep{Hochreiter+Schmidhuber-1997},
-their complexity may have been off-putting
-and, despite their impressive ability to learn very long-range temporal
-structure on toy problems, they were largely ignored until they achieved
-excellent results in a competition to read cursive
-hand-writing~\citep{Graves-et-al-2009}. They have subsequently proved to be very
-effective for tasks that have sequential input.  When they use several
-hidden layers, they are better than
-deep feedforward networks for speech recognition~\citep{Graves-et-al-ICASSP2013} and
-unlike feedforward nets they can implement an entire speech recognition
-system that goes all the way from coefficents derived from the soundwave to
-the sequence of characters in the transcription.  LSTMs or related forms of gated
-units are also what is currently used for the encoder and decoder networks that
-perform so well at machine
+\section{Recurrent Networks with Memory}
+%% YLC: I moved this up because it connects with the RNN section above.
+
+Recurrent networks, once unfolded in time, can be seen as very deep
+feed-forward networks in which all the layers share the same weights.
+While their main purpose is to learn long-term dependencies, they
+don't actually seem to be able to store any information for very
+long. The dynamics of the system is such that information about the
+initial state vanishes quickly.
+
+To correct for that, one idea is to augment the network with an
+explicit memory. The first proposal of this kind is the ``Long Short
+Term Memory'' networks (LSTMs) that use special hidden units whose
+natural behaviour is to remember inputs for a long
+time~\citep{Hochreiter+Schmidhuber-1997}.  A linear ``memory cell''
+interacts with other units via gated connections whose weights are
+multiplied by the state of a logistic gating unit that takes a value
+between $0$ and $1$. An input gating unit can learn when to isolate
+the cell from the effect of the outside world, and an output gating
+unit prevents the content of the memory cell to be broadcast to the
+rest of the neural network. More importantly, the memory cell itself
+acts like an accumulator or a gated leaky neuron: it has a connection
+to itself at the next time-step that has a weight of 1 so it copies
+its own real-valued state and accumulates the external signal, but
+this self-connection is {\em gated} by a ``forget gate'' so the stored
+value quickly leaks away unless the forget gate has an activation near
+$1$.  The gates themselves have their activities determined by
+weighted inputs that they receive from the input vectors and all of
+the memory cells.
+
+When LSTMs were first described \citep{Hochreiter+Schmidhuber-1997},
+their complexity may have been off-putting and, despite their
+impressive ability to learn very long-range temporal structure on toy
+problems, they were largely ignored until they were shown to yield
+excellent results in a cursive handwriting recognition
+competition~\citep{Graves-et-al-2009}. They have subsequently proved
+to be very effective for tasks that have sequential input.  When they
+use several hidden layers, they are better than deep feedforward
+networks for speech recognition~\citep{Graves-et-al-ICASSP2013} and
+unlike feedforward nets they can implement an entire speech
+recognition system that goes all the way from coefficents derived from
+the soundwave to the sequence of characters in the transcription.
+LSTMs or related forms of gated units are also what is currently used
+for the encoder and decoder networks that perform so well at machine
 translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
 
-Recent work~\citep{Chung-et-al-NIPSDL2014-small,Yao-et-al-SLU-workshop2014} 
-suggests that not all the elements of the LSTM
-architecture are required, the crucial element being the presence and learnability 
-of the forget gates which dynamically control the leaking rates  of
-the memory cells.
+Over the last year, several authors have made different proposals to
+augment recurrent nets with a memory module. This includes simplified
+version of the LSTM
+circuit~\citep{Chung-et-al-NIPSDL2014-small,Yao-et-al-SLU-workshop2014}.
+The results suggest that not all the elements of the LSTM architecture
+are required.
+
+Other proposals include the ``Neural Turing Machine'' in which the
+network is augmented by a ``tape-like'' memory that the recurrent net
+can chose to read from or write to~\citep{Graves-et-al-arxiv2014}, and
+the ``Memory Network'' in which a regular network is augmented by a
+kind of associative memory~\citep{weston-memorynet-2014}. Memory
+networks have yielded excellent performance on standard
+question-answering benchmarks. The memory is used to remember the
+story about which the network is asked to answer questions.
+
 %It is not yet clear which aspects of the complicated architecture of LSTMs
@@
 %predicting the next word in text\citep{LeHinton}.
+
+\section{Learning to reason and to manipulate symbols}
+
+Beyond simple memorization, Neural Turing Machines and Memory Networks
+are being used for task that would normally require to manipulate
+symbols and to reason.
+
+Neural Turing Machines can be taught ``algorithms''. Among other
+things, they can learn to ouput a sorted list of symbols when its
+input consists of an unsorted sequence in which each symbol is
+accompanied by a real value that indicates its priority in the list.
+To achieve this it must learn a sorting program purely from examples
+of the input and output sequences~\citep{Graves-et-al-arxiv2014}.
+
+Memory Networks can be trained to keep track of the state of the world
+when they are told a story, so that they can answer questions about
+the state of the world at the end of the story. Following a sequence
+of actions in a virtual world described in words, in a scenario
+similar to a text adventure game, the network is shown an
+automatically0generate question and taught to produce the
+corresponding answer (also generated automatically).  In one test
+example, the network is shown a 15-sentence version of the Lord of The
+Rings and correctly answers questions like ``where is Frodo now?''.
+~\citep{weston-memorynet-2014}
+
+A different system that also uses BPTT can learn to map the sequence
+of characters in a simple program written in Python to the sequence of
+characters that the program would
+print~\citep{Zaremba+Sutskever-arxiv2014}.  Rather surprisingly, BPTT
+can learn about {\it for loops} and can also learn to perform addition
+and multiplication of strings containing several decimal
+digits. Achieving this kind of behaviour seemed ludicrously optimistic
+when BPTT was first proposed.
+
+In a somewhat related work, a network is trained to evaluate whether a
+particular mathematical transformation of a formula is lilely to lead
+to a correct mathematical identity~\cite{zaremba-nips-2014}. The
+system was able to produce efficient identities for matrix expressions
+whose formula fill several pages.
+
+The renewal of interest in deep learning and recurrent neural nets is
+spurring a whole new generation of creative ideas.
 
@@
 
-\section{Learning to process symbols}
-
-Deep learning is now the method of choice for many perceptual tasks and is
-making major inroads in natural language processing where a huge amount of
-messy commonsense knowledge is required. Rapid progress is also being made
-in applying deep learning, in particular BPTT, to relatively pure symbol processing tasks.
-An impressive recent example is the ``Neural Turing Machine'' which is a
-recurrent neural network that can choose when and what to write to or read
-from an external memory~\citep{Graves-et-al-arxiv2014}. 
-Among other things, this can learn to
-ouput a sorted list of symbols when its input consists of an unsorted
-sequence in which each symbol is accompanied by a real value that indicates
-its priority in the list.  To achieve this it must learn a sorting program
-purely from examples of the input and output sequences.  A different system
-that also uses BPTT can learn to map the sequence of characters in a simple
-program written in Python to the sequence of characters that the program
-would print~\citep{Zaremba+Sutskever-arxiv2014}. 
-Rather surprisingly, BPTT can learn about {\it for
-  loops} and can also learn to perform addition and multiplication of
-strings containing several decimal digits. Achieving this kind of behaviour
-seemed ludicrously optimistic when BPTT was first proposed.
+\section{Overcoming the limitations of backpropagation}
+
+In the late 1990s, neural nets and backpropagation were largely
+forsaken by the machine learning community. At the time, people
+believed that learning many layers of non-linear feature detectors
+with no prior knowledge was much too difficult for a simple gradient
+descent procedures and that the networks were getting trapped in poor
+local minima, weight configurations for which no small change would
+reduce the average error, despite considerable empirical evidence to
+the contrary in the case of ConvNets~\citep{lecun-98}. We now know
+that the main reasons for the limited success of backpropagation in
+deep networks were quite different: labeled datasets that were often
+too small, computers that were too slow, non-linearities that were
+inappropriate, weights that were not properly
+initialised~\citep{lecun-98b,GlorotAISTATS2010-small}, and a lack of
+faith in the potential success.
+
+The obvious perceived weakness of stochastic gradient descent is that
+it can get trapped in poor local minima, but it turns out that local
+minima are rarely a problem with large networks. Regardless of the
+initial conditions, the system always reaches solutions of similar
+quality.  Recent results strongly suggest the landscape is packed with
+a combinatorially large number of {\em saddle points} where the
+gradient is zero, and the surface curves up in some dimensions and
+curves down in
+others~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
+The analysis seems to show that the local minima are present in very
+large numbers, but almost all of them have very similar error levels.
+Hence it doesn't seem to matter which one the algorithm
+finds. Avoiding being trapped near saddle points seems more important.
+
+It also took an inordinate amount of time for the community to adopt
+stochastic gradient methods for large-scale machine learning. key
+theoretical and practical work showed that, while SGD has terrible
+asymptotic convergence properties, it is extremely efficient for the
+kind of approximate optimization required in ML
+applications~\citep{bottou-bousquet-2008-small}. But when trainnig a
+large ConvNet can take days or weeks on the fastest GPUs, one is eager
+to discover more efficient optimization methods that use the curvature
+information the past update directions~\citep{lecun-98b,Montavon2012}.
+
+The parallelization of large neural nets over a single GPU has
+required to use ``mini-batches'' of samples over which the gradients
+are averaged. The purpose of this is to saturate the GPU. But one big
+unsolved question is how to scale up the training of very large neural
+nets over multiple GPU card across multiple machines. As of today, this
+question is unresolved.
+
+Interest in deep feedforward networks was revived around
+2006~\citep{IJCAI,Hinton06-small,Bengio-nips-2006-small,ranzato-07-small}
+by the introduction of unsupervised learning procedures that could
+create layers of feature detectors without requiring labeled data. The
+objective in learning each layer of feature detectors was to be able
+to reconstruct or model the activities of feature detectors (or raw
+inputs) in the layer below.  By ``pre-training'' several layers of
+progressively more complex feature detectors using this reconstruction
+objective, the weights of a deep network could be initialized to
+sensible values.  A final layer of output units could then be added to
+the top of the network and the whole deep system could be fine-tuned
+using standard backpropagation
+\citep{Hinton-Science2006,Bengio-nips-2006-small,ranzato-07-small}.
+This worked remarkably well for recognizing handwritten digits or for
+detecting pedestrians, especially when the amount of labeled data was
+very limited~\citep{Erhan+al-2010-small,sermanet-cvpr-13}.
+
+The first major application of this pre-training approach was in
+speech recognition and it was made possible by the advent of fast
+Graphics Processing Units (GPUs) that were convenient to
+program~\citep{RainaICML09-small} and allowed researchers to train
+networks 10 or 20 times faster.  In 2009, the approach was used to map
+short temporal windows of coefficients extracted from a soundwave to a
+set of probabilities for the various fragments of speech that might be
+represented by the frame in the center of the window.  It achieved
+record-breaking results on a standard speech recognition benchmark
+that used a small vocabulary~\citep{TIMITpaper} and was quickly
+developed to give record-breaking results on a large vocabulary
+task~\citep{Dahl2012}.  By 2012, versions of the deep net from 2009
+were being developed by many of the major speech
+groups~\citep{Hinton-et-al-2012} and were already being deployed in
+Android phones.  For smaller datasets, unsupervised pre-training helps
+to prevent overfitting~\citep{Erhan+al-2010-small}, leading to
+significantly better generalization when the number of labeled
+examples is small, or in a transfer setting where we have lots of
+examples for some ``source'' tasks but very few for some ``target''
+task. In this context, unsupervised representation learning helped to
+win several transfer learning
+competitions~\citep{UTLC+LISA-2011-small,Goodfellow-icml2012}.  Once
+deep learning had been rehabilitated, it turned out that large labeled
+datasets did not need the pre-training stage. Even better results,
+especially for very deep networks, could be obtained by using
+rectifying non-linearities rather than sigmoid or tanh non-linearities
+~\citep{Nair-2010-small,Glorot+al-AI-2011-small,Krizhevsky-2012-small,Dahl-et-al-ICASSP2013}.
 
@@
 
-Human vision is an active process in which intelligent fixations are used
-to ensure that only a tiny fraction of the optic array is ever processed at
-high resolution. We expect much of the future progress in vision,
-especially for video understanding or robotics, to come from systems that
-are trained end-to-end and combine deep convolutional networks with
-recurrent nets that use reinforcement learning to decide where to
-look.  Such systems already outperform passive vision
-systems~\citep{ba+mnih}.  Similarly, when understanding sentences or whole
-documents we expect future systems to learn strategies for selectively
-attending to one part at a time~\citep{Bahdanau-et-al-arxiv2014}. We also expect
-that in the next few years deep recurrent neural networks will become far
-better at understanding the meaning of documents, including the way in
-which the thoughts expressed by the sentences are related to one another.
+Human vision is an active process in which intelligent fixations are
+used to ensure that only a tiny fraction of the optic array is ever
+processed at high resolution. We expect much of the future progress in
+vision, especially for video understanding or robotics, to come from
+systems that are trained end-to-end and combine deep convolutional
+networks with recurrent nets that use reinforcement learning to decide
+where to look.  Such systems already outperform passive vision
+systems~\citep{ba+mnih}.  Similarly, when understanding sentences or
+whole documents we expect future systems to learn strategies for
+selectively attending to one part at a
+time~\citep{Bahdanau-et-al-arxiv2014}. We also expect that in the next
+few years deep recurrent neural networks will become far better at
+understanding the meaning of documents, including the way in which the
+thoughts expressed by the sentences are related to one another.
 
@@
 learning procedures for deep neural
-networks~\citep{Salakhutdinov2009-small,Hinton95,QuocLe-ICML2012,VincentPLarochelleH2008-small,ranzato-pami,Bengio-et-al-ICML-2014,Kingma-et-al-NIPS2014}. 
+networks~\citep{Salakhutdinov2009-small,Hinton95,QuocLe-ICML2012,VincentPLarochelleH2008-small,koray-nips-10,gregor-icml-10,ranzato-pami,Bengio-et-al-ICML-2014,Kingma-et-al-NIPS2014}.
 However, most data have not been hand-labeled by humans, and just like
-children are not told what they should have done at each instant, we need
-learning machines that can learn from mostly unlabeled examples.
-In the long run,
-we believe unsupervised learning will be essential for allowing
-deep networks to generalize well by creating internal representations that
-untangle the many different factors that interact to produce images, speech
-and documents~\citep{Bengio-Courville-Vincent-TPAMI2013}.
+children are not told what they should have done at each instant, we
+need learning machines that can learn from mostly unlabeled examples.
+In the long run, we believe unsupervised learning will be essential
+for allowing deep networks to generalize well by creating internal
+representations that untangle the many different factors that interact
+to produce images, speech and
+documents~\citep{Bengio-Courville-Vincent-TPAMI2013}.
+
+We expect to see more innovative ideas that marry deep learning with
+reasoning. Reasoning can be formulated in a number of different ways:
+energy minimization, contraint satisfaction, chains of operations
+applied to vectors representing objects or symbols. One is hoping for
+an architecture that unifies representation learning with
+reasoning~\citep{bottou-mlj-2014}.
 

### 2015-02-17 deepreview.tex
fixes

@@
 \addtolength{\evensidemargin}{-1cm}
+
+\newcommand{\deri}[2]{\frac{\partial  #1}{\partial #2}}
 
@@
 input $i$ were increased by a tiny amount $\Delta y_j =
-$\left(\deri{y}{x}\right)_{ij} \Delta x_i$. The above equation
+\left(\deri{y}{x}\right)_{ij} \Delta x_i$. The above equation
 constitutes a recurrence formula that can be used to propagate
@@
   \vspace{0.2cm}
-  % \includegraphics[width=0.99\linewidth]{convnet-diagram}
-  nice convnet figure goes here
+  \includegraphics[width=0.70\linewidth]{convnet-diagram}
 \end{center}

### 2015-02-17 deepreview.tex
typos

@@
 Deep Learning techniques allow a machine composed of multiple layers
-of processing to learn to represent data, so that each layer transform
+of processing to learn to represent data, so that each layer transforms
 the representation of the previous layer in a more abstract
@@
 data, such as images, video, speech and audio, while {\em recurrent
-  net} have shown excellent performance for analyzing sequential
+  nets} have shown excellent performance for analyzing sequential
 inputs and producing sequential outputs such as text and
-speech. Convolutional net has quickly become the leading for image
+speech. Convolutional nets have quickly become the leading technology for image
 recognition and speech recognition, while recurrent nets are
@@
 natural data in raw form. For decades, constructing a pattern
-recogntion or ML system has required engineering efforts and domain
+recognition or ML system has required engineering efforts and domain
 expertise to design a {\em feature extractor} that transforms the raw
@@
 of the edges.  The third layer would assemble motifs into parts of
-object, and subsequent layers would detect objects as combinations of
+object, and subsequent layers would to detect objects as combinations of
 parts. The key aspect of deep learning is that {\em these features and

### 2015-02-17 deepreview.tex
small corrections and integrating the two RNN sections

@@
 question-answering~\cite{bordes-14}, and language
-translation~\cite{Sutskever-et-al-NIPS2014}.
+translation~\cite{Jean-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
 
@@
 persons and pets, each labeled with its category. During training, the
-machine is shown an image and produces an output label in the form of
+machine is shown an image and produces an output in the form of
 a vector of scores, one for each category. We want the desired
@@
 The objective function, averaged over all the training examples, can
-be seen as a kind hilly landscape in the high-dimensional space of
+be seen as a kind of hilly landscape in the high-dimensional space of
 weight values. The negative gradient vector indicates the direction of
-steepest descent in this landscape, taking is closer to a minimum
+steepest descent in this landscape, taking it closer to a minimum
 where the output error is low on average.
@@
 close to a good minimum very quickly in
-practice~\cite{bottou-bousquet-2008-small} Following this, the
+practice~\cite{bottou-bousquet-2008-small}. Following this, the
 performance of the system is measured on a different set of examples,
@@
 the {\em generalization ability} of the machine, i.e. its ability to
-produce sensible answers on new inputs that is has never seen during
+produce sensible answers on new inputs that it has never seen during
 training.
@@
 learn to map a fixed-size input (e.g., an image) to a fixed-size
-output (e.g. a probability for each one of several categories). Each
-pair of layer can be seen as a set of units that compute a weighted
+output (e.g. a probability for each one of several categories). To go
+from one layer to the next, a set of units compute a weighted
 sum of their inputs from the previous layer and pass the result
@@
 GPU-optimized software for ConvNet made it possible to train these
-large networks in a few weeks (naw days) of computer time. Third, the
+large networks in a few weeks (now days) of computer time. Third, the
 use of rectifying non-linearities instead of sigmoids allowed to train
@@
 individual days remain different (for example one imagines that the
-neural network also learns to some features that distinguises week-end
+neural network also learns some features that distinguishes week-end
 days and weekdays, while Saturday and Sunday might have special roles
@@
 
-%% YLC: I moved this section down from the intro. It was redundant there.
-%% Some of it is probably redundant with much of what follows
+When backpropagation was first introduced, its most exciting use was
+for training recurrent neural networks. 
 For tasks that involve sequential inputs, such as speech and language,
@@
 ``state vector'' that implicitly contains information about the
-history all the past elements of the sequence.  Recurrent networks are
-very good at predicting the next character in
+history all the past elements of the sequence.  
+When we consider the outputs of
+the hidden units at different discrete time steps as if they were the
+outputs of different neurons in a deep multi-layer network (right of
+Figure~\ref{fig:rnn}), it becomes
+clear how we can apply backpropagation to train recurrent
+networks. The the gradient with respect to a recurrent weight,
+$w_{ij}$ that allows neuron $i$ at time $t$ to influence the activity
+of neuron $j$ at time $t+1$ can be computed by
+backpropagation-through-time (BPTT) of the error derivative for $j$ at
+time $t+1$. The full gradient with respect to $w_{ij}$ on a whole
+input sequence is just the sum of its gradients at each time-step.
+
+Recurrent neural networks are very powerful dynamical systems, but
+training them with BPTT proved to be problematic, because the
+backpropagated gradients either grow or shrink at each time step, so
+over many time-steps they typically explode or
+vanish~\citep{Hochreiter91-small,Bengio-et-al-TNN1994}.
+%% THE TEXT BELOW IS TOO COMPLICATED AND ALSO TOO SIMPLE: THERE ARE
+%% STABLE CONDITIONS IN WHICH THE GRADIENTS DO NOT VANISH (ORTHONORMAL
+%% RECURRENT MATRICES). I THINK ITS BETTER NOT TO GET INTO THIS.
+%% and the conditions which allow stable dynamics also correspond to
+%% vanishing gradients~\citep{Bengio-et-al-TNN1994},
+
+Thanks to advances in their architecture and ways of training them, recurrent networks 
+have been found to be very good at predicting the next character in
 text~\citep{Sutskever-et-al-ICML2011} or the next word in a
@@
 
-%% YLC: This is where the original recurrent net section begins.
-%% There should be some merging between the stuff above and the stuff below.
-
-When backpropagation was first introduced, its most exciting use was
-for training recurrent neural networks. These receive input vectors at
-each time-step and have one or more layers of hidden units that, in
-addition to having connections to higher layers or ouput units, also
-have connections that allow the neurons within a layer to receive
-weighted input from those same neurons at the previous time-step.  As
-illustrated in Figure~\ref{fig:rnn}, when we consider the outputs of
-the hidden units at different discrete time steps as if they were the
-outputs of different neurons in a deep multi-layer network, it becomes
-clear how we can apply backpropagation to train recurrent
-networks. The the gradient with respect to a recurrent weight,
-$w_{ij}$ that allows neuron $i$ at time $t$ to influence the activity
-of neuron $j$ at time $t+1$ can be computed by
-backpropagation-through-time (BPTT) of the error derivative for $j$ at
-time $t+1$. The full gradient with respect to $w_{ij}$ on a whole
-input sequence is just the sum of its gradients at each time-step.
-
-Recurrent neural networks are very powerful dynamical systems, but
-training them with BPTT proved to be problematic, because the
-backpropagated gradients either grow or shrink at each time step, so
-over many time-steps they typically explode or
-vanish~\citep{Hochreiter91-small,Bengio-et-al-TNN1994}.
-%% THE TEXT BELOW IS TOO COMPLICATED AND ALSO TOO SIMPLE: THERE ARE
-%% STABLE CONDITIONS IN WHICH THE GRADIENTS DO NOT VANISH (ORTHONORMAL
-%% RECURRENT MATRICES). I THINK ITS BETTER NOT TO GET INTO THIS.
-%% and the conditions which allow stable dynamics also correspond to
-%% vanishing gradients~\citep{Bengio-et-al-TNN1994},
-
-\begin{figure}[H]
+\begin{figure}[htp*]
 \fbox{

### 2015-02-17 deepreview.tex
added the attention examples for caption generation

@@
 \vspace{-0.6cm}
-\caption{A typical convolutional network architecture}
-\vspace{-0.8cm}
+\caption{The outputs of each layer of a typical convolutional network architecture
+applied on the image of a handwritten digit (on the top left). Each rectangular
+image corresponds to one of the features, detected at each of the image positions.
+Each column corresponds to a layer, with information flowing from the left to
+the right.
+}
+%\vspace{-0.8cm}
 \label{fig:convnet}
@@
 The architecture of a typical ConvNet is shown in
-figure~\ref{fig-convnet}.  It is structured as a series of stages. The
+figure~\ref{fig:convnet}.  It is structured as a series of stages. The
 first few stages are composed of two types of layers: convolutional
@@
 
-\begin{figure}[ht]
+\begin{figure}[htp*]
 \fbox{
@@
 \centerline{\includegraphics[width=\linewidth]{caption-generation-Vinyals-et-al.png}}
-\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}
-\includegraphics[width=0.49\linewidth]{Examples_im2txt_002.png}}
-\caption{Three examples of captions generated by a recurrent network taking
+%\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}}
+%\includegraphics[width=0.49\linewidth]{Examples_im2txt_002.png}}
+\centerline{\includegraphics[width=\linewidth]{good_capgen_examples.pdf}}
+%\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}
+\caption{Top: two examples of captions generated by a recurrent network taking
 as extra input the representation extracted by a deep convolutional
 net from a test image, with the recurrent network trained to ``translate''
-high-level representations of images into captions. GT stands for ``Ground Truth'' and was labeled by humans
-whereas the other text is machine-generated.
+high-level representations of images into captions. 
 Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
+Bottom: when the RNN is given the ability to focus its attention on a different
+location in the input image (lighter = more attention) as it generates each word (underlined), we
+found that it exploits this to achieve better ``translation'' of images into captions.
 }

### 2015-02-17 deepreview.tex
added the ref for caption generation with attention

@@
 location in the input image (lighter = more attention) as it generates each word (underlined), we
-found that it exploits this to achieve better ``translation'' of images into captions.
+found~\citep{Xu-et-al-arxiv2015} 
+that it exploits this to achieve better ``translation'' of images into captions.
 }

### 2015-02-17 deepreview.tex
cleaned up missing citations as best I could

@@
 particles~\citep{Melis-Higgs-boson-competition-2014},
-~\citep{YANNparticles}, classifying galaxies, ~\citep{YANNgalaxies},
+%~\citep{YANNparticles} % YB while waiting for bib entry, 
+% classifying galaxies, ~\citep{YANNgalaxies}, % YB while waiting for bib entry, 
 reconstructing brain circuits~\citep{helmstaedter-nature-2013}, and
@@
 understanding~\citep{collobert:2011b}, particularly topic
-classification~\citep{topic-classification}, sentiment
+classification, %~\citep{topic-classification},  % YB while waiting for bib entry
+sentiment
 analysis~\citep{Glorot+al-ICML-2011-small},
-question-answering~\cite{bordes-14}, and language
+question-answering~\cite{Bordes-et-al-EMNLP2014}, and language
 translation~\cite{Jean-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

### 2015-02-17 deepreview.tex
improved convnet caption

@@
 \vspace{-0.6cm}
-\caption{The outputs of each layer of a typical convolutional network architecture
+\caption{The outputs of each layer (column) and feature map (rectangular image)
+of a typical convolutional network architecture
 applied on the image of a handwritten digit (on the top left). Each rectangular
-image corresponds to one of the features, detected at each of the image positions.
+image is a feature map, corresponding to one of the learned features 
+detected at each of the image positions.
 Each column corresponds to a layer, with information flowing from the left to
-the right.
+the right. We see that lower level features mostly act as oriented edge detectors.
 }

### 2015-02-18 deepreview.tex
added the new Dahl journal paper on SAR

@@
 beaten other ML techniques at predicting the activity of potential
-drug molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental
+drug molecules~\citep{Dahl-et-al-arxiv2014,Ma-et-al-2015}, detecting new fundamental
 particles~\citep{Melis-Higgs-boson-competition-2014},

### 2015-02-23 deepreview.tex
added the Frey science paper and a putative YANNparticles ref

--- 2015-02-17 deepreview.tex
@@
 beaten other ML techniques at predicting the activity of potential
-drug molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental
-particles~\citep{Melis-Higgs-boson-competition-2014},
-%~\citep{YANNparticles} % YB while waiting for bib entry, 
+drug molecules~\citep{Dahl-et-al-arxiv2014,Ma-et-al-2015}, detecting new fundamental
+particles~\citep{Ciodaro-et-al-2012,Melis-Higgs-boson-competition-2014},
+% YB thinks that Ciodaro-et-al might be {YANNparticles} 
 % classifying galaxies, ~\citep{YANNgalaxies}, % YB while waiting for bib entry, 
@@
 predicting the effects of mutations in non-coding DNA on gene
-expression and disease~\citep{Keung-et-al-2014}.
+expression and disease~\citep{Keung-et-al-2014,Xiong-et-al-Frey-science2015}.
 

### 2015-03-24 deepreview.tex
removed the equations part for backprop from main text

@@
 {\em output} of that module (or the input of the subsequent
-module). Basically, take a module $M$ inside the network, whose input
-is $x$, weight vector is $w$, and output is $y = M(x,w)$. If one knows
-the gradient $\deri{E}{y}$ of the objective function $E$ with respect
-to the module's output $y$, one can blindly apply chain rule to
-compute the gradient with respect to the input:
-\[
-  \deri{E}{x} = \deri{y}{x}\deri{E}{y}
-\]
-This equation works even when $x$ and $y$ are vectors, in which case
-$\deri{E}{x}$ and $\deri{E}{y}$ are gradient vectors, and
-$\deri{y}{x}$ is a {\em Jacobian matrix} of $M$ where element $(i,j)$
-indicates by how much output $j$ would increase proportionally if
-input $i$ were increased by a tiny amount $\Delta y_j =
-\left(\deri{y}{x}\right)_{ij} \Delta x_i$. The above equation
+module), as illustrated in Figure~\ref{fig:backprop-box}.
+% Basically, take a module $M$ inside the network, whose input
+%is $x$, weight vector is $w$, and output is $y = M(x,w)$. If one knows
+%the gradient $\deri{E}{y}$ of the objective function $E$ with respect
+%to the module's output $y$, one can blindly apply chain rule to
+%compute the gradient with respect to the input:
+%\[
+%  \deri{E}{x} = \deri{y}{x}\deri{E}{y}
+%\]
+%This equation works even when $x$ and $y$ are vectors, in which case
+%$\deri{E}{x}$ and $\deri{E}{y}$ are gradient vectors, and
+%$\deri{y}{x}$ is a {\em Jacobian matrix} of $M$ where element $(i,j)$
+%indicates by how much output $j$ would increase proportionally if
+%input $i$ were increased by a tiny amount $\Delta y_j =
+%\left(\deri{y}{x}\right)_{ij} \Delta x_i$. 
+The backpropagation equation
 constitutes a recurrence formula that can be used to propagate
@@
 gradients are computed, one can similarly compute the gradients with
-respect to the weights of the module. Figure~\ref{fig:backprop-box})
-shows how this works out when the network is composed of layers of
-weight matrices interspersed with scalar non-linearities.
+respect to the weights of the module. 
+%Figure~\ref{fig:backprop-box})
+%shows how this works out when the network is composed of layers of
+%weight matrices interspersed with scalar non-linearities.
 %% the notation in the figure should be updated.

### 2015-03-24 deepreview.tex
modified figures and captions (1 and 2) following suggestions of referee 1

@@
 data. The name comes from the fact that deep learning machines are
-multilayer stacks of trainable modules, each of which turns the
+multilayer stacks of trainable modules, each of which (e.g. a neural
+network layer) turns the
 representation of the data in its input into a more abstract and
@@
 The equations used for computing the forward pass in a neural
-net with two hidden layers. At each layer, we first compute the total
+net with two hidden layers and one output layer, each constituting
+a module through which one can backpropagate gradients. At each layer, we first compute the total
 input $z$ to each unit, which is a weighted sum of the outputs of the
@@
   \vspace{0.2cm}
-  \includegraphics[width=0.70\linewidth]{convnet-diagram}
+  \includegraphics[width=0.70\linewidth]{convnet-diagram-labeled}
 \end{center}
 \vspace{-0.6cm}
-\caption{The outputs of each layer (column) and feature map (rectangular image)
+\caption{The {\em outputs} (not the filters) of each layer (each column of the figure) 
 of a typical convolutional network architecture
-applied on the image of a handwritten digit (on the top left). Each rectangular
-image is a feature map, corresponding to one of the learned features 
+applied on the image of a handwritten digit (left). Each rectangular
+image is a feature map, corresponding to the output for one of the learned features 
 detected at each of the image positions.

### 2015-03-24 deepreview.tex
reduced the preface to less than 100 words as requested

@@
 
-Deep Learning techniques allow a machine composed of multiple layers
-of processing to learn to represent data, so that each layer transforms
-the representation of the previous layer in a more abstract
-fashion. These methods have dramatically improved the state-of-the-art
+Deep Learning techniques allow a machine composed of multiple processing layers
+which define representations of the data that are more abstract for deeper layers.
+These methods have dramatically improved the state-of-the-art
 for speech recognition, object recognition, object detection,
@@
 learning discovers intricate structure in large datasets by using the
-{\em backpropagation algorithm} to compute how the machine should
-change in order to make less errors.  Deep {\em convolutional nets}
-have brought about dramatic improvements in the processing of array
-data, such as images, video, speech and audio, while {\em recurrent
-  nets} have shown excellent performance for analyzing sequential
-inputs and producing sequential outputs such as text and
-speech. Convolutional nets have quickly become the leading technology for image
-recognition and speech recognition, while recurrent nets are
-competitive in several natural language processing tasks, including
-language translation. Combinations of recurrent and convolutional nets
-can generate good textual captions of images.
+{\em backpropagation algorithm} to indicate how the machine should
+change.  Deep {\em convolutional nets}
+have brought about dramatic improvements to process
+images, video, speech and audio, while {\em recurrent
+  nets} have shone on sequential data such as text and
+speech.
+%Convolutional nets have quickly become the leading technology for image
+%recognition and speech recognition, while recurrent nets are
+%competitive in several natural language processing tasks, including
+%language translation. 
+%Combinations of recurrent and convolutional nets
+%can generate good textual captions of images.
 
@@
 the right. We see that lower level features mostly act as oriented edge detectors.
+The last layer is fully connected so it has no spatial structure.
 }

### 2015-03-25 deepreview.tex
implemented most of Bernhard's suggestions (ref 1)

@@
 % classical extensions of simple classifiers
+% YB: I propose to mostly remove the paragraph below (but keep the Scholkopf book ref)
 There have been numerous attempts, going back to the early 1960s, to
@@
 separate categories from one another by more complex surfaces than
-flat hyperplanes. A popular example is the class of kernel methods,
+flat hyperplanes. A popular example is the class of kernel methods~\citep{Scholkopf+Smola-2002book}, % added ref for BS
 particularly the Support Vector Machine~\citep{Boser92} which
@@
 a surprisingly long time to emerge. It was discovered independently by
-everal different groups in the 1970s and
+several different groups in the 1970s and
 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}, and became quite popular
@@
 If ConvNets existed in the 1990s, why did they become so popular and
-dominant only now? Certainly, the prejudice of the ML and vision
-communities against neural nets played a role. But a number of recent
-ideas and technologies enabled the success. First, ConvNets strive
+dominant only now? The idea of convexity of the objective function grew
+as an attractive property of learning algorithms, thanks to the success
+of kernel machines, which made neural nets (that involve a non-convex
+training criterion) less appealing. As discussed below, it is only
+in 2014 that theoretical and empirical evidence have suggested that local
+minima are probably not as much of a problem as was initially thought~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
+Much before that realization, though, a number of recent
+ideas and technologies enabled the success. First, supervised ConvNets work best
 when training data is abundant, and large labeled image datasets such
@@
 presented to the network as a one-of-N vector, i.e., in which one
-component has a value of $1$ and the rest are $0$. The output layer
-has one unit per word and is forced to produce a probability
-distribution over words by using the ``softmax'' function:
-\begin{equation}
-p_j = \frac{e^{z_j}}{\sum_i e^{z_i}}
-\end{equation}
-where $z_j$ is the total input received by output unit $j$. 
+component has a value of $1$ and the rest are $0$. 
+%The output layer
+%has one unit per word and is forced to produce a probability
+%distribution over words by using the ``softmax'' function:
+%\begin{equation}
+%p_j = \frac{e^{z_j}}{\sum_i e^{z_i}}
+%\end{equation}
+%where $z_j$ is the total input received by output unit $j$. 
 \iffalse
@@
 \fi
-
-The presence of a particular word at the $i$-th position in the input
-of the network sets the value of specialized hidden units to values
-that correspond to the representation of that word. Different words
-create a different pattern of activations, or word-vector, illustrated
+%
+%The presence of a particular word at the $i$-th position in the input
+%of the network sets the value of specialized hidden units to values
+%that correspond to the representation of that word. 
+In the first layer, each word
+creates a different pattern of activations, or word-vector, illustrated
 in Figure~\ref{fig:word-embeddings}.  In a language model, the other
@@
 active components each of which can be interpreted as a separate
-feature of the word. This was first demonstrated using sequences of
-three words, such as (Colin, Mother, Victoria), derived from two
-family trees~\citep{RHW}. The learned word vector for Colin contained
-``semantic'' features representing which family tree Colin was in,
-what branch of that tree he was in, and what generation he was from.
+feature of the word, as was first demonstrated by~\citet{RHW} in the context
+of learning distributed representations for symbols.
+%. This was first demonstrated using sequences of
+%three words, such as (Colin, Mother, Victoria), derived from two
+%family trees~\citep{RHW}. The learned word vector for Colin contained
+%``semantic'' features representing which family tree Colin was in,
+%what branch of that tree he was in, and what generation he was from.
 %% TWO NATIVE ENGLISH SPEAKERS AGREE THAT THE "FROM" SHOULD STAY AT THE
@@
 of factorizing the structured relationships between the input and output
-symbols into multiple ``micro-rules''.  For example, given Colin's
-generation and the learned feature of Mother that the output should be one
-generation higher than the input, we can already restrict the answer to
-people in a specific generation. The ability of deep learning to discover
-many such micro-rules allows it to generalize well to novel combinations of
-individually familiar context words.
-
-The family tree demonstration used a small, clean dataset in which the
-regularities could be captured by micro-rules that were never violated
-and the features were easy to interpret, but learned word-vectors also
+symbols into multiple ``micro-rules''.  
+%For example, given Colin's
+%generation and the learned feature of Mother that the output should be one
+%generation higher than the input, we can already restrict the answer to
+%people in a specific generation. The ability of deep learning to discover
+%many such micro-rules allows it to generalize well to novel combinations of
+%individually familiar context words.
+%
+%The family tree demonstration used a small, clean dataset in which the
+%regularities could be captured by micro-rules that were never violated
+%and the features were easy to interpret, 
+Learning word-vectors turned out to also
 work very well when the word sequences come from a large corpus of
@@
 ``state vector'' that implicitly contains information about the
-history all the past elements of the sequence.  
+history of all the past elements of the sequence.  
 When we consider the outputs of
@@
 clear how we can apply backpropagation to train recurrent
-networks. The the gradient with respect to a recurrent weight,
+networks. The gradient with respect to a recurrent weight,
 $w_{ij}$ that allows neuron $i$ at time $t$ to influence the activity
@@
 
-Thanks to advances in their architecture and ways of training them, recurrent networks 
+Thanks to advances in their architecture~\citep{Hochreiter+Schmidhuber-1997,ElHihi+Bengio-nips8-small} 
+and ways of training them~\citep{Sutskever-thesis2012}, recurrent networks 
 have been found to be very good at predicting the next character in
@@
 Many other architectures are possible, including a variant in which
-the sampled outputs (e.g. words) can be generated according to an output 
-distribution (the probability of the next word, for every possible next word) 
-and used as inputs for the next time step.
+the network can generate a sequence of outputs (e.g. words), each of which
+is used as inputs for the next time step.
 The back-propagation algorithm (Figure~\ref{fig:backprop-box})
@@
 similar to a text adventure game, the network is shown an
-automatically0generate question and taught to produce the
+automatically generate question and taught to produce the
 corresponding answer (also generated automatically).  In one test
@@
 
-In a somewhat related work, a network is trained to evaluate whether a
-particular mathematical transformation of a formula is lilely to lead
-to a correct mathematical identity~\cite{zaremba-nips-2014}. The
-system was able to produce efficient identities for matrix expressions
-whose formula fill several pages.
+% Bernhard was not impressed, so maybe just cut this off
+%In a somewhat related work, a network is trained to evaluate whether a
+%particular mathematical transformation of a formula is lilely to lead
+%to a correct mathematical identity~\citep{zaremba-nips-2014}. The
+%system was able to produce efficient identities for matrix expressions
+%whose formula fill several pages.
 
@@
 map from the pixels on the screen at several time steps to the
-expected future reward of each possible action and its only inputs are the
+expected future reward of each possible action (assuming some policy is followed
+in the future) and its only inputs are the
 pixels and the current score in the game. The learning works by minimizing
@@
 It also took an inordinate amount of time for the community to adopt
-stochastic gradient methods for large-scale machine learning. key
+stochastic gradient methods for large-scale machine learning. Key
 theoretical and practical work showed that, while SGD has terrible
@@
 kind of approximate optimization required in ML
-applications~\citep{bottou-bousquet-2008-small}. But when trainnig a
+applications~\citep{bottou-bousquet-2008-small}. But when training a
 large ConvNet can take days or weeks on the fastest GPUs, one is eager
@@
 unsolved question is how to scale up the training of very large neural
-nets over multiple GPU card across multiple machines. As of today, this
+nets over multiple GPU cards across multiple machines. As of today, this
 question is unresolved.
@@
 2006~\citep{IJCAI,Hinton06-small,Bengio-nips-2006-small,ranzato-07-small}
-by the introduction of unsupervised learning procedures that could
+thanks to the effort of a forward-looking Canadian funding agency 
+(the Canadian Institute for Advanced Research, or CIFAR), with
+the introduction of unsupervised learning procedures that could
 create layers of feature detectors without requiring labeled data. The

### 2015-03-25 deepreview.tex
discussion about convexity

@@
 
-If ConvNets existed in the 1990s, why did they become so popular and
-dominant only now? The idea of convexity of the objective function grew
-as an attractive property of learning algorithms, thanks to the success
-of kernel machines, which made neural nets (that involve a non-convex
-training criterion) less appealing. As discussed below, it is only
+If ConvNets and other neural nets existed in the 1990s, why did they become so popular and
+dominant only now? It was initially thought that training of neural nets
+would get stuck in poor solutions (called local minima) due to non-convexity
+of the training objective, whereas the alternative learning framework
+of {\em kernel machines}~\citep{Scholkopf+Smola-2002book} was more appealing
+because it involved a convex optimization. As discussed below, it is only
 in 2014 that theoretical and empirical evidence have suggested that local
-minima are probably not as much of a problem as was initially thought~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
-Much before that realization, though, a number of recent
-ideas and technologies enabled the success. First, supervised ConvNets work best
+minima are probably not as much of a problem as was initially 
+thought~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
+Much before that realization, though, a number of 
+ideas and technologies over the last decade enabled the success. 
+First, supervised ConvNets work best
 when training data is abundant, and large labeled image datasets such

### 2015-03-26 deepreview.tex
finishing ref 1 fixes (VC paper) and removing kernel paragraph

@@
 regression~\cite{Hastie2001}, and the linear support vector
-machine~\cite{Boser92}.
+machine~\cite{Vapnik+Chervonenkis-1964,Boser92}.
 
@@
 irrelevant aspects, like the pose of the animal.
-
+%
 % classical extensions of simple classifiers
 % YB: I propose to mostly remove the paragraph below (but keep the Scholkopf book ref)
-There have been numerous attempts, going back to the early 1960s, to
-increase the flexibility of classifiers so as to allow them to
-separate categories from one another by more complex surfaces than
-flat hyperplanes. A popular example is the class of kernel methods~\citep{Scholkopf+Smola-2002book}, % added ref for BS
-particularly the Support Vector Machine~\citep{Boser92} which
-represents inputs by a vector whose components are similarities to
-each of the training samples computed through a so-called {\em kernel
-  function}. While these methods are good general-purpose classifiers,
-they do not perform well for tasks such as image or speech recognition
-in which the input space is very high-dimensional and the input-output
-mapping is so complex that computing similarities using a simple
-kernel function is ineffective. Designing feature extractors that are
+% There have been numerous attempts, going back to the early 1960s, to
+% increase the flexibility of classifiers so as to allow them to
+% separate categories from one another by more complex surfaces than
+% flat hyperplanes. A popular example is the class of kernel methods~\citep{Scholkopf+Smola-2002book}, % added ref for BS
+%particularly the Support Vector Machine~\citep{Boser92} which
+% represents inputs by a vector whose components are similarities to
+% each of the training samples computed through a so-called {\em kernel
+%   function}. While these methods are good general-purpose classifiers,
+% they do not perform well for tasks such as image or speech recognition
+% in which the input space is very high-dimensional and the input-output
+% mapping is so complex that computing similarities using a simple
+% kernel function is ineffective. 
+Designing feature extractors that are
 sensitive to the relevant information and invariant to the irrelevant

### 2015-03-27 deepreview.tex
working on ref 2 comments

@@
 of the world is difficult when the rewards are delayed and the world has a
-very large number of possible states. One impressive recent development is
-a single system that can learn to play many different video
-games~\citep{Deepmind-atari-arxiv2013}. The system uses a deep convolutional network to
-map from the pixels on the screen at several time steps to the
+very large number of possible states. After neural networks won backgammon
+olympiads in the 90's~\citep{Tesauro90}, an impressive recent development is
+combines deep convolutional networks with reinforcement learning in a single system that can learn to play many different video
+games~\citep{Mnih-et-al-2015}. 
+Pixels on the screen at several time steps are mapped to the
 expected future reward of each possible action (assuming some policy is followed
@@
 networks~\citep{Salakhutdinov2009-small,Hinton95,QuocLe-ICML2012,VincentPLarochelleH2008-small,koray-nips-10,gregor-icml-10,ranzato-pami,Bengio-et-al-ICML-2014,Kingma-et-al-NIPS2014}.
-However, most data have not been hand-labeled by humans, and just like
+These papers often introduce a probabilistic view or the injection of randomness in the
+neural network computation, to allow neural networks to {\em generate} examples that are
+similar to training examples but generalize them.
+Unsupervised learning is important because most data have not been hand-labeled by humans, and just like
 children are not told what they should have done at each instant, we

### 2015-03-27 deepreview.tex
done almost all points of ref 2, now moving to ref 3

@@
 Thanks to advances in their architecture~\citep{Hochreiter+Schmidhuber-1997,ElHihi+Bengio-nips8-small} 
-and ways of training them~\citep{Sutskever-thesis2012}, recurrent networks 
+and ways of training them~\citep{Sutskever-thesis2012,Pascanu+al-ICML2013-small}, recurrent networks 
 have been found to be very good at predicting the next character in

### 2015-03-28 deepreview.tex
done with most suggestions of ref 3, now moving to space cutting
removed the RL section, merged a few other sections, and cut here and there

--- 2015-03-27 deepreview.tex
@@
 \usepackage{graphicx}
-\usepackage{natbib}
+\usepackage[numbers]{natbib}
 \usepackage{hyperref}
@@
 
-\begin{figure}
-\begin{center}
-  \vspace{0.2cm}
-  \includegraphics[width=0.70\linewidth]{convnet-diagram-labeled}
-\end{center}
-\vspace{-0.6cm}
-\caption{The {\em outputs} (not the filters) of each layer (each column of the figure) 
-of a typical convolutional network architecture
-applied on the image of a handwritten digit (left). Each rectangular
-image is a feature map, corresponding to the output for one of the learned features 
-detected at each of the image positions.
-Each column corresponds to a layer, with information flowing from the left to
-the right. We see that lower level features mostly act as oriented edge detectors.
-The last layer is fully connected so it has no spatial structure.
-}
-%\vspace{-0.8cm}
-\label{fig:convnet}
-\end{figure}
-
 One of the most widely used deep learning architectures is the
@@
 the next level to vary in relative position and appearance.
+
+\begin{figure}[htp*]
+\vspace{-4mm}
+\begin{center}
+  \includegraphics[width=0.70\linewidth]{convnet-diagram-labeled}
+\end{center}
+\vspace{-3mm}
+\caption{The {\em outputs} (not the filters) of each layer (each column of the figure) 
+of a typical convolutional network architecture
+applied on the image of a handwritten digit (left). Each rectangular
+image is a feature map, corresponding to the output for one of the learned features 
+detected at each of the image positions.
+Each column corresponds to a layer, with information flowing from the left to
+the right. We see that lower level features mostly act as oriented edge detectors.
+The last layer is fully connected so it has no spatial structure.
+}
+\vspace{-2mm}
+\label{fig:convnet}
+\end{figure}
 
@@
 
-\section{Distributed representations}
+\section{Distributed representations and language processing}
 
@@
 descent for updating the weights on connections has proved to be very
-good at learning appropriate distributed representations. These
-representations allow a lot of parallel computation at the feature
-level so that complex tasks can be solved in a very small number of
-sequential steps. In fact, it can be shown that a measure of the
-richness of the functions that can be represented by a deep neural
-network grows exponentially with
-depth~\citep{Montufar-et-al-NIPS2014}, whereas according to the same
-measure, classical (shallow) non-parametric statistical methods such
-as Gaussian Support Vector
-Machines~\citep{Bengio-localfailure-NIPS-2006-small} or decision
-trees~\citep{Bengio-decision-trees10} would require exponential
-storage
-%% an exponentially large number of training examples 
-%% THIS WOULD BE VERY CONFUSING FOR MOST READERS WHO WILL NOT KNOW
-%% ABOUT NON-PARAMETERIC METHODS.
-to represent the same kinds of functions.
-
-\section{Language processing with deep networks}
+good at learning appropriate distributed representations. 
+%These
+%representations allow a lot of parallel computation at the feature
+%level so that complex tasks can be solved in a very small number of
+%sequential steps.
+Deep learning theory shows that deep nets can benefit from
+an exponential advantage over classical learning algorithm not based on distributed 
+representations~\citep{Bengio-localfailure-NIPS-2006-small,Bengio-decision-trees10},
+ in two ways, both arising from the power of {\em composition}, so long as the
+underlying data generating distribution has such a componential structure, i.e.,
+is generated by a hierarchy of composition of factors~\citep{Bengio-Courville-Vincent-TPAMI2013}.
+First, learning distributed representations enables generalization to new
+combinations of the learned features (e.g. $2^n$ combinations, with $n$ features)
+~\citep{Bengio-2009-book,Montufar+Morton-2014}. 
+Second composing layers in a deep net brings the potential for another exponential
+advantage~\citep{Montufar-et-al-NIPS2014} (exponential in the depth).
 
@@
 models~\citep{BenDucVin01-short}, the standard approach to statistical
-modeling of language was based on counting frequencies of occurrences of
-short symbol sequences (called N-grams). One disadvantage of the n-gram
-approach is that the number of possible N-grams is $V^N$ where $V$ is
+modeling of language did not exploit distributed representations: it 
+was based on counting frequencies of occurrences of
+short symbol sequences (called N-grams). 
+The number of possible N-grams is $V^N$ where $V$ is
 the vocabulary size, so taking
 into account a context of more than a handful of words would require
-very large training corpora. Although this can
-be mitigated by only considering the more frequent of the longer sequences of 
-words, it means that less frequent sequences are modeled in a 
-context-blind way. N-grams
-treat each word as an atomic unit, so they cannot generalize across
-semantically related sequences of words. By contrast,
+very large training corpora. 
+N-grams treat each word as an atomic unit, so they cannot generalize across
+semantically related sequences of words, whereas
 neural language models can generalize across semantically related
@@
 words end up close to each other in that vector space, as illustrated in
-Figure~\ref{fig:word-embeddings}. The distributed representations of
-words are obtained by using backpropagation to jointly learn 
-a representation for each word and
-a function that predicts a target quantity such as the next word
-in a sequence (for language modeling) or a whole sequence
-of translated words (for machine translation)~\citep{Devlin-et-al-ACL2014,Bahdanau-et-al-arxiv2014,
-Sutskever-et-al-NIPS2014}.
+Figure~\ref{fig:word-embeddings}.
 
@@
 of phrases learned by an English-to-French encoder-decoder recurrent
-neural network~\citep{Cho-et-al-EMNLP2014}. One can observe that
+neural network or RNN~\citep{Cho-et-al-EMNLP2014}. One can observe that
 semantically similar words or sequences of words are  mapped to
-nearby representations.
+nearby representations. The distributed representations of
+words are obtained by using backpropagation to jointly learn 
+a representation for each word and
+a function that predicts a target quantity such as the next word
+in a sequence (for language modeling) or a whole sequence
+of translated words (for machine translation)~\citep{Devlin-et-al-ACL2014,Bahdanau-et-al-arxiv2014,
+Sutskever-et-al-NIPS2014}.
 }
@@
 
-Instead of translating the meaning of a French sentence to an English sentence,
-one can learn to ``translate'' the meaning of an image into an English sentence,
-as illustrated in Figure~\ref{fig:caption-generation}. The ``encoder'' here is
-a deep convolutional neural network which converts the pixels into an
-activity vector in its last hidden layer. The ``decoder''
-is a recurrent neural network similar to the ones used for machine translation
-and neural language modeling. There has been a surge of interest in such
-systems recently~\citep{Kiros-et-al-ICML2014,Mao+al-arxiv2014,Donahue-et-al-arxiv2014,
-              Vinyals-et-al-arxiv2014,Fang-et-al-arxiv2014,
-             Chen+Zitnick-arxiv2014,Karpathy+Li-arxiv2014,Venugopalan-et-al-arxiv2014}.
-
-\begin{figure}[htp*]
-\fbox{
-\begin{minipage}{\textwidth}
-\centerline{\includegraphics[width=\linewidth]{caption-generation-Vinyals-et-al.png}}
-%\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}}
-%\includegraphics[width=0.49\linewidth]{Examples_im2txt_002.png}}
-\centerline{\includegraphics[width=\linewidth]{good_capgen_examples.pdf}}
-%\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}
-\caption{Top: two examples of captions generated by a recurrent network taking
-as extra input the representation extracted by a deep convolutional
-net from a test image, with the recurrent network trained to ``translate''
-high-level representations of images into captions. 
-Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
-Bottom: when the RNN is given the ability to focus its attention on a different
-location in the input image (lighter = more attention) as it generates each word (underlined), we
-found~\citep{Xu-et-al-arxiv2015} 
-that it exploits this to achieve better ``translation'' of images into captions.
-}
-\label{fig:caption-generation}
-\end{minipage}
-}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%
@@
 When backpropagation was first introduced, its most exciting use was
-for training recurrent neural networks. 
+for training recurrent neural networks (RNNs). 
 For tasks that involve sequential inputs, such as speech and language,
-it is often better to use recurrent neural networks, illustrated in
-Figure~\ref{fig:rnn}.  Recurrent neural networks process an input
+it is often better to use RNNs, illustrated in
+Figure~\ref{fig:rnn}.  RNNs process an input
 sequence one element at a time, maintaining in their hidden units a
@@
 Figure~\ref{fig:rnn}), it becomes
-clear how we can apply backpropagation to train recurrent
-networks. The gradient with respect to a recurrent weight,
-$w_{ij}$ that allows neuron $i$ at time $t$ to influence the activity
-of neuron $j$ at time $t+1$ can be computed by
-backpropagation-through-time (BPTT) of the error derivative for $j$ at
-time $t+1$. The full gradient with respect to $w_{ij}$ on a whole
-input sequence is just the sum of its gradients at each time-step.
-
-Recurrent neural networks are very powerful dynamical systems, but
-training them with BPTT proved to be problematic, because the
+clear how we can apply backpropagation to train RNNs.
+% The gradient with respect to a recurrent weight,
+%$w_{ij}$ that allows neuron $i$ at time $t$ to influence the activity
+%of neuron $j$ at time $t+1$ can be computed by
+%backpropagation-through-time (BPTT) of the error derivative for $j$ at
+%time $t+1$. The full gradient with respect to $w_{ij}$ on a whole
+%input sequence is just the sum of its gradients at each time-step.
+
+RNNs are very powerful dynamical systems, but
+training them proved to be problematic, because the
 backpropagated gradients either grow or shrink at each time step, so
@@
 Thanks to advances in their architecture~\citep{Hochreiter+Schmidhuber-1997,ElHihi+Bengio-nips8-small} 
-and ways of training them~\citep{Sutskever-thesis2012}, recurrent networks 
+and ways of training them~\citep{Sutskever-thesis2012,Pascanu+al-ICML2013-small}, RNNs
 have been found to be very good at predicting the next character in
@@
 technique that people can eventually master, but the translation
-results with recurrent neural networks suggest that it may not be a good model of
-the way we normally reason or understand language. Recurrent neural
-networks that represent thoughts as big vectors are more compatible with
+results with RNNs suggest that it may not be a good model of
+the way we normally reason or understand language. RNNs
+that represent thoughts as big vectors are more compatible with
 the view that everyday reasoning involves many simultaneous analogies that
@@
 conclusion~\citep{Lakoff+Johnson-2008,Rogers+McClelland-book2004}.
+
+Instead of translating the meaning of a French sentence to an English sentence,
+one can learn to ``translate'' the meaning of an image into an English sentence,
+as illustrated in Figure~\ref{fig:caption-generation}. The ``encoder'' here is
+a deep convolutional neural network which converts the pixels into an
+activity vector in its last hidden layer. The ``decoder''
+is an RNN similar to the ones used for machine translation
+and neural language modeling. There has been a surge of interest in such
+systems recently~\citep{Kiros-et-al-ICML2014,Mao+al-arxiv2014,Donahue-et-al-arxiv2014,
+              Vinyals-et-al-arxiv2014,Fang-et-al-arxiv2014,
+             Chen+Zitnick-arxiv2014,Karpathy+Li-arxiv2014,Venugopalan-et-al-arxiv2014}.
+
+\begin{figure}[htp*]
+\fbox{
+\begin{minipage}{\textwidth}
+\centerline{\includegraphics[width=\linewidth]{caption-generation-Vinyals-et-al.png}}
+%\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}}
+%\includegraphics[width=0.49\linewidth]{Examples_im2txt_002.png}}
+\centerline{\includegraphics[width=\linewidth]{good_capgen_examples.pdf}}
+%\centerline{\includegraphics[width=0.49\linewidth]{Examples_im2txt_001.png}
+\caption{Top: two examples of captions generated by an RNN taking
+as extra input the representation extracted by a deep convolutional
+net from a test image, with the RNN trained to ``translate''
+high-level representations of images into captions. 
+Reproduced with permission from~\citet{Vinyals-et-al-arxiv2014}.
+Bottom: when the RNN is given the ability to focus its attention on a different
+location in the input image (lighter = more attention) as it generates each word (underlined), we
+found~\citep{Xu-et-al-arxiv2015} 
+that it exploits this to achieve better ``translation'' of images into captions.
+}
+\label{fig:caption-generation}
+\end{minipage}
+}
+\end{figure}
 
@@
 %to every node, i.e., the units value and the parameters).\newline
-A recurrent neural network (left) and the unfolding in time of the
-computation involved in its forward computation. The artificial neurons
+A recurrent neural network (RNN) and the unfolding in time of the
+computation involved in its forward computation (right). The artificial neurons
 (e.g. hidden units grouped under node $s$ with values $s_t$ at time $t$) 
@@
 (this is represented with the black square, standing for a delay of
-one time step, on the left). In this way, a recurrent net can map an input sequence
+one time step, on the left). In this way, an RNN can map an input sequence
 with elements $x_t$ into an output sequence with elements $o_t$, with
@@
 
-\section{Recurrent Networks with Memory}
+%\section{Recurrent Networks with Memory}
 %% YLC: I moved this up because it connects with the RNN section above.
 
-Recurrent networks, once unfolded in time, can be seen as very deep
+RNNs, once unfolded in time (Figure~\ref{fig:rnn}), can be seen as very deep
 feed-forward networks in which all the layers share the same weights.
@@
 natural behaviour is to remember inputs for a long
-time~\citep{Hochreiter+Schmidhuber-1997}.  A linear ``memory cell''
-interacts with other units via gated connections whose weights are
-multiplied by the state of a logistic gating unit that takes a value
-between $0$ and $1$. An input gating unit can learn when to isolate
-the cell from the effect of the outside world, and an output gating
-unit prevents the content of the memory cell to be broadcast to the
-rest of the neural network. More importantly, the memory cell itself
+time~\citep{Hochreiter+Schmidhuber-1997}.  
+%A linear ``memory cell''
+%interacts with other units via gated connections whose weights are
+%multiplied by the state of a logistic gating unit that takes a value
+%between $0$ and $1$. An input gating unit can learn when to isolate
+%the cell from the effect of the outside world, and an output gating
+%unit prevents the content of the memory cell to be broadcast to the
+%rest of the neural network. More importantly, 
+A special unit called the memory cell
 acts like an accumulator or a gated leaky neuron: it has a connection
@@
 its own real-valued state and accumulates the external signal, but
-this self-connection is {\em gated} by a ``forget gate'' so the stored
-value quickly leaks away unless the forget gate has an activation near
-$1$.  The gates themselves have their activities determined by
-weighted inputs that they receive from the input vectors and all of
-the memory cells.
-
-When LSTMs were first described \citep{Hochreiter+Schmidhuber-1997},
-their complexity may have been off-putting and, despite their
-impressive ability to learn very long-range temporal structure on toy
-problems, they were largely ignored until they were shown to yield
-excellent results in a cursive handwriting recognition
-competition~\citep{Graves-et-al-2009}. They have subsequently proved
+this self-connection is {\em multiplicatively gated} by another unit
+that learns to decide when to clear the content of the memory.
+
+%When LSTMs were first described \citep{Hochreiter+Schmidhuber-1997},
+%their complexity may have been off-putting and, despite their
+%impressive ability to learn very long-range temporal structure on toy
+%problems, they were largely ignored until they were shown to yield
+%excellent results in a cursive handwriting recognition
+%competition~\citep{Graves-et-al-2009}. 
+LSTM have subsequently proved
 to be very effective for tasks that have sequential input.  When they
@@
 Over the last year, several authors have made different proposals to
-augment recurrent nets with a memory module. This includes simplified
+augment RNNs with a memory module. This includes simplified
 version of the LSTM
@@
 are required.
-
 Other proposals include the ``Neural Turing Machine'' in which the
-network is augmented by a ``tape-like'' memory that the recurrent net
+network is augmented by a ``tape-like'' memory that the RNN
 can chose to read from or write to~\citep{Graves-et-al-arxiv2014}, and
@@
 
-The renewal of interest in deep learning and recurrent neural nets is
-spurring a whole new generation of creative ideas.
-
+% The renewal of interest in deep learning and recurrent neural nets is
+%spurring a whole new generation of creative ideas.
+
+\iffalse  
 \section{Deep reinforcement learning}
@@
 the currently legal actions.
+\fi
 
@@
 systems that are trained end-to-end and combine deep convolutional
-networks with recurrent nets that use reinforcement learning to decide
-where to look.  Such systems already outperform passive vision
+networks with RNNs that use reinforcement learning to decide
+where to look.  Combining reinforcement learning with deep learning
+has already given impressive results~\citep{Tesauro90,Mnih-et-al-2015}
+but much more remains to be explored.
+Such systems already outperform passive vision
 systems~\citep{ba+mnih}.  Similarly, when understanding sentences or
@@
 time~\citep{Bahdanau-et-al-arxiv2014}. We also expect that in the next
-few years deep recurrent neural networks will become far better at
+few years deep RNNs will become far better at
 understanding the meaning of documents, including the way in which the

### 2015-03-28 deepreview.tex
put the figures at the end to see clearly the page count for text
commenting out iffalse regions to make detex work properly

@@
 
-\begin{figure}[htp]
+%\begin{figure}[htp]
+\begin{figure}[b]
 \fbox{
@@
 
-\begin{figure}[htp*]
+%\begin{figure}[htp*]
+\begin{figure}[b]
 \vspace{-4mm}
@@
 %where $z_j$ is the total input received by output unit $j$. 
-\iffalse
-The hidden
-layers use a non-linearity such as the logistic, tanh, or rectifier, respectively:
-\begin{equation}
-y_j = \frac{1}{1+ e^{z_j}}, \ \ \ y_j = \frac{e^{z_j}-e^{-z_j}}{e^{z_j}+e^{-z_j}}, \ \ \
- y_j = \max(0, z_j).
-\end{equation}
-\fi
+%\iffalse
+%The hidden
+%layers use a non-linearity such as the logistic, tanh, or rectifier, respectively:
+%\begin{equation}
+%y_j = \frac{1}{1+ e^{z_j}}, \ \ \ y_j = \frac{e^{z_j}-e^{-z_j}}{e^{z_j}+e^{-z_j}}, \ \ \
+% y_j = \max(0, z_j).
+%\end{equation}
+%\fi
 %
@@
 
-\begin{figure}[ht]
+%\begin{figure}[ht]
+\begin{figure}[b]
 \fbox{
@@
 
-\begin{figure}[htp*]
+%\begin{figure}[htp*]
+\begin{figure}[b]
 \fbox{
@@
 
-\section{Learning to reason and to manipulate symbols}
+% \section{Learning to reason and to manipulate symbols}
 
@@
 symbols and to reason.
-
 Neural Turing Machines can be taught ``algorithms''. Among other
@@
 input consists of an unsorted sequence in which each symbol is
-accompanied by a real value that indicates its priority in the list.
-To achieve this it must learn a sorting program purely from examples
-of the input and output sequences~\citep{Graves-et-al-arxiv2014}.
-
+accompanied by a real value that indicates its priority in the list
+~\citep{Graves-et-al-arxiv2014}.
+% To achieve this it must learn a sorting program purely from examples
+% of the input and output sequences
 Memory Networks can be trained to keep track of the state of the world
 when they are told a story, so that they can answer questions about
-the state of the world at the end of the story. Following a sequence
-of actions in a virtual world described in words, in a scenario
-similar to a text adventure game, the network is shown an
-automatically generate question and taught to produce the
-corresponding answer (also generated automatically).  In one test
+the state of the world at the end of the story. 
+% Following a sequence
+%of actions in a virtual world described in words, in a scenario
+%similar to a text adventure game, the network is shown an
+%automatically generate question and taught to produce the
+%corresponding answer (also generated automatically).  
+In one test
 example, the network is shown a 15-sentence version of the Lord of The
@@
 
-A different system that also uses BPTT can learn to map the sequence
+Another RNN can learn to map the sequence
 of characters in a simple program written in Python to the sequence of
 characters that the program would
-print~\citep{Zaremba+Sutskever-arxiv2014}.  Rather surprisingly, BPTT
-can learn about {\it for loops} and can also learn to perform addition
-and multiplication of strings containing several decimal
-digits. Achieving this kind of behaviour seemed ludicrously optimistic
-when BPTT was first proposed.
+print~\citep{Zaremba+Sutskever-arxiv2014}, 
+learning about {\it for loops} and small integer arithmetic.
+Achieving this kind of behaviour seemed ludicrously optimistic
+in the 1990's.
 
@@
 
-\iffalse  
-\section{Deep reinforcement learning}
-
-Learning which action will maximize the total future reward given a state
-of the world is difficult when the rewards are delayed and the world has a
-very large number of possible states. After neural networks won backgammon
-olympiads in the 90's~\citep{Tesauro90}, an impressive recent development is
-combines deep convolutional networks with reinforcement learning in a single system that can learn to play many different video
-games~\citep{Mnih-et-al-2015}. 
-Pixels on the screen at several time steps are mapped to the
-expected future reward of each possible action (assuming some policy is followed
-in the future) and its only inputs are the
-pixels and the current score in the game. The learning works by minimizing
-the difference between the system's current estimate of future reward and a
-noisy but less biased estimate that is obtained by using the same deep
-network to estimate the future reward of the next state, where the next
-state is chosen stochastically based on the predicted future rewards of all
-the currently legal actions.
-\fi
-
-\section{Overcoming the limitations of backpropagation}
+%\iffalse  
+%\section{Deep reinforcement learning}
+%
+%Learning which action will maximize the total future reward given a state
+%of the world is difficult when the rewards are delayed and the world has a
+%% very large number of possible states. After neural networks won backgammon
+%% olympiads in the 90's~\citep{Tesauro90}, an impressive recent development is
+%% combines deep convolutional networks with reinforcement learning in a single system that can learn to play many different video
+%% games~\citep{Mnih-et-al-2015}. 
+%% Pixels on the screen at several time steps are mapped to the
+%% expected future reward of each possible action (assuming some policy is followed
+%% in the future) and its only inputs are the
+%% pixels and the current score in the game. The learning works by minimizing
+%% the difference between the system's current estimate of future reward and a
+%% noisy but less biased estimate that is obtained by using the same deep
+%% network to estimate the future reward of the next state, where the next
+%% state is chosen stochastically based on the predicted future rewards of all
+%% the currently legal actions.
+%% \fi
+
+\section{Historical perspective on backpropagation}
 
@@
 
-\iffalse
-\section{Boxes (TODO)}
-
-A box for the forward propagation and backpropagation equations.
-
-\section{Figures}
-
-1. Diagram showing:\\ a) A feedforward neural net.\\ b) A recurrent neural
-net expanded in time.\\ c) A convolutional neural net.
-
-DONE 2. Maps showing:\\ a) that word vectors learned by backpropagation are very
-good at capturing the meanings of the words.\\ b) that the thought vectors
-extracted by a translation model are very good at capturing the thoughts
-expressed by sentences.
-
-3. A diagram showing how unsupervised learning can be used to pre-train a
-deep feeedforward neural net.
-
-4. Some images with:\\ a) The opinions of a deep convolutional net about
-the classes of the prominent objects.\\ 
-DONE b) Several alternative captions
-generated by using the output of the convolutional net to determine the
-initial hidden state of a decoder recurrent net.
-
-5. Some images showing the localizations of parts of a person found by a
-deep convolutional neural net.\\
-
-\fi
+%% \iffalse
+%% \section{Boxes (TODO)}
+
+%% A box for the forward propagation and backpropagation equations.
+
+%% \section{Figures}
+
+%% 1. Diagram showing:\\ a) A feedforward neural net.\\ b) A recurrent neural
+%% net expanded in time.\\ c) A convolutional neural net.
+
+%% DONE 2. Maps showing:\\ a) that word vectors learned by backpropagation are very
+%% good at capturing the meanings of the words.\\ b) that the thought vectors
+%% extracted by a translation model are very good at capturing the thoughts
+%% expressed by sentences.
+
+%% 3. A diagram showing how unsupervised learning can be used to pre-train a
+%% deep feeedforward neural net.
+
+%% 4. Some images with:\\ a) The opinions of a deep convolutional net about
+%% the classes of the prominent objects.\\ 
+%% DONE b) Several alternative captions
+%% generated by using the output of the convolutional net to determine the
+%% initial hidden state of a decoder recurrent net.
+
+%% 5. Some images showing the localizations of parts of a person found by a
+%% deep convolutional neural net.\\
+
+%% \fi
 

### 2015-03-30 deepreview.tex
YB's changes + GH's changes

@@
 
-Deep Learning techniques allow a machine composed of multiple processing layers
-which define representations of the data that are more abstract for deeper layers.
+Deep Learning techniques allow machines that are composed of multiple processing layers
+to learn representations of data that become progressively more abstract in deeper layers.
 These methods have dramatically improved the state-of-the-art
@@
 {\em backpropagation algorithm} to indicate how the machine should
-change.  Deep {\em convolutional nets}
-have brought about dramatic improvements to process
+change the weights that are used to compute the representation in each layer from
+the representation in the previous layer.  Deep {\em convolutional nets}
+have brought about dramatic improvements in processing
 images, video, speech and audio, while {\em recurrent
@@
 Machine learning technology (ML) powers many aspects of modern
-society: from web search to content filtering on social networks, and
-to recommendations on e-commerce websites. It is also increasingly
+society: from web search to content filtering on social networks 
+to recommendations on e-commerce websites, and it is increasingly
 present in consumer products such as cameras and smartphones. ML
@@
 select relevant results of search. Increasingly, these applications
-make use a new class of techniques: {\em deep learning}.
+make use a new class of techniques called {\em deep learning}.
 
 % handcrafted features
-Traditional ML techniques are limited in their ability to process
+Traditional ML techniques were limited in their ability to process
 natural data in raw form. For decades, constructing a pattern
-recognition or ML system has required engineering efforts and domain
-expertise to design a {\em feature extractor} that transforms the raw
+recognition or ML system required careful engineering and considerable
+domain
+expertise to design a {\em feature extractor} that transformed the raw
 data, such as the pixel values of an image, into a suitable {\em
   internal representation}, or feature vector, from which the learning
-subsystem, often a classifier, can detect, recognize, or classify
+subsystem, often a classifier, could detect or classify
 patterns in the input.
@@
 Deep learning is a set of methods that allows a machine to be fed with
-raw data and to automatically discover suitable representations of the
-data. The name comes from the fact that deep learning machines are
-multilayer stacks of trainable modules, each of which (e.g. a neural
-network layer) turns the
-representation of the data in its input into a more abstract and
-higher-level representation on its output in which relevant
-characteristics and amplified and irrelevant variabilities are
-attenuated. For example, an image comes in the form of an array of
-pixel values, features in the first layer would represent the presence
-or absence of edges of a particular orientation at all locations in
-the image, the second layer would  detect motifs by spotting particular
-arrangements of edges, regardless of small variations in the positions
-of the edges.  The third layer would assemble motifs into parts of
-object, and subsequent layers would to detect objects as combinations of
-parts. The key aspect of deep learning is that {\em these features and
-  representations are not designed by human engineers, but are
-  automatically learned from data.}
+raw data and to automatically discover the representations needed for
+detection or classification. The name comes from the fact that deep learning machines are
+multilayer stacks of trainable modules, each of which turns the
+representation of the data on its input into a more abstract and
+higher-level representation on its output. The output representation
+amplifies aspects of the input that are important for discrimination
+and suppresses irrelevant variations. An image, for example, comes in the form of an array of
+pixel values and the learned features in the first layer of representation
+typically represent the presence
+or absence of edges at  particular orientations and locations in
+the image. The second layer typically detects motifs by spotting particular
+arrangements of edges, regardless of small variations in the edge positions
+The third layer  may assemble motifs into larger
+combinations that correspond to parts of familiar
+objects, and subsequent layers would to detect objects as combinations of
+these parts. The key aspect of deep learning is that these layers of
+features are not designed by human engineers: They are learned from
+data using a general purpose learning procedure.}
 
@@
 beaten other ML techniques at predicting the activity of potential
-drug molecules~\citep{Dahl-et-al-arxiv2014,Ma-et-al-2015}, detecting new fundamental
+drug molecules~\citep{Ma-et-al-2015}, detecting new fundamental
 particles~\citep{Ciodaro-et-al-2012,Melis-Higgs-boson-competition-2014},
@@
 
-Perhaps more interestingly, deep learning has produced extremely
+Perhaps more surprisingly, deep learning has produced extremely
 promising results for a variety of tasks in natural language
@@
 machine. In a typical deep learning system, there may be hundreds of
-millions of these adjustable weights, and tens or hundreds of millions
-of examples with which to train the machine.
+millions of these adjustable weights, and hundreds of millions
+of labeled examples with which to train the machine.
 
@@
 increased by a tiny amount. The weight vector is then adjusted in the
-opposite direction of the gradient vector.
+opposite direction to the gradient vector.
 
@@
 In practice, most practitioners use a procedure called {\em Stochastic
-  Gradient Descent} or SGD. It consists in showing an input, computing
-the output and the error, computing the gradient, and adjusting the
-weight vector right away. The process is repeated with each example in
-the training set, again and again, until the average of the objective
-function stops decreasing. It is called stochastic because each sample
-gives a noisy estimate of the average gradient. The procedure gets
-close to a good minimum very quickly in
-practice~\cite{bottou-bousquet-2008-small}. Following this, the
-performance of the system is measured on a different set of examples,
-not seen during training, called a {\em test set}. This serves to test
-the {\em generalization ability} of the machine, i.e. its ability to
+  Gradient Descent} or SGD. It consists in showing the input vector
+for a few examples, computing the outputs and the errors, computing
+the average gradient for those examples, and adjusting the
+weights. The process is repeated for many small sets of examples
+from the training set until the average of the objective function stops
+decreasing. It is called stochastic because each small set of examples
+gives a noisy estimate of the average gradient over all examples. In
+practice, this simple procedure usually finds a good set of weights
+surprisingly quickly when compared with far more elaborate
+optimization techniques~\cite{bottou-bousquet-2008-small}. After
+training, the performance of the system is measured on a different set
+of examples called a {\em test set}. This serves to test the {\em
+  generalization ability} of the machine, {\it i.e.} its ability to
 produce sensible answers on new inputs that it has never seen during
@@
 It has been known since the 1960s that linear classifiers can only
-carve out their input space into very simple regions, namely
+carve their input space into very simple regions, namely
 half-spaces separated by a hyperplane~\citep{Duda-Hart}.  But problems
@@
 variations in position, orientation or illumination of an object, or
-variation in pitch or accent of speech, while being very sensitive to
+variation in the pitch or accent of speech, while being very sensitive to
 particular minute variations, such as the difference between a white
@@
 of a samoyed and a wolf in the same position and on similar
-backgrounds may be quite similar to each other. A linear classifier,
-or any other ``shallow'' classifiers operating on raw pixels could not
+backgrounds may be very similar to each other. A linear classifier,
+or any other ``shallow'' classifier operating on raw pixels could not
 possibly distinguish the latter two, while putting the former two in
 the same category. This is why shallow classifiers require a good
-feature extractor that solve the {\em selectivity-invariance dilemma},
-i.e. that produces representations that are {\em selective} to the
-important aspects of the image, while being {\em invariant} to its
-irrelevant aspects, like the pose of the animal.
+feature extractor that solves the {\em selectivity-invariance dilemma},
+{\it i.e.} that produces representations that are {\em selective} to the
+aspects of the image that are important for discrimination while being {\em invariant} to 
+irrelevant aspects like the pose of the animal.
 %
@@
 % kernel function is ineffective. 
-Designing feature extractors that are
-sensitive to the relevant information and invariant to the irrelevant
-one is what generations of computer vision and pattern recognition
-researchers have spent their careers on.
+Designing good feature extractors traditionally required a lot of
+engineering skill and domain expertise, but this can all be avoided if
+good features can be learned automatically using a general purpose
+learning procedure.  That is the key idea of deep
+learning.
 
 % deep architectures
-But what if features could be learned? That is the key idea of deep
-learning.
 
@@
 non-linear input-output mappings.  Each module in the stack transforms
-its input so as to increase both the selectivity and the invariance a
-little bit.  With multiple non-linear layers, say 5 to 20, a system
+its input so as to increase both the selectivity and the invariance of
+the representation. With multiple non-linear layers, say 5 to 20, a system
 can implement extremely intricate functions of its inputs that are
 simultaneously sensitive to minute details-- distinguishing samoyeds
-from wolves-- and insensitive to irrelevant variations such as the
-background, the pose, the lighting, or the surrounding objects.
+from wolves-- and insensitive to large irrelevant variations such as the
+background, pose, lighting, and surrounding objects.
 
@@
 recognition~\citep{selfridge,Rosenblatt57} researchers have aimed to
-replace hand-engineered features by trainable multilayer networks.
-But despite its extreme simplicity, the solution did not appear until
+replace hand-engineered features by trainable multilayer networks,
+but despite its extreme simplicity, the solution was not widely
+understood until
 the mid 1980s and was largely forsaken by the ML community until
@@
 relatively smooth functions of their inputs and of their internal
-weights, one can compute gradients through the {\em backpropagation
+weights, one can compute gradients using the {\em backpropagation
   procedure}.
 
-The idea that this could be done, and that it could work at all, took
-a surprisingly long time to emerge. It was discovered independently by
+The idea that this could be done, and that it worked, was discovered independently by
 several different groups in the 1970s and
 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}, and became quite popular
-in the late 1980s, spurring the second wave of interest for neural
-networks. But for reasons that historians of science will have to
-explain, the method was the subject of scorn by the ML community
-between the late 1990s and the late 2000s.
+in the late 1980s, spurring a wave of interest for neural
+networks.
 
@@
 %\left(\deri{y}{x}\right)_{ij} \Delta x_i$. 
-The backpropagation equation
-constitutes a recurrence formula that can be used to propagate
-gradients from the top layer all the way to the bottom. Once all these
-gradients are computed, one can similarly compute the gradients with
-respect to the weights of the module. 
+The backpropagation equation can be used repeatedly to propagate
+gradients with respect to the outputs of a module from the top layer all the way to the bottom. Once these
+gradients have  computed, it is straightforward to compute the gradients with
+respect to the weights of each module. 
 %Figure~\ref{fig:backprop-box})
@@
 architectures, illustrated in Figure~\ref{fig:backprop-box}, which
-learn to map a fixed-size input (e.g., an image) to a fixed-size
-output (e.g. a probability for each one of several categories). To go
+learn to map a fixed-size input ({\it e.g.} an image) to a fixed-size
+output ({\it e.g.} a probability for each one of several categories). To go
 from one layer to the next, a set of units compute a weighted
@@
 half-wave rectifier $f(z) = \max(z,0)$. In past decades, neural nets
-used smoother non-linearities, such as $\tanh(z)$, but the ReLU seems
-to work better in networks with many layers. Units that are not in the
-output layer are traditionally called {\em hidden units}. The first
+used smoother non-linearities, such as $\tanh(z)$ or $1/(1+\exp(-z))$ but the ReLU 
+typically learns much faster in networks with many layers. Units that are not in the
+output layer are traditionally called {\em hidden units}. The early
 layers can be seen as distorting the input in a non-linear way so that
@@
 Tuesday and Wednesday are very similar, as are the word vectors for
-Vietnam and Iraq. This is not to say that Tuesday and Wednesday are
-put in the same category, but rather that they share many semantic and
-grammatical attributes, so that what the neural network learns from
-one day of the week can often be generalized to the other, while
-individual days remain different (for example one imagines that the
-neural network also learns some features that distinguishes week-end
-days and weekdays, while Saturday and Sunday might have special roles
-in some religions, etc.).  Such representations are called {\em
-  distributed representations} because its elements (the features) are
-not mutually exclusive and there exists a rich and vast set of
-possible configurations of values of these features, corresponding to
-the variations seen in the observed data.  These word-vectors are thus
+Vietnam and Iraq. 
+%% This is not to say that Tuesday and Wednesday are
+%% put in the same category, but rather that they share many semantic and
+%% grammatical attributes, so that what the neural network learns from
+%% one day of the week can often be generalized to the other, while
+%% individual days remain different (for example one imagines that the
+%% neural network also learns some features that distinguishes week-end
+%% days and weekdays, while Saturday and Sunday might have special roles
+%% in some religions, etc.).  
+Such representations are called {\em
+  distributed representations} because their elements (the features) are
+not mutually exclusive and their exponentially many configurations
+%% exists a rich and vast set of
+%% possible configurations of values of these features, 
+correspond to
+the variations seen in the observed data.  These word-vectors are
+composed of
 {\em learned features} that were not determined ahead of time by
-humans, but automatically discovered by the neural network.  Vector
+experts, but automatically discovered by the neural network.  Vector
 representations of words learned from text are now very widely used in
@@
 non-identical to other symbol instances. It has no internal structure
-that is relevant to its use. By contrast, neural networks use large
-activity vectors instead of symbols and parallel matrix operations
-followed by non-linearities instead of formal rules of inference. The
-elements of such vectors can be understood as the values of learned
-attributes (or features) associated with some object.  With sufficient
-data, computational power and depth, the combination of
-backpropagation for computing gradients and stochastic gradient
-descent for updating the weights on connections has proved to be very
-good at learning appropriate distributed representations. 
-%These
+that is relevant to its use and  to reason with symbols, they must be
+bound to the variables in judiciously chosen rules of inference.  By
+contrast, neural networks just use big activity vectors, big weight
+matrices and scalar non-linearities to perform the type of fast
+``intuitive'' inference that underpins effortless commonsense
+reasoning.
+
 %representations allow a lot of parallel computation at the feature
@@
 an exponential advantage over classical learning algorithm not based on distributed 
-representations~\citep{Bengio-localfailure-NIPS-2006-small,Bengio-decision-trees10},
+representations~\citep{Bengio-localfailure-NIPS-2006-small},
  in two ways, both arising from the power of {\em composition}, so long as the
@@
 of phrases learned by an English-to-French encoder-decoder recurrent
-neural network or RNN~\citep{Cho-et-al-EMNLP2014}. One can observe that
+neural network or RNN~\citep{Bahdanau-et-al-ICLR2015}. One can observe that
 semantically similar words or sequences of words are  mapped to
@@
 in a sequence (for language modeling) or a whole sequence
-of translated words (for machine translation)~\citep{Devlin-et-al-ACL2014,Bahdanau-et-al-arxiv2014,
-Sutskever-et-al-NIPS2014}.
+of translated words (for machine translation)~\citep{Bahdanau-et-al-ICLR2015,Sutskever-et-al-NIPS2014}.
 }
@@
 translation and so on until a full stop is
-chosen~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
+chosen~\citep{Bahdanau-et-al-ICLR2015,Sutskever-et-al-NIPS2014}.
 Overall, this process generates sequences of French words according to
@@
 whether understanding a sentence requires anything like internal symbolic
-expressions that are manipulated by binding their symbols to variables
-in judiciously chosen rules of inference, as was previously thought by
-many early AI researchers.  The encoder-decoder translation
-networks do not work by using internal rules of inference in which
-variables get bound to symbols. They just use big state vectors, big weight
-matrices and scalar non-linearities to get the job done.  
-Deriving conclusions using only the form of the premises (which is
-what formal logic does) is a powerful
-technique that people can eventually master, but the translation
-results with RNNs suggest that it may not be a good model of
-the way we normally reason or understand language. RNNs
-that represent thoughts as big vectors are more compatible with
+expressions that are manipulated by using inference rules.
+%%binding their symbols to variables
+%%in judiciously chosen rules of inference, as was previously thought by
+%%many early AI researchers. It is 
+%%The encoder-decoder translation
+%%networks do not work by using internal rules of inference in which
+%%variables get bound to symbols. They just use big state vectors, big weight
+%%matrices and scalar non-linearities to get the job done. 
+%%Deriving conclusions using only the form of the premises (which is
+%%what formal logic does) is a powerful
+%%technique that people can eventually master, but the translation
+%%results with RNNs suggest that it may not be a good model of
+%%the way we normally reason or understand language.
+%%RNNs that represent thoughts as big vectors are 
+It is more compatible with
 the view that everyday reasoning involves many simultaneous analogies that
@@
 for the encoder and decoder networks that perform so well at machine
-translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
+translation~\citep{Bahdanau-et-al-ICLR2015,Sutskever-et-al-NIPS2014}.
 
 Over the last year, several authors have made different proposals to
-augment RNNs with a memory module. This includes simplified
-version of the LSTM
-circuit~\citep{Chung-et-al-NIPSDL2014-small,Yao-et-al-SLU-workshop2014}.
-The results suggest that not all the elements of the LSTM architecture
-are required.
-Other proposals include the ``Neural Turing Machine'' in which the
+augment RNNs with a memory module. 
+%% This includes simplified
+%% version of the LSTM
+%% circuit~\citep{Chung-et-al-NIPSDL2014-small,Yao-et-al-SLU-workshop2014}.
+%% The results suggest that not all the elements of the LSTM architecture
+%% are required.
+Proposals include the ``Neural Turing Machine'' in which the
 network is augmented by a ``tape-like'' memory that the RNN
 can chose to read from or write to~\citep{Graves-et-al-arxiv2014}, and
-the ``Memory Network'' in which a regular network is augmented by a
+``Memory Networks'' in which a regular network is augmented by a
 kind of associative memory~\citep{weston-memorynet-2014}. Memory
@@
 
-Another RNN can learn to map the sequence
-of characters in a simple program written in Python to the sequence of
-characters that the program would
-print~\citep{Zaremba+Sutskever-arxiv2014}, 
-learning about {\it for loops} and small integer arithmetic.
-Achieving this kind of behaviour seemed ludicrously optimistic
-in the 1990's.
+%% Another RNN can learn to map the sequence
+%% of characters in a simple program written in Python to the sequence of
+%% characters that the program would
+%% print~\citep{Zaremba+Sutskever-arxiv2014}, 
+%% learning about {\it for loops} and small integer arithmetic.
+%% Achieving this kind of behaviour seemed ludicrously optimistic
+%% in the 1990's.
 
@@
 
-%\iffalse  
-%\section{Deep reinforcement learning}
-%
-%Learning which action will maximize the total future reward given a state
-%of the world is difficult when the rewards are delayed and the world has a
-%% very large number of possible states. After neural networks won backgammon
-%% olympiads in the 90's~\citep{Tesauro90}, an impressive recent development is
-%% combines deep convolutional networks with reinforcement learning in a single system that can learn to play many different video
-%% games~\citep{Mnih-et-al-2015}. 
-%% Pixels on the screen at several time steps are mapped to the
-%% expected future reward of each possible action (assuming some policy is followed
-%% in the future) and its only inputs are the
-%% pixels and the current score in the game. The learning works by minimizing
-%% the difference between the system's current estimate of future reward and a
-%% noisy but less biased estimate that is obtained by using the same deep
-%% network to estimate the future reward of the next state, where the next
-%% state is chosen stochastically based on the predicted future rewards of all
-%% the currently legal actions.
-%% \fi
-
 \section{Historical perspective on backpropagation}
@@
 with no prior knowledge was much too difficult for a simple gradient
-descent procedures and that the networks were getting trapped in poor
+descent procedure and that the networks were getting trapped in poor
 local minima, weight configurations for which no small change would
@@
 that the main reasons for the limited success of backpropagation in
-deep networks were quite different: labeled datasets that were often
-too small, computers that were too slow, non-linearities that were
-inappropriate, weights that were not properly
-initialised~\citep{lecun-98b,GlorotAISTATS2010-small}, and a lack of
-faith in the potential success.
-
-The obvious perceived weakness of stochastic gradient descent is that
-it can get trapped in poor local minima, but it turns out that local
-minima are rarely a problem with large networks. Regardless of the
-initial conditions, the system always reaches solutions of similar
-quality.  Recent results strongly suggest the landscape is packed with
-a combinatorially large number of {\em saddle points} where the
-gradient is zero, and the surface curves up in some dimensions and
-curves down in
+deep networks were that the labeled datasets were typically much too
+small and the computers were much too slow. Additional reasons were
+that researchers were not initializing the weights properly and were
+not using the type of non-linearlity that learns most
+rapidly~\citep{lecun-98b,GlorotAISTATS2010-small}.
+
+Poor local minima are rarely a problem with large networks. Regardless
+of the initial conditions, the system nearly always reaches solutions
+of very similar quality.  Recent results strongly suggest the
+landscape is packed with a combinatorially large number of {\em saddle
+  points} where the gradient is zero, and the surface curves up in
+some dimensions and curves down in
 others~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
-The analysis seems to show that the local minima are present in very
+The analysis seems to show that saddle points with predominantly
+upward curving directions are present in very
 large numbers, but almost all of them have very similar error levels.
 Hence it doesn't seem to matter which one the algorithm
-finds. Avoiding being trapped near saddle points seems more important.
-
-It also took an inordinate amount of time for the community to adopt
-stochastic gradient methods for large-scale machine learning. Key
-theoretical and practical work showed that, while SGD has terrible
-asymptotic convergence properties, it is extremely efficient for the
-kind of approximate optimization required in ML
-applications~\citep{bottou-bousquet-2008-small}. But when training a
-large ConvNet can take days or weeks on the fastest GPUs, one is eager
-to discover more efficient optimization methods that use the curvature
-information the past update directions~\citep{lecun-98b,Montavon2012}.
+finds. 
+%% Avoiding being trapped near saddle points seems more important.
+
+%% It also took an inordinate amount of time for the community to adopt
+%% stochastic gradient methods for large-scale machine learning. Key
+%% theoretical and practical work showed that, while SGD has terrible
+%% asymptotic convergence properties, it is extremely efficient for the
+%% kind of approximate optimization required in ML
+%% applications~\citep{bottou-bousquet-2008-small}. But when training a
+%% large ConvNet can take days or weeks on the fastest GPUs, one is eager
+%% to discover more efficient optimization methods that use the curvature
+%% information the past update directions~\citep{lecun-98b,Montavon2012}.
 
@@
 are averaged. The purpose of this is to saturate the GPU. But one big
-unsolved question is how to scale up the training of very large neural
+unsolved question is the best way to scale up the training of very large neural
 nets over multiple GPU cards across multiple machines. As of today, this
-question is unresolved.
+question is still unresolved.
 
@@
 2006~\citep{IJCAI,Hinton06-small,Bengio-nips-2006-small,ranzato-07-small}
-thanks to the effort of a forward-looking Canadian funding agency 
-(the Canadian Institute for Advanced Research, or CIFAR), with
-the introduction of unsupervised learning procedures that could
+by a group of researchers brought together by the Canadian Institute
+for Advanced Research who introduced
+unsupervised learning procedures that could
 create layers of feature detectors without requiring labeled data. The
@@
 detecting pedestrians, especially when the amount of labeled data was
-very limited~\citep{Erhan+al-2010-small,sermanet-cvpr-13}.
+very limited~\citep{sermanet-cvpr-13}.
 
@@
 Android phones.  For smaller datasets, unsupervised pre-training helps
-to prevent overfitting~\citep{Erhan+al-2010-small}, leading to
+to prevent overfitting~\citep{Bengio-Courville-Vincent-TPAMI2013}, leading to
 significantly better generalization when the number of labeled
@@
 vision, especially for video understanding or robotics, to come from
-systems that are trained end-to-end and combine deep convolutional
-networks with RNNs that use reinforcement learning to decide
-where to look.  Combining reinforcement learning with deep learning
-has already given impressive results~\citep{Tesauro90,Mnih-et-al-2015}
-but much more remains to be explored.
-Such systems already outperform passive vision
-systems~\citep{ba+mnih}.  Similarly, when understanding sentences or
+systems that are trained end-to-end and combine deep ConvNets with
+RNNs that use reinforcement learning to decide where to look.
+Combining reinforcement learning with deep learning has already
+produced an impressive system that can learn to play many different
+video games~\citep{Mnih-et-al-2015}.  Such systems already outperform passive vision
+systems~\citep{ba+mnih} at classification tasks, but much more remains to be
+explored.   Similarly, when understanding sentences or
 whole documents we expect future systems to learn strategies for
@@
 time~\citep{Bahdanau-et-al-arxiv2014}. We also expect that in the next
-few years deep RNNs will become far better at
-understanding the meaning of documents, including the way in which the
-thoughts expressed by the sentences are related to one another.
+few years deep RNNs will become far better at understanding the
+meaning of documents, including the way in which the thoughts
+expressed by nearby sentences are related to one another.
 
@@
 
-We expect to see more innovative ideas that marry deep learning with
-reasoning. Reasoning can be formulated in a number of different ways:
-energy minimization, contraint satisfaction, chains of operations
-applied to vectors representing objects or symbols. One is hoping for
-an architecture that unifies representation learning with
-reasoning~\citep{bottou-mlj-2014}.
+%% We expect to see more innovative ideas that marry deep learning with
+%% reasoning. Reasoning can be formulated in a number of different ways:
+%% energy minimization, contraint satisfaction, chains of operations
+%% applied to vectors representing objects or symbols. One is hoping for
+%% an architecture that unifies representation learning with
+%% reasoning~\citep{bottou-mlj-2014}.
 

### 2015-03-30 deepreview.tex
missing {

@@
 these parts. The key aspect of deep learning is that these layers of
-features are not designed by human engineers: They are learned from
+features are not designed by human engineers: {\em They are learned from
 data using a general purpose learning procedure.}

### 2015-03-30 deepreview.tex
merge of deepreview4.tex from GH

@@
 learning algorithm such as backpropagation plus stochastic gradient
-descent.
+descent. A primitive one-dimensional convnet called a time-delay
+neural net was used for phoneme recognition~cite{WaibelIEEE}.
 
@@
 
-Depsite these successes, ConvNets were largely forsaken by the
+Despite these successes, ConvNets were largely forsaken by the
 mainstream computer vision and machine learning communities until the
@@
 results, almost halving the error rates of the best competing
-approaches~\citep{Krizhevsky-2012-small}.  Recent ConvNet
+approaches~\citep{Krizhevsky-2012-small}.  This success came from efficient use of GPUs,
+ReLUs, and a new regularization technique called ``dropout''
+~\citep{Srivastava14} which prevented overfitting in the fully
+connected layers. 
+Recent ConvNet
 architectures have 10 to 20 layers of ReLUs, hundreds of millions of
@@
 
-If ConvNets and other neural nets existed in the 1990s, why did they become so popular and
-dominant only now? It was initially thought that training of neural nets
-would get stuck in poor solutions (called local minima) due to non-convexity
-of the training objective, whereas the alternative learning framework
-of {\em kernel machines}~\citep{Scholkopf+Smola-2002book} was more appealing
-because it involved a convex optimization. As discussed below, it is only
-in 2014 that theoretical and empirical evidence have suggested that local
-minima are probably not as much of a problem as was initially 
-thought~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
-Much before that realization, though, a number of 
-ideas and technologies over the last decade enabled the success. 
-First, supervised ConvNets work best
-when training data is abundant, and large labeled image datasets such
-as ImageNet~\citep{imagenet_cvpr09} only appeared recently. Second,
-the appearance of GPU cards capable of over 1Tflops, together with
-GPU-optimized software for ConvNet made it possible to train these
-large networks in a few weeks (now days) of computer time. Third, the
-use of rectifying non-linearities instead of sigmoids allowed to train
-much deeper networks than previously possible without unsupervised
-pre-training~\cite{jarrett-iccv-09,Glorot+al-AI-2011-small}. Lastly,
-the ``drop out'' regularization method~\citep{Srivastava14} was shown
-to mitigate the effects of over-fitting in very large networks.
-
-Deep convolutional nets have quickly replaced the carefully engineered
-pipelines full of hand-designed features developed by the computer
-vision community and they are now deployed in many computer vision
-applications, including image tagging, object detection, face
-recognition, facial expression recognition and scene parsing.
+%% If ConvNets and other neural nets existed in the 1990s, why did they become so popular and
+%% dominant only now? It was initially thought that training of neural nets
+%% would get stuck in poor solutions (called local minima) due to non-convexity
+%% of the training objective, whereas the alternative learning framework
+%% of {\em kernel machines}~\citep{Scholkopf+Smola-2002book} was more appealing
+%% because it involved a convex optimization. As discussed below, it is only
+%% in 2014 that theoretical and empirical evidence have suggested that local
+%% minima are probably not as much of a problem as was initially 
+%% thought~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
+%% Much before that realization, though, a number of 
+%% ideas and technologies over the last decade enabled the success. 
+%% First, supervised ConvNets work best
+%% when training data is abundant, and large labeled image datasets such
+%% as ImageNet~\citep{imagenet_cvpr09} only appeared recently. Second,
+%% the appearance of GPU cards capable of over 1Tflops, together with
+%% GPU-optimized software for ConvNet made it possible to train these
+%% large networks in a few weeks (now days) of computer time. Third, the
+%% use of rectifying non-linearities instead of sigmoids allowed to train
+%% much deeper networks than previously possible without unsupervised
+%% pre-training~\cite{jarrett-iccv-09,Glorot+al-AI-2011-small}. Lastly,
+%% the ``drop out'' regularization method was shown
+%% to mitigate the effects of over-fitting in very large networks. 
+%% 
+%% Deep convolutional nets have quickly replaced the carefully engineered
+%% pipelines full of hand-designed features developed by the computer
+%% vision community and they are now deployed in many computer vision
+%% applications, including image tagging, object detection, face
+%% recognition, facial expression recognition and scene parsing.
+%%
 
@@
 Field-Programmable Gate Arrays~\citep{farabet-suml-11} and are
-currently the subject of chip development in several companies that
+currently the subject of chip development that
 could enable real-time vision applications in embedded devices,
@@
   distributed representations} because their elements (the features) are
-not mutually exclusive and their exponentially many configurations
+not mutually exclusive and their many configurations
 %% exists a rich and vast set of
@@
 selectively attending to one part at a
-time~\citep{Bahdanau-et-al-arxiv2014}. We also expect that in the next
+time~\citep{Bahdanau-et-al-ICLR2015}. We also expect that in the next
 few years deep RNNs will become far better at understanding the

### 2015-03-30 deepreview.tex
reducing Bengio citations
removing citations

@@
 sentiment
-analysis~\citep{Glorot+al-ICML-2011-small},
+analysis, %~\citep{Glorot+al-ICML-2011-small}, % YB cutting self-citations
 question-answering~\cite{Bordes-et-al-EMNLP2014}, and language
@@
 regression~\cite{Hastie2001}, and the linear support vector
-machine~\cite{Vapnik+Chervonenkis-1964,Boser92}.
+machine~\cite{Vapnik+Chervonenkis-1964}. %{Boser92}.
 
@@
 applications~\citep{Schwenk-2007,collobert:2011b,Socher-2011,
-  Mikolov-et-al-NIPS2013,Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
+  Mikolov-et-al-NIPS2013,Bahdanau-et-al-ICLR2015,Sutskever-et-al-NIPS2014}.
 
@@
 and neural language modeling. There has been a surge of interest in such
-systems recently~\citep{Kiros-et-al-ICML2014,Mao+al-arxiv2014,Donahue-et-al-arxiv2014,
-              Vinyals-et-al-arxiv2014,Fang-et-al-arxiv2014,
-             Chen+Zitnick-arxiv2014,Karpathy+Li-arxiv2014,Venugopalan-et-al-arxiv2014}.
+systems recently (see citations in~\citet{Xu-et-al-arxiv2015}).
+%Kiros-et-al-ICML2014,Mao+al-arxiv2014,Donahue-et-al-arxiv2014,
+%              Vinyals-et-al-arxiv2014,Fang-et-al-arxiv2014,
+%             Chen+Zitnick-arxiv2014,Karpathy+Li-arxiv2014,Venugopalan-et-al-arxiv2014}.
 

### 2015-03-30 deepreview.tex
removed 2 more Bengio citations

@@
 win several transfer learning
-competitions~\citep{UTLC+LISA-2011-small,Goodfellow-icml2012}.  Once
+competitions~\citep{Bengio-Courville-Vincent-TPAMI2013}.  Once
 deep learning had been rehabilitated, it turned out that large labeled

### 2015-03-30 deepreview.tex
Fixes, rewording, citation suppression, not much compression, I'm afraid.

@@
 
-Deep Learning techniques allow machines that are composed of multiple processing layers
-to learn representations of data that become progressively more abstract in deeper layers.
+Deep Learning allows machines that are composed of multiple processing 
+layers to learn representions of data with multiple levels of abstraction.
 These methods have dramatically improved the state-of-the-art
-for speech recognition, object recognition, object detection,
-predicting the activity of drug molecules, and many other tasks. Deep
+in speech recognition, object recognition, object detection,
+and many other domains such as drug discovery and genommics. Deep
 learning discovers intricate structure in large datasets by using the
@@
 select relevant results of search. Increasingly, these applications
-make use a new class of techniques called {\em deep learning}.
+make use of a new class of techniques called {\em deep learning}.
 
@@
 combinations that correspond to parts of familiar
-objects, and subsequent layers would to detect objects as combinations of
+objects, and subsequent layers would detect objects as combinations of
 these parts. The key aspect of deep learning is that these layers of
@@
 for many years. It has turned out to be very good at discovering
-intricate structure in large, high-dimensional data and is therefore
+intricate structures in large, high-dimensional data and is therefore
 applicable to many domains of science, business, and government. In
@@
 variations in position, orientation or illumination of an object, or
-variation in the pitch or accent of speech, while being very sensitive to
+variations in the pitch or accent of speech, while being very sensitive to
 particular minute variations, such as the difference between a white
@@
 % kernel function is ineffective. 
-Designing good feature extractors traditionally required a lot of
-engineering skill and domain expertise, but this can all be avoided if
+Designing good feature extractors traditionally required considerable amounts
+of engineering skills and domain expertise. But this can all be avoided if
 good features can be learned automatically using a general purpose
-learning procedure.  That is the key idea of deep
+learning procedure.  That is the key advantage of deep
 learning.
@@
 several different groups in the 1970s and
-1980s~\citep{Werbos74,Parker85,LeCun85,RHW}, and became quite popular
-in the late 1980s, spurring a wave of interest for neural
-networks.
+1980s~\citep{Werbos74,Parker85,LeCun85,RHW}. It spurred a wave of interest for
+neural networks, becoming quite popular in the late 1980s, but fell out of favor
+between the mid 1990s until its rennaissance in the late 2000s.
 
@@
 %\left(\deri{y}{x}\right)_{ij} \Delta x_i$. 
-The backpropagation equation can be used repeatedly to propagate
-gradients with respect to the outputs of a module from the top layer all the way to the bottom. Once these
-gradients have  computed, it is straightforward to compute the gradients with
-respect to the weights of each module. 
+The backpropagation equation can be applied repeatedly to propagate
+gradients through all modules, starting from the output at the top 
+all the way to the bottom. Once these
+gradients have been computed, it is straightforward to compute the gradients 
+with respect to the weights of each module. 
 %Figure~\ref{fig:backprop-box})
@@
 four key ideas behind ConvNets that take advantage of the properties
-of many such natural signals: local connections, shared weights,
+of natural signals: local connections, shared weights,
 pooling, and the use of many layers.
@@
 set of weights called a {\em filter bank}. The result of this local
-weight sum is then passed through a non-linearity such as a ReLU.  All
+weighted sum is then passed through a non-linearity such as a ReLU.  All
 units in a feature map share the same filter bank. Different feature
@@
 array.  Mathematically, the filtering operation performed by a feature
-map is a {\em discrete convolution}, hence the name ConvNet.
+map is a {\em discrete convolution}, hence the name.
 
@@
 
-This ConvNet architecture exploits the property that many natural
+ConvNets exploit the property that many natural
 signals are {\em compositional hierarchies}. In images, local
 combinations of edges form motifs, motifs assemble into parts, and
-parts form objects in various combinations.  Similar hierarchies
+parts form objects.  Similar hierarchies
 exists in speech and text from sounds to phones, phonemes, syllables,
-words, and sentences. The pooling allow features that form elements at
-the next level to vary in relative position and appearance.
+words, and sentences. The pooling allows representations to vary very little
+when elements in the previous layer vary in position and appearance.
 
@@
 neuroscience~\citep{Hubel62}, and the overall architecture is
-reminiscent of the LGN-V1-V2-V4-IT visual cortex ventral pathway
-hierarchy ~\citep{Felleman+VanEssen-1991}. When ConvNet models and
+reminiscent of the LGN-V1-V2-V4-IT hierarchy in the visual cortex 
+ventral pathway~\citep{Felleman+VanEssen-1991}. When ConvNet models and
 monkeys are shown the same picture, the activations of high-level
 units in the ConvNet explains half the variance of random sets of 160
-neurans in the monkey's
-inferro-temporal~\citep{cadieu-plos-2014}. ConvNets have their roots
+neurons in the monkey's
+inferro-temporal cortex~\citep{cadieu-plos-2014}. ConvNets have their roots
 in Fukushima's Neocognitron~\citep{fukushima-82} whose architecture
 was somewhat similar but did not have an end-to-end supervised
-learning algorithm such as backpropagation plus stochastic gradient
-descent. A primitive one-dimensional convnet called a time-delay
-neural net was used for phoneme recognition~cite{WaibelIEEE}.
+learning algorithm such as backpropagation. 
+A primitive one-dimensional convnet called a time-delay
+neural net was used for the recognition of phonemes and simple 
+words~\cite{WaibelIEEE,bottou-89}
 
@@
 were later deployed by
-Microsoft~\citep{simard-03,chellapilla-ist-06,chellapilla-iwfhr-06b}
+Microsoft~\citep{simard-03}
 including for Arabic and
 Chinese~\citep{abdulkader-iwfhr-06,chellapilla-iwfhr-06a}. ConvNets
-have also been used since the early 1990s for object detection in
-natural images, including faces and
+were also experimented with in the early 1990s for object detection 
+in natural images, including faces and
 hands~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95} and for face
@@
 
-Since the early 2000s, there have been many tasks in which ConvNets
-have been applied with success to the detection, segmentation, and
-recognition of objects and regions in images. These were invariably
+Since the early 2000s, ConvNets have been applied with great 
+success to the detection, segmentation, and
+recognition of objects and regions in images. These were all
 tasks in which labeled data was relatively abundant, such as traffic
@@
 images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,
-  nasse-09,fan-pami-2010,Jain-et-al-arxiv2013,sermanet-cvpr-13,Tompson-et-al-arxiv2014}.
+  nasse-09,fan-pami-2010,sermanet-cvpr-13,Tompson-et-al-arxiv2014}.
 A major practical success of ConvNets in recent times is face
-detection~\citep{taigman-cvpr-2014}.
+recognition~\citep{taigman-cvpr-2014}. A recent stunning demonstration
+of ConvNets is the generation of image captions, as shown in
+Figure~\ref{fig:caption-generation}.
 
 An important application is the labeling of images at the pixel level,
-whose applications include driving mobile robots and self-driving
-cars~\cite{hadsell-jfr-09,farabet-icml-12}. Other applications of
-ConvNets that are gaining importance concern natural language
+the applications of which include autonomous mobile robots and
+self-driving cars~\cite{hadsell-jfr-09,farabet-icml-12}. Companies
+such as Mobileye and NVIDIA are using ConvNets in their upcoming
+vision systems for cars.  Other applications of ConvNets that are
+gaining importance concern natural language
 understanding~\citep{collobert:2011b}, and speech
@@
 results, almost halving the error rates of the best competing
-approaches~\citep{Krizhevsky-2012-small}.  This success came from efficient use of GPUs,
-ReLUs, and a new regularization technique called ``dropout''
-~\citep{Srivastava14} which prevented overfitting in the fully
-connected layers. 
-Recent ConvNet
-architectures have 10 to 20 layers of ReLUs, hundreds of millions of
-weights, and billions of connections between units. The same kind of
-network is used to generate the captions shown in
-Figure~\ref{fig:caption-generation}. This success has brought about a
+approaches~\citep{Krizhevsky-2012-small}.  This success came from
+efficient use of GPUs, ReLUs, and a new regularization technique
+called ``dropout'' ~\citep{Srivastava14} which prevented overfitting
+in the fully connected layers.  This success has brought about a
 revolution in computer vision, and ConvNets are now the dominant
@@
 tasks~\citep{sermanet-iclr-14,girshick-cvpr-2014,taigman-cvpr-2014,simonyan-arxiv-2014,szegedy-2014,Tompson-et-al-arxiv2014}.
+
+Recent ConvNet architectures have 10 to 20 layers of ReLUs, hundreds
+of millions of weights, and billions of connections between
+units. While training such large networks could take weeks only two
+years ago, progress in hardware, software, and algorithm
+parallelization have reduced training times to a few hours.
+
 The performance of ConvNet-based vision systems has caused most major
-technology companies to start research and development projects and to
-deploy ConvNet-based image understanding products and services
-including Google, Facebook, Microsoft, Yahoo!, Twitter, Adobe, and a
-quickly-growing number of startups.
+technology companies to initiate research and development projects and
+to deploy ConvNet-based image understanding products and services
+including Google, Facebook, Microsoft, IBM, Yahoo!, Twitter, Adobe,
+and a quickly-growing number of startups.
 
@@
 
-ConvNets have been implemented in efficient hardware, using
-Field-Programmable Gate Arrays~\citep{farabet-suml-11} and are
-currently the subject of chip development that
-could enable real-time vision applications in embedded devices,
-robots, cars, and mobile devices.
+ConvNets are easily ammenable to efficient hardware implementations in
+chips or Field-Programmable Gate
+Arrays~\cite{boser-92,farabet-suml-11}.  A number of companies such as
+NVIDIA, Mobileye, Intel, Qualcomm and Samsung are developing ConvNet
+chips to enable real-time vision applications in smart phones,
+cameras, robots, and self-driving cars.
 
@@
 %competition~\citep{Graves-et-al-2009}. 
-LSTM have subsequently proved
-to be very effective for tasks that have sequential input.  When they
-use several hidden layers, they are better than deep feedforward
-networks for speech recognition~\citep{Graves-et-al-ICASSP2013} and
-unlike feedforward nets they can implement an entire speech
-recognition system that goes all the way from coefficents derived from
-the soundwave to the sequence of characters in the transcription.
-LSTMs or related forms of gated units are also what is currently used
-for the encoder and decoder networks that perform so well at machine
+LSTM have subsequently proved to be very effective for tasks that have
+sequential input.  When they use several hidden layers, they are
+better than deep feedforward networks for speech
+recognition~\citep{Graves-et-al-ICASSP2013} and unlike feedforward
+nets they can implement an entire speech recognition system that goes
+all the way from coefficents derived from the soundwave to the
+sequence of characters in the transcription.  LSTMs or related forms
+of gated units are also what is currently used for the encoder and
+decoder networks that perform so well at machine
 translation~\citep{Bahdanau-et-al-ICLR2015,Sutskever-et-al-NIPS2014}.
@@
 question-answering benchmarks. The memory is used to remember the
-story about which the network is asked to answer questions.
+story about which the network is later asked to answer questions.
 
@@
 Memory Networks can be trained to keep track of the state of the world
-when they are told a story, so that they can answer questions about
-the state of the world at the end of the story. 
+in a setting similar to a text adventure game. After reading a story,
+they can answer questions that require complex 
+inference~\cite{weston-babi-arxiv-2015}.
 % Following a sequence
@@
 Rings and correctly answers questions like ``where is Frodo now?''.
-~\citep{weston-memorynet-2014}
+~\citep{weston-memorynet-2014}.
 
@@
 % Bernhard was not impressed, so maybe just cut this off
+% YLC: it's true that the basic concept is more impressive than the end result.
 %In a somewhat related work, a network is trained to evaluate whether a
@@
 In the late 1990s, neural nets and backpropagation were largely
-forsaken by the machine learning community. At the time, people
-believed that learning many layers of non-linear feature detectors
-with no prior knowledge was much too difficult for a simple gradient
-descent procedure and that the networks were getting trapped in poor
-local minima, weight configurations for which no small change would
-reduce the average error, despite considerable empirical evidence to
+forsaken by the machine learning community and ignored by the computer
+vision and speech recognition communities.  It was widely believed
+that learning multi-stage feature extractors with little prior knowledge 
+was neither useful nor effective. In particular, it was commonly
+believed that simple gradient descent procedure get trapped in poor
+local minima -- weight configurations for which no small change would
+reduce the average error-- despite considerable empirical evidence to
 the contrary in the case of ConvNets~\citep{lecun-98}. We now know
@@
 others~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
-The analysis seems to show that saddle points with predominantly
-upward curving directions are present in very
-large numbers, but almost all of them have very similar error levels.
-Hence it doesn't seem to matter which one the algorithm
-finds. 
+The analysis seems to show that local minima (and saddle points with
+only a few downward curving directions) are present in very large
+numbers, but almost all of them have very similar values of the
+objective function.  Hence it doesn't seem to matter which local
+minimum the algorithm finds.
 %% Avoiding being trapped near saddle points seems more important.
@@
 Android phones.  For smaller datasets, unsupervised pre-training helps
-to prevent overfitting~\citep{Bengio-Courville-Vincent-TPAMI2013}, leading to
-significantly better generalization when the number of labeled
-examples is small, or in a transfer setting where we have lots of
-examples for some ``source'' tasks but very few for some ``target''
+to prevent overfitting~\citep{Bengio-Courville-Vincent-TPAMI2013},
+leading to significantly better generalization when the number of
+labeled examples is small, or in a transfer setting where we have lots
+of examples for some ``source'' tasks but very few for some ``target''
 task. In this context, unsupervised representation learning helped to
 win several transfer learning
-competitions~\citep{Bengio-Courville-Vincent-TPAMI2013}.  Once
-deep learning had been rehabilitated, it turned out that large labeled
+competitions~\citep{Bengio-Courville-Vincent-TPAMI2013}.  Once deep
+learning had been rehabilitated, it turned out that large labeled
 datasets did not need the pre-training stage. Even better results,
@@
 rectifying non-linearities rather than sigmoid or tanh non-linearities
-~\citep{Nair-2010-small,Glorot+al-AI-2011-small,Krizhevsky-2012-small,Dahl-et-al-ICASSP2013}.
+~\citep{jarrett-iccv-09,Nair-2010-small,Glorot+al-AI-2011-small,Krizhevsky-2012-small,Dahl-et-al-ICASSP2013}. Deep
+neural nets are now the dominant method for acoustic modeling in
+speech, and ConvNet are dominant for image and video recognition.
 
@@
 
-Human vision is an active process in which intelligent fixations are
-used to ensure that only a tiny fraction of the optic array is ever
-processed at high resolution. We expect much of the future progress in
-vision, especially for video understanding or robotics, to come from
-systems that are trained end-to-end and combine deep ConvNets with
-RNNs that use reinforcement learning to decide where to look.
-Combining reinforcement learning with deep learning has already
-produced an impressive system that can learn to play many different
-video games~\citep{Mnih-et-al-2015}.  Such systems already outperform passive vision
-systems~\citep{ba+mnih} at classification tasks, but much more remains to be
-explored.   Similarly, when understanding sentences or
-whole documents we expect future systems to learn strategies for
-selectively attending to one part at a
-time~\citep{Bahdanau-et-al-ICLR2015}. We also expect that in the next
-few years deep RNNs will become far better at understanding the
-meaning of documents, including the way in which the thoughts
-expressed by nearby sentences are related to one another.
+Human vision is an active process in which the few degrees around the
+center of the visual field that is processed at high resolution is
+intelligently moved around through saccades.
+
+We expect much of the future progress in vision to come from systems
+that are trained end-to-end and combine ConvNets with RNNs that decide
+where to look through reinfoercement learning. Such systems already
+outperform passive vision systems~\citep{ba+mnih} at classification
+tasks, but much more remains to be explored.  Combining reinforcement
+learning with ConvNets has already produced impressive results in
+learning to play many different video games~\citep{Mnih-et-al-2015}.
+Similarly, when understanding sentences or whole documents we expect
+future systems to learn strategies for selectively attending to one
+part at a time~\citep{Bahdanau-et-al-ICLR2015}.
+
+Certainly, natural language understanding is an area in which deep
+learning is poised to make a large impact over the next few years,
+using combinations of RNN, ConvNets, reasoning, and unsupervised
+learning.
 
@@
 learning procedures for deep neural
-networks~\citep{Salakhutdinov2009-small,Hinton95,QuocLe-ICML2012,VincentPLarochelleH2008-small,koray-nips-10,gregor-icml-10,ranzato-pami,Bengio-et-al-ICML-2014,Kingma-et-al-NIPS2014}.
-These papers often introduce a probabilistic view or the injection of randomness in the
-neural network computation, to allow neural networks to {\em generate} examples that are
-similar to training examples but generalize them.
-Unsupervised learning is important because most data have not been hand-labeled by humans, and just like
-children are not told what they should have done at each instant, we
-need learning machines that can learn from mostly unlabeled examples.
-In the long run, we believe unsupervised learning will be essential
-for allowing deep networks to generalize well by creating internal
-representations that untangle the many different factors that interact
-to produce images, speech and
-documents~\citep{Bengio-Courville-Vincent-TPAMI2013}.
-
-%% We expect to see more innovative ideas that marry deep learning with
-%% reasoning. Reasoning can be formulated in a number of different ways:
-%% energy minimization, contraint satisfaction, chains of operations
-%% applied to vectors representing objects or symbols. One is hoping for
-%% an architecture that unifies representation learning with
-%% reasoning~\citep{bottou-mlj-2014}.
-
-Finally, we believe that the recent successes of deep learning using
-stochastic gradient descent are only the beginning.  As pointed out
-in~\citet{Bengio+Lecun-chapter2007-small}, end-to-end learning
-systems that have many parameters and few built-in assumptions can make use
-of more data and more compute power far more easily than systems that
-require hand-engineering of domain-specific knowledge.
+networks~\citep{Salakhutdinov2009-small,Hinton95,VincentPLarochelleH2008-small,koray-nips-10,gregor-icml-10,ranzato-pami,Bengio-et-al-ICML-2014,Kingma-et-al-NIPS2014}.
+Human and animal learning is largely unsupervised: we discover the
+structure of the world by observing it, not by being told the name of
+every object. Most current methods are built around the idea of {\em
+  reconstruction} of the input, and can sometimes be used to {\em
+  generate} inputs that are similar to training examples. The
+principle is to disentangle the independent factors that interact to
+produce images, speech and
+documents~\citep{Bengio-Courville-Vincent-TPAMI2013}.  But some
+researchers view the essence of intelligence as the ability to predict
+the future, leading them to propose new paradigms based on
+{\em prediction}.
+
+%% YLC: I uncommented this. I really think this is were things are going.
+Ultimately, progress in AI will come about through systems that
+combine representation learning with complex {\em reasoning}. While
+deep learning and simple resoning has been used for speech and
+handwriting recognition for a long time, new paradigms are needed to
+replace symbolic logic manipulation by operations on
+vectors~\citep{bottou-mlj-2014}.
+
+With new paradigms for unsupervised learning and reasoning on the
+horizon, and with new learning algorithms that transcend simple
+gradient descent, deep learning is poised to progress quickly over the
+next few years, as the research community grows, the machines become
+more powerful, and the datasets become larger.
 

### 2015-03-30 deepreview.tex
a coupe more fixes

@@
 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}. It spurred a wave of interest for
-neural networks, becoming quite popular in the late 1980s, but fell out of favor
-between the mid 1990s until its rennaissance in the late 2000s.
+neural networks, becoming quite popular in the late 1980s, but falling out of favor
+between the mid 1990s until their rennaissance in the late 2000s.
 
@@
 A major practical success of ConvNets in recent times is face
-recognition~\citep{taigman-cvpr-2014}. A recent stunning demonstration
-of ConvNets is the generation of image captions, as shown in
-Figure~\ref{fig:caption-generation}.
+recognition~\citep{taigman-cvpr-2014}.
 
@@
 self-driving cars~\cite{hadsell-jfr-09,farabet-icml-12}. Companies
-such as Mobileye and NVIDIA are using ConvNets in their upcoming
-vision systems for cars.  Other applications of ConvNets that are
-gaining importance concern natural language
+such as Mobileye and NVIDIA are using such ConvNets-based methods in
+their upcoming vision systems for cars.  Other applications of
+ConvNets that are gaining importance concern natural language
 understanding~\citep{collobert:2011b}, and speech
@@
 approach for almost all recognition and detection
-tasks~\citep{sermanet-iclr-14,girshick-cvpr-2014,taigman-cvpr-2014,simonyan-arxiv-2014,szegedy-2014,Tompson-et-al-arxiv2014}.
+tasks~\citep{sermanet-iclr-14,girshick-cvpr-2014,taigman-cvpr-2014,simonyan-arxiv-2014,szegedy-2014,Tompson-et-al-arxiv2014}
+and approach human performance on some tasks. A recent stunning
+demonstration combines ConvNets and neural net modules for the
+generation of image captions, as shown in
+Figure~\ref{fig:caption-generation}.
 

### 2015-03-30 deepreview.tex
fixed a citation

@@
 pedestrians, and human bodies in natural
-images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,
-  nasse-09,fan-pami-2010,sermanet-cvpr-13,Tompson-et-al-arxiv2014}.
+images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,nasse-09,fan-pami-2010,sermanet-cvpr-13,tompson-cvpr-15}.
 A major practical success of ConvNets in recent times is face
@@
 approach for almost all recognition and detection
-tasks~\citep{sermanet-iclr-14,girshick-cvpr-2014,taigman-cvpr-2014,simonyan-arxiv-2014,szegedy-2014,Tompson-et-al-arxiv2014}
+tasks~\citep{sermanet-iclr-14,girshick-cvpr-2014,taigman-cvpr-2014,simonyan-arxiv-2014,szegedy-2014,tompson-cvpr-15}
 and approach human performance on some tasks. A recent stunning

### 2015-03-30 deepreview.tex
small changes and trying to reduce more

--- 2015-03-30 deepreview.tex
@@
 
-\addtolength{\textwidth}{2cm}
+\addtolength{\textwidth}{2.5cm}
 %\addtolength{\textwidth}{1cm}
-\addtolength{\textheight}{2cm}
+\addtolength{\textheight}{2.5cm}
 %\addtolength{\textheight}{1cm}
-\addtolength{\topmargin}{-1cm}
-\addtolength{\oddsidemargin}{-1cm}
-\addtolength{\evensidemargin}{-1cm}
+\addtolength{\topmargin}{-1.25cm}
+\addtolength{\oddsidemargin}{-1.25cm}
+\addtolength{\evensidemargin}{-1.25cm}
 
@@
 
-\parskip 0.1in
+\parskip 0.085in
 \parindent 0.0in
@@
 
-\author{Yann LeCun, Yoshua Bengio, Geoffrey Hinton}
-
+\author{Yann LeCun, Yoshua Bengio and Geoffrey Hinton}
+\date{}
 \begin{document}
@@
 \maketitle
+\vspace*{-6mm}
 
@@
 in speech recognition, object recognition, object detection,
-and many other domains such as drug discovery and genommics. Deep
+and many other domains such as drug discovery and genomics. Deep
 learning discovers intricate structure in large datasets by using the
@@
 above a threshold, the input is classified as belonging to a
-particular category. Multiple classes can be handled by using multiple
-two-class classifiers, each classifying one class against all
-others. The learning procedure adjusts the weights of the linear
-combination, for which the gradient is very easy to compute. Linear
-classifiers have been widely used for many decades, starting with the
-Perceptron model~\citep{Rosenblatt57}, the still-popular logistic
-regression~\cite{Hastie2001}, and the linear support vector
-machine~\cite{Vapnik+Chervonenkis-1964}. %{Boser92}.
+particular category. % Multiple classes can be handled by using multiple
+% two-class classifiers, each classifying one class against all
+% others. 
+% The learning procedure adjusts the weights of the linear
+% combination, for which the gradient is very easy to compute. Linear
+% classifiers have been widely used for many decades, starting with the
+% Perceptron model~\citep{Rosenblatt57} and the still-popular logistic
+% regression~\cite{Hastie2001}.
+% , and the linear support vector
+% machine~\cite{Vapnik+Chervonenkis-1964}. %{Boser92}.
 
@@
 neural networks, becoming quite popular in the late 1980s, but falling out of favor
-between the mid 1990s until their rennaissance in the late 2000s.
+between the mid 1990s until their renaissance in the late 2000s.
 
@@
 ConvNet is as simple as through a regular deep network, allowing all
-the weights in all the filter banks to be trained. The error
-derivative for a shared weight is just the sum the derivatives for all
-the instances of that weight.
+the weights in all the filter banks to be trained. 
+% The error
+% derivative for a shared weight is just the sum the derivatives for all
+% the instances of that weight.
 
@@
 learning algorithm such as backpropagation. 
-A primitive one-dimensional convnet called a time-delay
-neural net was used for the recognition of phonemes and simple 
-words~\cite{WaibelIEEE,bottou-89}
+% A primitive one-dimensional convnet called a time-delay
+% neural net was used for the recognition of phonemes and simple 
+% words~\cite{WaibelIEEE,bottou-89}
 
 There have been numerous applications of convolutional networks going
-back to the early 1990s. A major early success was a bank check
-reading system which was commercially deployed by AT\&T and NCR in
-1994~\citep{lecun-98}. The system used a ConvNet trained jointly with
+back to the early 1990s, starting with time-delay neural networks for
+speech recognition and document reading~\citep{lecun-98}. 
+The document reading system used a ConvNet trained jointly with
 a probabilistic model that implemented language constraints. By the
@@
 were later deployed by
-Microsoft~\citep{simard-03}
-including for Arabic and
-Chinese~\citep{abdulkader-iwfhr-06,chellapilla-iwfhr-06a}. ConvNets
-were also experimented with in the early 1990s for object detection 
-in natural images, including faces and
-hands~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95} and for face
-recognition~\cite{lawrence-tnn-1997}.
+Microsoft~\citep{simard-03}.
+% including for Arabic and
+% Chinese~\citep{abdulkader-iwfhr-06,chellapilla-iwfhr-06a}. 
+% ConvNets
+% were also experimented with in the early 1990s for object detection 
+% in natural images, including faces and
+% hands~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95} and for face
+% recognition~\cite{lawrence-tnn-1997}.
 
@@
 
-ConvNets are easily ammenable to efficient hardware implementations in
+ConvNets are easily amenable to efficient hardware implementations in
 chips or Field-Programmable Gate
@@
 applications~\citep{Schwenk-2007,collobert:2011b,Socher-2011,
-  Mikolov-et-al-NIPS2013,Bahdanau-et-al-ICLR2015,Sutskever-et-al-NIPS2014}.
+  Mikolov-et-al-NIPS2013,Bahdanau-et-al-ICLR2015-small,Sutskever-et-al-NIPS2014}.
 
@@
 of phrases learned by an English-to-French encoder-decoder recurrent
-neural network or RNN~\citep{Bahdanau-et-al-ICLR2015}. One can observe that
+neural network or RNN~\citep{Bahdanau-et-al-ICLR2015-small}. One can observe that
 semantically similar words or sequences of words are  mapped to
@@
 in a sequence (for language modeling) or a whole sequence
-of translated words (for machine translation)~\citep{Bahdanau-et-al-ICLR2015,Sutskever-et-al-NIPS2014}.
+of translated words (for machine translation)~\citep{Bahdanau-et-al-ICLR2015-small,Sutskever-et-al-NIPS2014}.
 }
@@
 translation and so on until a full stop is
-chosen~\citep{Bahdanau-et-al-ICLR2015,Sutskever-et-al-NIPS2014}.
+chosen~\citep{Bahdanau-et-al-ICLR2015-small,Sutskever-et-al-NIPS2014}.
 Overall, this process generates sequences of French words according to
@@
 %competition~\citep{Graves-et-al-2009}. 
-LSTM have subsequently proved to be very effective for tasks that have
-sequential input.  When they use several hidden layers, they are
-better than deep feedforward networks for speech
-recognition~\citep{Graves-et-al-ICASSP2013} and unlike feedforward
-nets they can implement an entire speech recognition system that goes
-all the way from coefficents derived from the soundwave to the
+LSTM have subsequently proved to be more effective than vanilla RNNs
+especially when they have several layers for each time step~\citep{Graves-et-al-ICASSP2013},
+enabling an entire speech recognition system that goes
+all the way from acoustics to the
 sequence of characters in the transcription.  LSTMs or related forms
@@
 decoder networks that perform so well at machine
-translation~\citep{Bahdanau-et-al-ICLR2015,Sutskever-et-al-NIPS2014}.
+translation~\citep{Bahdanau-et-al-ICLR2015-small,Sutskever-et-al-NIPS2014}.
 
@@
 
-The parallelization of large neural nets over a single GPU has
-required to use ``mini-batches'' of samples over which the gradients
-are averaged. The purpose of this is to saturate the GPU. But one big
-unsolved question is the best way to scale up the training of very large neural
-nets over multiple GPU cards across multiple machines. As of today, this
-question is still unresolved.
-
 Interest in deep feedforward networks was revived around
@@
 very limited~\citep{sermanet-cvpr-13}.
+
+% The parallelization of large neural nets over a single GPU has
+% required to use ``mini-batches'' of samples over which the gradients
+% are averaged. The purpose of this is to saturate the GPU. 
+% But one big
+% unsolved question is the best way to scale up the training of very large neural
+% nets over multiple GPU cards across multiple machines. As of today, this
+% question is still unresolved.
 
@@
 that are trained end-to-end and combine ConvNets with RNNs that decide
-where to look through reinfoercement learning. Such systems already
+where to look through reinforcement learning. Such systems already
 outperform passive vision systems~\citep{ba+mnih} at classification
@@
 future systems to learn strategies for selectively attending to one
-part at a time~\citep{Bahdanau-et-al-ICLR2015}.
+part at a time~\citep{Bahdanau-et-al-ICLR2015-small}.
 

### 2015-03-30 response.txt
preparing a rebuttal

We have made the following main changes:

- reduced the preface to less than 100 words
- followed most of the suggestions from the reviewers
- reduced the text considerably, especially in the more technical parts
  (but still 3/4 pages over the target of 8 pages)
- reduced the number of citations from 139 to 105 (target was 100)

Regarding the reviews, please find more detailed comments below.

---------

From Editor:
-----------

> In doing so, we should endeavour to preserve both accessibility and topicality. Deep Learning being red hot as it is, it is hard for us to select which places to cut in priority, for most areas covered seem quite vibrant indeed (with most references dating from just the past couple of years). We are, therefore, discussing the issue further with our peer-reviewers...

We have reduced parts on kernel machines and shallow learners, to keep our core business of deep learning.

From Referee 1:
--------------

> The text starting 'the backpropagation procedure' on p 4 is the key case in point.  The mathematical ideas here, particularly their presentation in equation form, might derail some readers.  Do we need to know that the matrix of partial derivatives is called the Jacobian Matrix?  Do we need to see the expression for dE/dx = dy/dx dE/dy?  I think all of the symbols and definitions can be placed in figure 1.

Done. We have removed the equations and their mathematical explanation from the text.

> - The naive reader sould be a touch confused by the reversal of order of z and y from the top right of the figure to the rest of the figure lower down.  I understand that in the top right x, y, and z are strictly generic, whereas they have specific roles as neural network values in the bottom panels.  It would be easy to avoid confusion here by using a, b, and c to refer to generic variables and using x, z, and y to refer to neural network variables.  Of course the neural network variables could be more memorable than x, y, and z -- z could be replaced by n for net input for example.

We suggest that the Nature staff redoes the chain-rule.pdf figure with y and z swapped for one another, to match the other figures. Alternatively, if redoing all the figures, use x for input h for hidden and y for output.

> - The abstract discussion of modules could be related more explicitly to the layers in the neural network.  For example, the network on the bottom of the figure could be decribed explicitly as having three modules, where the role of x for the top module is played by the y's at h2, etc.

Done, in the caption of the backprop figure-box 1.

> - The interpretation of Fig 2 is far from obvious.  I assume that the second column of the figure is the first set of filters, and the third column is the result of pooling/downsampling the first set of filters, but it is not perfectly clear that the column are arranged so that the pooled result corresponds to the unpooled input.  Lines connecting layers appropriately would help (first col goes to all six first-layer arrays, each of these to just one downsampling array -- then I assume each of these goes to all of the arrays at the next later, which in turn each project to the corresponding down-sampled array in the 5th column.  The sixth column appears to be different -- is this convolutional?  Another improvement would be to label the columns 'first filter layer' first pooling layer, etc.

Done, labels added in the figure and caption clarifying as requested the meaning of what is displayed.

> A major issue that needs to be dealt with in revision is the handling of the material related to unsupervised learning.  The paper currently reflects the differing perspectives of the authors on this issue and a greater degree of integration is necessary to avoid the perception that the different parts were written independently and without cognizance of each other.  The story about unsupervised learning can and should be woven into the main narrative of the paper.  Giving credit to the wave of unsupervised learning work of the early part of the new century together with the subsequent breakthroughs in back propagation methods as currenrly described.  In this way the material on pages 15-16 will not seem so starkly out of place as it does at present.

We are not sure to understand the concern; we do talk about unsupervised pre-training in the section on 'overcoming the limitations of backprop'. What differing perspectives?

> In addition to the above comments, I had several minor comments that I provide on an annotated copy of the ms.  These comments relate to literature that I think needs to be cited from early work in the 1980s on neural networks, as well as issues of rhetorical stance.  As written the article does not do justice to the reasons why some were not convinced to use neural networks -- these people had reasons that seemed to make sense, and should not simply be characterized as hostile or biased.  The better way to present the work is that it suggests that these reasons, however legitimate, might have stood in the way of progress, which is possible if these considerations are downweighted.  A more concilliatory tone to past researchers will also minimize future resentments, which will not serve the current authors well when the current approaches hit an inevitable plateau.

We have changed the tone in many places to make it sound more positive.
We rephrased the sentence about "prejudice" to talk in positive terms about the appeal of convexity

> Finally the paper has many typos and infelicities of writing and deserves a careful proof-reading by a person who is careful and graceful in the use of English definite and indefinite articles. I have provided a commented version of the ms that highlights several such issues, though far from all of the ones that need attention in the ms.

We have made a lot of changes in the form and English, to clarify and remove errors.
We have made the other changes suggested by the referee, however, some suggestions
had to be dropped in order to reduce the size of the manuscript and the number of citations.

From Referee 2:
--------------

We have not re-organized the text around the suggested questions, but we have
made sure to address most of them.

> 1) should features be hand-designed vs. learned?

This is central in the paper and we argue about it.

> 2) do optimizations need to be convex?
> The text talks about the local minimum problem, but never mentions the word "convex." This was and is still a big deal in machine learning. Surprisingly,  neural nets suggest that convex optimization is a distraction.

A brief discussion of convexity and saddle points vs local minima has been added.

> 3) should we imitate the brain or not?

This is an interesting topic and we could talk more about it, but we have been
asked to cut 1/4th of the paper, so adding more material seems difficult. There
is a discussion of the inspiration from the visual cortex, for ConvNets.

> 4) symbolic or connectionist?

That question is covered.

> 5)  probabilistic or deterministic?
> People have argued that probabilistic representations are crucial for
> machine learning and AI.  The innards of backprop-trained neural nets
> are fully deterministic.  Isn't that surprising?

Backprop-trained neural nets (a) can compute conditional probabilities (e.g. for machine translation)
and (b) can easily incorporate "injected noise", as in recent work with Variational Auto-Encoders cited
at the end (Kingma et al NIPS 2014). We have not included an explicit discussion about this, though.

> 6)  modular or unstructured?
> Minsky criticized connectionists as naively hoping that intelligence would emerge spontaneously from unstructured networks.  In reality, modern neural nets do have modular structure. A ConvNet has the convolution and pooling layers.  LSTM is another module for recurrent nets. And ConvNets are being used as modules in larger systems.

ML theory clearly states that some form of prior (what you call structure) is necessary, so it's
not a black and white question. We believe that it is more valuable to uncover general-purpose
priors that cover a lot of tasks, and convolutions or LSTMs are of that kind. As the amount
of available training increases, this helps the most the more generic types of structure imposed
on the neural network.

> When the authors compare deep learning with the rest of machine learning, they emphasize support vector machines.  What about Markov random fields and other graphical models?  These terms have been and still are a big deal, yet they don't even appear in the text. Why have they been pushed aside by neural nets?

The space devoted to kernel machines has been mostly removed, in the interest of the requested 
length cuts.
Graphical models are very present in the unsupervised side of deep learning
but much more space would be needed to give justice to that aspect.
Two sentences were added in the middle of this paragraph to clarify this:
"This review has focused on supervised learning because of its recent
achievements and we have ignored much interesting work on unsupervised
learning procedures for deep neural
networks~\citep{Salakhutdinov2009-small,Hinton95,QuocLe-ICML2012,VincentPLarochelleH2008-small,koray-nips-10,gregor-icml-10,ranzato-pami,Bengio-et-al-ICML-2014,Kingma-et-al-NIPS2014}.
These papers often introduce a probabilistic view or the injection of randomness in the
neural network computation, to allow neural networks to {\em generate} examples that are
similar to training examples but generalize them.
Unsupervised learning is important because most data have not been hand-labeled by humans, and just like
children are not told what they should have done at each instant, we
need learning machines that can learn from mostly unlabeled examples."

> On a related note, the authors choose to emphasize the divide between
> deep learning and the rest of machine learning. One could also stress

We have rephrased in a few places to make the tone less confrontational

> the similarities.  A support vector machine is similar to a two-layer
> perceptron.

Indeed, except that its first layer is fixed (the feature space view)
or parametrized directly by the training examples themselves (the kernel function view).
Again we would like to discuss that at greater length but space is heavily limited.

> ConvNets have emerged victorious for visual object recognition, both
> in academic research and in industry.  From the review it's less clear
> whether recurrent nets have reached the same level of success. For
> example, do these nets outperform conventional language models based
> on quantitative metrics like BLEU?  Are recurrent nets used by
> industry in commercial systems?

You are right that although there is enormous enthusiasm 
with recurrent nets, the industrial adoption in products is less well known;
we informally know of several products (e.g. for speech recognition)
based on recurrent nets, but convolutional nets clearly have a head start
(and of course it's not a competition, they make sense in different contexts
and have been combined, e.g., in the recent work on caption generation from images).

> Mnih et al. has appeared in Nature, so perhaps that should be cited rather than the tech report.

Done

> Tesauro's work on TD-gammon could be cited as an important precedent.

Done

> Another bit of history: the earlier demonstrations that GPU implementations of neural nets could beat other machine learning techniques on the MNIST dataset were perhaps an important precursor to the ImageNet-winning ConvNet.

Yes, cited.

From Referee 3:
--------------

> The first one is that the article comes across as every so slightly "whiny" (this might not be the right English term) - it sounds like the authors feel upset that their methods did not get the recognition they deserved. When I entered the field, (multi-layer) neural networks were actually the standard method on many tasks - so I think that impression probably only refers to the local minimum in between the current waves and the wave in the early nineties. Maybe it need not be discussed in as much detail..

The tone has been changed to be less confrontational.

> The second one is that many of the impressive features and successes of deep learning applications, as described in the paper, I would really view more generally as successes of applying high capacity models to huge datasets - i.e., they are the result of taking a rich and well-designed class of models and training it on a huge training set. The current deep learning hype may wane again one day, but whatever we will be using next will probably also involve data-driven rich models. To me, that is the underlying long term trend - but again, this is my personal opinion, and it is fine if the review expresses the subjective point of view of the authors.

Ou opinion is different: yes high capacity flexible models trained on lots of data is an important ingredient,
but by far it is insufficient. Consider for example the fate of decision trees. They can be trained quickly,
they can efficiently handle very high capacity much better than neural nets, but they don't generalize well.
Deep learning is motivated by ideas about {\em non-local} generalization that is possible thanks to the
compositional effects arising out of (a) distributed representations (composition of features) and (b) depth
(composition of layers).

> Section 3: I was surprized you're not mentioning the work of Bruna.

Which Bruna do you refer to?

> page 3: the linear support vector machine is much older and was developed by Vapnik and Chervonenkis

We wanted to add the 1964 reference but we have been asked to cut citations, so it's going to
be up to the editor to decide.

> page 3: kernel methods: I think Boser et al. is a good reference for SVMs, but for Kernel Methods in general (including the general kernel trick, and the fact that kernels can be defined on non-vectorial data, etc.), it would be better to cite the "Learning with Kernels" book, or the Cristianini book, or - if you prefer original articles - the kernel PCA paper.

Done, we are citing the "Learning with Kernels" book.

> page 12: "Thanks to advances..." - it would be interesting to know which advances you mean.

We added references for both the architecture advances and the training advances

> - I think section 5 is a bit tedious. Some of it will be lost on the readers. The points made in the last paragraph are very good though and should certainly be made somewhere. The section is also somewhat oddly placed in between the sections on image understanding and language processing.

The old Section 5 has been reduced (note that section numbers have changed to due reorganizations
to cut pieces out).

### 2015-03-30 deepreview.tex
as sent to Tanguy today

@@
 sentiment
-analysis, %~\citep{Glorot+al-ICML-2011-small}, % YB cutting self-citations
+analysis~\citep{Glorot+al-ICML-2011-small}, 
 question-answering~\cite{Bordes-et-al-EMNLP2014}, and language
@@
 % Chinese~\citep{abdulkader-iwfhr-06,chellapilla-iwfhr-06a}. 
-% ConvNets
-% were also experimented with in the early 1990s for object detection 
-% in natural images, including faces and
-% hands~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95} and for face
-% recognition~\cite{lawrence-tnn-1997}.
+ConvNets
+were also experimented with in the early 1990s for object detection 
+in natural images, including faces and
+hands~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95} and for face
+recognition~\cite{lawrence-tnn-1997}.
 
@@
 tasks in which labeled data was relatively abundant, such as traffic
-sign recognition~\cite{sermanet-ijcnn-11,Ciresan-et-al-2012}, the
+sign recognition~\cite{Ciresan-et-al-2012}, the
 segmentation of biological images~\cite{ning-05} particularly for
@@
 pedestrians, and human bodies in natural
-images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,nasse-09,fan-pami-2010,sermanet-cvpr-13,tompson-cvpr-15}.
+images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,sermanet-cvpr-13,tompson-cvpr-15}.
 A major practical success of ConvNets in recent times is face

### 2015-04-23 deepreview.tex
put back the TDNN citation
Geoff's new version with couple of typos fixed

@@
 these parts. The key aspect of deep learning is that these layers of
-features are not designed by human engineers: {\em They are learned from
-data using a general purpose learning procedure.}
+features are not designed by human engineers: They are learned from
+data using a general purpose learning procedure.
 
@@
 for many years. It has turned out to be very good at discovering
-intricate structures in large, high-dimensional data and is therefore
+intricate structures in high-dimensional data and is therefore
 applicable to many domains of science, business, and government. In
@@
 expression and disease~\citep{Keung-et-al-2014,Xiong-et-al-Frey-science2015}.
-
 Perhaps more surprisingly, deep learning has produced extremely
@@
 translation~\cite{Jean-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
+
+% our conclusions (which Nature insists on having at the end of the intro).
+We believe that deep learning will have many more successes in the
+near future because it requires very little hand-engineering so it can
+easily take advantage of increases in the amount of available
+computation and data. New learning algorithms and new architectures
+that are currently being developed for deep neural networks will only
+accelerate this progress.
+
+\section{Supervised Learning}
 
@@
 % shallow linear classifiers
-Many practical applications of ML use linear classifiers on top of
+Many of the current practical applications of ML use linear classifiers on top of
 hand-engineered features. A two-class linear classifier computes a
@@
 % kernel function is ineffective. 
-Designing good feature extractors traditionally required considerable amounts
-of engineering skills and domain expertise. But this can all be avoided if
+Designing good feature extractors traditionally required a considerable amount
+of engineering skill and domain expertise. But this can all be avoided if
 good features can be learned automatically using a general purpose
@@
 % backprop
-How do we train multilayer architectures?
+\section{Backpropagation: a simple way to train multilayer architectures}
 
@@
 replace hand-engineered features by trainable multilayer networks,
-but despite its extreme simplicity, the solution was not widely
+but despite its simplicity, the solution was not widely
 understood until
-the mid 1980s and was largely forsaken by the ML community until
-recently. As it turns out, training multilayer architectures can be
-done by simple stochastic gradient descent. As long as the modules are
+the mid 1980s.  
+% and was largely forsaken by the ML community until recently.
+As it turns out, multilayer architectures can be
+trained by simple stochastic gradient descent. As long as the modules are
 relatively smooth functions of their inputs and of their internal
@@
   procedure}.
-
 The idea that this could be done, and that it worked, was discovered independently by
 several different groups in the 1970s and
-1980s~\citep{Werbos74,Parker85,LeCun85,RHW}. It spurred a wave of interest for
-neural networks, becoming quite popular in the late 1980s, but falling out of favor
-between the mid 1990s until their renaissance in the late 2000s.
-
-The backpropagtion procedure to compute gradient of an objective
+1980s~\citep{Werbos74,Parker85,LeCun85,RHW}. 
+% It spurred a wave of interest for
+% neural networks, becoming quite popular in the late 1980s, but falling out of favor
+% between the mid 1990s until their renaissance in the late 2000s.
+
+The backpropagation procedure to compute the gradient of an objective
 function with respect to the weights of a multilayer stack of modules
@@
 learn to map a fixed-size input ({\it e.g.} an image) to a fixed-size
-output ({\it e.g.} a probability for each one of several categories). To go
-from one layer to the next, a set of units compute a weighted
-sum of their inputs from the previous layer and pass the result
-through a non-linear function. The most common non-linear function is
-the so-called {\em ReLU} (Rectified Linear Unit) which is simply the
-half-wave rectifier $f(z) = \max(z,0)$. In past decades, neural nets
-used smoother non-linearities, such as $\tanh(z)$ or $1/(1+\exp(-z))$ but the ReLU 
-typically learns much faster in networks with many layers. Units that are not in the
-output layer are traditionally called {\em hidden units}. The early
-layers can be seen as distorting the input in a non-linear way so that
-categories become linearly separable by the last layer, as indicated
-in figure~\ref{fig:backprop-box}.
+output ({\it e.g.} a probability for each of several categories). To
+go from one layer to the next, a set of units compute a weighted sum
+of their inputs from the previous layer and pass the result through a
+non-linear function. Currently, the most popular non-linear function
+is the rectified linear unit ({\em ReLU}) which is simply
+the half-wave rectifier $f(z) = \max(z,0)$. In past decades, neural
+nets used smoother non-linearities, such as $\tanh(z)$ or
+$1/(1+\exp(-z))$ but the ReLU typically learns much faster in networks
+with many layers. Units that are not in the input or output layer are
+traditionally called {\em hidden units}. The hidden layers can be seen
+as distorting the input in a non-linear way so that categories become
+linearly separable by the last layer, as indicated in
+figure~\ref{fig:backprop-box}.
 
@@
 
-\section{Convolutional networks}
-
-One of the most widely used deep learning architectures is the
-convolutional neural network, or
-ConvNet~\cite{lecun-90c,lecun-98}. ConvNets are designed to process
+% the reduced history section
+
+In the late 1990s, neural nets and backpropagation were largely
+forsaken by the machine learning community and ignored by the computer
+vision and speech recognition communities.  It was widely believed
+that learning useful, multi-stage, feature extractors with little prior knowledge 
+was infeasible. In particular, it was commonly
+believed that simple gradient descent would get trapped in poor
+local minima -- weight configurations for which no small change would
+reduce the average error.
+%We now know
+%that the main reasons for the limited success of backpropagation in
+%deep networks with dense connectivity were that the labeled datasets were typically much too
+%small and the computers were much too slow. Additional reasons were
+%that researchers were not initializing the weights properly and were
+%not using the type of non-linearlity that learns most
+%rapidly~\citep{lecun-98b,GlorotAISTATS2010-small}.
+
+In practice, poor local minima are rarely a problem with large networks. Regardless
+of the initial conditions, the system nearly always reaches solutions
+of very similar quality.  Recent results strongly suggest the
+landscape is packed with a combinatorially large number of {\em saddle
+  points} where the gradient is zero, and the surface curves up in
+most dimensions and curves down in the 
+remainder~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
+The analysis seems to show that saddle points with
+only a few downward curving directions are present in very large
+numbers, but almost all of them have very similar values of the
+objective function.  Hence it doesn't much matter which of these
+saddle points the algorithm gets stuck at.
+%% Avoiding being trapped near saddle points seems more important.
+
+%% It also took an inordinate amount of time for the community to adopt
+%% stochastic gradient methods for large-scale machine learning. Key
+%% theoretical and practical work showed that, while SGD has terrible
+%% asymptotic convergence properties, it is extremely efficient for the
+%% kind of approximate optimization required in ML
+%% applications~\citep{bottou-bousquet-2008-small}. But when training a
+%% large ConvNet can take days or weeks on the fastest GPUs, one is eager
+%% to discover more efficient optimization methods that use the curvature
+%% information the past update directions~\citep{lecun-98b,Montavon2012}.
+
+Interest in deep feedforward networks was revived around
+2006~\citep{IJCAI,Hinton06-small,Bengio-nips-2006-small,ranzato-07-small}
+by a group of researchers brought together by the Canadian Institute
+for Advanced Research who introduced
+unsupervised learning procedures that could
+create layers of feature detectors without requiring labeled data. The
+objective in learning each layer of feature detectors was to be able
+to reconstruct or model the activities of feature detectors (or raw
+inputs) in the layer below.  By ``pre-training'' several layers of
+progressively more complex feature detectors using this reconstruction
+objective, the weights of a deep network could be initialized to
+sensible values.  A final layer of output units could then be added to
+the top of the network and the whole deep system could be fine-tuned
+using standard backpropagation
+\citep{Hinton-Science2006,Bengio-nips-2006-small,ranzato-07-small}.
+This worked remarkably well for recognizing handwritten digits or for
+detecting pedestrians, especially when the amount of labeled data was
+very limited~\citep{sermanet-cvpr-13}.
+
+% The parallelization of large neural nets over a single GPU has
+% required to use ``mini-batches'' of samples over which the gradients
+% are averaged. The purpose of this is to saturate the GPU. 
+% But one big
+% unsolved question is the best way to scale up the training of very large neural
+% nets over multiple GPU cards across multiple machines. As of today, this
+% question is still unresolved.
+
+The first major application of this pre-training approach was in
+speech recognition and it was made possible by the advent of fast
+Graphics Processing Units (GPUs) that were convenient to
+program~\citep{RainaICML09-small} and allowed researchers to train
+networks 10 or 20 times faster.  In 2009, the approach was used to map
+short temporal windows of coefficients extracted from a soundwave to a
+set of probabilities for the various fragments of speech that might be
+represented by the frame in the center of the window.  It achieved
+record-breaking results on a standard speech recognition benchmark
+that used a small vocabulary~\citep{TIMITpaper} and was quickly
+developed to give record-breaking results on a large vocabulary
+task~\citep{Dahl2012}.  By 2012, versions of the deep net from 2009
+were being developed by many of the major speech
+groups~\citep{Hinton-et-al-2012} and were already being deployed in
+Android phones.  For smaller datasets, unsupervised pre-training helps
+to prevent overfitting~\citep{Bengio-Courville-Vincent-TPAMI2013},
+leading to significantly better generalization when the number of
+labeled examples is small, or in a transfer setting where we have lots
+of examples for some ``source'' tasks but very few for some ``target''
+task.  Once deep learning had been rehabilitated, it turned out the
+pre-training stage was only needed for small datasets.
+
+There was, however, one particular type of deep, feed-forward network
+that was much easier to train and generalized much better than
+networks with full connectivity between adjacent layers. This was the
+convolutional neural network~\cite{lecun-90c,lecun-98}. It achieved
+many practical successes during the period when neural networks were
+out of favor and it has recently been widely adopted by the computer
+vision community.
+
+\section{Convolutional neural networks}
+
+Convolutional neural networks (ConvNets) are designed to process
 data that come in the form of multiple arrays, for example a color
@@
 learning algorithm such as backpropagation. 
-% A primitive one-dimensional convnet called a time-delay
-% neural net was used for the recognition of phonemes and simple 
-% words~\cite{WaibelIEEE,bottou-89}
+A primitive one-dimensional convnet called a time-delay
+neural net was used for the recognition of phonemes and simple 
+words~\cite{WaibelIEEE,bottou-89}
 
@@
 back to the early 1990s, starting with time-delay neural networks for
-speech recognition and document reading~\citep{lecun-98}. 
+speech recognition~\citep{TDNN} and document reading~\citep{lecun-98}. 
 The document reading system used a ConvNet trained jointly with
@@
 %sequential steps.
-Deep learning theory shows that deep nets can benefit from
-an exponential advantage over classical learning algorithm not based on distributed 
-representations~\citep{Bengio-localfailure-NIPS-2006-small},
- in two ways, both arising from the power of {\em composition}, so long as the
-underlying data generating distribution has such a componential structure, i.e.,
-is generated by a hierarchy of composition of factors~\citep{Bengio-Courville-Vincent-TPAMI2013}.
-First, learning distributed representations enables generalization to new
-combinations of the learned features (e.g. $2^n$ combinations, with $n$ features)
-~\citep{Bengio-2009-book,Montufar+Morton-2014}. 
-Second composing layers in a deep net brings the potential for another exponential
-advantage~\citep{Montufar-et-al-NIPS2014} (exponential in the depth).
+Deep learning theory shows that deep nets have two different
+exponential advantages over classical learning algorithms that do not
+use distributed
+representations~\citep{Bengio-localfailure-NIPS-2006-small}. Both of
+these advantages arise from the power of {\em composition} and depend
+on the underlying data generating distribution having an appropriate
+componential structure~\citep{Bengio-Courville-Vincent-TPAMI2013}.
+First, learning distributed representations enables generalization to
+new combinations of the learned features (e.g. $2^n$ combinations,
+with $n$ features) ~\citep{Bengio-2009-book,Montufar+Morton-2014}.
+Second composing layers in a deep net brings the potential for another
+exponential advantage~\citep{Montufar-et-al-NIPS2014} (exponential in
+the depth).
 
@@
 a probability distribution that depends on the English sentence.
-
 This rather naive way of performing machine translation has quickly become
@@
 expressions that are manipulated by using inference rules.
+
 %%binding their symbols to variables
@@
 Beyond simple memorization, Neural Turing Machines and Memory Networks
-are being used for task that would normally require to manipulate
-symbols and to reason.
+are being used for tasks that would normally require reasoning and symbol manipulation.
 Neural Turing Machines can be taught ``algorithms''. Among other
-things, they can learn to ouput a sorted list of symbols when its
+things, they can learn to output a sorted list of symbols when their
 input consists of an unsorted sequence in which each symbol is
@@
 Memory Networks can be trained to keep track of the state of the world
-in a setting similar to a text adventure game. After reading a story,
+in a setting similar to a text adventure game and after reading a story,
 they can answer questions that require complex 
@@
 
-\section{Historical perspective on backpropagation}
-
-In the late 1990s, neural nets and backpropagation were largely
-forsaken by the machine learning community and ignored by the computer
-vision and speech recognition communities.  It was widely believed
-that learning multi-stage feature extractors with little prior knowledge 
-was neither useful nor effective. In particular, it was commonly
-believed that simple gradient descent procedure get trapped in poor
-local minima -- weight configurations for which no small change would
-reduce the average error-- despite considerable empirical evidence to
-the contrary in the case of ConvNets~\citep{lecun-98}. We now know
-that the main reasons for the limited success of backpropagation in
-deep networks were that the labeled datasets were typically much too
-small and the computers were much too slow. Additional reasons were
-that researchers were not initializing the weights properly and were
-not using the type of non-linearlity that learns most
-rapidly~\citep{lecun-98b,GlorotAISTATS2010-small}.
-
-Poor local minima are rarely a problem with large networks. Regardless
-of the initial conditions, the system nearly always reaches solutions
-of very similar quality.  Recent results strongly suggest the
-landscape is packed with a combinatorially large number of {\em saddle
-  points} where the gradient is zero, and the surface curves up in
-some dimensions and curves down in
-others~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
-The analysis seems to show that local minima (and saddle points with
-only a few downward curving directions) are present in very large
-numbers, but almost all of them have very similar values of the
-objective function.  Hence it doesn't seem to matter which local
-minimum the algorithm finds.
-%% Avoiding being trapped near saddle points seems more important.
-
-%% It also took an inordinate amount of time for the community to adopt
-%% stochastic gradient methods for large-scale machine learning. Key
-%% theoretical and practical work showed that, while SGD has terrible
-%% asymptotic convergence properties, it is extremely efficient for the
-%% kind of approximate optimization required in ML
-%% applications~\citep{bottou-bousquet-2008-small}. But when training a
-%% large ConvNet can take days or weeks on the fastest GPUs, one is eager
-%% to discover more efficient optimization methods that use the curvature
-%% information the past update directions~\citep{lecun-98b,Montavon2012}.
-
-Interest in deep feedforward networks was revived around
-2006~\citep{IJCAI,Hinton06-small,Bengio-nips-2006-small,ranzato-07-small}
-by a group of researchers brought together by the Canadian Institute
-for Advanced Research who introduced
-unsupervised learning procedures that could
-create layers of feature detectors without requiring labeled data. The
-objective in learning each layer of feature detectors was to be able
-to reconstruct or model the activities of feature detectors (or raw
-inputs) in the layer below.  By ``pre-training'' several layers of
-progressively more complex feature detectors using this reconstruction
-objective, the weights of a deep network could be initialized to
-sensible values.  A final layer of output units could then be added to
-the top of the network and the whole deep system could be fine-tuned
-using standard backpropagation
-\citep{Hinton-Science2006,Bengio-nips-2006-small,ranzato-07-small}.
-This worked remarkably well for recognizing handwritten digits or for
-detecting pedestrians, especially when the amount of labeled data was
-very limited~\citep{sermanet-cvpr-13}.
-
-% The parallelization of large neural nets over a single GPU has
-% required to use ``mini-batches'' of samples over which the gradients
-% are averaged. The purpose of this is to saturate the GPU. 
-% But one big
-% unsolved question is the best way to scale up the training of very large neural
-% nets over multiple GPU cards across multiple machines. As of today, this
-% question is still unresolved.
-
-The first major application of this pre-training approach was in
-speech recognition and it was made possible by the advent of fast
-Graphics Processing Units (GPUs) that were convenient to
-program~\citep{RainaICML09-small} and allowed researchers to train
-networks 10 or 20 times faster.  In 2009, the approach was used to map
-short temporal windows of coefficients extracted from a soundwave to a
-set of probabilities for the various fragments of speech that might be
-represented by the frame in the center of the window.  It achieved
-record-breaking results on a standard speech recognition benchmark
-that used a small vocabulary~\citep{TIMITpaper} and was quickly
-developed to give record-breaking results on a large vocabulary
-task~\citep{Dahl2012}.  By 2012, versions of the deep net from 2009
-were being developed by many of the major speech
-groups~\citep{Hinton-et-al-2012} and were already being deployed in
-Android phones.  For smaller datasets, unsupervised pre-training helps
-to prevent overfitting~\citep{Bengio-Courville-Vincent-TPAMI2013},
-leading to significantly better generalization when the number of
-labeled examples is small, or in a transfer setting where we have lots
-of examples for some ``source'' tasks but very few for some ``target''
-task. In this context, unsupervised representation learning helped to
-win several transfer learning
-competitions~\citep{Bengio-Courville-Vincent-TPAMI2013}.  Once deep
-learning had been rehabilitated, it turned out that large labeled
-datasets did not need the pre-training stage. Even better results,
-especially for very deep networks, could be obtained by using
-rectifying non-linearities rather than sigmoid or tanh non-linearities
-~\citep{jarrett-iccv-09,Nair-2010-small,Glorot+al-AI-2011-small,Krizhevsky-2012-small,Dahl-et-al-ICASSP2013}. Deep
-neural nets are now the dominant method for acoustic modeling in
-speech, and ConvNet are dominant for image and video recognition.
-
 \section{The future of deep learning}
 
-Human vision is an active process in which the few degrees around the
-center of the visual field that is processed at high resolution is
-intelligently moved around through saccades.
-
-We expect much of the future progress in vision to come from systems
-that are trained end-to-end and combine ConvNets with RNNs that decide
-where to look through reinforcement learning. Such systems already
-outperform passive vision systems~\citep{ba+mnih} at classification
-tasks, but much more remains to be explored.  Combining reinforcement
-learning with ConvNets has already produced impressive results in
-learning to play many different video games~\citep{Mnih-et-al-2015}.
-Similarly, when understanding sentences or whole documents we expect
-future systems to learn strategies for selectively attending to one
-part at a time~\citep{Bahdanau-et-al-ICLR2015-small}.
-
-Certainly, natural language understanding is an area in which deep
-learning is poised to make a large impact over the next few years,
-using combinations of RNN, ConvNets, reasoning, and unsupervised
-learning.
-
-This review has focused on supervised learning because of its recent
-achievements and we have ignored much interesting work on unsupervised
-learning procedures for deep neural
-networks~\citep{Salakhutdinov2009-small,Hinton95,VincentPLarochelleH2008-small,koray-nips-10,gregor-icml-10,ranzato-pami,Bengio-et-al-ICML-2014,Kingma-et-al-NIPS2014}.
-Human and animal learning is largely unsupervised: we discover the
+Unsupervised learning~\citep{Salakhutdinov2009-small,Hinton95,VincentPLarochelleH2008-small,
+koray-nips-10,gregor-icml-10,ranzato-pami,Bengio-et-al-ICML-2014,Kingma-et-al-NIPS2014} 
+had a catalytic effect in reviving interest in
+deep learning but has since been overshadowed by the successes of
+purely supervised learning. Even though we have not focused on it
+here, we expect unsupervised learning to become far more important in
+the longer term.  Human and animal learning is largely unsupervised: we discover the
 structure of the world by observing it, not by being told the name of
-every object. Most current methods are built around the idea of {\em
-  reconstruction} of the input, and can sometimes be used to {\em
-  generate} inputs that are similar to training examples. The
-principle is to disentangle the independent factors that interact to
-produce images, speech and
-documents~\citep{Bengio-Courville-Vincent-TPAMI2013}.  But some
-researchers view the essence of intelligence as the ability to predict
-the future, leading them to propose new paradigms based on
-{\em prediction}.
-
-%% YLC: I uncommented this. I really think this is were things are going.
-Ultimately, progress in AI will come about through systems that
+every object.
+
+Human vision is an active process which sequentially samples the optic
+array in an intelligent, task-specific way using a small,
+high-resolution fovea with a large, low-resolution surround. We expect
+much of the future progress in vision to come from systems that are
+trained end-to-end and combine ConvNets with RNNs that use
+reinforcement learning to decide where to look. Such systems are in
+their infancy but they already outperform passive vision
+systems~\citep{ba+mnih} at classification tasks and produce impressive
+results in learning to play many different video
+games~\citep{Mnih-et-al-2015}.
+
+Natural language understanding is another area in which deep learning
+is poised to make a large impact over the next few years and we expect
+systems that use RNNs to understand sentences or whole documents to learn
+strategies for selectively attending to one part at a
+time~\citep{Bahdanau-et-al-ICLR2015-small,Xu-et-al-arxiv2015}.
+
+Ultimately, major progress in AI will come about through systems that
 combine representation learning with complex {\em reasoning}. While
-deep learning and simple resoning has been used for speech and
+deep learning and simple reasoning have been used for speech and
 handwriting recognition for a long time, new paradigms are needed to
-replace symbolic logic manipulation by operations on
-vectors~\citep{bottou-mlj-2014}.
-
-With new paradigms for unsupervised learning and reasoning on the
-horizon, and with new learning algorithms that transcend simple
-gradient descent, deep learning is poised to progress quickly over the
-next few years, as the research community grows, the machines become
-more powerful, and the datasets become larger.
+replace rule-based manipulation of symbolic expressions by operations on
+large vectors~\citep{bottou-mlj-2014}.
 

### 2015-04-23 deepreview.tex
fixed the TDNN citation into the existing Waibel89b

@@
 back to the early 1990s, starting with time-delay neural networks for
-speech recognition~\citep{TDNN} and document reading~\citep{lecun-98}. 
+speech recognition~\citep{Waibel89b} and document reading~\citep{lecun-98}. 
 The document reading system used a ConvNet trained jointly with

### 2015-04-23 deepreview.tex
cleaning up references and other minor changes

@@
 % kernel function is ineffective. 
-Designing good feature extractors traditionally required a considerable amount
+To make classifiers more powerful, one can use generic non-linear features,
+as with kernel methods~\citep{Scholkopf+Smola-2002book}, but generic
+features such as those arising with the Gaussian kernel do not allow the learner to
+generalize well far from the training examples~\citep{Bengio-localfailure-NIPS-2006-small}.
+The traditional option is to design good feature extractors, which requires a considerable amount
 of engineering skill and domain expertise. But this can all be avoided if
@@
 by a group of researchers brought together by the Canadian Institute
-for Advanced Research who introduced
+for Advanced Research (CIFAR) who introduced
 unsupervised learning procedures that could
@@
 neural net was used for the recognition of phonemes and simple 
-words~\cite{WaibelIEEE,bottou-89}
+words~\cite{Waibel89b,bottou-89}
 
@@
 were later deployed by
-Microsoft~\citep{simard-03}.
+Microsoft~\citep{simard-03-small}.
 % including for Arabic and
@@
 A major practical success of ConvNets in recent times is face
-recognition~\citep{taigman-cvpr-2014}.
+recognition~\citep{Taigman-et-al-CVPR2014}.
 
@@
 approach for almost all recognition and detection
-tasks~\citep{sermanet-iclr-14,girshick-cvpr-2014,taigman-cvpr-2014,simonyan-arxiv-2014,szegedy-2014,tompson-cvpr-15}
+tasks~\citep{sermanet-iclr-14,girshick-cvpr-2014,Taigman-et-al-CVPR2014,simonyan-arxiv-2014,szegedy-2014,tompson-cvpr-15}
 and approach human performance on some tasks. A recent stunning
@@
 \fbox{
-\begin{minipage}{\textwidth}
+\begin{minipage}{0.9\textwidth}
 \centerline{\includegraphics[width=\linewidth]{caption-generation-Vinyals-et-al.png}}
@@
 \fbox{
-\begin{minipage}{\textwidth}
+\begin{minipage}{0.5\textwidth}
 %\centerline{
@@
 \centerline{
-\includegraphics[width=0.89\linewidth]{fig-hidden-recurrence-rnn.pdf}
+\includegraphics[width=0.9\linewidth]{fig-hidden-recurrence-rnn.pdf}
 }
@@
 
-Unsupervised learning~\citep{Salakhutdinov2009-small,Hinton95,VincentPLarochelleH2008-small,
+Unsupervised learning~\citep{Hinton95,Salakhutdinov2009-small,VincentPLarochelleH2008-small,
 koray-nips-10,gregor-icml-10,ranzato-pami,Bengio-et-al-ICML-2014,Kingma-et-al-NIPS2014} 

### 2015-04-24 deepreview.tex
more changes to respond to comments from Mikolov and Ranzato

@@
 \usepackage{graphicx}
-\usepackage[numbers]{natbib}
+%\usepackage[numbers]{natbib}
 \usepackage{hyperref}
@@
 \newcommand{\deri}[2]{\frac{\partial  #1}{\partial #2}}
-
+\newcommand{\citep}[1]{\cite{#1}}
+\newcommand{\citet}[1]{\cite{#1}}
 \parskip 0.085in
@@
 
-Deep Learning allows machines that are composed of multiple processing 
+Deep Learning allows computational models that are composed of multiple processing 
 layers to learn representions of data with multiple levels of abstraction.
 These methods have dramatically improved the state-of-the-art
-in speech recognition, object recognition, object detection,
+in speech recognition, visual object recognition, object detection,
 and many other domains such as drug discovery and genomics. Deep
 learning discovers intricate structure in large datasets by using the
-{\em backpropagation algorithm} to indicate how the machine should
-change the weights that are used to compute the representation in each layer from
+{\em backpropagation algorithm} to indicate how a machine should
+change its internal parameters that are used to compute the representation in each layer from
 the representation in the previous layer.  Deep {\em convolutional nets}
@@
 select relevant results of search. Increasingly, these applications
-make use of a new class of techniques called {\em deep learning}.
+make use of a class of techniques called {\em deep learning}.
 
@@
 % deep learning: learning representations
-Deep learning is a set of methods that allows a machine to be fed with
+Representation learning is a set of methods that allows a machine to be fed with
 raw data and to automatically discover the representations needed for
-detection or classification. The name comes from the fact that deep learning machines are
-multilayer stacks of trainable modules, each of which turns the
-representation of the data on its input into a more abstract and
-higher-level representation on its output. The output representation
+detection or classification. Deep learning methods are representation
+learning methods with multiple levels of representation, obtained
+by composing simple but non-linear modules that each transform the representation at one level
+(starting with the raw input) into a representation at a higher, slightly more 
+abstract level. With the composition of enough such transformations, very complex functions
+can be learned. For classification tasks, the output representation
 amplifies aspects of the input that are important for discrimination
@@
 and speech
-recognition~\citep{Hinton-et-al-2012,Sainath-et-al-ICASSP2013}, it has
+recognition~\citep{Mikolov-ASRU-2011,Hinton-et-al-2012,Sainath-et-al-ICASSP2013}, it has
 beaten other ML techniques at predicting the activity of potential
@@
 gradients through all modules, starting from the output at the top 
-all the way to the bottom. Once these
+(where the network produces its prediction) all the way to the bottom
+(where the external input is fed). Once these
 gradients have been computed, it is straightforward to compute the gradients 
@@
 
-ConvNets exploit the property that many natural
-signals are {\em compositional hierarchies}. In images, local
+Deep neural networks exploit the property that many natural
+signals are {\em compositional hierarchies}, where higher-level
+features are obtained by composing lower-level ones. In images, local
 combinations of edges form motifs, motifs assemble into parts, and
@@
 approaches~\citep{Krizhevsky-2012-small}.  This success came from
-efficient use of GPUs, ReLUs, and a new regularization technique
-called ``dropout'' ~\citep{Srivastava14} which prevented overfitting
-in the fully connected layers.  This success has brought about a
+efficient use of GPUs, ReLUs, a new regularization technique
+called ``dropout'' ~\citep{Srivastava14}, and techniques to
+generate more training examples by deforming the existing ones.
+This success has brought about a
 revolution in computer vision, and ConvNets are now the dominant
@@
 
+%representations allow a lot of parallel computation at the feature
+%level so that complex tasks can be solved in a very small number of
+%sequential steps.
+Deep learning theory shows that deep nets have two different
+exponential advantages over classical learning algorithms that do not
+use distributed
+representations~\citep{Bengio-localfailure-NIPS-2006-small}. Both of
+these advantages arise from the power of {\em composition} and depend
+on the underlying data generating distribution having an appropriate
+componential structure~\citep{Bengio-Courville-Vincent-TPAMI2013}.
+First, learning distributed representations enables generalization to
+new combinations of the values learned features beyond those seen
+during training (e.g. $2^n$ combinations
+are possible with $n$ binary features) ~\citep{Bengio-2009-book,Montufar+Morton-2014}.
+Second composing layers of representation in a deep net brings the potential for another
+exponential advantage~\citep{Montufar-et-al-NIPS2014} (exponential in
+the depth).
+
 The hidden layers of a multi-layer neural network learn to represent
@@
 In the first layer, each word
-creates a different pattern of activations, or word-vector, illustrated
+creates a different pattern of activations, or word-vector, a distributed
+representation illustrated
 in Figure~\ref{fig:word-embeddings}.  In a language model, the other
@@
 Tuesday and Wednesday are very similar, as are the word vectors for
-Vietnam and Iraq. 
+Sweden and Norway.
 %% This is not to say that Tuesday and Wednesday are
@@
 reasoning.
-
-%representations allow a lot of parallel computation at the feature
-%level so that complex tasks can be solved in a very small number of
-%sequential steps.
-Deep learning theory shows that deep nets have two different
-exponential advantages over classical learning algorithms that do not
-use distributed
-representations~\citep{Bengio-localfailure-NIPS-2006-small}. Both of
-these advantages arise from the power of {\em composition} and depend
-on the underlying data generating distribution having an appropriate
-componential structure~\citep{Bengio-Courville-Vincent-TPAMI2013}.
-First, learning distributed representations enables generalization to
-new combinations of the learned features (e.g. $2^n$ combinations,
-with $n$ features) ~\citep{Bengio-2009-book,Montufar+Morton-2014}.
-Second composing layers in a deep net brings the potential for another
-exponential advantage~\citep{Montufar-et-al-NIPS2014} (exponential in
-the depth).
 
@@
 network is augmented by a ``tape-like'' memory that the RNN
-can chose to read from or write to~\citep{Graves-et-al-arxiv2014}, and
+can choose to read from or write to~\citep{Graves-et-al-arxiv2014}, and
 ``Memory Networks'' in which a regular network is augmented by a
@@
 \bibliography{strings,ml,aigaion,convnetyann,lecun}
-\bibliographystyle{natbib}
+%\bibliographystyle{natbib}
+\bibliographystyle{ieeetr}
 

### 2015-04-25 deepreview.tex
few more fixes, plus new figure for phrase embeddings

@@
 to $z$ to get the output of the unit.  For simplicity, we have omitted
-bias terms. The non-linear function used in neural networks include 
+bias terms. The non-linear functions used in neural networks include 
 the rectified linear unit (ReLU) $f(z) = max(0, z)$, commonly-used in 
@@
 of a typical convolutional network architecture
-applied on the image of a handwritten digit (left). Each rectangular
+applied to the image of a handwritten digit (left). Each rectangular
 image is a feature map, corresponding to the output for one of the learned features 
@@
 \includegraphics[width=0.49\linewidth]{word-embeddings.png}
-\includegraphics[width=0.49\linewidth]{phrase_zoom2.png}
+\includegraphics[width=0.49\linewidth]{phrase-embeddings.pdf}
 }
@@
 expressions that are manipulated by using inference rules.
-
+%
 %%binding their symbols to variables
@@
 (e.g., the log-probability of generating the right sequence of
-words) with respect to all the states $s_t$ and all the parameters.
+outputs) with respect to all the states $s_t$ and all the parameters.
 }

### 2015-09-23 deepreview.tex
final small changes

@@
 sentiment
-analysis~\citep{Glorot+al-ICML-2011-small}, 
+analysis,% cutting citations ~\citep{Glorot+al-ICML-2011-small}, 
 question-answering~\cite{Bordes-et-al-EMNLP2014}, and language
-translation~\cite{Jean-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
+translation~\cite{Jean-et-al-ACL2015,Sutskever-et-al-NIPS2014}.
 
@@
 $1/(1+\exp(-z))$ but the ReLU typically learns much faster in networks
-with many layers. Units that are not in the input or output layer are
+with many layers, allowing to train a deep supervised network
+without unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
+Units that are not in the input or output layer are
 traditionally called {\em hidden units}. The hidden layers can be seen

### 2015-09-23 Deep-Learning-Nature-Review.tex

@@
 \usepackage{graphicx}
-\usepackage{xspace}
-\usepackage{tabularx}
-\usepackage{algorithm}
-\usepackage{algorithmic}
+
+\addtolength{\textwidth}{2cm}
+%\addtolength{\textwidth}{1cm}
+\addtolength{\textheight}{2cm}
+%\addtolength{\textheight}{1cm}
+\addtolength{\topmargin}{-1cm}
+\addtolength{\oddsidemargin}{-1cm}
+\addtolength{\evensidemargin}{-1cm}
 
@@
 distribution of sequences of words (sentences from a corpus of text). We
-have basically the composition two components: (1) mapping each of the
+have basically the composition of two components: (1) mapping each of the
 previous words in the sentence to their representation, and (2) mapping the
@@
 between the representation for ``king'' and the representation for ``queen'' is
-very close the difference between the representation for ``man'' and the
+very close to the difference between the representation for ``man'' and the
 representation for ``woman'' \citep{Mikolov-et-al-ICLR2013}. The system has discovered
@@
 recurrent network can also have outputs, corresponding to actions or
-predictions, and the training objective for a recurrent networks is
+predictions, and the training objective for a recurrent network is
 obtained from these values. For example, a recurrent network for language
@@
 
-\begin{verbatim}
+{\em
 We will explain the breakthrough that occurred in 2006 when an effective
  new way of learning multiple layers of features was discovered. The idea
- is to learn the layers one a time with the objective being to build a good
+ is to learn the layers one at a time with the objective being to build a good
  model of the correlations between feature activities in the previous layer
@@
 humans) and even data associated with other tasks (see ‘transfer learning’,
-below).  These unsupervised learning algorithm produce a new representation
+below).  These unsupervised learning algorithms produce a new representation
 of the data by learning the synaptic weights of a new layer of neurons that
